### 2022-11-21 draft/2018 rejected - Nat Human behavior/GRAPHICALMODEL/Siyu_EEHorizon_2sigma_final.tex
Initial commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template graphical model %
% Author: Tom Lodewyckx    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% 0. Begin document
%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{pst-all}
\usepackage{com.braju.graphicalmodels}
\usepackage{ifpdf}
\ifpdf
\pdfpaperheight=10.5in
\pdfpaperwidth=5.5in
\else
\special{papersize=10.5in,5.5in}
\fi
\catcode`\@=11%
\pagestyle{empty}
\begin{document}
	%\TeXtoEPSepsi
	\psset{unit=14mm,arrowscale=1.5}
	\SpecialCoor

% 1. Grid
%%%%%%%%%

\begin{pspicture}(4,2)(19,8)
%\showgrid

% 2. Nodes
%%%%%%%%%%

%\rput(1,1){\GM@node[NodeOptions]{name}}
%                 \GM@label[LabelOptions]{name}{label}

% dm and dI
\rput(5,3.25){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{isg}$}
\rput(7,3.25){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{isg}$}

% choices
\rput(6,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{c}}\GM@label[offset=-3mm]{c}{$c_{isgr}$}

% subject parameters
\rput(4,8.5){\GM@node[nodeSize=12mm]{A}}\GM@label[offset=-3mm]{A}{$A_{is}$}
\rput(6,8.5){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$b_{is}$}
\rput(8,8.5){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma^n_{is}$}
\rput(10,8.5){\GM@node[nodeSize=12mm]{epsilon}}\GM@label[offset=-2mm]{epsilon}{$\sigma^\epsilon_{is}$}

% hyperparameters
\rput(3,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{i}$}
\rput(4.07,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{i}$}

\rput(5.14,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^b_{i}$}
\rput(6.21,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^b_{i}$}

\rput(7.28,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^n_{i}$}
\rput(8.35,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^n_{i}$}

\rput(9.42,10){\GM@node[nodeSize=12mm]{kEpsilon}}\GM@label[offset=-2mm]{kEpsilon}{$k^\epsilon_{i}$}
\rput(10.49,10){\GM@node[nodeSize=12mm]{lambdaEpsilon}}\GM@label[offset=-2mm]{lambdaEpsilon}{$\lambda^\epsilon_{i}$}

\rput(8,6){\GM@node[nodeSize=12mm]{sigmar}}\GM@label[offset=-4mm]{sigmar}{$n_{isgr}$}
\rput(10,4.5){\GM@node[nodeSize=12mm]{epsilonr}}\GM@label[offset=-3mm]{epsilonr}{$\epsilon_{isg}$}

% differences
%\rput(3,13){\GM@node[nodeSize=12mm]{dmuAOUT}}
%\rput(3,13){\GM@node[nodeSize=10mm]{dmuA}}\GM@label[offset=-3mm]{dmuA}{$\delta\mu^A_{u}$}
%\rput(5,13){\GM@node[nodeSize=12mm]{dkSigmaOUT}}
%\rput(5,13){\GM@node[nodeSize=10mm]{dkSigma}}\GM@label[offset=-3mm]{dkSigma}{$\delta k^\sigma_{u}$}
%\rput(7,13){\GM@node[nodeSize=12mm]{dmubOUT}}
%\rput(7,13){\GM@node[nodeSize=10mm]{dmub}}\GM@label[offset=-3mm]{dmub}{$\delta\mu^b_{u}$}

% 3. Arrows
%%%%%%%%%%%
%
\ncline[arrows=->]{dm}{c}
\ncline[arrows=->]{dI}{c}

\ncline[arrows=->]{A}{c}
\ncline[arrows=->]{sigmar}{c}
\ncline[arrows=->]{b}{c}
\ncline[arrows=->]{epsilonr}{c}

\ncline[arrows=->]{muA}{A}
\ncline[arrows=->]{sigmaA}{A}

\ncline[arrows=->]{kSigma}{sigma}
\ncline[arrows=->]{lambdaSigma}{sigma}

\ncline[arrows=->]{kEpsilon}{epsilon}
\ncline[arrows=->]{lambdaEpsilon}{epsilon}

\ncline[arrows=->]{mub}{b}
\ncline[arrows=->]{sigmab}{b}

\ncline[arrows=->]{sigma}{sigmar}
\ncline[arrows=->]{epsilon}{epsilonr}

%\ncline[arrows=->]{muA}{dmuAOUT}
%\ncline[arrows=->]{kSigma}{dkSigmaOUT}
%\ncline[arrows=->]{mub}{dmubOUT}

% 4. Plates
%%%%%%%%%%%

\rput(4.25,4.25){\GM@plate{4.5}{2.5}{repeat $r=1$:$2$   }}
\rput(3.25,2.5){\GM@plate{7.5}{5.25}{game $g=1$:$G$   }}
\rput(2.25,2){\GM@plate{9.5}{7.25}{subject $s=1$:$S$   }}
\rput(1.25,1.5){\GM@plate{11.5}{9.25}{condition $i=1$:$N$   }}
%\rput(1,1.5){\GM@plate{10}{10}{$u\in\left\{ [1\mbox{ }3], [2\mbox{ }2]\right\}$   }}

% 5. Model specification
%%%%%%%%%%%%%%%%%%%%%%%%

% priors
\rput(12.75,10.5){\pnode{priors}}\GM@label{priors}{Priors}
\rput(12.75,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^A_{i} \sim \mbox{Exponential}(0.01)$}
\rput(12.75,9.5){\pnode{mub}}\GM@label{mub}{$\mu^b_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^b_{i} \sim \mbox{Exponential}(0.01)$}
\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^n_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^n_{i} \sim \mbox{Exponential}(10)$}
\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^\epsilon_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^\epsilon_{i} \sim \mbox{Exponential}(10)$}

% subject level params
\rput(12.75,8){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
\rput(12.75,7.5){\pnode{A}}\GM@label{A}{$A_{is} \sim \mbox{Gaussian}(\mu^A_{i}, \sigma^A_{i})$}
\rput(12.75,7){\pnode{b}}\GM@label{b}{$B_{is} \sim \mbox{Gaussian}(\mu^B_{i}, \sigma^B_{i})$}
\rput(12.75,6.5){\pnode{sigma}}\GM@label{sigma}{$\sigma^n_{is} \sim \mbox{Gamma}(k^n_{i}, \lambda^n_{i})$}
\rput(12.75,6){\pnode{epsilon}}\GM@label{epsilon}{$\sigma^\epsilon_{is} \sim \mbox{Gamma}(k^\epsilon_{i}, \lambda^\epsilon_{i})$}

% game level params
\rput(12.75,5){\pnode{game}}\GM@label{game}{Stimulus driven noise for repeated game}
\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$\epsilon_{isg} \sim \mbox{Logistic}(0, \sigma^\epsilon_{is})$}

% repeated game level params
\rput(12.75,4){\pnode{repeat}}\GM@label{repeat}{Free noise for each game}
\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n_{isgr} \sim \mbox{Logistic}(0, \sigma^n_{is})$}

% choices!
%\rput(9,5.5){\pnode{c}}
%\GM@label{c}{$p \leftarrow }  $}

\rput(12.75,2.5){\pnode{choices}}\GM@label{choices}{Observed choices}
\rput(12.75,2){\pnode{q}}
\GM@label{q}{$\Delta Q_{isgr} \leftarrow
\Delta R_{isg}+A_{is}\Delta I_{isg} + B_{is} + n_{isgr} + \epsilon_{isg}$}
\rput(12.75,1.5){\pnode{c}}
\GM@label{c}{$c_{isgr} \sim \mbox{Bernoulli}
\left( Q_{isgr} > 0 \right)$}

% 6. End document
%%%%%%%%%%%%%%%%%

\end{pspicture}

%\endTeXtoEPS
\end{document}

### 2022-11-21 draft/2020 rejected - JEP General/RandomDeterministicNoise/Title page.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[authoryear, round,sort,comma,numbers]{natbib}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage[font={sf,small}]{caption}
\usepackage{amssymb}
\usepackage{float}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\newcommand*{\email}[1]{%
    \normalsize\href{mailto:#1}{#1}\par
    }

\title{The nature of decision noise in random exploration}

\author[1]{Siyu Wang\thanks{sywangr@email.arizona.edu}}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
	\maketitle

\end{document}

### 2022-11-21 draft/Older drafts/Older version/V1/Internal-External Noise.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
%\usepackage[maxbibnames=99]{biblatex}
%\usepackage{setspace}
%\usepackage{geometry}
\usepackage[font={sf,small}]{caption}
%\usepackage{setspace}
%\usepackage{geometry}
%\usepackage{hyperref}
%\hypersetup{
%    colorlinks,
%    citecolor=black,
%    filecolor=black,
%    linkcolor=black,
%    urlcolor=black
%}
%\geometry{letterpaper}

\usepackage{amssymb}
%\usepackage{epstopdf}
\usepackage{float}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\title{What is the nature of decision noise in random exploration?}

\author[1]{Siyu Wang}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
	\maketitle

	\newpage
	\begin{abstract}

		The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms. Should we explore new options in the hopes of finding a better meal, a better house or a better mate, or should we exploit the options we currently believe to be best? Striking the right balance between exploration and exploitation is hard computational problem and there is significant interest in how humans and other animals make explore-exploit decisions in practice. One particularly effective strategy for solving the explore-exploit dilemma is choice randomization.  In this strategy, the decision process is noisy meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration', can be surprisingly effective in explore-exploit problems and, if implemented correctly, can come close to optimal performance. Recent work suggests that humans actually use random exploration to solve simple explore-exploit problems. Despite this progress a number of questions remain about the nature of random exploration as there are a number of ways in which seemingly stochastic choices could be generated. In one strategy, that we call the “external noise strategy”, participants could rely on stochasticity in the world and allow irrelevant features of the stimulus to drive choice. In another strategy called “internal noise strategy”, people could rely on stochastic processes within their own brains.  In this work, we modified our recently published `Horizon Task' in such a way as to distinguish these two strategies. Using both a model-free and model-based analysis of human behavior we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control. 
	\end{abstract}
	\newpage

	\section*{Introduction}

	Imagine  trying to decide where to go to dinner with a friend ... EXPLORE-EXPLOIT EXAMPLE.  More generally, such decisions are known as explore-exploit decisions ... there's considerable interest in how humans and animals solve it CITE 10 PAPERS INCLUDING WILSON ET AL. 2014...

	%The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms. Should we explore new options in the hopes of finding a better meal, a better house or a better mate, or should we exploit the options we currently believe to be best? Striking the right balance between exploration and exploitation is hard computational problem and there is significant interest in how humans and other animals make explore-exploit decisions in practice.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \cite{}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. IN OUR RESTAURANT EXAMPLE ... In theory, such “random exploration”, is surprisingly effective and, if implemented correctly, can come close to optimal performance \cite{}.

	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \cite{}.  The key manipulation in the task is the horizon condition, i.e. the number of decisions remaining to make. THE IDEA BEHIND THIS MANIPULATION IS THAT PEOPLE SHOULD EXPLORE MORE IN THE LONG HORIZON CONDITION (LINK BACK TO EXAMPLE).  USING SUCH A HORIZON MANIPULATION WE FOUND THAT  people have greater decision noise in the long versus the short horizon condition.

	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown. Of particular interest is whether the adaptive decision noise that is linked to exploration is generated internally, within the brain, or arises externally, in the input from the world. Previous work makes a strong case for both types of noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \cite{}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds \cite{} and the generation and control of this internal noise has been linked to specific neural structures.

	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \cite{}. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely externally driven, then people choices should be identical both times, that is their choices should be consistent since the stimulus is the same both times. Meanwhile, if noise is internally driven, the extent to which their choices are consistent should be determined by the level of the internal noise. By analyzing behavior on this task in both a model-free and model-based manner, we were able to show that, while both types of noise are present in explore-exploit decisions, the contribution of internal noise to random exploration far exceeds that contributed by the stimulus.

	\section*{Results}

	\subsection*{The Repeated-Games Horizon Task}
	We used a modified version of our previously published `Horizon Task' (Figure \ref{fig:task}) to show the influence of internal vs external noise on people's decisions (CITE). THE KEY MANIPULATION WAS REPEATED GAMES SAME DECISION TWICE - LINK INTO EXAMPLE

	MORE SPECIFICALLY, 
	In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

	%Participants played a series of games of a sequential two-armed bandit gambling task. 
	In each game they made multiple decisions between two options. Each option paid out a random reward between 0 and 100 points sampled from a Gaussian distribution. The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game. One of the bandit will always have a higher mean than the other. Participants were instructed to maximize the points earned over the entire task.

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).

	THESE CONDITIONS ALLOW US TO MEASURED DIRECTED AND RNADOM EXPLORATION IN A MODEL-FREE WAY.  REPLICATION OF PREVIOUS RESULTS

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfig.PNG}
			\caption{A key additional manipulation here is repeated games. Each pair of repeated games with identical example trials will appear twice during the experiment. We setup the repeated games such that they are at least 5 games apart from each other. A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of disagreement for repeated games.}
			\label{fig:task2}
		\end{center}
	\end{figure}

	Finally, the crucial additional manipulation in this task is repeated games (Figure \ref{fig:task2}). In each pair of repeated games, the four forced-choice trials were yoked, meaning that on the first free choice trial participants were faced with identical stimuli.  After the first free choice trial, the outcomes on the repeated games were not yoked and the outcomes were sampled independently from the underlying Gaussian distribution.  Not yoking the later trials made it harder for participants to detect repeated games.  In addition, the presentation of repeated games was controlled so that each repeated pair was at least five games away from each other.

	\subsection*{Random exploration is dominated by internal noise}

	In this section we use both model-free and model-based analyses to show that both internal and external noise contribute to the behavioral variability in random exploration. Using the model-based hierarchical Bayesian analysis, we also show that the effect of internal noise is the main source of noise in random exploration.

	\subsubsection*{Model-free analysis}

	In the model-free analysis we asked whether participants' choices were  consistent or inconsistent in the two repetitions of each game.  The idea behind this measure is that purely external noise should lead to consistent choices as the external stimulus is identical both time.  Conversely, internal noise should lead to inconsistent choices both times.  SPELL OUT MORE

	%used choice inconsistency in the repeated games as a measure of internal noise.

	%GOT TO INTRODUCE DIRECTED EXPLORATION - MAYBE BEFORE THIS

	%MOVE THIS PARAGRAPH UP As in our previous paper \cite{}, the influence of directed exploration is measured as the probability of choosing the more informative option, i.e. the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.

	In addition, whether participants make consistent choices in repeated games is used as a model-free measure of internal noise, since in repeated trials, external noise should be identical on both trials. So only internal noise can differ and drive the choice inconsistency. The degree to which people make consistent choices in repeated trials can reflect the internal noise.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/Modelfree.png}
			\caption{Both direct (P $<$ 0.001) and random exploration (P $<$ 0.001 for both [1 3] and [2 2]) increases with horizon.(A,B) Choice consistency in both [1 3] and [2 2] condition increases with horizon(P([1 3]) $<$ 0.001, P([2 2])$<$0.001).(C)}
			\label{fig:mf}
		\end{center}
	\end{figure}

	%In line with previous findings, both directed and random exploration increase with horizon (Figure \ref{fig:mf}). In addition, 
	Choice inconsistency between repeated games was non-zero in both horizon conditions, suggesting that not all of the noise was stimulus driven.  In addition, choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (Figure \ref{fig:mf}), suggesting that at least some of the horizon dependent noise is internal.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/modelfree_intext.png}
			\caption{For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into the data we computed predicted values of the choice inconsistency for the purely external and purely internal noise cases.  For the purely externally driven noise case, then people should make the exact same decisions each time in repeated games, which corresponds to the light blue line in Figure \ref{fig:mf2}. If noise is however purely internally driven, then there should be no stimulus-dependent noise component such that people should in principle treat repeated games independently, hence 
	\begin{equation*}
	\begin{split}
	P(\mbox{consistency}) &= P(\mbox{low mean})^2 + P(\mbox{high mean})^2\\
	&= P(\mbox{low mean})^2 + (1-P(\mbox{low mean}))^2\\ \mbox{hence},\quad P(\mbox{inconsistency}) &= 1 - P(\mbox{consistency})
	\end{split}
	\end{equation*},

	we can make the prediction of p(disagree) if there is only internal noise. However, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction, suggesting that both external and internal noise are present in driving this choice inconsistency.

	Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon, however we can not draw any obvious conclusions about whether external noise is horizon dependent or not.

	\subsubsection*{Model-based analysis}
	To more precisely quantify the size of internal and external noise in this task, we turned to model fitting.

	\paragraph{Overview of model}
	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the information bias between the two bandits. To model participants’ choices on this first free-choice trial, we assume that they make decisions by computing the difference of value $\Delta Q$ between the two options:

	$$\Delta Q= \Delta R+A \Delta I+b+n_{ext}+n_{int}$$

	Where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial (for simplicity we define $\Delta I$ to be +1, -1 or 0: +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0). $n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.

	The subject-and-condition-specific parameters are: the spatial bias, b, which determines the extent to which participants prefer the option on the right, the information bonus A, which controls the level of directed exploration, ϵ denotes the external, external noise, which is identical on the repeat versions of each game, and n denotes internal noise, which is uncorrelated between repeat plays. For repeated games, the set of forced-choice trials are exact the same, so the external noise should be the same, but the free decision noise may be different. This is exactly how we distinguish external noise from internal noise. In symbolic terms, for repeated games i and j,  $\sigma_{ext}^i=\sigma_{ext}^j$  and $\sigma_{int}^i \neq \sigma_{int}^j$.
	The decision model assumes that participants choose right if $\Delta Q > 0$, otherwise the participant will choose the left option.

	We used hierarchical Bayesian analysis to fit the parameters (see Figure \ref{fig:model} for an graphical representation of the model in the style of [5]). In particular, we fit values of the information bonus A, spatial bias B, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon and uncertainty condition. The mean and standard deviation of information bonus A and spatial bias B are sampled from a Gaussian prior and an exponential prior respectively. The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter k and λ for the gamma distribution are sampled from exponential priors. At the game level, we sampled one stimulus driven noise for each pair of repeated games and we sampled two free decision noises for each individual game in the pair.  The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS \cite{}.

	\begin{table}[H]
		\small
		\begin{center}
			\begin{tabular}{|c|c|c|c|}
				\hline
				Parameter &
				Horizon dependent? &
				Uncertainty dependent? &
				Game dependent? \\
				\hline
				information bonus, $A$ &
				yes &
				n/a &
				no \\
				\hline
				spatial bias, $B$ &
				yes &
				yes &
				no \\
				\hline
				external decision noise, $\sigma_{ext}$ &
				yes & 
				yes &
				no \\
				\hline
				internal decision noise, $\sigma_{int}$ &
				yes & 
				yes &
				yes \\
				\hline
			\end{tabular}
		\end{center}
		\caption{Model parameters. }
		\label{table1}
	\end{table}%

	%Before presenting the results of the model-based analysis we begin with a brief overview of the most salient points of the model.  A full description of the model can be found in the Methods and code to implement the model can be found in the Supplementary Material.

	%Conceptually, the model breaks the explore-exploit choice down into two components: a learning component, in which participants estimate the mean payoff of each option from the rewards they see, and a decision component, in which participants use this estimated payoff to guide their choice.  The learning component assumes that participants compute an estimate of the average payoff for each slot machine, $R^i_t$, using a simple delta rule update equation (based on a Kalman filter \cite{Kalman1960-uo}, see Methods):
	%\begin{equation}
	%R^i_{t+1}= R^i_t + \alpha_t^i (r_t - R^i_t)
	%\end{equation}
	%where $r_t$ is the reward on trial $t$ and $\alpha_t^i$ is the time-varying learning rate that determines the extent to which the prediction error, $(r_t - R^i_t)$, updates the estimate of the mean of bandit $i$. The learning process is described by three free parameters: the initial value of the estimated payoff, $R_0$, and two learning rates, the initial learning rate, $\alpha_1$, and the asymptotic learning rate, $\alpha_{\inf}$, which together describe the evolution of the actual learning rate, $\alpha_t$, over time. For simplicity, we assume that these parameters are independent of horizon and uncertainty condition (Table 1).

	%The decision component of the model assumes that participants choose between the two options (left and right) probabilistically according to 
	%\begin{equation}
	%	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
	%\end{equation}
	%where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition). The decision process is described by three free parameters: the information bonus $A$, the spatial bias $B$, and the decision noise $\sigma$.  We estimate separate values of the decision parameters for each horizon and (since the information bonus is only used in the [1 3] condition) separate values of only the bias and decision noise for each uncertainty condition.

	%Overall, subject's behavior in each session (vertex vs RFPC stimulation) is described by 13 free parameters (Table \ref{table1}): three describing learning ($R_0$, $\alpha_1$ and $\alpha_{\infty}$) and 10 describing the decision process ($A$ in the two horizon conditions, $B$ and $\sigma$ in the four horizon-x-uncertainty conditions). These 13 parameters were fit to each subject in each stimulation condition using a hierarchical Bayesian approach \cite{Lee2014-yh} (see Methods).

	\paragraph{Model fitting results} 
	%Posterior distributions over the group-level means are shown in the left column of Figure \ref{fig:mb1}, while posteriors over the TMS-related change in parameters are shown in the right column.  Both columns suggest a selective effect of RFPC stimulation on the information bonus in horizon 6.

	Only internal noise increases  with horizon (Figure \ref{fig:modelbased}). 
	%In both horizon 1 and horizon 6, internal noises are significantly higher than internal noises (for both horizon 1 and horizon 6, P $<$ 0.001).

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.5\textwidth]{figures/bayes.png}
			\caption{The change of external and internal decision noises between horizons. The errorbars are confidence intervals at $95\%$. Only internal noise increases significantly with horizon.}
			\label{fig:modelbased}
		\end{center}
	\end{figure}

	Figure 3 also demonstrates the existence of both external noise and internal noise. Hierarchical Bayesian analysis provides us a way to quantify the magnitude of external and internal noise respectively, from the result, internal noise is the main source of random exploration.

	\section*{Discussion}
	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment.  We find horizon dependent changes in both internal and external sources of noise, but that the effect of internal noise is much greater. 
	One limitation of this work is in the interpretation of the different noises as being internal and external. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control all stimuli the participant received which, for example, would vary based on where they were looking.  Thus, our estimate of external noise is likely a lower bound. Likewise, our estimate of internal noise is likely an upper bound as these “missing” sources of stimulus driven noise would be interpreted as internal noise in our model. Despite this, the sheer magnitude of the difference between internal and external noise (internal noise is 2-3 times the size of external noise, Figure 3), suggests that our interpretation may be safe as an awful lot of external noise would have to be explained by variables not relevant to the task.  
	The horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain. In addition, this work is consistent with a recent report from Ebitz et al [7] in which behavioral variability in an “explore” state was also tied to internal rather than external sources of noise. Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory [4] suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.  While there is some evidence that NE plays a role in explore-exploit behavior [8], this link has been questioned [9].

	\section*{Methods}
	\subsection*{Participants}
	78 Participants (37 male, 41 female, ages 17-25) from the UA subject pool, played the task.  27 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 51 for the analysis.
	%31 healthy right-handed, adult volunteers (19 female, 12 male; ages 19-32). An initial sample size of 16 was chosen based on two studies using a very similar cTBS design that stimulated lateral FPC (Costa et. al. 2011; 2013) and this was augmented to 31 on the basis of feedback from reviewers. 5 participants (5 female, 0 male) were excluded from the analysis due to chance-level performance in both experimental sessions. One (female) participant failed to return for the second (vertex stimulation condition) session and is excluded from the model-free analyses but not the model-based analyses as that can handle missing data more gracefully. Thus our final data set consisted of 25 participants (13 female, 12 male, ages 19-32) with complete data and one participant (female, aged 20) with data from the RFPC session only.

	%All participants were informed about potential risks connected to TMS and signed a written consent. The study was approved by University of Social Sciences and Humanities ethics committee.

	%\subsection*{Procedure}
	%There were two experimental TMS sessions and a preceding MRI session.  On the first session T1 structural images were acquired using a 3T Siemens TRIO scanner. The scanning session lasted up to 10 minutes. Before the first two sessions, participants filled in standard safety questionnaires regarding MRI scanning and TMS. During the experimental sessions, prior to the stimulation participants went through 16 training games to get accustomed to the task. Afterwards, resting motor thresholds were obtained and the stimulation took place. Participants began the main task immediately after stimulation. The two experimental sessions were performed with an intersession interval of at least 5 days. The order of stimulation conditions was counterbalanced across subjects. All sessions took place at Nencki Institute of Experimental Biology in Warsaw.

	%\subsection*{Limitations}

	%Defining stimulation target by peak coordinates based on findings from previous studies did not allow to account for individual differences in either brain anatomy or the impact of TMS on brain networks \cite{Gratton2013-wz}. However, a study by Volman and colleagues \cite{Volman2011-fz} that used the same theta-burst protocol on the left frontopolar cortex has shown bilateral inhibitory effects on blood perfusion in the frontal pole. This suggests that both right and left parts of the frontopolar cortex might have been inhibited in our experiment, which is consistent with imaging results indicating bilateral involvement of the frontal pole in exploratory decisions.

	\subsection*{Task}
	The task was a modified version of the Horizon Task \cite{}. As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points. Participants were informed that in every game one of the bandits always has a higher mean reward than the other. Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20. One of the means was always equal to either 40 or 60 and the second was set accordingly. The order of games was randomized. Mean sizes and order of presentation were counterbalanced. The number of games participants play depends on how well they perform, the better they perform, the sooner the task will end. On average, participants played 128 games and the whole task lasted between 39 and 50 minutes (mean 43.4 minutes).

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfig1.PNG}
			\caption{Timeline of a game}
			\label{fig:modelbased}
		\end{center}
	\end{figure}

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game. For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome. From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 500ms. The task was programmed using Psychtoolbox in MATLAB. \cite{}.

	%Finally, the random seeds were not perfectly controlled between subjects.  The first 16 subjects ran the task with identical random seeds and thus all 16 saw the same sequence of forced-choice trials in both vertex and RFPC sessions. For the remaining subjects the random seed was unique for each subject and each session, thus these subjects had unique a series of forced-choice trials for each session.  Despite this limitation we saw no evidence of different behavior across the two groups.

	\subsection*{Data and code}
	%Behavioral data as well as Matlab code to recreate the main figures from this paper can be found on the Dataverse website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.

	\subsection*{Model-based analysis}
	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \cite{} that was modified to differentiate internal and external noises. In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.

	%\subsubsection*{Decision component}
	%Once the payoffs of each option, $R^i_t$, have been estimated from the outcomes of the forced-choice trials, the model makes a decision using a simple logistic choice rule:
	%\begin{equation}
	%	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
	%\end{equation}
	%where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition). The three free parameters of the decision process are: the information bonus, $A$, the spatial bias, $B$, and the decision noise $\sigma$.  We assume that these three decision parameters can take on different values in the different horizon and uncertainty conditions (with the proviso that $A$ is undefined in the [2 2] information condition since $\Delta I = 0$). Thus the decision component of the model has 10 free parameters ($A$ in the two horizon conditions, and $B$ and $\sigma$ in the 4 horizon x uncertainty conditions).  Directed exploration is then quantified as the change in information bonus with horizon, while random exploration is quantified as the change in decision noise with horizon.

	\subsection*{Model Fitting}
	\subsubsection*{Hierarchical Bayesian Model}
	Each subject's behavior is described by 4 free parameters. These parameters are: the information bonus, $A$, in both horizon conditions, the spatial bias, $B$, in the four horizon x uncertainty conditions, and external decision noise, $n_{int}$, and internal decision noise, $n_{ext}$ in the four horizon x uncertainty conditions (Table \ref{tab:pars2}, Figure \ref{fig:model}).

	Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \cite{}.  In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure. The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.

	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}. For example, we assume that the information bonus, $A^{is}$, for each information condition x horizon condition $i$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$. These prior parameters are sampled in turn from  their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.

	\begin{table}[H]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.1)}		\\
			\hline
			spatial bias, $B_{is}$ 
			& $B_{is} \sim $  Gaussian($\mu_i^{B}$, $\sigma_i^{B}$) 
			& $\theta_{i}^{B} = (\mu_i^{B}, \sigma_i^{B}) $
			& \specialcell{
				$\mu_i^{B} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{B} \sim $ Exponential(0.1)}		\\
			\hline
			external decision noise, $\epsilon_{isg}$ 
			& $\epsilon_{isg} \sim $  Gaussian(0, $\sigma_{is}^{ext}$) 
			& $\theta_{i}^{ext} = (k_i^{ext}, \lambda_i^{ext}) $
			& \specialcell{
				$k_i^{ext} \sim $ Gaussian(0.1) \\ 
				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
			\hline
			internal decision noise, $\sigma_{isgr}$ 
			& $\sigma_{isg} \sim $  Gaussian(0, $\sigma_{is}^{int}$) 
			& $\theta_{i}^{int} = (k_i^{int}, \lambda_i^{int}) $
			& \specialcell{
				$k_i^{int} \sim $ Gaussian(0.1) \\ 
				$\lambda_i^{int} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors.}
		\label{tab:pars2}	
	\end{table}

	\subsubsection*{Model fitting using MCMC}
	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \cite{} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags/).  This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.

	In particular we used 4 independent Markov chains to generate 40000 samples from the posterior distribution over parameters (10000 samples per chain).  Each chain had a burn in period of 1000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.  Convergence of the Markov chains was confirmed post hoc by eye. Code and data to replicate our analysis and reproduce our Figures is provided as part of the Supplementary Materials.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/Figure3.pdf}
			\caption{Hierarchical Bayesian model}
			\label{fig:model}
		\end{center}
	\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.

	\bibliographystyle{apalike}
	\bibliography{references}

\end{document}

### 2022-11-21 draft/Older drafts/Older version/V3/Internal-External Noise.tex
Initial commit

--- 2022-11-21 draft/Older drafts/Older version/V1/Internal-External Noise.tex
@@
 \documentclass[12pt]{article}
+\usepackage{hyperref}
+\usepackage[authoryear, round,sort,comma,numbers]{natbib}
 \usepackage{times}
@@
 \usepackage{amsmath}
+
 %\usepackage[maxbibnames=99]{biblatex}
@@
 	\begin{abstract}
-
-		The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms.
-Should we explore new options in the hopes of finding a better meal, a better house or a better mate, or should we exploit the options we currently believe to be best?
-Striking the right balance between exploration and exploitation is hard computational problem and there is significant interest in how humans and other animals make explore-exploit decisions in practice.
+		The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms, from everyday life decisions like deciding for a meal to important life decisions like finding a life partner.
+It is computationally a hard problem to find the right balance between exploration and exploitation and hence there is significant interest in how humans and other animals solve the explore-exploit dilemma.
 One particularly effective strategy for solving the explore-exploit dilemma is choice randomization.
@@
 
-	Imagine  trying to decide where to go to dinner with a friend ...
-EXPLORE-EXPLOIT EXAMPLE.
-More generally, such decisions are known as explore-exploit decisions ...
-there's considerable interest in how humans and animals solve it CITE 10 PAPERS INCLUDING WILSON ET AL.
-2014...
+	Imagine trying to decide where to go to dinner with a friend, you can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened a few days ago right across the street to the other restaurant which may end up being your newly favorite.
+More generally, such decisions are known as explore-exploit decisions.
+The explore-exploit decision refers to deciding between exploiting the best known option so far, like going to your old favorite restaurant, and exploring other options for potential better decisions in the future, like trying the new restaurant.
+There's considerable interest in how humans and animals solve it.\citep{eegittins74, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}
 
-	%The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms.
-Should we explore new options in the hopes of finding a better meal, a better house or a better mate, or should we exploit the options we currently believe to be best?
-Striking the right balance between exploration and exploitation is hard computational problem and there is significant interest in how humans and other animals make explore-exploit decisions in practice.
+	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}.
+In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
+In our restaurant example, your restaurant decision does not always depend on the quality of the food, you are very likely to go to the new restaurant if you happen to see another old friend walking right in, or you wait until the last moment to make a split second decision about where to go as if you flipped a mental coin in your head and decide to go to your old favorite if it's heads up.
+In theory, such “random exploration”, is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eebridle90}.
 
-	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \cite{}.
-In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-IN OUR RESTAURANT EXAMPLE ...
-In theory, such “random exploration”, is surprisingly effective and, if implemented correctly, can come close to optimal performance \cite{}.
-
-	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \cite{}.
+	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.
 The key manipulation in the task is the horizon condition, i.e.
 the number of decisions remaining to make.
-THE IDEA BEHIND THIS MANIPULATION IS THAT PEOPLE SHOULD EXPLORE MORE IN THE LONG HORIZON CONDITION (LINK BACK TO EXAMPLE).
-USING SUCH A HORIZON MANIPULATION WE FOUND THAT  people have greater decision noise in the long versus the short horizon condition.
+The idea behind this manipulation is that people should explore more in the long horizon condition.
+If you are leaving town tomorrow for vacation, you probably want to go to your old favorite to guarantee a good last meal, but if you are not going anywhere, you would be more likely to try the new restaurant.
+Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.
 
@@
 Of particular interest is whether the adaptive decision noise that is linked to exploration is generated internally, within the brain, or arises externally, in the input from the world.
+In the restaurant example, an old friend walking by would be a source of external noise, but flipping a mental coin would be an internal noise.
 Previous work makes a strong case for both types of noise being relevant to behavior.
-For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \cite{}.
-Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds \cite{} and the generation and control of this internal noise has been linked to specific neural structures.
+For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}.
+Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures.
 
-	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \cite{}.
+	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \citep{wilson2014}.
 To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice.
@@
 
+	%\begin{figure}[H]
+	%	\begin{center}
+	%		\includegraphics[width=0.7\textwidth]{figs/bayes_s.png}
+	%			\caption[Subject level estimated mean of external and internal noise variance.]{Subject level estimated mean of external and internal noise variance.
+Both internal and external noise increase with horizon, internal noise is dominant.}
+	%			\label{fig:mb3}
+	%		\end{center}
+	%	\end{figure}
+
 	\section*{Results}
@@
 	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of our previously published `Horizon Task' (Figure \ref{fig:task}) to show the influence of internal vs external noise on people's decisions (CITE).
-THE KEY MANIPULATION WAS REPEATED GAMES SAME DECISION TWICE - LINK INTO EXAMPLE
+	We used a modified version of our previously published `Horizon Task' \citep{wilson2014} to show the influence of internal vs external noise on people's decisions (Figure \ref{fig:taskfig}).
+The key manipulation was to use repeated games to let people make the same decision twice.
+In the restaurant example, if your decision is mainly driven by external noise, then a few months later when you see the friend walking in front of you into the restaurant, you are very likely to make the same decision and follow him into that restaurant and says hi.
+However, if your decision is mainly driven by internal noise, then next time you make a split-second decision about where to go, you are equally likely to go to the other restaurant.
 
-	MORE SPECIFICALLY, 
+	More specifically, we look at the contribution of external and internal noise by providing with participants the same decision problem twice in the task.
 In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
@@
 
-	%Participants played a series of games of a sequential two-armed bandit gambling task. 
 	In each game they made multiple decisions between two options.
-Each option paid out a random reward between 0 and 100 points sampled from a Gaussian distribution.
+Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution.
 The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game.
@@
 After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
+In unequal uncertainty condition, people are more likely to choose the option that they know less about - the more informative option - to explicitly explore that option more.
+This type of information driven exploration is known as directed exploration.
 
-	THESE CONDITIONS ALLOW US TO MEASURED DIRECTED AND RNADOM EXPLORATION IN A MODEL-FREE WAY.  REPLICATION OF PREVIOUS RESULTS
+	As in our previous paper \citep{wilson2014}, These conditions allow us to measure directed and random exploration in a model-free way.
+Directed exploration is measured as the probability of choosing the more informative option in [1 3] condition whereas random exploration is measured as the probability of choosing the low mean option in [2 2] condition.
+In line with previous results \citep{wilson2014}, we showed that both directed and random exploration increase with horizon.
+Both direct (t(59)=6.88, p $<$ 0.001) and random exploration (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]) increases with horizon.
+(Figure \ref{fig:modelfree}, A,B).
+Directed exploration is considered to be driven by information bias and random exploration is considered to be driven by decision noise, in this work, we are investigating which source of decision noise, external vs internal, drives random exploration.
+
+	%the influence of directed exploration is measured as the probability of choosing the more informative option, i.e.
+the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.
 
@@
 We setup the repeated games such that they are at least 5 games apart from each other.
-A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of disagreement for repeated games.}
-			\label{fig:task2}
+A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
+			\label{fig:taskfig}
 		\end{center}
@@
 
-	Finally, the crucial additional manipulation in this task is repeated games (Figure \ref{fig:task2}).
+	Finally, the crucial additional manipulation in this task is repeated games (Figure \ref{fig:taskfig}).
 In each pair of repeated games, the four forced-choice trials were yoked, meaning that on the first free choice trial participants were faced with identical stimuli.
@@
 	\subsubsection*{Model-free analysis}
-
 	In the model-free analysis we asked whether participants' choices were  consistent or inconsistent in the two repetitions of each game.
 The idea behind this measure is that purely external noise should lead to consistent choices as the external stimulus is identical both time.
-Conversely, internal noise should lead to inconsistent choices both times.
-SPELL OUT MORE
-
-	%used choice inconsistency in the repeated games as a measure of internal noise.
-
-	%GOT TO INTRODUCE DIRECTED EXPLORATION - MAYBE BEFORE THIS
-
-	%MOVE THIS PARAGRAPH UP As in our previous paper \cite{}, the influence of directed exploration is measured as the probability of choosing the more informative option, i.e.
-the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.
+Conversely, internal noise should lead to independent choices, and hence possible inconsistent choices both times.
+More specifically, we look at the proportion of times that a participant make inconsistent decisions in repeated game.
+Choice inconsistency is defined as the proportions of inconsistent choices for repeated games.
+(See Figure \ref{fig:taskfig})
 
@@
 The degree to which people make consistent choices in repeated trials can reflect the internal noise.
+Since choice inconsistency in both [1 3](t(59) = 5.10, P([1 3]) $<$ 0.001) and [2 2](t(59) = 6.08, P([2 2])$<$0.001) condition increase with horizon (Figure \ref{fig:modelfree} C), this is a behavioral evidence that internal noise increases with horizon.
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/Modelfree.png}
-			\caption{Both direct (P $<$ 0.001) and random exploration (P $<$ 0.001 for both [1 3] and [2 2]) increases with horizon.(A,B) Choice consistency in both [1 3] and [2 2] condition increases with horizon(P([1 3]) $<$ 0.001, P([2 2])$<$0.001).(C)}
-			\label{fig:mf}
+			\includegraphics[width=\textwidth]{figures/modelfree.png}
+			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions. }
+			\label{fig:modelfree}
 		\end{center}
@@
 
-	%In line with previous findings, both directed and random exploration increase with horizon (Figure \ref{fig:mf}). In addition, 
 	Choice inconsistency between repeated games was non-zero in both horizon conditions, suggesting that not all of the noise was stimulus driven.
-In addition, choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (Figure \ref{fig:mf}), suggesting that at least some of the horizon dependent noise is internal.
+In addition, choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (Figure \ref{fig:modelfree}), suggesting that at least some of the horizon dependent noise is internal.
 
@@
 			\includegraphics[width=\textwidth]{figures/modelfree_intext.png}
-			\caption{For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
+			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
+For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
 Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently.
@@
 	To gain more quantitative insight into the data we computed predicted values of the choice inconsistency for the purely external and purely internal noise cases.
-For the purely externally driven noise case, then people should make the exact same decisions each time in repeated games, which corresponds to the light blue line in Figure \ref{fig:mf2}.
-If noise is however purely internally driven, then there should be no stimulus-dependent noise component such that people should in principle treat repeated games independently, hence
+For the purely externally driven noise case, then people should make the exact same decisions each time in repeated games, so $P(\mbox{inconsistent})$ should be 0, which corresponds to the light blue line in Figure \ref{fig:mf2} in [1 3] and [2 2] respectively.
+If noise is however purely internally driven, then there should be no stimulus-dependent noise component such that people should in principle treat repeated games independently, and the extent to which they are consistent with themselves in repeated games can be predicted by choosing the low mean option twice or choosing the high mean option twice, which corresponds to the red line in Figure \ref{fig:mf2}.
+
 	\begin{equation*}
 	\begin{split}
-	P(\mbox{consistency}) &= P(\mbox{low mean})^2 + P(\mbox{high mean})^2\\
-	&= P(\mbox{low mean})^2 + (1-P(\mbox{low mean}))^2\\ \mbox{hence},\quad P(\mbox{inconsistency}) &= 1 - P(\mbox{consistency})
+	P(\mbox{consistent}) &= P(\mbox{low mean})^2 + P(\mbox{high mean})^2\\
+	&= P(\mbox{low mean})^2 + (1-P(\mbox{low mean}))^2\\ \mbox{hence},\quad P(\mbox{inconsistent}) &= 1 - P(\mbox{consistent})
 	\end{split}
-	\end{equation*},
+	\end{equation*}
 
-	we can make the prediction of p(disagree) if there is only internal noise.
-However, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction, suggesting that both external and internal noise are present in driving this choice inconsistency.
-
-	Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon, however we can not draw any obvious conclusions about whether external noise is horizon dependent or not.
+	However, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Figure \ref{fig:mf2}), suggesting that both external and internal noise are present in driving this choice inconsistency.
+Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.
 
@@
 	To more precisely quantify the size of internal and external noise in this task, we turned to model fitting.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noises.
+In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.
 
@@
 	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the information bias between the two bandits.
-To model participants’ choices on this first free-choice trial, we assume that they make decisions by computing the difference of value $\Delta Q$ between the two options:
+To model participants’ choices on this first free-choice trial, we assume that they make decisions by computing the difference of value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
+Specifically, we write
 
-	$$\Delta Q= \Delta R+A \Delta I+b+n_{ext}+n_{int}$$
+	\begin{equation}
+	\Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
+	\end{equation}
 
-	Where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial (for simplicity we define $\Delta I$ to be +1, -1 or 0: +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0).
+	Where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
+For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
 $n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.
 
-	The subject-and-condition-specific parameters are: the spatial bias, b, which determines the extent to which participants prefer the option on the right, the information bonus A, which controls the level of directed exploration, ϵ denotes the external, external noise, which is identical on the repeat versions of each game, and n denotes internal noise, which is uncorrelated between repeat plays.
-For repeated games, the set of forced-choice trials are exact the same, so the external noise should be the same, but the free decision noise may be different.
+	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game; and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays and changes every game.
+
+	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the external noise, $n_{ext}$, should be the same while the internal noise, $n_{int}$ may be different.
 This is exactly how we distinguish external noise from internal noise.
-In symbolic terms, for repeated games i and j,  $\sigma_{ext}^i=\sigma_{ext}^j$  and $\sigma_{int}^i \neq \sigma_{int}^j$.
-	The decision model assumes that participants choose right if $\Delta Q > 0$, otherwise the participant will choose the left option.
+In symbolic terms, for repeated games i and j,  $n_{ext}^i=n_{ext}^j$  and $n_{int}^i \neq n_{int}^j$.
 
-	We used hierarchical Bayesian analysis to fit the parameters (see Figure \ref{fig:model} for an graphical representation of the model in the style of [5]).
-In particular, we fit values of the information bonus A, spatial bias B, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon and uncertainty condition.
-The mean and standard deviation of information bonus A and spatial bias B are sampled from a Gaussian prior and an exponential prior respectively.
-The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter k and λ for the gamma distribution are sampled from exponential priors.
-At the game level, we sampled one stimulus driven noise for each pair of repeated games and we sampled two free decision noises for each individual game in the pair.
-The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS \cite{}.
+	\paragraph{Model fitting}
+	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
+In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon.
+The mean and standard deviation of information bonus $A$ and spatial bias $B$ are sampled from a Gaussian prior and an exponential prior respectively.
+The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter $k$ and $\lambda$ for the gamma distribution are sampled from exponential priors.
 
-	\begin{table}[H]
-		\small
-		\begin{center}
-			\begin{tabular}{|c|c|c|c|}
-				\hline
-				Parameter &
-				Horizon dependent? &
-				Uncertainty dependent? &
-				Game dependent? \\
-				\hline
-				information bonus, $A$ &
-				yes &
-				n/a &
-				no \\
-				\hline
-				spatial bias, $B$ &
-				yes &
-				yes &
-				no \\
-				\hline
-				external decision noise, $\sigma_{ext}$ &
-				yes & 
-				yes &
-				no \\
-				\hline
-				internal decision noise, $\sigma_{int}$ &
-				yes & 
-				yes &
-				yes \\
-				\hline
-			\end{tabular}
-		\end{center}
-		\caption{Model parameters. }
-		\label{table1}
-	\end{table}%
+	To capture the idea that external noise should be identical on repeated games, we sampled one value of the external noise, $n_{ext}$ for each pair of repeated games.
+Conversely, because internal noise is expected to change between games we sampled two values of the internal noise, $n_{int}^1$ and $n_{int}^2$, i.e.
+one for each individual game.
+
+	The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS software \citep{jags, matjags}.
 
@@
 	\paragraph{Model fitting results}
-	%Posterior distributions over the group-level means are shown in the left column of Figure \ref{fig:mb1}, while posteriors over the TMS-related change in parameters are shown in the right column.
-Both columns suggest a selective effect of RFPC stimulation on the information bonus in horizon 6.
 
-	Only internal noise increases  with horizon (Figure \ref{fig:modelbased}). 
+	%Group-level estimates for the variances of internal and external noise are shown in Figure \ref{fig:modelbased}A.
+While both variances are non-zero, this shows that the internal noise is much larger than external noise.
+This horizon-based change is probed further in Figure XXXB in which we plot the posterior distributions over the change in internal and external noise with horizon.
+This clearly shows that only internal noise varies with horizon (STATS XXX\% of the samples above zero for internal noise, XXX\% of samples above zero for external noise).
+
+	Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:mb1}.
+While both variances are non-zero, this shows that the internal noise is much larger than external noise.
+This horizon-based change is probed further in Figure \ref{fig:mb2} in which we plot the posterior distributions over the change in internal and external noise with horizon.
+As is clear from this plot, internal noise increases dramatically with horizon whereas there is a slight increase of external noise with horizon as well (0 just falls in the critical region at level of 5\% for external noise but lies right at the border, although significant, it's not as strong as the internal noise).
+This clearly shows that internal noise dominates and significantly varies with horizon  (100\% of the samples above zero for internal noise, 96.33\% of samples above zero for external noise).
+
+	%Furthermore, we plotted the subject level estimates of both internal and external noises in a similar fashion of the model-free plot of choice-inconsistency for comparison (Figure \ref{fig:mb3}).
+We showed again that both internal and external noises increase significantly with horizon.
+But in terms of magnitude, internal noise is what' really dominant here (See also Figure \ref{fig:rcurvepd6} in Appendix \ref{ch:appendicies:reward}).
+Subject-level estimates also allow us to look at the relationship between internal and external noises, and we showed that internal noise correlates with external noise significantly in both horizon conditions ($R^2 = 0.32, p < 0.001$ in horizon 1, $R^2 = 0.24, p < 0.001$ in horizon 6, see Figure \ref{fig:corrintext} in Appendix \ref{ch:appendix:corrintext}).
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoiseh.png}
+			\caption[Posterior distributions over the group-level means of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise variance.
+Both internal and external noises are nonzero, and internal noise has a much greater magnitude.}
+			\label{fig:mb1}
+		\end{center}
+	\end{figure}
+
+	%LEFT EACH TYPE OF NOISE IN EACH HORIZON
+	%RIGHT IS DISTRIBUTION OVER CHANGE IN NOISE WITH HORIZON
+
+	%Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:modelbased}.
+As is clear from this plot, only internal noise increases  with horizon.
 	%In both horizon 1 and horizon 6, internal noises are significantly higher than internal noises (for both horizon 1 and horizon 6, P $<$ 0.001).
@@
 		\begin{center}
-			\includegraphics[width=0.5\textwidth]{figures/bayes.png}
-			\caption{The change of external and internal decision noises between horizons. The errorbars are confidence intervals at $95\%$. Only internal noise increases significantly with horizon.}
-			\label{fig:modelbased}
+			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
+			\caption[Posterior distributions over the group-level means of the change of external and internal noise variance.]{Posterior distributions over the group-level means of the change of external and internal noise variance.
+Internal noise increases significantly with horizon, external noise increases with horizon as well.}
+			\label{fig:mb2}
 		\end{center}
 	\end{figure}
-
-	Figure 3 also demonstrates the existence of both external noise and internal noise.
-Hierarchical Bayesian analysis provides us a way to quantify the magnitude of external and internal noise respectively, from the result, internal noise is the main source of random exploration.
 
@@
 We find horizon dependent changes in both internal and external sources of noise, but that the effect of internal noise is much greater.
+
 	One limitation of this work is in the interpretation of the different noises as being internal and external.
@@
 Thus, our estimate of external noise is likely a lower bound.
-Likewise, our estimate of internal noise is likely an upper bound as these “missing” sources of stimulus driven noise would be interpreted as internal noise in our model.
-Despite this, the sheer magnitude of the difference between internal and external noise (internal noise is 2-3 times the size of external noise, Figure 3), suggests that our interpretation may be safe as an awful lot of external noise would have to be explained by variables not relevant to the task.
+Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of stimulus driven noise would be interpreted as internal noise in our model.
+Despite this, the sheer magnitude of the difference between internal and external noise (internal noise is 2-3 times the size of external noise, Figure \ref{fig:mb3}), suggests that our interpretation may be safe as an awful lot of noise would have to be explained by variables not relevant to the task.
+
 	The horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain.
-In addition, this work is consistent with a recent report from Ebitz et al [7] in which behavioral variability in an “explore” state was also tied to internal rather than external sources of noise.
-Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory [4] suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
-While there is some evidence that NE plays a role in explore-exploit behavior [8], this link has been questioned [9].
+This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
+In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.
+
+	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
+While there is some evidence that NE plays a role in explore-exploit behavior \citep{jepma2012}, this link has been questioned \citep{nieuwenhuis05}.
+
+	More generally, our finding that internal noise dominates behavioral varibility over external noise, is consistent with findings of \cite{drugowitsch16}.
+In particular these authors show that choice suboptimality might arise from imperfections in mental inference rather than in peripheral stages such as sensory processing and response selection.
+This suggests that internal noise may come from computational errors in computing the correct strategy, especially in long horizon conditions.
+	%MORE GENERALLY, OUT FINDING THAT INTERNAL DOMINATES BEHAVIORAL VARIABILITY OVER EXTERNAL NOISE, IS CONSISTENT WITH FINDINGS OF DRUGOWITSCH ET AL.  IN PARTICULAR THESE AUTHORS SHOW THAT 
+	%	LINK IN THE Jan Drugowitsch stuff on variability.  ALEX POUGET STUFF. INTERNAL NOISE COMES FROM COMPUTATIONAL ERRORS IN COMPUTING THE CORRECT STRATEGY IN LONG  HORIZON CONDITIONS.
+
+	%THREE IDEAS
+	%NOISE ADDED TO THE PROCESS
+	%ATTENTION = DECREASE SIGNAL VS INCREASE NOISE
+	%NOT NOISY BUT WRONG
+	%PLANNING WITH FEW SAMPLES DEEP EXPLORATION
+
+	There are many ways that internal noise can be implemented to cause the computational error and ultimately drive different behaviors.
+As far as behavior, choice variability would go up as long as the signal-noise ratio goes up.
+But neurally there are two ways that signal-noise ratio can go up.
+It can be that people are just paying less attention in a long horizon game and devaluing the signal, it can also be that it's really the noise that goes up.
+We can not tell these two apart easily just with behavior.
+One way that may get at it at a behavioral level is to fit both the choices and the reaction times using drift-diffusion model, then we are able to get whether it's the drift rate that goes down (signal), or it's the noise that goes up (noise).
+It would also be interesting to see how external and internal noises are coded in the DDM model.
+Another way to get at this question is to do an EEG study which directly looks at how noise is represented in the electrical signals and see how it changes with horizon.
+
+	Another possible source of noise is the imperfectness of our model.
+It's not that noise is added to the process, it's simply that the model is wrong and only accounted for part of behavior.
+In our case, the stimulus people get has two components - reward, and the order of rewards (in which order two bandits are played), our model only accounts for the difference in mean rewards between the two bandit, the pattern of choices and the individual reward could have an influence on people's behavior.
+But this only explains for external noise, but not internal.
+Since in repeated games, people are facing the same stimuli, even if the model is suboptimal, it's suboptimal in the same way for repeated pairs, so this idea of 'not noisy but wrong' only explains possible sources for external noise but not internal.
+
+	Also, the theory of deep exploration is a possible mechanism through which decision noise can be generated when planning with fewer samples.
+It also has a potential to explain how noise is adapted to horizon.
+If implemented well with a well-designed task, two hypothesis under this model can be tested: Does the model quantitatively explain the change of decision noise across horizons?
+Do the number of samples each participants use correlate with their individual reaction times since in theory the more sampling you do, the more planning you do, the longer it should take you to make the decision.
+
+	In summary, current works shows that both external and internal noises are guilty for the behavioral variability, but internal noise dominates in driving random exploration.
+Further work needs to be done to understand how noices are generated or implemented in our brains.
 
@@
 	\subsection*{Participants}
-	78 Participants (37 male, 41 female, ages 17-25) from the UA subject pool, played the task.
-27 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 51 for the analysis.
+
+	A total of 84 participants from the UA subject pool participated in the experiment.
+60 out of 84 participants are included for analysis.
+4 were excluded because of young age, 20 were excluded on the basis of performance (using the same exclusion criterion as in \citep{wilson2014}) leaving 60 for the analysis.\\
+
+	No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities.
+(\ref{appendix:participants})
+
+	%There are no differences of p(low mean) or p(inconsistent) across studies. No differences were found between ages, races and ethnicities.
+
 	%31 healthy right-handed, adult volunteers (19 female, 12 male; ages 19-32).
@@
 	\subsection*{Task}
-	The task was a modified version of the Horizon Task \cite{}.
+	The task was a modified version of the Horizon Task \citep{wilson2014}.
+In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
+In each game they made multiple decisions between two options.
+Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution.
+The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game.
+One of the bandit will always have a higher mean than the other.
+Participants were instructed to maximize the points earned over the entire task.
+To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.
+The number of games participants play depends on how well they perform, the better they perform, the sooner the task will end.
+On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 10.6 and 29.2 minutes (mean 19.7 minutes).
+%(\ref{appendixb})
+
 	As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points.
-Participants were informed that in every game one of the bandits always has a higher mean reward than the other.
 Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20.
 One of the means was always equal to either 40 or 60 and the second was set accordingly.
+Participants were informed that in every game one of the bandits always has a higher mean reward than the other.
 The order of games was randomized.
 Mean sizes and order of presentation were counterbalanced.
-The number of games participants play depends on how well they perform, the better they perform, the sooner the task will end.
-On average, participants played 128 games and the whole task lasted between 39 and 50 minutes (mean 43.4 minutes).
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfig1.PNG}
-			\caption{Timeline of a game}
-			\label{fig:modelbased}
-		\end{center}
-	\end{figure}
-
 Each game consisted of 5 or 10 choices.
@@
 During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit.
-The score feedback was presented for 500ms.
-The task was programmed using Psychtoolbox in MATLAB.
-\cite{}.
+The score feedback was presented for 300ms.
+The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.
+(See Figure \ref{fig:taskfig1})
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/taskfig1.PNG}
+			\caption{Timeline of a game}
+			\label{fig:taskfig1}
+		\end{center}
+	\end{figure}
+
+	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
+We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
+The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
+After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
+(See Figure \ref{fig:taskfig2})
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
+			\caption{Task conditions}
+			\label{fig:taskfig2}
+		\end{center}
+	\end{figure}
 
@@
 	\subsection*{Model-based analysis}
-	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \cite{} that was modified to differentiate internal and external noises.
+	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noises.
 In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.
-
-	%\subsubsection*{Decision component}
-	%Once the payoffs of each option, $R^i_t$, have been estimated from the outcomes of the forced-choice trials, the model makes a decision using a simple logistic choice rule:
-	%\begin{equation}
-	%	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
-	%\end{equation}
-	%where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition).
-The three free parameters of the decision process are: the information bonus, $A$, the spatial bias, $B$, and the decision noise $\sigma$.
-We assume that these three decision parameters can take on different values in the different horizon and uncertainty conditions (with the proviso that $A$ is undefined in the [2 2] information condition since $\Delta I = 0$).
-Thus the decision component of the model has 10 free parameters ($A$ in the two horizon conditions, and $B$ and $\sigma$ in the 4 horizon x uncertainty conditions).
-Directed exploration is then quantified as the change in information bonus with horizon, while random exploration is quantified as the change in decision noise with horizon.
-
-	\subsection*{Model Fitting}
 	\subsubsection*{Hierarchical Bayesian Model}
 	Each subject's behavior is described by 4 free parameters.
-These parameters are: the information bonus, $A$, in both horizon conditions, the spatial bias, $B$, in the four horizon x uncertainty conditions, and external decision noise, $n_{int}$, and internal decision noise, $n_{ext}$ in the four horizon x uncertainty conditions (Table \ref{tab:pars2}, Figure \ref{fig:model}).
+These parameters are: the information bonus $A$ and the spatial bias $b$ in both horizon conditions, the external decision noise, $n_{int}$, and internal decision noise, $n_{ext}$ (Table \ref{para1}, Figure \ref{fig:model}).
 
-	Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \cite{}.
+	Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.
 In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure.
@@
 	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}.
-For example, we assume that the information bonus, $A^{is}$, for each information condition x horizon condition $i$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$.
-These prior parameters are sampled in turn from  their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.
+For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$.
+These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.
+	\begin{table}[H]
+		\small
+		\begin{center}
+			\begin{tabular}{|c|c|c|c|}
+				\hline
+				Parameter &
+				Horizon dependent? &
+				Uncertainty dependent? &
+				Repeated game dependent? \\
+				\hline
+				information bonus, $A$ &
+				yes &
+				n/a &
+				no \\
+				\hline
+				spatial bias, $B$ &
+				yes &
+				no &
+				no \\
+				\hline
+				external decision noise, $\sigma_{ext}$ &
+				yes & 
+				yes &
+				no \\
+				\hline
+				internal decision noise, $\sigma_{int}$ &
+				yes & 
+				yes &
+				yes \\
+				\hline
+			\end{tabular}
+		\end{center}
+		\caption{Model parameters. }
+		\label{para1}
+	\end{table}%
 
@@
 			Parameter & Prior & Hyperparameters & Hyperpriors \\
+			\hline
 			information bonus, $A_{is}$ 
@@
 	\subsubsection*{Model fitting using MCMC}
-	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \cite{} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags/).
+	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags).
 This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.
 
-	In particular we used 4 independent Markov chains to generate 40000 samples from the posterior distribution over parameters (10000 samples per chain).
-Each chain had a burn in period of 1000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
+	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).
+Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
 Convergence of the Markov chains was confirmed post hoc by eye.
-Code and data to replicate our analysis and reproduce our Figures is provided as part of the Supplementary Materials.
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/Figure3.pdf}
+			\includegraphics[width=\textwidth]{figures/hbm.pdf}
 			\caption{Hierarchical Bayesian model}
 			\label{fig:model}
+		\end{center}
+	\end{figure}
+
+	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}
+
+	Here we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and re-fit the simulated the choices to see whether we can recover the parameters that we use to simulate the choices.
+The true values of the parameters are plotted against the recovered value of the same parameter in Figure \ref{fig:pararecover1} and Figure \ref{fig:pararecover2} respectively.
+
+	From these two figures, we can see that the recovery for information bonus is consistenly well for both horizon 1 and horizon 6.
+The recovery for internal noise is consistently well but better at horizon 1 than horizon 6.
+This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6.
+The same phenomenon can be seen for external noises too that it fits better in horizon 1 than in horizon 6.
+The recovery for external noise is fine but not as good as it is for internal noise.
+This is because we have only half as many trials as for internal noise since we are only generating one sample of external noise for each repeated game pair.
+At last, we don't have a good recovery of bias in horizon 1, but this doesn't hurt our result since the biases are all very close to 0.
+We don't have a good recovery of bias in horizon 6 either, again, we are not really interested in biases here and the magnitude of bias is also small compared to other parameters.
+Overall, we are able to recover both external and internal noises using our model to a satisfactory extent.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/paramrecover_1.png}
+			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 1 games}
+			\label{fig:pararecover1}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
+			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
+			\label{fig:pararecover2}
 		\end{center}
@@
 
-	\bibliographystyle{apalike}
-	\bibliography{references}
+	\bibliographystyle{plainnat}
+	%\bibliographystyle{unsrt}
+	\bibliography{Refs/refs}
+
+	% add the Bibliography to the Table of Contents
+	\cleardoublepage
+	\ifdefined\phantomsection
+	\phantomsection  % makes hyperref recognize this section properly for pdf link
+	\else
+	\fi
+	\addcontentsline{toc}{chapter}{Bibliography}
 

### 2022-11-21 draft/Older drafts/Siyu#final/Internal-external noise Siyu#final/Internal-External Noise.tex
Initial commit

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[authoryear, round,sort,comma,numbers]{natbib}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}

%\usepackage[maxbibnames=99]{biblatex}
%\usepackage{setspace}
%\usepackage{geometry}
\usepackage[font={sf,small}]{caption}
%\usepackage{setspace}
%\usepackage{geometry}
%\usepackage{hyperref}
%\hypersetup{
%    colorlinks,
%    citecolor=black,
%    filecolor=black,
%    linkcolor=black,
%    urlcolor=black
%}
%\geometry{letterpaper}

\usepackage{amssymb}
%\usepackage{epstopdf}
\usepackage{float}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\title{What is the nature of decision noise in random exploration?}

\author[1]{Siyu Wang}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
	\maketitle

	\newpage
	\begin{abstract}
	[150 words for Nature Human Behavior, currently 162]
	Human decision making appears to contains components that are inherently random. While this behavioral variability has traditionally been seen as noise, recent work suggests that random choices may actually be adaptive. An example arises when deciding between exploring unknown options or exploiting options we know well. Theory shows that a little randomness in explore-exploit decisions is remarkably effective, leading to greater rewards overall. Meanwhile, experiments show that people use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore. Despite this progress, the nature of adaptive decision noise is unknown and it is unclear whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world? Here we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control.

	%From everyday decisions, like deciding where to dine, to major life milestones, like deciding whom to marry, many decisions involve a tradeoff between exploring options that are unknown and exploiting options we know well.

	%Recently we have shown that humans make simple explore-exploit choices using two distinct strategies. One of these strategies, known as random exploration, involves choice randomization

	%While making optimal explore-exploit choices is a hard computational problem, we have recently shown that humans make effective explore-exploit choices using heuristic strategies. One particularly simple strategy is `random exploration'.  In this strategy, the decision process is noisy meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.

	%Despite this progress, questions remain about the nature of random exploration in particular, is the noise in behavior generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world?

	%as there are a number of ways in which seemingly stochastic choices could be generated. In one strategy, the `external noise strategy,' participants could rely on stochasticity in the world and allow irrelevant features of the stimulus to drive choice. In another strategy called `internal noise strategy,' people could rely on stochastic processes within their own brains.

	%In this work, we modified our recently published `Horizon Task' in such a way as to distinguish these two strategies. Using both a model-free and model-based analysis of human behavior we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control. 
	\end{abstract}
	\newpage

	%In theory, such `random exploration', can be surprisingly effective in explore-exploit problems and, if implemented correctly, can come close to optimal performance.
	%Moreover, recent work suggests that humans actually use random exploration in simple explore-exploit tasks.

	\section*{Introduction}

	Imagine trying to decide where to go to dinner with a friend. You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing. Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.

	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.  The key manipulation in the task is the horizon condition, i.e. the number of decisions remaining to make. Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards. For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new place. Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.

	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown. Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain. In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants. Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Previous work makes a strong case for both types of noise being relevant to behavior. For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}. Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.

	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \citep{wilson2014}. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely externally driven, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times. Meanwhile, if noise is internally driven their choices should be inconsistent, since the internal noise is different both times. By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the contribution of internal noise to random exploration far exceeds that contributed by the stimulus.

	%\begin{figure}[H]
	%	\begin{center}
	%		\includegraphics[width=0.7\textwidth]{figs/bayes_s.png}
	%			\caption[Subject level estimated mean of external and internal noise variance.]{Subject level estimated mean of external and internal noise variance. Both internal and external noise increase with horizon, internal noise is dominant.}
	%			\label{fig:mb3}
	%		\end{center}
	%	\end{figure}

	\section*{Results}

	\subsection*{The Repeated-Games Horizon Task}
	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of internal vs external noise on people's decisions (Figure \ref{fig:taskfig}). In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards. Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.

	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit, in the form of four example plays distributed either unequally between bandits (i.e. 1 play of one bandit, 3 plays of the other) or equally (2 plays each). These example plays allow us to manipulate exactly what people know about each option before they make their first choice. Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.

	These `repeated games' are the key manipulation in this paper and allow us to distinguish between external and internal sources of noise.  Specifically, if noise is externally driven, then choices on repeated games should be consistent. Conversely if noise is internally driven, then choices on repeated games should be inconsistent.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{A key additional manipulation here is repeated games. Each pair of repeated games with identical example trials will appear twice during the experiment. We setup the repeated games such that they are at least 5 games apart from each other. A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

	\subsection*{Both behavioral variability and information seeking increase with horizon}

	%This further allowed us to have people make the exact same

	%The key manipulation was to use repeated games to let people make the same decision twice. In the restaurant example, if your decision is mainly driven by external noise, then a few months later when you see the friend walking in front of you into the restaurant, you are very likely to make the same decision and follow him into that restaurant to say hi. However, if your decision is mainly driven by internal noise, then next time you make a split-second decision about where to go, you are equally likely to go to the other restaurant.

	%More specifically, we look at the contribution of external and internal noise by providing with participants the same decision problem twice in the task. In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

	%In each game they made multiple decisions between two options. Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution. The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game. One of the bandit will always have a higher mean than the other. Participants were instructed to maximize the points earned over the entire task.

	%The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions). In unequal uncertainty condition, people are more likely to choose the option that they know less about - the more informative option - to explicitly explore that option more. This type of information driven exploration is known as directed exploration.

	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}. As in our previous work, we find evidence for two types of exploration in the Horizon Task.  Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.

	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition. This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]).  Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition. Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(59)=6.88, p $<$ 0.001).

	%These conditions allow us to measure directed and random exploration in a model-free way. Directed exploration is measured as the probability of choosing the more informative option in [1 3] condition whereas random exploration is measured as the probability of choosing the low mean option in [2 2] condition. In line with previous results \citep{wilson2014}, we showed that both directed and random exploration increase with horizon. Both direct (t(59)=6.88, p $<$ 0.001) and random exploration (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]) increases with horizon. (Figure \ref{fig:modelfree}, A,B).   Directed exploration is considered to be driven by information bias and random exploration is considered to be driven by decision noise, in this work, we are investigating which source of decision noise, external vs internal, drives random exploration.

	%the influence of directed exploration is measured as the probability of choosing the more informative option, i.e. the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree}
		\end{center}
	\end{figure}

	%Finally, the crucial additional manipulation in this task is repeated games (Figure \ref{fig:taskfig}). In each pair of repeated games, the four forced-choice trials were yoked, meaning that on the first free choice trial participants were faced with identical stimuli.  After the first free choice trial, the outcomes on the repeated games were not yoked and the outcomes were sampled independently from the underlying Gaussian distribution.  Not yoking the later trials made it harder for participants to detect repeated games.  In addition, the presentation of repeated games was controlled so that each repeated pair was at least five games away from each other.

	\subsection*{Model-free analysis shows that random exploration may involve both internal and external noise}

	%In this section we use both model-free and model-based analyses to show that both internal and external noise contribute to the behavioral variability in random exploration. Using the model-based hierarchical Bayesian analysis, we also show that the effect of internal noise is the main source of noise in random exploration.

	%\subsubsection*{Model-free analysis}
	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.  The idea behind this measure is that purely external noise should lead to consistent choices as the external stimulus is identical both times. Conversely, purely internal noise should lead to independent choices, and hence more inconsistent choices both times.

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven. In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.

	%As is clear from the plots, in both the unequal ([1 3], panel A) and equal ([2 2], panel B) conditions,

	%In addition, whether participants make consistent choices in repeated games is used as a model-free measure of internal noise, since in repeated trials, external noise should be identical on both trials. So only internal noise can differ and drive the choice inconsistency. The degree to which people make consistent choices in repeated trials can reflect the internal noise.  Since choice inconsistency in both [1 3](t(59) = 5.10, P([1 3]) $<$ 0.001) and [2 2](t(59) = 6.08, P([2 2])$<$0.001) condition increase with horizon (Figure \ref{fig:modelfree} C), this is a behavioral evidence that internal noise increases with horizon.

	%Choice inconsistency between repeated games was non-zero in both horizon conditions, suggesting that not all of the noise was stimulus driven.  In addition, choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (Figure \ref{fig:modelfree}), suggesting that at least some of the horizon dependent noise is internal.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely external and purely internal noise cases.  For purely external noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case. For purely internal noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
	\begin{equation*}
	\begin{split}
	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean})(1-p(\mbox{low mean}))
	\end{split}
	\end{equation*}
	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(59) = 5.10, p $<$ 0.001 for horizon 1, t(59) = 5.65, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 3.49, p $<$ 0.001 for horizon 1, t(59) = 3.31, p $<$ 0.001 for horizon 6. Behavior is different from pure external noise prediction in both [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency. Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.

	\subsection*{Model-based analysis shows that random exploration is dominated by internal noise}

	To more precisely quantify internal and external noise, we turned to model fitting. We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noises. In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.

	\subsubsection*{Overview of model}
	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the information bias between the two bandits. To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.

	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game; and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays and changes every game.

	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the external noise, $n_{ext}$, should be the same while the internal noise, $n_{int}$ may be different. This is exactly how we distinguish external noise from internal noise. In symbolic terms, for repeated games $i$ and $j$,  $n_{ext}^i=n_{ext}^j$  and $n_{int}^i \neq n_{int}^j$.

	\subsubsection*{Model fitting}
	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}). In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon. Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.   %The mean and standard deviation of information bonus $A$ and spatial bias $B$ are sampled from a Gaussian prior and an exponential prior respectively. The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter $k$ and $\lambda$ for the gamma distribution are sampled from exponential priors.

	%To capture the idea that external noise should be identical on repeated games, we sampled one value of the external noise, $n_{ext}$ for each pair of repeated games.  Conversely, because internal noise is expected to change between games we sampled two values of the internal noise, $n_{int}^1$ and $n_{int}^2$, i.e. one for each individual game.

	%The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS software \citep{jags, matjags}.

	%Before presenting the results of the model-based analysis we begin with a brief overview of the most salient points of the model.  A full description of the model can be found in the Methods and code to implement the model can be found in the Supplementary Material.

	%Conceptually, the model breaks the explore-exploit choice down into two components: a learning component, in which participants estimate the mean payoff of each option from the rewards they see, and a decision component, in which participants use this estimated payoff to guide their choice.  The learning component assumes that participants compute an estimate of the average payoff for each slot machine, $R^i_t$, using a simple delta rule update equation (based on a Kalman filter \cite{Kalman1960-uo}, see Methods):
	%\begin{equation}
	%R^i_{t+1}= R^i_t + \alpha_t^i (r_t - R^i_t)
	%\end{equation}
	%where $r_t$ is the reward on trial $t$ and $\alpha_t^i$ is the time-varying learning rate that determines the extent to which the prediction error, $(r_t - R^i_t)$, updates the estimate of the mean of bandit $i$. The learning process is described by three free parameters: the initial value of the estimated payoff, $R_0$, and two learning rates, the initial learning rate, $\alpha_1$, and the asymptotic learning rate, $\alpha_{\inf}$, which together describe the evolution of the actual learning rate, $\alpha_t$, over time. For simplicity, we assume that these parameters are independent of horizon and uncertainty condition (Table 1).

	%The decision component of the model assumes that participants choose between the two options (left and right) probabilistically according to 
	%\begin{equation}
	%	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
	%\end{equation}
	%where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition). The decision process is described by three free parameters: the information bonus $A$, the spatial bias $B$, and the decision noise $\sigma$.  We estimate separate values of the decision parameters for each horizon and (since the information bonus is only used in the [1 3] condition) separate values of only the bias and decision noise for each uncertainty condition.

	%Overall, subject's behavior in each session (vertex vs RFPC stimulation) is described by 13 free parameters (Table \ref{table1}): three describing learning ($R_0$, $\alpha_1$ and $\alpha_{\infty}$) and 10 describing the decision process ($A$ in the two horizon conditions, $B$ and $\sigma$ in the four horizon-x-uncertainty conditions). These 13 parameters were fit to each subject in each stimulation condition using a hierarchical Bayesian approach \cite{Lee2014-yh} (see Methods).

	\subsubsection*{Model fitting results}

	%Group-level estimates for the variances of internal and external noise are shown in Figure \ref{fig:modelbased}A.  While both variances are non-zero, this shows that the internal noise is much larger than external noise. This horizon-based change is probed further in Figure XXXB in which we plot the posterior distributions over the change in internal and external noise with horizon.  This clearly shows that only internal noise varies with horizon (STATS XXX\% of the samples above zero for internal noise, XXX\% of samples above zero for external noise).

	Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:mb1}. Consistent with our model-free results, we see that both internal and external noise variances are non-zero and that internal noise is about 2-3 times larger than the external noise. In addition, we find that internal noise increases dramatically with horizon (M = 4.55, p $<$ 0.001) whereas there is only a slight increase of external noise (M = 1.78, p = 0.009).  Taken together, these results suggest that random exploration is dominated by internal noise.

	%internal noise changes

	%appear to increase with horizon (Figure \ref{fig:mb2}, 100\% of the samples above zero for internal noise, 96.33\% of samples above zero for external noise), the change in internal noise is much larger (STATS).

	%With regard to the change in noise between horizons,

	%While both variances are non-zero, this shows that the internal noise is much larger than external noise. This horizon-based change is probed further in Figure \ref{fig:mb2} in which we plot the posterior distributions over the change in internal and external noise with horizon. As is clear from this plot, internal noise increases dramatically with horizon whereas there is a slight increase of external noise with horizon as well (0 just falls in the critical region at level of 5\% for external noise but lies right at the border, although significant, it's not as strong as the internal noise). This clearly shows that internal noise dominates and significantly varies with horizon  (100\% of the samples above zero for internal noise, 96.33\% of samples above zero for external noise).

	%Furthermore, we plotted the subject level estimates of both internal and external noises in a similar fashion of the model-free plot of choice-inconsistency for comparison (Figure \ref{fig:mb3}). We showed again that both internal and external noises increase significantly with horizon. But in terms of magnitude, internal noise is what' really dominant here (See also Figure \ref{fig:rcurvepd6} in Appendix \ref{ch:appendicies:reward}). Subject-level estimates also allow us to look at the relationship between internal and external noises, and we showed that internal noise correlates with external noise significantly in both horizon conditions ($R^2 = 0.32, p < 0.001$ in horizon 1, $R^2 = 0.24, p < 0.001$ in horizon 6, see Figure \ref{fig:corrintext} in Appendix \ref{ch:appendix:corrintext}).

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb1}
		\end{center}
	\end{figure}

	%LEFT EACH TYPE OF NOISE IN EACH HORIZON
	%RIGHT IS DISTRIBUTION OVER CHANGE IN NOISE WITH HORIZON

	%Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:modelbased}.  As is clear from this plot, only internal noise increases  with horizon. 
	%In both horizon 1 and horizon 6, internal noises are significantly higher than internal noises (for both horizon 1 and horizon 6, P $<$ 0.001).

%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
%			\caption[]{}			
%			\label{fig:mb2}
%		\end{center}
%	\end{figure}

	\section*{Discussion}
	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment. Using a version of the Horizon Task with repeated games, we found that evidence for both types of noise in explore-exploit decisions. In addition, we see that both internal and external noise increase with horizon, but that the horizon effect is much larger for internal noise.  Taken together our results suggest that random exploration, i.e. the use and adaptation of decision noise to drive exploration, is driven primarily by internal noise.

	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being internal and external. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose. Thus, our estimate of external noise is likely a lower bound. Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of external noise would be interpreted as internal noise in our model. Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb1}, where internal noise is 2-3 times the size of external noise.

	Taken at face value, the horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}. In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise. Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

	More generally, our finding that internal noise dominates behavioral variability over external noise, is consistent with findings of \cite{drugowitsch16}. In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection. This suggests that most noise in behavior is generated internally and that this may arise from  computational errors in computing the correct strategy. In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.
	%MORE GENERALLY, OUT FINDING THAT INTERNAL DOMINATES BEHAVIORAL VARIABILITY OVER EXTERNAL NOISE, IS CONSISTENT WITH FINDINGS OF DRUGOWITSCH ET AL.  IN PARTICULAR THESE AUTHORS SHOW THAT 
	%	LINK IN THE Jan Drugowitsch stuff on variability.  ALEX POUGET STUFF. INTERNAL NOISE COMES FROM COMPUTATIONAL ERRORS IN COMPUTING THE CORRECT STRATEGY IN LONG  HORIZON CONDITIONS.

	%THREE IDEAS
	%NOISE ADDED TO THE PROCESS
	%ATTENTION = DECREASE SIGNAL VS INCREASE NOISE
	%NOT NOISY BUT WRONG
	%PLANNING WITH FEW SAMPLES DEEP EXPLORATION

	%There are many ways that internal noise can be implemented to cause the computational error and ultimately drive different behaviors. As far as behavior, choice variability would go up as long as the signal-noise ratio goes up. But neurally there are two ways that signal-noise ratio can go up. It can be that people are just paying less attention in a long horizon game and devaluing the signal, it can also be that it's really the noise that goes up. We can not tell these two apart easily just with behavior. One way that may get at it at a behavioral level is to fit both the choices and the reaction times using drift-diffusion model, then we are able to get whether it's the drift rate that goes down (signal), or it's the noise that goes up (noise). It would also be interesting to see how external and internal noises are coded in the DDM model. Another way to get at this question is to do an EEG study which directly looks at how noise is represented in the electrical signals and see how it changes with horizon.

	%Another possible source of noise is the imperfectness of our model. It's not that noise is added to the process, it's simply that the model is wrong and only accounted for part of behavior. In our case, the stimulus people get has two components - reward, and the order of rewards (in which order two bandits are played), our model only accounts for the difference in mean rewards between the two bandit, the pattern of choices and the individual reward could have an influence on people's behavior. But this only explains for external noise, but not internal. Since in repeated games, people are facing the same stimuli, even if the model is suboptimal, it's suboptimal in the same way for repeated pairs, so this idea of 'not noisy but wrong' only explains possible sources for external noise but not internal.

	%Also, the theory of deep exploration is a possible mechanism through which decision noise can be generated when planning with fewer samples. It also has a potential to explain how noise is adapted to horizon. If implemented well with a well-designed task, two hypothesis under this model can be tested: Does the model quantitatively explain the change of decision noise across horizons? Do the number of samples each participants use correlate with their individual reaction times since in theory the more sampling you do, the more planning you do, the longer it should take you to make the decision.

	%In summary, current works shows that both external and internal noises are guilty for the behavioral variability, but internal noise dominates in driving random exploration. Further work needs to be done to understand how noises are generated or implemented in our brains.

	\section*{Methods}
	\subsection*{Participants}

	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment. 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 60 for the main analysis. Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
			\label{fig:modelfree2}
		\end{center}
	\end{figure}

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration. For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games. Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently. }
			\label{fig:mf22}
		\end{center}
	\end{figure}
	\newpage
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation. Both internal and external noises are nonzero, and internal noise has a much greater magnitude. Posterior distributions over the group-level means of the change of external and internal noise standard deviation. Internal noise increases significantly with horizon, external noise increases with horizon as well.}
			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
			\label{fig:mb12}
		\end{center}
	\end{figure}
	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities. (\ref{appendix:participants})

	\subsection*{Task}
	The task was a modified version of the Horizon Task \citep{wilson2014}. In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. In each game they made multiple decisions between two options. Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution. The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game. One of the bandits always had a higher mean than the other. Participants were instructed to maximize the points earned over the entire task. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?  I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME

	As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points. Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20. One of the means was always equal to either 40 or 60 and the second was set accordingly. Participants were informed that in every game one of the bandits always has a higher mean reward than the other. The order of games was randomized. Mean sizes and order of presentation were counterbalanced.

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game. For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome. From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 300ms. The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}. (See Figure \ref{fig:taskfigb})

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
			\caption{Time line of a horizon 6 game.  The game begins with four forced choice trials in which participants are instructed which option to play.   After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
			I WILL SEND YOU THE POWERPOINTS TOO}
			\label{fig:taskfigb}
		\end{center}
	\end{figure}

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions). (See Figure \ref{fig:taskfigb})

%	\begin{figure}[H]
%		\begin{center}
%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
%			\caption{Task conditions}
%			\label{fig:taskfig2}
%		\end{center}
%	\end{figure}

	%Finally, the random seeds were not perfectly controlled between subjects.  The first 16 subjects ran the task with identical random seeds and thus all 16 saw the same sequence of forced-choice trials in both vertex and RFPC sessions. For the remaining subjects the random seed was unique for each subject and each session, thus these subjects had unique a series of forced-choice trials for each session.  Despite this limitation we saw no evidence of different behavior across the two groups.

	\subsection*{Data and code}
	%Behavioral data as well as Matlab code to recreate the main figures from this paper can be found on the Dataverse website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.

	\subsection*{Model-based analysis}
	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noise. In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.

	\subsubsection*{Hierarchical Bayesian Model}

	To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.

	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game; and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays and changes every game.

	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the external noise, $\sigma_{ext}$, and the standard deviation of the internal noise, $\sigma_{int}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}). Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.  In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure. The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.

	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}. For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$. These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.
	%\begin{table}[h]
	%	\small
	%	\begin{center}
	%		\begin{tabular}{|c|c|c|c|}
	%			\hline
	%			Parameter &
	%			Horizon dependent? &
	%			Uncertainty dependent? &
	%			Repeated game dependent? \\
	%			\hline
	%			information bonus, $A$ &
	%			yes &
	%			n/a &
	%			no \\
	%			\hline
	%			spatial bias, $B$ &
	%			yes &
	%			no &
	%			no \\
	%			\hline
	%			external decision noise, $\sigma_{ext}$ &
	%			yes & 
	%			yes &
	%			no \\
	%			\hline
	%			internal decision noise, $\sigma_{int}$ &
	%			yes & 
	%			yes &
	%			yes \\
	%			\hline
	%		\end{tabular}
	%	\end{center}
	%	\caption{Model parameters and their dependence on horizon, uncertainty condition and repeated games }
	%	\label{para1}
	%\end{table}%

	\begin{table}[h]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			\hline
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
			\hline
			spatial bias, $b_{is}$ 
			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
			& \specialcell{
				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
			\hline
			deviation of external noise, $\sigma^{ext}_{isg}$ 
			& $\sigma_{i} \sim $  Gamma($k_i^{ext}$, $\lambda_{i}^{ext}$) 
			& $\theta_{i}^{ext} = (k_i^{ext}, \lambda_i^{ext}) $
			& \specialcell{
				$k_i^{ext} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
			\hline
			deviation of internal noise, $\sigma^{int}_{isgr}$ 
			& $\sigma_{is} \sim $  Gamma($k_i^{int}$, $\lambda_{i}^{int}$) 
			& $\theta_{i}^{int} = (k_i^{int}, \lambda_i^{int}) $
			& \specialcell{
				$k_i^{int} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{int} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
		\label{tab:pars2}	
	\end{table}

	\subsubsection*{Model fitting using MCMC}
	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags). This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.

	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).  Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.  Convergence of the Markov chains was confirmed {\it post hoc} by eye.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
			\caption{Hierarchical Bayesian model}
			\label{fig:model}
		\end{center}
	\end{figure}

	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}

	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.  In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated the choices to see whether we can recover the parameters.

	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}. As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C, D).  The recovery for the noise parameters, $\sigma_{ext}$ and $\sigma_{int}$, is slightly better for horizon 1 than horizon 6. This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6. In addition we see better recovery for internal noise than external noise because we effectively have half as many trials for external noise since we are only generating one sample of external noise for each repeated game pair. Overall, we are able to recover both external and internal noises using our model to a satisfactory extent.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/precover.png}
			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.  REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.  DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.  RED LINES CHANGE TO DOTTED BLACK LINES.  
			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
			%USE ADDABCS FUNCTION TO ADD ABCs!
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
	%\begin{figure}[H]-
	%	\begin{center}
	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
	%		\label{fig:pararecover2}
	%	\end{center}
	%\end{figure}
	%References
	%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
	%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.

	\bibliographystyle{plainnat}
	%\bibliographystyle{unsrt}
	\bibliography{Refs/refs}

	% add the Bibliography to the Table of Contents
	\cleardoublepage
	\ifdefined\phantomsection
	\phantomsection  % makes hyperref recognize this section properly for pdf link
	\else
	\fi
	\addcontentsline{toc}{chapter}{Bibliography}

\end{document}

### 2022-11-21 codes/EEplot_2noise_PureRanDet.m

@@ -1,0 +1,48 @@
+function plt = EEplot_2noise_PureRanDet(plt, sp, sp2)
+names = {'Random noise only - \sigma_{ran}','','Deterministic noise only - \sigma_{det}',...
+    ''};
+
+ymax = 2.5;
+plt.figure(2,2,'istitle',0);
+plt.setup_pltparams('fontsize_leg', 20, 'linewidth', 2);
+plt.setfig('xlim', {[-1 22],[-3 12], [-1 22], [-3 12]}, 'ylim', [0 ymax], ...
+    'ytick', [0:0.5:ymax], 'xtick',{0:4:20,-3:3:12,0:4:20,-3:3:12});
+color = {'AZblue','AZred'};
+fns = {'NoiseRan', 'NoiseDet'};
+stepsize = 0.02;
+xbins = [-10:stepsize:30];
+plt.setfig('color', color, 'legend', {'h = 1','h = 6'},...
+    'title', {'Random noise - \sigma_{ran}','Deterministic noise - \sigma_{det}','',''}, ...
+    'xlabel', {'','','noise standard deviation [points]', 'noise standard deviation [points]'}, ...
+    'ylabel', names, 'legloc', 'NorthEast');
+for i = 1:2
+    fn = fns{i};
+    plt.new;
+    for hi = 1:2
+        td = sp.(fn);
+        td = squeeze(td(:,:,hi));
+        td = reshape(td, 1, []);
+        tl = hist(td, xbins)/(length(td)*stepsize);
+        plt.lineplot(tl, [], xbins);
+    end
+    hold on;
+    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
+    hold off;
+end
+for i = 1:2
+    fn = fns{i};
+    plt.new;
+    for hi = 1:2
+        td = sp2.(fn);
+        td = squeeze(td(:,:,hi));
+        td = reshape(td, 1, []);
+        tl = hist(td, xbins)/(length(td)*stepsize);
+        plt.lineplot(tl, [], xbins);
+    end
+    hold on;
+    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
+    hold off;
+end
+plt.update;
+plt.save(['pure_ran_det_simu_hyperprior']);
+end

### 2022-11-21 codes/EEplot_2noise_parameter_recovery_hyperprior.m

@@ -1,0 +1,57 @@
+function plt = EEplot_2noise_parameter_recovery_hyperprior(plt, sp, sp2)
+names = {'Random noise - \sigma_{ran}','Random noise - \sigma_{ran}',...
+    'Deterministic noise - \sigma_{det}','Deterministic noise - \sigma_{det}'};
+plt.figure(2,2,'title');
+plt.setup_pltparams('fontsize_leg', 20, 'linewidth', 2);
+plt.setfig('xlim', {[-1 22],[-3 12], [-1 22], [-3 12]}, 'ylim', [0 1.2], ...
+    'ytick', [0:0.2:1.2], 'xtick',{0:4:20,-3:3:12,0:4:20,-3:3:12});
+color = {{'AZblue','AZblue50','AZred','AZred50'},{'black','black50'}, ...
+    {'AZblue','AZblue50','AZred','AZred50'},{'black','black50'}};
+fns = {'NoiseRan', 'NoiseDet'};
+stepsize = 0.02;
+xbins = [-10:stepsize:30];
+lg1 = {'original, h = 1','recovered, h = 1','original, h = 6', 'recovered, h = 6'};
+lg2 = {'original, change', 'recovered change'};
+plt.setfig('color', color, 'legend', ...
+    {lg1,lg2,lg1,lg2},...
+    'ylabel', {'posterior density','','posterior density',''}, ...
+    'xlabel', {'','','noise standard deviation [points]', 'noise standard deviation [points]'}, ...
+    'title', names);
+ymax = 1.2;
+plt.param_fig.colordelete = 1;
+for i = 1:2
+    fn = fns{i};
+    plt.new;
+    for hi = 1:2
+        td = sp.(fn);
+        td = squeeze(td(:,:,hi));
+        td = reshape(td, 1, []);
+        tl = hist(td, xbins)/(length(td)*stepsize);
+        plt.lineplot(tl, [], xbins);
+        
+        td = sp2.(fn);
+        td = squeeze(td(:,:,hi));
+        td = reshape(td, 1, []);
+        tl = hist(td, xbins)/(length(td)*stepsize);
+        plt.lineplot(tl, [], xbins);
+    end
+    hold on;
+    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
+    hold off;
+    
+    plt.new;
+    td = sp.(['d' fn]);
+    td = reshape(td, 1, []);
[…]

### 2022-11-21 codes/EEsimulate_bayes_2noise.m

@@ -1,0 +1,44 @@
+function out = EEsimulate_bayes_2noise(data, As, bs, nrans, ndets)
+    disp(['Generating simulation']);
+    if size(ndets, 2) == 1
+        ndets = repmat(ndets, 1, 2);
+    end
+    if size(nrans, 2) == 1
+        nrans = repmat(nrans, 1, 2);
+    end
+    
+    for si = 1:data.nSubject
+        choice = [];
+        nT = data.nTrial(si);
+        dR = data.dR(si,1:nT)';
+        dI = data.dI(si,1:nT)';
+        repeatID = data.repeatID(si,1:nT)';
+        horizon = data.horizon(si,1:nT)';
+        n_det = NaN(1, data.nrepeatID(si));
+        for gi = 1:nT
+            if nrans(si, horizon(gi)) > 0
+                pdni = makedist('Logistic',0, nrans(si, horizon(gi)));
+                dni = random(pdni);
+            else
+                dni = 0;
+            end
+            if isnan(n_det(repeatID(gi)))
+                if ndets(si, horizon(gi)) > 0
+                    pdne = makedist('Logistic',0, ndets(si, horizon(gi)));
+                    n_det(repeatID(gi)) = random(pdne);
+                else
+                    n_det(repeatID(gi)) = 0;
+                end
+            end
+            dne = n_det(repeatID(gi));
+            dQ = dR(gi) + As(si, horizon(gi)) * dI(gi) + bs(si, horizon(gi)) + dne + dni;
+            if dQ > 0
+                choice(gi) = 1;
+            else
+                choice(gi) = 0;
+            end
+        end
+        data.choice(si, :) = W.extend_nan(choice, size(data.choice, 2));
+    end
+    out = data;
+end

### 2022-11-21 codes/mainBOTH_ANALYSIS_bayes2noise.m

@@ -1,0 +1,223 @@
+%% get version
+ver = 'all';
+% ver = 'paironly';
+mdpath = './HBI_models';
+%%
+cd(W.funcout('fileparts', 1, matlab.desktop.editor.getActiveFilename));
+outputdir = fullfile('../bayesoutput', ver);
+outputdir = GetFullPath(outputdir);
+wj = W_JAGS;
+%% set up chains
+nchains = 4; % How Many Chains?
+nburnin = 5000; % How Many Burn-in Samples?
+nsamples = 5000; % How Many Recorded Samples?
+HBItest = '';
+wj.setup_params(nchains, nburnin, nsamples);
+%% test
+% HBItest = 'test_';
+% wj.setup_params;
+%%
+nchains = wj.bayes_params.nchains;
+%% main bayes
+files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
+suffix = {'','_all'};
+for fi = 1:2
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains, mdpath);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise' suffix{fi}]);
+    wj.run;
+end
+%% model no bias
+% data = load(fullfile('../data', 'bayes_2noise.mat')).(['bayes_data']);
+% data.modelname = ['2noisemodel_nobias'];
+% wj.setup_data_dir(data, outputdir);
+% wjinfo = EEbayes_analysis(data, nchains);
+% wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_nobias']);
+% wj.run;
+%% model paironly
+% data = load(fullfile('../data', 'bayes_2noise_paironly.mat')).(['bayes_data_paironly']);
+% data.modelname = ['2noisemodel'];
+% wj.setup_data_dir(data, outputdir);
+% wjinfo = EEbayes_analysis(data, nchains);
+% wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_paironly']);
+% wj.run;
+%% 6 model comparison (rest of the 5 models)
+data = load(fullfile(fullfile('../data',ver), 'bayes_2noise.mat')).(['bayes_data']);
+for mi = 2:6
+    data.modelname = ['2noisemodel' char(64 + mi)];
[…]

### 2022-11-21 codes/mainBOTH_FIGURE_parameterrecovery.m

@@ -1,0 +1,253 @@
+%% setup
+ver = 'all';
+verout = 'all_50chains';
+% verout = ver;
+% ver = 'paironly';
+outputdir = fullfile('../bayesoutput', verout);
+figdir = fullfile('../figures', verout);
+plt = W_plt;
+plt.setup_W_plt('fig_dir', figdir, 'fig_suffix', '','fig_projectname', 'RanDetNoise');
+vv = ''; vv2 = '';
+%% parameter recovery
+%% figures - parameter recovery
+for i = 1:1
+    simunum = num2str(i);
+    plt.setup_W_plt('fig_dir', fullfile(figdir,'parameter_recovery'), ...
+        'fig_suffix', simunum,'fig_projectname', ['RanDetNoise' vv], ...
+        'isshow', false);
+    p1 = load(fullfile(outputdir, ['HBI_DetRanNoise' vv '_stat.mat'])).stats.mean;
+    p2 = load(fullfile(outputdir, ['HBI_DetRanNoise' vv '_simu' simunum '_stat.mat'])).stats.mean;
+    mx = {[-25 25],[-25 25],[-3 3],[-3 3],[-1 20],[-1 30],[-1 5],[-1 15]};
+    mxtick = {[-20:10:20],[-20:10:20],[-2:2:2],[-2:2:2],[0:5:30],[0:5:30],[0:5:30],[0:5:30]};
+    plt.setfig(8, 'xtick', mxtick, ...
+        'ytick', mxtick);
+    plt.param_fig.islocked = 1;
+    EEplot_2noise_parameter_recovery(plt, p1, p2);
+end
+%% figures - hyperprior recovery
+for i = 1:1
+    simunum = num2str(i);
+    plt.setup_W_plt('fig_dir', fullfile(figdir,'parameter_recovery'),...
+        'fig_suffix', [simunum],'fig_projectname', ['RanDetNoise' vv],...
+        'isshow', false);
+    sp = load(fullfile(outputdir, ['HBI_DetRanNoise' vv '_samples.mat'])).samples;
+    sp2 = load(fullfile(outputdir, ['HBI_DetRanNoise' vv '_simu' simunum '_samples.mat'])).samples;
+    plt = EEplot_2noise_parameter_recovery_hyperprior(plt, sp, sp2);
+end
+%% figures - hyperprior recovery (50 simulations)
+for i = 1:1
[…]

### 2022-11-21 codes/mainPAIRONLY_ANALYSIS_FIGURE_RanDetNoise.m

@@ -1,0 +1,60 @@
+%% main analysis 2016
+rawdatadir = '../data';
+savedir = '../data/paironly';
+files = dir(fullfile(rawdatadir,'Imported_EE_blinkCuriosity_2016S.csv'));
+filename = fullfile(files.folder, files.name);
+%% basic analysis
+opt_sub = {'subjectID', 'date', 'time'}; 
+opt_preprocess = 'EEpreprocess_game_basic';
+opt_game_sub = {'EEpreprocess_game_sub_repeatedpaironly'};
+opt_analysis = {'EEanalysis_sub_basic'};
+[sub0, game, idxsub] = W_sub.analysis_pipeline(filename, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, savedir);
+W.writetable(game, fullfile(savedir,'data_paironly.csv'));
+%% exclude participants
+sub0 = readtable('../data/paironly/SUB_EE_blinkCuriosity_2016S.csv', 'Delimiter', ',');
+load('../data/paironly/idxsub_EE_blinkCuriosity_2016S.mat');
+[suball, idall] = W_sub.excludesubject(sub0, 'age', [18 NaN]);
+[sub, id] = W_sub.excludesubject(sub0, 'age', [18 NaN], 'pvalue_ac', [NaN 0.05]);
+save(fullfile(savedir, 'idxsub_exclude.mat'), 'id', 'idall','idxsub');
+%% analyze group
+clear gp;
+additionalcompare = {{'p_inconsistent13'},{'p_inconsistent22'},...
+    {'p_inconsistent13','p_inconsistent13_randomtheory'},{'p_inconsistent22','p_inconsistent22_randomtheory'}};
+gp{1} = W_sub.analysis_1group(sub, true, additionalcompare);
+gp{2} = W_sub.analysis_1group(suball, true, additionalcompare);
+suffix = {'','_all'};
+%%
+plt = W_plt('isshow', 1);
+for gi = 1:2
+%% model-free plots
+    plt.setup_W_plt('fig_dir', '../figures/paironly','fig_suffix', suffix{gi},'fig_projectname', 'RanDetNoise');
+    plt.param_fig.islocked = true;
+    plt.setfig(2, 'color', {{'AZsand','AZcactus'}, {'AZcactus'}}, ...
+        'ylim', {[0, 0.3] + gi*0.05,[0.35 0.65]}, ...
+        'ytick', {0.1:.1:0.4, 0.4:.1:0.6});
[…]

### 2022-11-21 codes/main_bayes_40chains.m

@@ -1,0 +1,212 @@
+%% get version
+ver = 'all';
+versave = 'all_50chains';
+% ver = 'paironly';
+%%
+cd(W.funcout('fileparts', 1, matlab.desktop.editor.getActiveFilename));
+outputdir = fullfile('../bayesoutput', versave);
+outputdir = GetFullPath(outputdir);
+wj = W_JAGS;
+%% set up chains
+nchains = 50; % How Many Chains?
+nburnin = 2000; % How Many Burn-in Samples?
+nsamples = 1000; % How Many Recorded Samples?
+HBItest = '';
+wj.setup_params(nchains, nburnin, nsamples);
+wj.openpool;
+%% test
+% HBItest = 'test_';
+% wj.setup_params;
+%%
+nchains = wj.bayes_params.nchains;
+%% main bayes
+files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
+suffix = {'','_all'};
+for fi = 1:2
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise' suffix{fi}]);
+    wj.run;
+end
+%% 6 model comparison (rest of the 5 models)
+data = load(fullfile(fullfile('../data',ver), 'bayes_2noise.mat')).(['bayes_data']);
+for mi = 2:6
+    data.modelname = ['2noisemodel' char(64 + mi)];
+    wj.setup_data_dir(data, outputdir);
+    wjinfo = EEbayes_analysis(data, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise_' char(64 + mi)]);
+    wj.run;
+end
+%% reduced model
+files = {'bayes_2noise_dRonly.mat', 'bayes_2noise_0model.mat', 'bayes_2noise_dIonly.mat'};
+suffix = {'_dRonly','_0model','_dIonly'};
+for fi = 1:1 % just the first one
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/AAA_075.m

@@ -1,0 +1,14 @@
+
+close all;
+clear all;
+clc
+option_time = 40;
+% 0 with time limit off
+% any positive number, with time limit on to be that number
+option_test = 1;
+% 0 when test mode is off
+% 1 when test mode is on
+option_eye = 1;
+% 0 when eye tracker if off;
+% 1 when eye tracker is on;
+main(option_time,option_test,option_eye);

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/bandit.m

@@ -1,0 +1,107 @@
+classdef bandit < handle
+    properties
+        %window
+        window
+        %draw
+        w
+        h
+        w_lever
+        h_lever
+        left
+        top
+        horizon
+        rect
+        color
+        color_highlight
+        dotradius
+        penwidth
+        %font
+        textSize
+        textColour
+        textWrapat
+        %task
+        numbers
+        lever_side
+    end
+    methods
+        function obj = bandit(window, textSize, textColour, textWrapat)
+            obj.window = window;
+            obj.setTextParameters(textSize, textColour, textWrapat);
+        end
+        function setTextParameters(obj, textSize, textColour, textWrapat)
+            Screen('TextFont', obj.window, 'Arial');
+            Screen('TextSize', obj.window, textSize);
+            Screen('TextStyle', obj.window, 1);
+            obj.textWrapat = textWrapat;
+            obj.textSize = textSize;
+            obj.textColour = textColour;
+        end
+        function setup(obj,w,h,w_lever,h_lever,horizon,color,color_highlight,lever_side)
+            obj.color = color;
+            obj.color_highlight = color_highlight;
+            obj.w = w;
+            obj.h = h;
+            obj.w_lever = w_lever;
+            obj.h_lever = h_lever;
+            obj.penwidth = 5;
+            obj.dotradius = 30;
+            obj.horizon = horizon;
+            obj.settppos(w/2,0);
+            obj.lever_side = lever_side;
+        end
+        function settppos(obj,cleft,top)
+            w = obj.w;
+            h = obj.h;
+            left = cleft - w/2;
+            obj.left = left;
+            obj.top = top;
+            for i = 1:obj.horizon
+                obj.rect(:,i) = [left+0 top+h*(i-1) left+w top+h*i];
+            end
+        end
+        function flush(obj)
+            obj.numbers = {};
+        end
+        function addreward(obj,new)
+            obj.numbers = {obj.numbers{:} new};
+        end
+        function draw(obj,played,forcedchoice)
+            if exist('forcedchoice') ~= 1
+                forcedchoice = [];
+            end
+            window = obj.window;
+            numbers = obj.numbers;
+            rect = obj.rect;
+            if (length(forcedchoice) == 0 && length(numbers) < size(rect,2))
+                Screen('FillRect', window, obj.color_highlight, rect(:,length(numbers)+1));
+            end
+            Screen('FrameRect', window, obj.color, rect, obj.penwidth);
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/boxQuestion.m

@@ -1,0 +1,55 @@
+classdef boxQuestion < handle
+    properties
+        window
+        qString
+        qTextSize
+        qTextColor
+        qPos
+        margin
+        bx
+        legalkey
+    end
+    methods
+        function obj = boxQuestion(window)
+            obj.window = window;
+        end
+        function setup(obj,qString,nx,ny)
+            obj.qString     = qString;
+            obj.qTextSize   = 30;
+            colorscale = 1;
+            obj.qTextColor  = [1 1 1]*colorscale;
+            
+            % to keep this simple have it be a set size
+            pos = [100 500];
+            
+            edgeWidth_on = 5;
+            edgeWidth_off = 2;
+            edgeColor_off = [1 1 1]*colorscale;
+            edgeColor_on = [1 1 1]*colorscale;
+            textSize = 50;
+            textColor = [1 1 1]*colorscale;
+            bgColor = [0 0 0]*colorscale;
+            pad = 40;
+            obj.margin = [pad/2,obj.qTextSize*2.2];
+            obj.bx = inputTextBox(obj.window,obj.legalkey);
+            obj.bx.setup(pos+obj.margin, nx, ny, ...
+                edgeWidth_on, edgeWidth_off, edgeColor_on, edgeColor_off, ...
+                textSize, textColor, bgColor, pad);
+        end
+        function setCornerPosition(obj, pos)
+            obj.qPos = pos;
+            obj.bx.setCornerPosition(pos+obj.margin);
+        end
+        function draw(obj)
+            sx = obj.qPos(1);
+            sy = obj.qPos(2);
+            oldTextSize = Screen('TextSize', obj.window, obj.qTextSize);
+            DrawFormattedText(obj.window, obj.qString, sx, sy, obj.qTextColor);
+            Screen('TextSize', obj.window, oldTextSize);
+            obj.bx.draw;
+        end        
+        function check(obj, x, y)
+            obj.bx.check(x,y);
+        end
+    end
+end

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/buttonQuestion.m

@@ -1,0 +1,78 @@
+classdef buttonQuestion < handle
+    properties
+        window
+        qString
+        qTextSize
+        qTextColor
+        qPos
+        multichoice
+        choice
+        bt
+    end
+    methods
+        function obj = buttonQuestion(window,multichoice)
+            obj.window = window;
+            obj.multichoice = multichoice;
+            obj.choice = -1;
+        end
+        function default(obj,value)
+            obj.choice = value;
+        end
+        function setup(obj,qString, aString)
+            obj.qString     = qString;
+            obj.qTextSize   = 30;
+            colorscale = 1;
+            obj.qTextColor  = [1 1 1]*colorscale;
+
+            textSize    = 20;
+            textColor   = [1 1 1]*colorscale;
+            edgeWidth   = 2;
+            edgeColor   = [1 1 1]*colorscale;
+            onColor     = [1 1 1]*colorscale;
+            offColor    = [0 0 0]*colorscale;
+            margin      = 10;
+            radius      = 20;
+            fillFrac    = 0.6;
+            for i = 1:length(aString)
+                string = aString{i};
+                if i == 1
+                    obj.bt = pushButtonText(obj.window);
+                else
+                    obj.bt(i) = pushButtonText(obj.window);
+                end
+                obj.bt(i).setup(textSize,textColor,string,edgeColor,onColor,offColor,edgeWidth,radius,fillFrac,margin);
+            end
+        end
+        function setCornerPosition(obj, pos)
+            obj.qPos = pos;
+            for i = 1:length(obj.bt)
+                obj.bt(i).setCornerPosition([pos(1) pos(2)+obj.qTextSize*1.1+i*obj.bt(i).textSize*1.1]);
+            end
+        end
+        function draw(obj)
+            sx = obj.qPos(1);
+            sy = obj.qPos(2);
+            oldTextSize = Screen('TextSize', obj.window, obj.qTextSize);
+            DrawFormattedText(obj.window, obj.qString, sx, sy, obj.qTextColor);
+            Screen('TextSize', obj.window, oldTextSize);
+            for i = 1:length(obj.bt)
+                obj.bt(i).draw;
+            end
+        end
+        function check(obj, x, y)
+            for i = 1:length(obj.bt)
+                obj.bt(i).check(x,y);
+                if obj.bt(i).bt.isOn
+                   if  ~obj.multichoice && obj.choice ~= -1 && obj.choice ~= i
+                        temp = obj.choice;
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/demographic.m

@@ -1,0 +1,105 @@
+function demo = demographic(window,subID)
+    qString{1} = 'Gender';
+    aString{1} = {
+        'Female' 
+        'Male' 
+        'Decline to answer'};
+    qString{2} = 'Race';
+    aString{2} = {
+        'American Indian or Alaska Native' 
+        'Asian' 
+        'Black or African-American' 
+        'Native Hawaiian or Other Pacific Islander' 
+        'White' 
+        'More than one race'
+        'Unknown'
+        'Decline to answer'};
+    qString{3} = 'Ethnicity';
+    aString{3} = {
+        'Hispanic or Latino'
+        'Not Hispanic or Latino'
+        'Unknown'
+        'Decline to answer'};
+    qString{4} = 'Are you a native English speaker?';
+    aString{4} = {
+        'Yes'
+        'No'
+        'Decline to answer'};
+    
+    qString{5} = 'How many hours approximately did you sleep last evening?';
+    aString{5} = {
+        '10 or more'
+        '8 to 10'
+        '6 to 8'
+        '4 to 6'
+        '2 to 4'
+        '2 or less'
+        'Decline to answer'
+    };
+    for i = 1:length(qString)
+        bq(i) = buttonQuestion(window,0);
+        bq(i).setup(qString{i}, aString{i});
+    end
+    bq(1).setCornerPosition([50 100]);
+    bq(2).setCornerPosition([400 100]);
+    bq(3).setCornerPosition([1050 100]);
+    bq(4).setCornerPosition([400 500]);
+    bq(5).setCornerPosition([1050 500]);
+
+    nx = 3;ny = 1;
+    bxq = boxQuestion(window);
+    bxq.legalkey = [KbName('1'):KbName('0'),KbName('1!'):KbName('0)')];
+    bxq.setup('Age', nx, ny)
+    bxq.setCornerPosition([50 500]);
+    if subID == 999
+    bxq.bx.string = '0';
+    end
+    string = 'Submit';
+    colorscale = 1;
+    cb = clickBox(window);
+    textColor = [0 0 0]*colorscale;
+    edgeColor = [1 1 1]*colorscale;
+    fillColor_on = [1 0 0]*colorscale;
+    fillColor_off = [0 1 0]*colorscale;
+    textSize = 30;
+    rect = [0 0 300 100];
+    cb.setup(string,textColor,textSize,edgeColor,fillColor_on,fillColor_off,rect);
+    cb.setCornerPosition([900 800]);
+    while true
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/getKbCharOrClick.m

@@ -1,0 +1,33 @@
+function out = getKbCharOrClick(window, untilTime)
+    yieldInterval = 0.01;
+    if nargin < 2
+        untilTime = inf;
+    end
+    if isempty(untilTime)
+        untilTime = inf;
+    end
+    secs = -inf;
+    while secs < untilTime
+        [isDown, secs, keyCode, deltaSecs] = KbCheck();
+        if isDown || (secs >= untilTime)
+            out.what        = 'keyPress';
+            out.secs        = secs;
+            out.keyCode     = keyCode;
+            out.deltaSecs   = deltaSecs;
+            out.ch          = find(keyCode,1);
+%             disp(out.ch);
+            break;
+        end    
+        [x,y,buttons,focus,valuators,valinfo] = GetMouse(window);
+        if any(buttons)
+            out.what        = 'click';
+            out.secs        = secs;
+            out.x           = x;
+            out.y           = y;
+            out.buttons     = buttons;
+            break;
+        end
+%         secs = WaitSecs('YieldSecs', yieldInterval);
+    end
+    secs = WaitSecs(yieldInterval);    
+end

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/getsubjectID.m

@@ -1,0 +1,108 @@
+function sub = getsubjectID(window,subjectID)
+    qString{1} = 'Eye make-ups';
+    aString{1} = {
+    'None' 
+    'Eyeliner' 
+    'Mascara'
+    'Mascara & Eyeliner'
+    'Other'
+    };
+    
+    qString{2} = 'Eyetracker calibration quality';
+    aString{2} = {
+        '5'
+        '4'
+        '3'
+        '2'
+        '1'
+        'not calibrated'
+    };
+    qString{3} = 'Sight';
+    aString{3} = {
+        'Normal sight'
+        'Corrected sight(wear glasses)'
+        'Short-sighted'
+        'Far-sighted'
+        'Astigmatic'
+        'Other'
+    };
+    nx = 5;
+    ny = 1;
+    bxq = boxQuestion(window);
+    bxq.legalkey = [KbName('1'):KbName('0'),KbName('1!'):KbName('0)')];
+    bxq.setup('Subject ID', nx, ny);
+    bxq.bx.string = num2str(subjectID);
+%     nx = 30;
+%     ny = 1;
+%     bxq2 = boxQuestion(window);
+%     bxq2.legalkey = [KbName('a'):KbName('z')];
+%     bxq2.setup('Last Name', nx, ny);
+%     nx = 30;
+%     ny = 1;
+%     bxq3 = boxQuestion(window);
+%     bxq3.legalkey = [KbName('a'):KbName('z')];
+%     bxq3.setup('First Name', nx, ny);
+    for i = 1:length(qString)
+        if i == 1
+            bq = buttonQuestion(window,0);
+        else
+            bq(i) = buttonQuestion(window,0);
+        end
+        bq(i).setup(qString{i}, aString{i});
+    end
+    bq(1).setCornerPosition([400 100]);
+    bq(2).setCornerPosition([700 100]);
+    bq(3).setCornerPosition([50 400]);
+    bxq.setCornerPosition([50 100]);
+%     bxq2.setCornerPosition([400 100]);
+%     bxq3.setCornerPosition([400 400]);
+    string = 'Continue';
+    colorscale = 1;
+    cb = clickBox(window);
+    textColor = [0 0 0]*colorscale;
+    edgeColor = [1 1 1]*colorscale;
+    fillColor_on = [1 0 0]*colorscale;
+    fillColor_off = [0 1 0]*colorscale;
+    textSize = 30;
+    rect = [0 0 300 100];
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/inputTextBox.m

@@ -1,0 +1,142 @@
+classdef inputTextBox < handle
+    properties
+        window
+        nx
+        ny
+        edgeWidth_on
+        edgeWidth_off
+        edgeColor_on
+        edgeColor_off
+        textSize
+        textColor
+        bgColor
+        string
+        max_string
+        edgeRect
+        borderRect
+        pad
+        cornerPosition
+        isOn
+        legalkey
+    end
+    methods
+        function obj = inputTextBox(window,legalkey)
+            obj.window = window;
+            obj.legalkey = legalkey;
+        end
+        function setup(obj,pos, nx, ny, ...
+                edgeWidth_on, edgeWidth_off, edgeColor_on, edgeColor_off, ...
+                textSize, textColor, bgColor, pad)
+            obj.cornerPosition  = pos;
+            obj.nx              = nx;
+            obj.ny              = ny;
+            obj.edgeWidth_on    = edgeWidth_on;
+            obj.edgeWidth_off   = edgeWidth_off;
+            obj.textSize        = textSize;
+            obj.textColor       = textColor;
+            obj.edgeColor_on    = edgeColor_on;
+            obj.edgeColor_off   = edgeColor_off;
+            obj.bgColor         = bgColor;
+            obj.isOn            = false;
+            obj.string = [];
+            % make max_string to get bounding box
+            max_string(1:nx,1:ny) = 'X';
+            obj.max_string = max_string(:)';
+            oldTextSize = Screen('TextSize', obj.window, obj.textSize);
+            [~,~,textbounds] = DrawFormattedText(obj.window, obj.max_string, pos(1), pos(2), obj.textColor);
+            Screen('TextSize', obj.window, oldTextSize);
+            obj.edgeRect = textbounds;
+            obj.pad = pad;
+            obj.borderRect = obj.edgeRect + obj.pad/2*[-1 -1 1 1];
+        end
+        function setCornerPosition(obj, pos)
+            obj.cornerPosition = pos;
+            edgeRect = obj.edgeRect;
+            edgeRect = edgeRect - [edgeRect(1:2) edgeRect(1:2)];
+            obj.edgeRect = edgeRect + [pos pos];
+            obj.borderRect = obj.edgeRect + obj.pad/2*[-1 -1 1 1];
+        end
+        function draw(obj)
+            switch obj.isOn
+                case 0
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/instructions.m

@@ -1,0 +1,84 @@
+classdef instructions < handle
+    properties
+        window
+        iStr
+        text
+        N
+    end
+    methods
+        function obj = instructions(window)
+            obj.window = window;
+        end
+        function setup(obj,text)
+            obj.text.textSize = text.textSize;
+            obj.text.textWrapat = text.textWrapat;
+            obj.text.textColor = text.textColor;
+            obj.text.textFont = text.textFont;
+        end
+        function update(obj,iStr)
+            obj.iStr = iStr;
+            obj.N = size(iStr,2);
+        end
+        function run(obj,option)
+            count = 1;
+            while count <= obj.N
+            obj.talkAndFlip(obj.iStr{count},option);
+            keyspace = KbName('space');
+                            keybackspace =  KbName('delete');
+                            % press button to move on or not
+                            endFlag = false;
+                            while ~endFlag
+
+                                [KeyNum, when] = waitForInput([keyspace,keybackspace], Inf);
+                                switch KeyNum
+
+                                    case keyspace % go forwards
+                                        count = count +1;
+                                            endFlag = true;
+
+                                    case keybackspace % go backwards
+                                        count = count - 1;
+                                        if count < 1
+                                            count = 1;
+                                        end
+                                            endFlag = true;
+
+            %                         case 5 % skip through
+            %                             endFlag = true;
+            %                             
+            %                         case 6 % quit
+            %                             sca
+            %                             error('User requested escape!  Bye-bye!');
+
+                                end
+
+                            end
+            end
+        end
+        function talkAndFlip(obj, str, option, pTime)
+                    Screen('TextFont', obj.window, 'Arial');
+                    Screen('TextSize', obj.window, obj.text.textSize);
+                    [A, B] = Screen('WindowSize', obj.window);
+                    if exist('pTime') ~= 1
+                        pTime = 0.3;
+                    end
+                    if option == 0
+                        [nx, ny] = DrawFormattedText(obj.window, ...
+                            ['' str], ...
+                            0.05*A,[B*0.02], obj.text.textColor, obj.text.textWrapat);
+                        Screen('TextSize', obj.window,round(obj.text.textSize));
+                        DrawFormattedText(obj.window, ...
+                            ['' ...
+                            'Press space to continue or delete to go back'], ...
+                            'center', [B*0.93], [150 150 150], obj.text.textWrapat);
+                    elseif option == 1
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/pushButtonText.m

@@ -1,0 +1,45 @@
+classdef pushButtonText < handle
+    properties
+        window
+        
+        textSize
+        textColor
+        textPosition
+        cornerPosition
+        string
+        margin
+        
+%         isOn
+        bt
+    end
+    methods
+        function obj = pushButtonText(window)
+            obj.window = window;
+            obj.bt = pushButton(window);
+        end
+        function setup(obj,textSize,textColor,string,edgeColor,onColor,offColor,edgeWidth,radius,fillFrac,margin)
+            obj.textSize = textSize;
+            obj.textColor = textColor;
+            obj.string = string;
+            obj.margin = margin;
+            obj.bt.setup(edgeColor,onColor,offColor,edgeWidth,radius,fillFrac);
+        end
+        function setCornerPosition(obj, pos)
+            obj.cornerPosition  = pos;            
+            obj.bt.setCornerPosition(pos);
+            obj.textPosition    = obj.cornerPosition + [obj.bt.radius + obj.margin -0.2*obj.textSize];
+        end
+        function draw(obj)
+            sx = obj.textPosition(1);
+            sy = obj.textPosition(2);
+            oldTextSize = Screen('TextSize', obj.window, obj.textSize);
+            DrawFormattedText(obj.window, obj.string, sx, sy, obj.textColor);
+            Screen('TextSize', obj.window, oldTextSize);
+            obj.bt.draw;
+        end
+        function check(obj,x,y)
+            obj.bt.check(x,y);
+%             obj.isOn = obj.bt.isOn;
+        end
+    end
+end

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/python_source/EyeTribe_Matlab_server (Bryan Kromenacker's conflicted copy 2015-10-22).py

@@ -1,0 +1,326 @@
+# native
+import time
+import socket
+
+# external
+from pytribe import EyeTribe
+
+
+# # # # #
+# CONSTANTS
+
+DEBUG = False
+
+MLIP = 'localhost'
+MLPORT = 5666
+
+
+# # # # #
+# INIT CONNECTION
+
+# start socket connection
+print("Starting new socket connection (ip=%s, port=%d)." % (MLIP, MLPORT))
+sock = socket.socket(socket.AF_INET, socket.SOCK_STREAM)
+sock.bind((MLIP,MLPORT))
+
+# set server timeout to two minutes
+sock.settimeout(120)
+
+# start listening for new connections (max = 10)
+print("Listening for connections...")
+sock.listen(10)
+
+# wait for the connection to be made
+matlabconnected = False
+while not matlabconnected:
+	# accept any new connection
+	try:
+		conn, addr = sock.accept()
+	except socket.timeout:
+		sock.close()
+		raise Exception("ERROR: socket timed out, no connections available.")
+	print("Found new connection (%s, %d)" % (addr[0], addr[1]))
+	# set the timeout for this connection to 10 seconds
+	conn.settimeout(10)
+	# see whether the new connection is Matlab
+	try:
+		message = conn.recv(1024)
+		print("Client says: %s" % message)
+	except socket.timeout:
+		print("Timeout while listening to (%s, %d)." % (addr[0], addr[1]))
+		message = None
+	if message == "Hi, this is Matlab!":
+		matlabconnected = True
+# return success message
+conn.send('success')
+
+
+# # # # #
+# RUN EYETRIBE
+
+# run until the stop command is received
+result = None
+tracker = None
+stopped = False
+calibrationmode = False
+while not stopped:
+	
+	# READ INCOMING MESSAGES
+	try:
+		# get a new incoming message
+		message = conn.recv(1024)
+		if DEBUG:
+			print("Client says: %s" % message)
+	except socket.timeout:
+		# on a timeout, set message to None
+		message = None
+	
+	# if there was no new message, skip further processing
+	# (which will make us go back to checking for new messages)
+	if message == None:
+		continue
+	
+	# CALIBRATION MODE
+	if calibrationmode:
+		
+		# POINT START
+		if "Calibration pointstart" in message:
+			try:
+				# parse message for point coordinates
+				xpos = message.find('x=')
+				ypos = message.find('y=')
+				x = int(message[xpos+2:ypos-1])
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/python_source/setup.py

@@ -1,0 +1,4 @@
+from distutils.core import setup
+import py2exe
+
+setup(console=['EyeTribe_Matlab_server.py'], options={'py2exe':{"bundle_files":1}}, zipfile=None)

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/python_source/setup_OLD.py

@@ -1,0 +1,19 @@
+"""
+This is a setup.py script generated by py2applet
+
+Usage:
+    python setup.py py2app
+"""
+
+from setuptools import setup
+
+APP = ['EyeTribe_Matlab_server.py']
+DATA_FILES = []
+OPTIONS = {'argv_emulation': True}
+
+setup(
+    app=APP,
+    data_files=DATA_FILES,
+    options={'py2app': OPTIONS},
+    setup_requires=['py2app'],
+)

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/surveys_copy.m

@@ -1,0 +1,193 @@
+classdef surveys < handle
+    properties
+        subjectID
+        surveydir
+        surveyname
+        surveyansname
+        qnum
+        survey
+        datadir
+        savename
+        window
+        bgColour
+        screenWidth
+        screenHeight
+        screenCenter
+        text
+        legalkey
+    end
+    methods
+        function obj = surveys(window,subjectID)
+            obj.subjectID = subjectID;
+            obj.window = window;
+%             obj.setup;
+        end
+        function obj = setupsurvey(obj,surveydir,datadir,surveyname,surveyansname)
+            obj.datadir = datadir;
+            obj.surveydir = surveydir;
+            obj.surveyname = surveyname;
+            if ~exist('surveyansname')
+                obj.surveyansname = [];
+            else
+                obj.surveyansname = surveyansname;
+            end
+            obj.readSurvey;
+        end
+        function setup(obj,savename,text)
+            obj.savename = savename;
+            obj.setwindow;
+            obj.text.textSize = text.textSize;
+            obj.text.textWrapat = text.textWrapat;
+            obj.text.textColor = text.textColor;
+            obj.text.textFont = text.textFont;
+            obj.setTextParameters;
+        end
+        function setwindow(obj)
+            [screenWidth,screenHeight] = Screen('WindowSize',obj.window);
+            screenCenter = [screenWidth,screenHeight]/2;
+            colorscale = 1;
+            obj.bgColour = [0 0 0]*colorscale;
+            obj.screenWidth = screenWidth;
+            obj.screenHeight = screenHeight;
+            obj.screenCenter = screenCenter;
+        end
+        function setTextParameters(obj)
+            Screen('TextFont', obj.window, obj.text.textFont);
+            Screen('TextSize', obj.window, obj.text.textSize);
+            Screen('TextColor', obj.window, obj.text.textColor);
+%             Screen('TextStyle', obj.window, 1);
+        end
+        function [KeyNum, when, deltawhen] = waitForInput(obj, validKeys, timeOut)
+            % wait for a keypress for TimeOut seconds
+            Pressed = 0;
+            while ~Pressed && (GetSecs < timeOut)
+                [Pressed, when, KeyNum, deltawhen] = KbCheck;
+                KeyNum = find(KeyNum,1);
+                if ~ismember(KeyNum,validKeys) % && length(validKeys) > 0
+                    Pressed = 0;
+                end
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/task_horizon.m

@@ -1,0 +1,990 @@
+classdef task_horizon < handle
+%horizon task    
+    properties
+        %subject
+        subjectID
+        savename
+        %task parameter
+        sig_risk
+        nBlocks
+        ShowForced
+        game
+        %bandit
+        ban_L5
+        ban_R5
+        ban_L10
+        ban_R10
+        cross
+        crosscenter
+        %window
+        window
+        bgColor
+        screenHeight
+        screenWidth
+        screenCenter
+        %font
+        text
+        textSize
+        textColor
+        textWrapat
+        %Beep
+        repetitions
+        startCue
+        waitForDeviceStart
+        beepLengthSecs
+        pahandle
+        myBeep
+        starttime
+        startclock
+        endtime
+        instime
+        mtime
+        eye
+        iStr
+        iEv
+        maxtimeallowed
+    end
+    methods
+        function obj = task_horizon(window,savename,bgColour)
+            obj.setwindow(window,bgColour);
+            obj.savename = savename;
+            obj.maxtimeallowed = 40;
+        end
+        function setup(obj,subjectID,eye,textColour)            
+            obj.subjectID = subjectID;
+            rand('seed',subjectID);
+            obj.setTextParameters(40,textColour, 70);
+            obj.makebandit;
+            obj.makecross;
+            obj.taskparameter;
+            obj.setupsound;
+            obj.eye = eye;
+        end
+        function setwindow(obj,window,bgColour)
+%             screens = Screen('Screens');
+%             screenNumber = max(screens);
+            colorscale = 1;
+            obj.bgColor = bgColour;
+%             [window, windowRect] = PsychImaging('OpenWindow', screenNumber, obj.bgColor);
+            obj.window = window;
+            [sw, sh] = Screen('WindowSize', window);
+            windowRect = [0 0 sw sh];
+            obj.screenWidth = sw;
+            obj.screenHeight = sh;
+            [xCenter, yCenter] = RectCenter(windowRect);
+            obj.screenCenter = [xCenter, yCenter];
+        end
+        function makebandit(obj)
+            window = obj.window;
+            w = 80;
+            h = 60;
+            w_lever = 80;
+            h_lever = 60;
+            colorscale = 1;
+            color_highlight = [ 55    85    55 ]/255;
+            horizon = 5;
+            color_l = [ 71    71   101 ]/255;
+            color_r = [91    71    41]/255;
[…]

### 2022-11-21 data/EE_blinkCuriosity_2016S/functions/waitForInput.m

@@ -1,0 +1,15 @@
+function [KeyNum, when, deltawhen] = waitForInput(validKeys, timeOut)
+            % wait for a keypress for TimeOut seconds
+            Pressed = 0;
+            while ~Pressed && (GetSecs < timeOut)
+                [Pressed, when, KeyNum, deltawhen] = KbCheck;
+                KeyNum = find(KeyNum,1);
+                if ~ismember(KeyNum,validKeys)
+                    Pressed = 0;
+                end
+            end
+            if Pressed == 0
+                KeyNum = [];
+                when = [];
+            end
+        end

### 2022-11-21 data/EE_blinkCuriosity_2016S/main.m

@@ -1,0 +1,187 @@
+function main(option_time, option_test, option_eye)
+if ~exist('option_time') || option_time < 0
+    option_time = 0;
+end
+if ~exist('option_test') 
+    option_test = 0;
+end
+if ~exist('option_eye')
+    option_eye = 1;
+end
+if option_test ~= 0
+    disp('Warning: Test mode is on!');
+    input('Type Enter to continue!');
+    tbaseline = 1;
+else
+    tbaseline = 300;
+end
+if option_eye ~= 1
+    disp('Warning: Eye tracker is off!');
+    input('Type Enter to continue!');
+end
+
+% directories
+mainpath = [pwd '/'];
+addpath('./functions');
+temppath = [mainpath 'temp/subjectID.mat'];
+savepath = [mainpath 'data/'];
+
+if option_test == 0
+    sublist = importdata(temppath);
+    subjectID = sublist(end);
+    subjectID = subjectID + 1;
+    save(temppath,'subjectID');
+    sub.subjectID = input('SubjectID:');
+else
+    sub.subjectID = 999;
+end
+% initial parameters
+projectname = 'P075_';  
+time = datestr(now,30);      
+starttaskclock = clock;
+% parameters
+text.textSize = 40;
+text.textColor = [1 1 1]*1;
+text.textWrapat = 70;
+text.textFont = 'Arial';
+colour{1} = [ 97    67    67 ]/255;                           
+colour{2} = [ 55    85    55 ]/255;
+colour{3} = [ 71    71   101 ]/255;
+colour{4} = [ 91    71    41 ]/255;
+textColour = [255 255 255]/255;  
+bgColour = [1 1 1]/255;
+
+% eyetracker setup
+eyetribedir  = ['~/Desktop/python_source'];
+eyetribeSourcedir = ['~/Desktop/EyeTribe_for_Matlab/'];
+eye = eyetracker(eyetribedir,eyetribeSourcedir);
+
+if option_eye ~= 1
+    eye.flag = false;
+else
+    eye.flag = true;
+end
+
+filename = [projectname, 'eye_raw_sub', num2str(sub.subjectID), '_',time];
+eye.setup(savepath,filename);
+eye.open;
+eye.calibrate;
+eye.startmatlabserver;
+eye.connect;
+
+cd(mainpath);
+
+% start screen
+Screen('Preference', 'SkipSyncTests', 1); 
+PsychDefaultSetup(2);
+KbName('UnifyKeyNames');
+screens = Screen('Screens');
+screenNumber = max(screens);
[…]

### 2022-11-23 codes/main_2_bayes2noise.m
bayes basic

@@ -1,0 +1,54 @@
+%% get version
+ver = 'all';
+%%
+W.cd();
+outputdir = fullfile('../bayesoutput', ver);
+outputdir = W.get_fullpath(W.mkdir(outputdir));
+wj = W_JAGS;
+%% set up chains
+nchains = 10; % How Many Chains?
+nburnin = 5000; % How Many Burn-in Samples?
+nsamples = 5000; % How Many Recorded Samples?
+HBItest = '';
+wj.setup_params(nchains, nburnin, nsamples);
+%% test
+% HBItest = 'test_';
+% wj.setup_params;
+%%
+nchains = wj.bayes_params.nchains;
+W.library_wang('Wang_EEHorizon', 1);
+W.library('JAGS', 1);
+%% main bayes
+files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
+suffix = {'','_all'};
+for fi = 1:2
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise' suffix{fi}]);
+    wj.run;
+end
+%% model no bias
+% data = load(fullfile('../data', 'bayes_2noise.mat')).(['bayes_data']);
+% data.modelname = ['2noisemodel_nobias'];
+% wj.setup_data_dir(data, outputdir);
+% wjinfo = EEbayes_analysis(data, nchains);
+% wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_nobias']);
+% wj.run;
+%% model paironly
+d = load(fullfile(fullfile('../data',ver), 'bayes_2noise_paironly.mat')).(['bayes_data_paironly']);
+d.modelname = ['2noisemodel'];
+wj.setup_data_dir(d, outputdir);
+wjinfo = EEbayes_analysis(d, nchains);
+wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise_paironly']);
+wj.run;
+%% reduced model
+files = {'bayes_2noise_dRonly.mat', 'bayes_2noise_0model.mat', 'bayes_2noise_dIonly.mat'};
+suffix = {'_dRonly','_0model','_dIonly'};
[…]

### 2022-12-12 manuscript/2020 rejected - JEP General/RandomDeterministicNoise/Ran-DetNoise.tex
updates

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[authoryear, round,sort,comma,numbers]{natbib}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage[font={sf,small}]{caption}
\usepackage{amssymb}
\usepackage{float}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\title{The nature of decision noise in random exploration}

\author[1]{Siyu Wang}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{}

\begin{document}
	\maketitle
\setcounter{page}{28}
	\newpage
	\begin{abstract}

Human decision making is inherently variable. While this variability is often seen as a sign of suboptimality in human behavior, recent work suggests that randomness can actually be adaptive. An example  arises when we must choose between exploring unknown options or exploiting options we know well. A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore. Moreover, people appear to use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore. From a modeling perspective, behavioral variability is essentially the variance that can not be explained by a model and is modeled as the level of decision noise. However, what we have called "decision noise" in previous researches could actually just be missing deterministic components from the model, it is difficult to tell whether decision noise truly arises from a stochastic process. Here we show that, while both random and deterministic noise drive variability in behavior, the noise driving random exploration is predominantly random. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.

\end{abstract}
	\newpage
	\section*{Introduction}

	Imagine trying to decide where to go to dinner. You can go to your favorite restaurant, the one you really enjoy and always go to, or you can try a new restaurant that you know nothing about. Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.

	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.  In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.

	It has recently been shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems and can increase such decision noise when it's more beneficial to explore \citep{Gershman2018, wilson2014}. In one of these tasks, known as the Horizon Task, the key manipulation is the horizon condition, i.e. the number of decisions remaining for the participant to make. Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards. For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant. Using such a horizon manipulation it has been shown that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.

	It is however difficult in these tasks to tell whether what is measured as decision noise is really random, what we have called 'noise' in previous researches could actually just be some missing deterministic components from the model. Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model. In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants. If the model did not consider that the agent is favoring the behavior of following a friend in a deterministic way, the behavior of going to a less favorable restaurant because of a friend will appear to be 'random' when it is really a deterministic effect. hence this deterministic factor will be modeled as a random decision noise. Crucially, however, this `deterministic noise' is very much in the stimulus and if you saw the same friend go into the same restaurant at a later date you might follow them again. Conversely, truly 'random noise' would arise from stochastic mental processes tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.

	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.  In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.

    % The idea here is behavior is less predictable in horizon 6 than horizon 1 even when all stimuli are the same.  Thus anything that deterministically depends on stimuli - i.e. most modeling effects - are ruled out as an explanation for increased noise.  In that light, what your model calls deterministic noise is that amount of variance that could be explained by deterministic effects of stimulus - i.e. modeling errors.

    %* Difficulty with measuring random exploration, how do you know something is random, previous work model
    %* Model all the deterministic factor (contribute to noise)
    %* Have repeated games
    %* Missing deterministic dependent, much smaller than non-deterministic component
    %* Stochastic process…

	%In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants. Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in. Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again.

	%Conversely, random noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.

	%Previous work makes a strong case for both deterministic and random noise being relevant to behavior. For instance, deterministic, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than random noise \citep{eeBrunton13}. Conversely random, neural noise is thought to drive exploratory singing behavior in song birds and the generation and control of this random noise has been linked to specific neural structures \citep{songbird2}.

	In this paper, we investigate which source of noise, deterministic vs random, drives random exploration in humans in a modified version of the Horizon Task. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely deterministic noise, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times. Meanwhile, if decision noise is truly random their choices should be less consistent, since random noise can be different both times. By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the variability related to random exploration is dominated by random noise. The missing deterministic component is much smaller than the non-deterministic component in random exploration.

	\section*{Results}

	\subsection*{The Repeated-Games Horizon Task}
	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of random vs deterministic noise on people's decisions (Figure \ref{fig:taskfig}). In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards. Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.

	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e. 1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition). These example plays allow us to manipulate exactly what people know about each option before they make their first choice. Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.

	These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.  Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent. Conversely if noise is randomly driven, then choices on repeated games should be independent and can be inconsistent.

	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{ 
			Schematic of the experiment. (A) Dynamics of an example horizon 6 game.  Here the first four trials are forced trials in which participants are instructed which option to play.  After the forced trials, participants are free to choose between the two options for the remainder of the game.  (B) Different possible states of the game after the first free choice over the course of the experiment. Overall participants play about 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.  In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.  These repeated games allow us to compute the relative contribution of deterministic and random noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

	\subsection*{Both behavioral variability and information seeking increase with horizon}

	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}. As in our previous work, we find evidence for two types of exploration in the Horizon Task.  Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.

	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition. This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(64) = 6.55, p $<$ 0.001 for [1 3], t(64) = 7.99, p $<$ 0.001 for [2 2]).  Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition. Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(64) = 6.92, p $<$ 0.001).

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.png}
			\caption{
			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
			\label{fig:modelfree}
		\end{center}
	\end{figure}

	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}

	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.  The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times. Conversely, purely random noise should lead to independent choices, and hence more inconsistent choices both times.

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was deterministic (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions.  For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6). In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] (t(64) = 5.41, p $<$ 0.001) and [2 2] condition (t(64) = 6.26, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is random.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/RanDetNoise_2noise.png}
			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration. For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1. However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely deterministic and purely random noise cases.  For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case. For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
	\begin{equation*}
	\begin{split}
	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
	\mbox{hence},\quad p(\mbox{inconsistent}) &=  
	1 - p(\mbox{consistent}) = 
	2 p(\mbox{low mean})(1-p(\mbox{low mean}))
	\end{split}
	\end{equation*}
	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction (Behavior is different from pure random noise prediction in both [1 3] condition, t(64) = 8.66, p $<$ 0.001 for horizon 1, t(64) = 9.48, p $<$ 0.001 for horizon 6; and [2 2] condition, t(64) = 6.94, p $<$ 0.001 for horizon 1, t(64) = 7.47, p $<$ 0.001 for horizon 6. Behavior is different from pure deterministic noise prediction in both [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; and [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6.), suggesting that both deterministic and random noise are present in driving this choice inconsistency. Since choice inconsistency only reflects random noise, Figure \ref{fig:mf2} suggests that random noise increases with horizon.

	\subsection*{Model-based analysis shows that random exploration is dominated by random noise}

	To more precisely quantify random and deterministic noise, we turned to model fitting. We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate random and deterministic noises. In particular, we assume that in repeated games, deterministic noise remains the same whereas random noise can change.

	\subsubsection*{Overview of model}
	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits. To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta    I+b+n_{det}+n_{ran}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference between information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively which are assumed to come from logistic distribution with mean 0.

	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.

	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the deterministic noise, $n_{det}$, should be the same while the random noise, $n_{ran}$ may be different. This is exactly how we distinguish deterministic noise from random noise. In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.

	\subsubsection*{Model fitting}
	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}). In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon. Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.

	\subsubsection*{Model fitting results}

	Posterior distributions over the group-level means of the deterministic and random noise variance are shown in Figure \ref{fig:mb1}. Consistent with our model-free results, we see that both random and deterministic noise variances are non-zero and that random noise is about 2-3 times larger than the deterministic noise. In addition, we find that random noise increases dramatically with horizon (M = 4.55, 100\% of samples showed an increase in random noise with horizon) whereas the increase in deterministic noise is smaller (M = 1.78, 98.12\% of samples showed an increase in deterministic noise with horizon).  Taken together, these results suggest that random exploration is dominated by random noise.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/RanDetNoise_hyperprior.png}
			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
			\label{fig:mb1}
		\end{center}
	\end{figure}

	%\begin{figure}[H]
	%	\begin{center}
%			\includegraphics[width=1\textwidth]{figures/pinconsistent_4panel_age_acchance_2noisemodelA_2noisemodelA.png}
%			\caption{}
%			\label{fig:mb2}
%		\end{center}
%	\end{figure}

	\subsubsection*{Model comparison}
	Previous section suggests that behavioral variability in random exploration is dominated by random noise. To test this more explicitly, we build a series of models in which different assumptions are made regarding the presence and absence of both types of noise and whether each type of noise if exists is horizon dependent (See Table \ref{tab:models}). In model A-D, we assumed the existence of both random and deterministic noise, in model A and B, random noise is assumed to be horizon-dependent, whereas in model A and C, deterministic noise is assumed to be horizon dependent. In model E, we assumed no deterministic noise. In model F, we assumed no random noise.

	\begin{table}[h]
		\small
		\centering
		\begin{tabular}{|c|c|c|}
			\hline
			Model & Deterministic noise & Random noise \\
			\hline
			A & Horizon dependent & Horizon dependent\\
			\hline
			B & Fixed & Horizon dependent\\
			\hline
			C & Horizon dependent & Fixed\\
			\hline
			D & Fixed & Fixed\\
			\hline
			E & None & Horizon dependent\\
			\hline
			F & Horizon dependent & None\\
			\hline
		\end{tabular}
		\caption{Model description. }%CHECK UPDATE}
		\label{tab:models}	
	\end{table}

	To evaluate and compare between models, we simulated choice behavior by taking the subject-level parameters from the Hierarchical Bayesian fits using each model. The same model-free analysis as described in the previous session is applied to all 6 sets of simulated data for the 6 models respectively. (See Figure  \ref{fig:mb3}).

	The original measure of random exploration, p(low mean), as used in \cite{wilson2014} can be explained by having deterministic noise alone (Figure \ref{fig:mb3}, Panel F2) or having random noise alone (Figure \ref{fig:mb3}, Panel E2). That participants qualitatively exploit the high-mean option less and choose the low-mean option more in horizon 6, can be explained by having both pure deterministic noise and pure random noise, as long as noise is horizon dependent. If both deterministic and random noise are assumed to be the same for both horizons (Figure \ref{fig:mb3}, Panel D), p(low mean) becomes completely flat and no horizon dependent random exploration is observed.

	On the other hand, by looking at the percentage of inconsistent choices in the repeated pair of game, p(inconsistent), deterministic noise alone can not account for behavior any more (Figure \ref{fig:mb3}, Panel F3, F4). Moreover, model C and D are disqualified that the increase of choice inconsistency with horizon can only be qualitatively accounted for when random noise is horizon-dependent (Figure \ref{fig:mb3}, Panel A, B, E).

	Among the models A, B and E, where random noise is horizon dependent, model A provides the best quantitative fit. If there is no deterministic noise (Model E), then we overestimate the level of choice inconsistency in both horizons by a constant. In addition, horizon dependent deterministic noise (Model A) gives slightly better model fits than if deterministic noise is assumed to be the same in both horizons (Model B). Overall, these model simulations confirmed that the horizon dependence of random noise is the main source of random exploration.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/RanDetNoise_2noise_6modelcomparison.png}
			\caption{Model comparison: model A - both deterministic and random noise are horizon dependent, model B - only random noise is horizon dependent, model C - only deterministic noise is horizon dependent, model D - neither random nor deterministic noise is horizon dependent, model E - only random noise is assumed to be present, model F - only deterministic noise is assumed to be present.}
			\label{fig:mb3}
		\end{center}
	\end{figure}

	\section*{Discussion}
	In this paper, we investigated whether random exploration is driven by random noise, putatively arising in the brain, or deterministic noise, arising from the environment. Using a version of the Horizon Task with repeated games, we found evidence for both types of noise in explore-exploit decisions. In addition, we see that both random and deterministic noise increase with horizon, but that the horizon effect is much larger for random noise.  Taken together our results suggest that random exploration, i.e. the use and adaptation of decision noise to drive exploration, is primarily driven by random noise.

	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being random and deterministic. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose. Thus, our estimate of deterministic noise is likely a lower bound. Likewise, our estimate of random noise is likely an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model. Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between random and deterministic noise that we found in Figure \ref{fig:mb1}, where random noise is 2-3 times the size of deterministic noise.

	Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}. In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.

	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise. Indeed, changes in the NE system have been associated with changes in behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

	More generally, our finding that random noise dominates behavioral variability over deterministic noise, is consistent with findings of \cite{drugowitsch16}. In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection. This suggests that most noise in behavior is generated randomly and that this may arise from  computational errors in computing the correct strategy. In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.

	\section*{Methods}
	\subsection*{Participants}

	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment. 15 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. This left 65 for the main analysis. Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	\subsection*{Task}
	The task was a modified version of the Horizon Task \citep{wilson2014}. In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. In each game they made multiple decisions between two options. Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution. The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game. One of the bandits always had a higher mean than the other. Participants were instructed to maximize the points earned over the entire task. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 153.7 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.12 minutes (mean 22.75 minutes).

	As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points. Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20. One of the means was always equal to either 40 or 60 and the second was set accordingly. Participants were informed that in every game one of the bandits always has a higher mean reward than the other. The order of games was randomized. Mean sizes and order of presentation were counterbalanced.

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game. For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome. From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 300ms. The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}. (See Figure \ref{fig:taskfig})

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfig}.

	\subsection*{Data and code}
	Behavioral data as well as Matlab code to recreate the main figures from this paper will be made available on the Dataverse website by publication. %website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.

	\subsection*{Model-based analysis}
	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate random and deterministic noise. In particular, we assume that in repeated games, deterministic noise remains the same whereas random noise can change.

	\subsubsection*{Hierarchical Bayesian Model}

	To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta    I+b+n_{det}+n_{ran}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference in information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.

	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.

	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the deterministic noise, $\sigma_{det}$, and the standard deviation of the random noise, $\sigma_{ran}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}). Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.  In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure. The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.

	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}. For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$. These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.

	\begin{table}[h]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			\hline
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
			\hline
			spatial bias, $b_{is}$ 
			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
			& \specialcell{
				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
			\hline
			deviation of deterministic noise, $\sigma^{det}_{is}$ 
			& $\sigma^{det}_{is} \sim $  Gamma($k_i^{det}$, $\lambda_{i}^{det}$) 
			& $\theta_{i}^{det} = (k_i^{det}, \lambda_i^{det}) $
			& \specialcell{
				$k_i^{det} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{det} \sim $ Exponential(10)}		\\
			\hline
			deviation of random noise, $\sigma^{ran}_{is}$ 
			& $\sigma^{ran}_{is} \sim $  Gamma($k_i^{ran}$, $\lambda_{i}^{ran}$) 
			& $\theta_{i}^{ran} = (k_i^{ran}, \lambda_i^{ran}) $
			& \specialcell{
				$k_i^{ran} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{ran} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
		\label{tab:pars2}	
	\end{table}

	\subsubsection*{Model fitting using MCMC}
	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags). This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.

	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).  Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.  Convergence of the Markov chains was confirmed {\it post hoc} by eye.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/EEHorizon_2sigma.pdf}
    			\caption{Schematic of the hierarchical Bayesian model using notation of \cite{lee_wagenmakers_2014}}
			\label{fig:model}
		\end{center}
	\end{figure}

	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}

	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.  In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters.

	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}. As is clear from this figure, parameter recovery is good for all parameters  (Figure \ref{fig:pararecover1}).  The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6. This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6. In addition we see better recovery for random noise than deterministic noise because we effectively have half as many trials for deterministic noise since we are only generating one sample of deterministic noise for each repeated game pair. Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.

	% apart from the bias, which is likely due to this parameter being so close to zero
	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=0.9\textwidth]{figures/RanDetNoise_parameter_recovery1.png}
			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games. 
			}
			\label{fig:pararecover1}
		\end{center}
	\end{figure}

	\bibliographystyle{plainnat}
	%\bibliographystyle{unsrt}
	\bibliography{Refs/refs}

	% add the Bibliography to the Table of Contents
	\cleardoublepage
	\ifdefined\phantomsection
	\phantomsection  % makes hyperref recognize this section properly for pdf link
	\else
	\fi
	\addcontentsline{toc}{chapter}{Bibliography}

\end{document}

### 2022-12-12 manuscript/2020 rejected - JEP General/SupplementaryFigures_RanDetNoise/DetRanNoise_Supplementary.tex
updates

--- 2022-12-12 manuscript/2020 rejected - JEP General/RandomDeterministicNoise/Ran-DetNoise.tex
@@
 \usepackage{amsmath}
+
+%\usepackage[maxbibnames=99]{biblatex}
+%\usepackage{setspace}
+%\usepackage{geometry}
 \usepackage[font={sf,small}]{caption}
+%\usepackage{setspace}
+%\usepackage{geometry}
+%\usepackage{hyperref}
+%\hypersetup{
+%    colorlinks,
+%    citecolor=black,
+%    filecolor=black,
+%    linkcolor=black,
+%    urlcolor=black
+%}
+%\geometry{letterpaper}
+
 \usepackage{amssymb}
+%\usepackage{epstopdf}
 \usepackage{float}
+%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
 
@@
 \renewcommand{\baselinestretch}{1.5}
+%\renewcommand{\figurename}{Supplementary Figure}
+\renewcommand{\thefigure}{S\arabic{figure}}
 
-\title{The nature of decision noise in random exploration}
+\title{Supplementary Figures: The nature of decision noise in random exploration}
 
@@
 
-\date{}
+\date{\today}
 
@@
 	\maketitle
-\setcounter{page}{28}
+
 	\newpage
-	\begin{abstract}
 
-Human decision making is inherently variable.
-While this variability is often seen as a sign of suboptimality in human behavior, recent work suggests that randomness can actually be adaptive.
-An example  arises when we must choose between exploring unknown options or exploiting options we know well.
-A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore.
-Moreover, people appear to use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore.
-From a modeling perspective, behavioral variability is essentially the variance that can not be explained by a model and is modeled as the level of decision noise.
-However, what we have called "decision noise" in previous researches could actually just be missing deterministic components from the model, it is difficult to tell whether decision noise truly arises from a stochastic process.
-Here we show that, while both random and deterministic noise drive variability in behavior, the noise driving random exploration is predominantly random.
-This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.
-
-\end{abstract}
-	\newpage
-	\section*{Introduction}
-
-	Imagine trying to decide where to go to dinner.
-You can go to your favorite restaurant, the one you really enjoy and always go to, or you can try a new restaurant that you know nothing about.
-Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.
-From a theoretical perspective, making optimal explore-exploit choices, i.e.
-choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}.
-In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.
-
-	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}.
-In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
-
-	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.
-In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.
-
-	It has recently been shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems and can increase such decision noise when it's more beneficial to explore \citep{Gershman2018, wilson2014}.
-In one of these tasks, known as the Horizon Task, the key manipulation is the horizon condition, i.e.
-the number of decisions remaining for the participant to make.
-Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards.
-For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant.
-Using such a horizon manipulation it has been shown that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.
-
-	It is however difficult in these tasks to tell whether what is measured as decision noise is really random, what we have called 'noise' in previous researches could actually just be some missing deterministic components from the model.
-Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model.
-In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants.
-If the model did not consider that the agent is favoring the behavior of following a friend in a deterministic way, the behavior of going to a less favorable restaurant because of a friend will appear to be 'random' when it is really a deterministic effect.
-hence this deterministic factor will be modeled as a random decision noise.
-Crucially, however, this `deterministic noise' is very much in the stimulus and if you saw the same friend go into the same restaurant at a later date you might follow them again.
-Conversely, truly 'random noise' would arise from stochastic mental processes tossing a metaphorical coin in your head.
-Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.
-
-	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.
-In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.
-
-    % The idea here is behavior is less predictable in horizon 6 than horizon 1 even when all stimuli are the same.
-Thus anything that deterministically depends on stimuli - i.e.
-most modeling effects - are ruled out as an explanation for increased noise.
-In that light, what your model calls deterministic noise is that amount of variance that could be explained by deterministic effects of stimulus - i.e.
-modeling errors.
-
-    %* Difficulty with measuring random exploration, how do you know something is random, previous work model
-    %* Model all the deterministic factor (contribute to noise)
-    %* Have repeated games
-    %* Missing deterministic dependent, much smaller than non-deterministic component
-    %* Stochastic process…
-
-	%In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants.
-Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in.
-Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again.
-
-	%Conversely, random noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
-Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.
-
-	%Previous work makes a strong case for both deterministic and random noise being relevant to behavior.
-For instance, deterministic, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than random noise \citep{eeBrunton13}.
-Conversely random, neural noise is thought to drive exploratory singing behavior in song birds and the generation and control of this random noise has been linked to specific neural structures \citep{songbird2}.
-
-	In this paper, we investigate which source of noise, deterministic vs random, drives random exploration in humans in a modified version of the Horizon Task.
-To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice.
-If decision noise is purely deterministic noise, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times.
-Meanwhile, if decision noise is truly random their choices should be less consistent, since random noise can be different both times.
-By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the variability related to random exploration is dominated by random noise.
-The missing deterministic component is much smaller than the non-deterministic component in random exploration.
-
-	\section*{Results}
-
-	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of random vs deterministic noise on people's decisions (Figure \ref{fig:taskfig}).
-In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
-Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
-
-	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
-1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
-These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
-Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.
-
-	These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.
-Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent.
-Conversely if noise is randomly driven, then choices on repeated games should be independent and can be inconsistent.
-
-	\begin{figure}[hp]
+	\begin{figure}[H]
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
-			\caption{ 
-			Schematic of the experiment.
-(A) Dynamics of an example horizon 6 game.
-Here the first four trials are forced trials in which participants are instructed which option to play.
-After the forced trials, participants are free to choose between the two options for the remainder of the game.
-(B) Different possible states of the game after the first free choice over the course of the experiment.
-Overall participants play about 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
-In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.
-These repeated games allow us to compute the relative contribution of deterministic and random noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}
-			\label{fig:taskfig}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree_all.png}
+			\caption{Replication of previous findings with data from all participants (i.e.
+no exclusions).
+Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.
+}
+			\label{fig:modelfree2}
 		\end{center}
@@
 
-	\subsection*{Both behavioral variability and information seeking increase with horizon}
-
-	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}.
-As in our previous work, we find evidence for two types of exploration in the Horizon Task.
-Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.
-
-	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
-This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(64) = 6.55, p $<$ 0.001 for [1 3], t(64) = 7.99, p $<$ 0.001 for [2 2]).
-Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition.
-Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(64) = 6.92, p $<$ 0.001).
-
 	\begin{figure}[H]
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.png}
-			\caption{
-			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
-			\label{fig:modelfree}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_2noise_all.png}
+			\caption{Model-free analysis with data from all participants (i.e.
+no exclusions) suggests that both deterministic and random noise contribute to the choice variability in random exploration.
+For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
+However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
+			\label{fig:mf22}
+		\end{center}
+	\end{figure}
+	\newpage
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RanDetNoise_hyperprior_all.png}
+			\caption{Model based analysis with data from all participants (i.e.
+no exclusions) showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
+			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
+			\label{fig:mb12}
 		\end{center}
@@
 
-	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}
-
-	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.
-The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times.
-Conversely, purely random noise should lead to independent choices, and hence more inconsistent choices both times.
-
-	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was deterministic (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions.
-For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6).
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] (t(64) = 5.41, p $<$ 0.001) and [2 2] condition (t(64) = 6.26, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is random.
-
-	\begin{figure}[h]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_2noise.png}
-			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration.
-For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
-However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
-			\label{fig:mf2}
-		\end{center}
-	\end{figure}
-
-	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely deterministic and purely random noise cases.
-For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case.
-For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
-	\begin{equation*}
-	\begin{split}
-	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
-	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
-	\mbox{hence},\quad p(\mbox{inconsistent}) &=  
-	1 - p(\mbox{consistent}) = 
-	2 p(\mbox{low mean})(1-p(\mbox{low mean}))
-	\end{split}
-	\end{equation*}
-	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction (Behavior is different from pure random noise prediction in both [1 3] condition, t(64) = 8.66, p $<$ 0.001 for horizon 1, t(64) = 9.48, p $<$ 0.001 for horizon 6; and [2 2] condition, t(64) = 6.94, p $<$ 0.001 for horizon 1, t(64) = 7.47, p $<$ 0.001 for horizon 6.
-Behavior is different from pure deterministic noise prediction in both [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; and [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6.), suggesting that both deterministic and random noise are present in driving this choice inconsistency.
-Since choice inconsistency only reflects random noise, Figure \ref{fig:mf2} suggests that random noise increases with horizon.
-
-	\subsection*{Model-based analysis shows that random exploration is dominated by random noise}
-
-	To more precisely quantify random and deterministic noise, we turned to model fitting.
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate random and deterministic noises.
-In particular, we assume that in repeated games, deterministic noise remains the same whereas random noise can change.
-
-	\subsubsection*{Overview of model}
-	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
-To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
-	\begin{equation}
-	\Delta Q= \Delta R+A \Delta    I+b+n_{det}+n_{ran}
-	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference between information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively which are assumed to come from logistic distribution with mean 0.
-
-	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.
-
-	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the deterministic noise, $n_{det}$, should be the same while the random noise, $n_{ran}$ may be different.
-This is exactly how we distinguish deterministic noise from random noise.
-In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.
-
-	\subsubsection*{Model fitting}
-	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
-In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon.
-Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
-
-	\subsubsection*{Model fitting results}
-
-	Posterior distributions over the group-level means of the deterministic and random noise variance are shown in Figure \ref{fig:mb1}.
-Consistent with our model-free results, we see that both random and deterministic noise variances are non-zero and that random noise is about 2-3 times larger than the deterministic noise.
-In addition, we find that random noise increases dramatically with horizon (M = 4.55, 100\% of samples showed an increase in random noise with horizon) whereas the increase in deterministic noise is smaller (M = 1.78, 98.12\% of samples showed an increase in deterministic noise with horizon).
-Taken together, these results suggest that random exploration is dominated by random noise.
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_hyperprior.png}
-			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
-Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
-However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
-			\label{fig:mb1}
-		\end{center}
-	\end{figure}
-
-	%\begin{figure}[H]
-	%	\begin{center}
-%			\includegraphics[width=1\textwidth]{figures/pinconsistent_4panel_age_acchance_2noisemodelA_2noisemodelA.png}
-%			\caption{}
-%			\label{fig:mb2}
-%		\end{center}
-%	\end{figure}
-
-	\subsubsection*{Model comparison}
-	Previous section suggests that behavioral variability in random exploration is dominated by random noise.
-To test this more explicitly, we build a series of models in which different assumptions are made regarding the presence and absence of both types of noise and whether each type of noise if exists is horizon dependent (See Table \ref{tab:models}).
-In model A-D, we assumed the existence of both random and deterministic noise, in model A and B, random noise is assumed to be horizon-dependent, whereas in model A and C, deterministic noise is assumed to be horizon dependent.
-In model E, we assumed no deterministic noise.
-In model F, we assumed no random noise.
-
-	\begin{table}[h]
-		\small
-		\centering
-		\begin{tabular}{|c|c|c|}
-			\hline
-			Model & Deterministic noise & Random noise \\
-			\hline
-			A & Horizon dependent & Horizon dependent\\
-			\hline
-			B & Fixed & Horizon dependent\\
-			\hline
-			C & Horizon dependent & Fixed\\
-			\hline
-			D & Fixed & Fixed\\
-			\hline
-			E & None & Horizon dependent\\
-			\hline
-			F & Horizon dependent & None\\
-			\hline
-		\end{tabular}
-		\caption{Model description. }%CHECK UPDATE}
-		\label{tab:models}	
-	\end{table}
-
-	To evaluate and compare between models, we simulated choice behavior by taking the subject-level parameters from the Hierarchical Bayesian fits using each model.
-The same model-free analysis as described in the previous session is applied to all 6 sets of simulated data for the 6 models respectively.
-(See Figure  \ref{fig:mb3}).
-
-	The original measure of random exploration, p(low mean), as used in \cite{wilson2014} can be explained by having deterministic noise alone (Figure \ref{fig:mb3}, Panel F2) or having random noise alone (Figure \ref{fig:mb3}, Panel E2).
-That participants qualitatively exploit the high-mean option less and choose the low-mean option more in horizon 6, can be explained by having both pure deterministic noise and pure random noise, as long as noise is horizon dependent.
-If both deterministic and random noise are assumed to be the same for both horizons (Figure \ref{fig:mb3}, Panel D), p(low mean) becomes completely flat and no horizon dependent random exploration is observed.
-
-	On the other hand, by looking at the percentage of inconsistent choices in the repeated pair of game, p(inconsistent), deterministic noise alone can not account for behavior any more (Figure \ref{fig:mb3}, Panel F3, F4).
-Moreover, model C and D are disqualified that the increase of choice inconsistency with horizon can only be qualitatively accounted for when random noise is horizon-dependent (Figure \ref{fig:mb3}, Panel A, B, E).
-
-	Among the models A, B and E, where random noise is horizon dependent, model A provides the best quantitative fit.
-If there is no deterministic noise (Model E), then we overestimate the level of choice inconsistency in both horizons by a constant.
-In addition, horizon dependent deterministic noise (Model A) gives slightly better model fits than if deterministic noise is assumed to be the same in both horizons (Model B).
-Overall, these model simulations confirmed that the horizon dependence of random noise is the main source of random exploration.
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_2noise_6modelcomparison.png}
-			\caption{Model comparison: model A - both deterministic and random noise are horizon dependent, model B - only random noise is horizon dependent, model C - only deterministic noise is horizon dependent, model D - neither random nor deterministic noise is horizon dependent, model E - only random noise is assumed to be present, model F - only deterministic noise is assumed to be present.}
-			\label{fig:mb3}
-		\end{center}
-	\end{figure}
-
-	\section*{Discussion}
-	In this paper, we investigated whether random exploration is driven by random noise, putatively arising in the brain, or deterministic noise, arising from the environment.
-Using a version of the Horizon Task with repeated games, we found evidence for both types of noise in explore-exploit decisions.
-In addition, we see that both random and deterministic noise increase with horizon, but that the horizon effect is much larger for random noise.
-Taken together our results suggest that random exploration, i.e.
-the use and adaptation of decision noise to drive exploration, is primarily driven by random noise.
-
-	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being random and deterministic.
-In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
-the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose.
-Thus, our estimate of deterministic noise is likely a lower bound.
-Likewise, our estimate of random noise is likely an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
-Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between random and deterministic noise that we found in Figure \ref{fig:mb1}, where random noise is 2-3 times the size of deterministic noise.
-
-	Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
-In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
-
-	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
-Indeed, changes in the NE system have been associated with changes in behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
-In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}
-
-	More generally, our finding that random noise dominates behavioral variability over deterministic noise, is consistent with findings of \cite{drugowitsch16}.
-In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
-This suggests that most noise in behavior is generated randomly and that this may arise from  computational errors in computing the correct strategy.
-In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.
-
-	\section*{Methods}
-	\subsection*{Participants}
-
-	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment.
-15 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-This left 65 for the main analysis.
-Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
-
-	\subsection*{Task}
-	The task was a modified version of the Horizon Task \citep{wilson2014}.
-In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
-In each game they made multiple decisions between two options.
-Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution.
-The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game.
-One of the bandits always had a higher mean than the other.
-Participants were instructed to maximize the points earned over the entire task.
-To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.
-
-	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task.
-Thus, the better participants performed, the sooner they got to leave the experiment.
-On average, participants played 153.7 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.12 minutes (mean 22.75 minutes).
-
-	As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points.
-Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20.
-One of the means was always equal to either 40 or 60 and the second was set accordingly.
-Participants were informed that in every game one of the bandits always has a higher mean reward than the other.
-The order of games was randomized.
-Mean sizes and order of presentation were counterbalanced.
-
-	Each game consisted of 5 or 10 choices.
-Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game.
-For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome.
-From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision.
-There was no time limit for decisions.
-During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit.
-The score feedback was presented for 300ms.
-The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.
-(See Figure \ref{fig:taskfig})
-
-	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
-The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfig}.
-
-	\subsection*{Data and code}
-	Behavioral data as well as Matlab code to recreate the main figures from this paper will be made available on the Dataverse website by publication.
-%website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.
-
-	\subsection*{Model-based analysis}
-	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate random and deterministic noise.
-In particular, we assume that in repeated games, deterministic noise remains the same whereas random noise can change.
-
-	\subsubsection*{Hierarchical Bayesian Model}
-
-	To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
-	\begin{equation}
-	\Delta Q= \Delta R+A \Delta    I+b+n_{det}+n_{ran}
-	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference in information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.
-
-	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.
-
-	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the deterministic noise, $\sigma_{det}$, and the standard deviation of the random noise, $\sigma_{ran}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}).
-Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.
-In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure.
-The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.
-
-	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}.
-For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$.
-These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.
-
-	\begin{table}[h]
-		\small
-		\begin{tabular}{|c|c|c|c|}
-			\hline
-			Parameter & Prior & Hyperparameters & Hyperpriors \\
-			\hline
-			information bonus, $A_{is}$ 
-			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
-			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
-			& \specialcell{
-				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
-			\hline
-			spatial bias, $b_{is}$ 
-			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
-			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
-			& \specialcell{
-				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
-			\hline
-			deviation of deterministic noise, $\sigma^{det}_{is}$ 
-			& $\sigma^{det}_{is} \sim $  Gamma($k_i^{det}$, $\lambda_{i}^{det}$) 
-			& $\theta_{i}^{det} = (k_i^{det}, \lambda_i^{det}) $
-			& \specialcell{
-				$k_i^{det} \sim $ Exponential(0.01) \\ 
-				$\lambda_i^{det} \sim $ Exponential(10)}		\\
-			\hline
-			deviation of random noise, $\sigma^{ran}_{is}$ 
-			& $\sigma^{ran}_{is} \sim $  Gamma($k_i^{ran}$, $\lambda_{i}^{ran}$) 
-			& $\theta_{i}^{ran} = (k_i^{ran}, \lambda_i^{ran}) $
-			& \specialcell{
-				$k_i^{ran} \sim $ Exponential(0.01) \\ 
-				$\lambda_i^{ran} \sim $ Exponential(10)}		\\
-			\hline
-		\end{tabular}
-		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
-		\label{tab:pars2}	
-	\end{table}
-
-	\subsubsection*{Model fitting using MCMC}
-	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags).
-This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.
-
-	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).
-Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
-Convergence of the Markov chains was confirmed {\it post hoc} by eye.
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/EEHorizon_2sigma.pdf}
-    			\caption{Schematic of the hierarchical Bayesian model using notation of \cite{lee_wagenmakers_2014}}
-			\label{fig:model}
-		\end{center}
-	\end{figure}
-
-	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}
-
-	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.
-In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters.
-
-	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}.
-As is clear from this figure, parameter recovery is good for all parameters  (Figure \ref{fig:pararecover1}).
-The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
-This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6.
-In addition we see better recovery for random noise than deterministic noise because we effectively have half as many trials for deterministic noise since we are only generating one sample of deterministic noise for each repeated game pair.
-Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.
-
-	% apart from the bias, which is likely due to this parameter being so close to zero
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=0.9\textwidth]{figures/RanDetNoise_parameter_recovery1.png}
-			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
-			}
-			\label{fig:pararecover1}
-		\end{center}
-	\end{figure}
-
-	\bibliographystyle{plainnat}
+	%\bibliographystyle{plainnat}
 	%\bibliographystyle{unsrt}
-	\bibliography{Refs/refs}
+	%\bibliography{Refs/refs}
 
 	% add the Bibliography to the Table of Contents
-	\cleardoublepage
-	\ifdefined\phantomsection
-	\phantomsection  % makes hyperref recognize this section properly for pdf link
-	\else
-	\fi
-	\addcontentsline{toc}{chapter}{Bibliography}
+	%\cleardoublepage
+	%\ifdefined\phantomsection
+	%\phantomsection  % makes hyperref recognize this section properly for pdf link
+	%\else
+	%\fi
+	%\addcontentsline{toc}{chapter}{Bibliography}
 

### 2022-12-12 manuscript/2022 draft - Plos Comp Bio/OlderVersion/08-05-20/Internal-External Noise.tex
updates

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[authoryear, round,sort,comma,numbers]{natbib}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage[font={sf,small}]{caption}
\usepackage{amssymb}
\usepackage{float}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\title{The nature of decision noise in random exploration}

\author[1]{Siyu Wang}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
	\maketitle

	\newpage
	\begin{abstract}

Human decision making is inherently variable. While this variability is often seen as a sign of suboptimality behavior, recent work suggests that variability can actually be adaptive. An example arises when we must choose between exploring unknown options or exploiting options we know well. A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore. Recent work suggests that people may actually use such `random exploration' in practice, increasing their behavioral variability when it is more valuable to explore. 
%Other work, however, suggests that variability in human behavior is not always random, and is instead driven by deterministic factors (such as repeating past actions) 
A key question is whether the variability in random exploration is actually random. That is, is random exploration driven by stochastic processes in the brain or by some unobserved deterministic process that we have failed to account for when measuring behavioral variability? By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question. In particular, we find evidence that at least part (IT SOUNDS HARD TO ME TO ARGUE THAT THIS NUMBER IS NOT TASK SPECIFIC. - BOB SAYS: CAN WE PUT A NUMBER ON THIS E.G. 25\%?) of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus. This still leaves open the possibility that much of random exploration is truly `random,' but narrows the window of opportunity for stochastic processes to explain this behavior.
%suggests that at least some `random' exploration is not random, but leaves open the possibility that the 
%However, it is unclear whether this variability is truly random
%From a modeling perspective, behavioral variability is essentially the variance that can not be explained by a model and is modeled as the level of decision noise. However, what we have called "decision noise" in previous researches could actually just be missing deterministic components from the model, it is difficult to tell whether decision noise truly arises from a stochastic process. Here we show that, while both random and deterministic noise drive variability in behavior, the noise driving random exploration is predominantly random. This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.

\end{abstract}
	\newpage
	\section*{Introduction}

	%BOB SAYS: IN GENERAL, NEED TO GO THROUGH THE PAPER AND MAKE SURE THAT `RANDOM' NOISE IS REPLACES WITH `NOT A DETERMINISTIC FUNCTION OF THE STIMULUS' OR SOMETHING LIKE THAT

	%(BOB SAYS: TAKE OR LEAVE THIS CHANGE TO DATE, BUT SETTING UP AS A DATE ALLOWS YOU TO TALK ABOUT AVOID YOUR EX LATER ON, WHICH MAY BE A BETTER EXAMPLE OF A DETERMINISTIC PROCESS!).
	Imagine trying to decide where to go to dinner on a date, you can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try a new restaurant that you know nothing about. Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.  From a theoretical perspective, making optimal explore-exploit choices, i.e. choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}. In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.

	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}. In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance. In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.

	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.  In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.

	It has recently been shown that humans appear to use random exploration and can increase such decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014}. In one of these tasks, known as the Horizon Task, the key manipulation is the horizon condition, i.e. the number of decisions remaining for the participant to make. Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards. For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant. Using such a horizon manipulation we found that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.

	One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random. That is, whether behavioral variability is due to intrinsic stochastic processes in the brain or whether it is due to deterministic processes that we simply have not observed.  For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them. Avoiding an ex is a deterministic process, but if we fail to take ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.

	%This difficulty arises in how noise is defined in these previous tasks. For example, in the Horizon Task, variability is defined as the inverse temperature parameter of a softmax choice function.

	%In particular, in most cases noise is defined by fitting a softmax decision rule to behavior.

	%because noise in previous work is defined as the component of the decision process that cannot be accounted for by a model. 
	%, what we have called 'noise' in previous research could actually just be some missing deterministic components from the model. Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model. 
	%For example, in the restaurant case above, if you happen to spot an old friend walking in to one of the restaurants, you may be more likely to follow them in. Such a process would be deterministic, in that it is driven by an observable stimulus, but unless a model of your behavior incorporated the presence of your friend,

	%. If the model did not consider that the agent is favoring the behavior of following a friend in a deterministic way, the behavior of going to a less favorable restaurant because of a friend will appear to be `random' when it is really a deterministic effect. Hence this deterministic factor will be modeled as random decision noise. Crucially, however, this deterministic source of variability is very much in the stimulus and if you saw the same friend go into the same restaurant at a later date you might follow them again. Conversely, truly `random' noise would arise from stochastic mental processes,

	%tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.

	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.  In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.

    % The idea here is behavior is less predictable in horizon 6 than horizon 1 even when all stimuli are the same.  Thus anything that deterministically depends on stimuli - i.e. most modeling effects - are ruled out as an explanation for increased noise.  In that light, what your model calls deterministic noise is that amount of variance that could be explained by deterministic effects of stimulus - i.e. modeling errors.

    %* Difficulty with measuring random exploration, how do you know something is random, previous work model
    %* Model all the deterministic factor (contribute to noise)
    %* Have repeated games
    %* Missing deterministic dependent, much smaller than non-deterministic component
    %* Stochastic process…

	%In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants. Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in. Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again.

	%Conversely, random noise would arise from stochastic neural processes tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.

	%Previous work makes a strong case for both deterministic and random noise being relevant to behavior. For instance, deterministic, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than random noise \citep{eeBrunton13}. Conversely random, neural noise is thought to drive exploratory singing behavior in song birds and the generation and control of this random noise has been linked to specific neural structures \citep{songbird2}.

	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by such deterministic processing of the stimulus. In particular, we modify the Horizon Task to have people face the exact same explore-exploit choice twice. If the decision is a purely deterministic function of the stimulus, then people's choices should be identical both times. That is, their choices should be consistent, since the stimulus is the same both times. Conversely, the more their decision is driven by other processes, including both stochastic and unobserved deterministic processes, the less consistent their behavior should be. By analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from  deterministic processing of the stimulus.  This does not prove that random exploration is entirely deterministic or stochastic, but provides a lower bound on how much deterministic processes contribute to random exploration. This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.

	%which source of noise, deterministic vs random, drives random exploration in humans in a modified version of the Horizon Task. To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice. If decision noise is purely deterministic noise, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times. Meanwhile, if decision noise is truly random their choices should be less consistent, since random noise can be different both times. By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the variability related to random exploration is dominated by random noise. The missing deterministic component is much smaller than the non-deterministic component in random exploration.

		\section*{Methods}
	\subsection*{Participants}

	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment. 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}. In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved driven by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p > 0.001$ were excluded in the analysis. This left 66 for the main analysis. Note that including the 14 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS

	\subsection*{Task}
	The task was a modified version of the Horizon Task \citep{wilson2014} (Figure \ref{fig:taskfig}). In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. In each game they made multiple decisions between two options. Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution. The means of the underlying Gaussians were different for the two bandit options, remained the same within a game, but changed with each new game. One of the bandits always had a higher mean than the other. Participants were instructed to maximize the points earned over the entire task. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

		\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
			\caption{ 
			Schematic of the experiment. (A) Dynamics of an example horizon 6 game.  Here the first four trials are forced trials in which participants are instructed which option to play.  After the forced trials, participants are free to choose between the two options for the remainder of the game.  (B) Different possible states of the game after the first free choice over the course of the experiment. Overall participants play 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.  In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.  These repeated games allow us to compute the relative contribution of deterministic and random noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}
			\label{fig:taskfig}
		\end{center}
	\end{figure}

	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task. Thus, the better participants performed, the sooner they got to leave the experiment. On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes). Participants played an average of 67.31 repeated pairs of games (minimum = 25 repeated pairs, maximum =  100 repeated pairs).
	%BOB SAYS: ADD HOW MANY REPEATED GAMES THEY PLAYED - GIVE A MEAN AND RANGE FOR THIS NUMBER

	As in the original paper \citep{wilson2014}, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points. Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20. One of the means was always equal to either 40 or 60 and the second was set accordingly. Participants were informed that in every game one of the bandits always has a higher mean reward than the other. The order of games was randomized. Mean sizes and order of presentation were counterbalanced.

	Each game consisted of 5 or 10 choices. Every game started with a fixation cross, then a bar of boxes appeared indicating the horizon for that game. For the first 4 trials - the instructed trials, we highlight the box on one of the bandits to instruct the participant to choose that option. On these trials, they have to press the corresponding key to reveal the outcome. From the fifth trial, boxes on both bandits will be highlighted and they are free to make their own decision. There was no time limit for decisions. During free choices participants could press either the left arrow key or right arrow key to indicate their choice of left or right bandit. The score feedback was presented for 300ms. The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.

	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose. We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice. The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice. After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfig}.

	\subsection*{Model-based analysis}
	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').  Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.

	\subsubsection*{Hierarchical Bayesian Model}

	To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta I+b+n_{det}+n_{ran}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of the rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.

	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.

	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the deterministic noise, $\sigma_{det}$, and the standard deviation of the random noise, $\sigma_{ran}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}). Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.  In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure. The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.

	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}. For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$. These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.

	\begin{table}[h]
		\small
		\begin{tabular}{|c|c|c|c|}
			\hline
			Parameter & Prior & Hyperparameters & Hyperpriors \\
			\hline
			information bonus, $A_{is}$ 
			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
			& \specialcell{
				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
			\hline
			spatial bias, $b_{is}$ 
			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
			& \specialcell{
				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
			\hline
			deviation of deterministic noise, $\sigma^{det}_{isg}$ 
			& $\sigma_{i} \sim $  Gamma($k_i^{det}$, $\lambda_{i}^{det}$) 
			& $\theta_{i}^{det} = (k_i^{det}, \lambda_i^{det}) $
			& \specialcell{
				$k_i^{det} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{det} \sim $ Exponential(10)}		\\
			\hline
			deviation of random noise, $\sigma^{ran}_{isgr}$ 
			& $\sigma_{is} \sim $  Gamma($k_i^{ran}$, $\lambda_{i}^{ran}$) 
			& $\theta_{i}^{ran} = (k_i^{ran}, \lambda_i^{ran}) $
			& \specialcell{
				$k_i^{ran} \sim $ Exponential(0.01) \\ 
				$\lambda_i^{ran} \sim $ Exponential(10)}		\\
			\hline
		\end{tabular}
		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
		\label{tab:pars2}	
	\end{table}

	\subsubsection*{Model fitting using MCMC}
	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags). This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.

	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).  Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.  Convergence of the Markov chains was confirmed {\it post hoc} by eye.

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
    			\caption{Schematic of the hierarchical Bayesian model using notation of \cite{lee_wagenmakers_2014}}
			\label{fig:model}
		\end{center}
	\end{figure}

	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}

	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.  In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated choices to see whether we can recover the parameters.

	Results of this parameter recovery procedure are shown in Supplementary Figure S4. As is clear from this figure, parameter recovery is good in terms of correlation between the true vs fitted parameters for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Supplementary Figure S4 Panel C). The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6. This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6. In addition we see better recovery for random noise than deterministic noise because we effectively have half as many trials for deterministic noise since we are only generating one sample of deterministic noise for each repeated game pair.

	One limitation of our model is that we observed a systematic underestimation of deterministic noise (despite the strong correlation). This is further illustrated in Supplementary Figure S5. In thee simulation with 0 random noise and full deterministic noise, our model perfectly recovered both random and deterministic noise, however in the simulation with full random noise and 0 deterministic noise, our model recovered only XXX \% of deterministic noise, and the remaining shows up as random noise instead (See Supplementary Figure S5). Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.

	Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.

	\subsection*{Data and code}
	Behavioral data as well as Matlab code to recreate the main figures from this paper will be made available on the Dataverse website upon publication. %website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.

	\section*{Results}

	\subsection*{The Repeated-Games Horizon Task}
	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus driven `random noise' on people's decisions (Figure \ref{fig:taskfig}). In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards. Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.

	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e. 1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition). These example plays allow us to manipulate exactly what people know about each option before they make their first choice.

	Relative to the original Horizon Task, the key modification here is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games (separated by several minutes in time so as to avoid detection). By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice. Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their behavior on the two versions of each game will be consistent. Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games should be inconsistent.

	On average participants played 67.31 repeated games each allowing us to quantify the extent to which their behavior was a deterministic function of the stimulus or not.

	%, the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.

	%These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.  Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent. Conversely if noise is randomly driven, then choices on repeated games should be inconsistent.

	\subsection*{Both behavioral variability and information seeking increase with horizon}

	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}. As in our previous work, we find evidence for two types of exploration in the Horizon Task.  Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.

	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition. This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(65) = 6.60, p $<$ 0.001 for [1 3], t(65) = 8.14, p $<$ 0.001 for [2 2]).  Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition. Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(65) = 7.00, p $<$ 0.001).

	\begin{figure}[H]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
			\caption{
			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
			\label{fig:modelfree}
		\end{center}
	\end{figure}

	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}

	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.  The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times. Conversely, if choice is not a deterministic function of the stimulus, participants' choices should be independent, and hence more inconsistent across the repetitions of the game.

	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}). Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was stimulus driven (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions.  For [1 3] condition, t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6; For [2 2] condition, t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6). In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 196) = 61.19, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is not a deterministic function of the stimulus.

	\begin{figure}[h]
		\begin{center}
			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration. For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1. However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
			\label{fig:mf2}
		\end{center}
	\end{figure}

	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely deterministic and purely random noise cases.  For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case. For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
	\begin{equation*}
	\begin{split}
	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
	\mbox{hence},\quad p(\mbox{inconsistent}) &=  
	1 - p(\mbox{consistent}) = 
	2 p(\mbox{low mean})(1-p(\mbox{low mean}))
	\end{split}
	\end{equation*}
	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction Specifically, behavior is different from pure random noise prediction in the both the [1 3] condition (t(65) = 5.60, p $<$ 0.001 for horizon 1, t(65) = 5.62 p $<$ 0.001 for horizon 6) and [2 2] condition (t(65) = 3.86, p $<$ 0.001 for horizon 1, t(65) = 3.42, p $<$ 0.001 for horizon 6). Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition ( t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6). This suggests that at least some of the `noise' is a deterministic function of the stimulus, although from this analysis it is not clear whether this deterministic noise increases with horizon or not.

	\subsection*{Model-based analysis shows deterministic noise changes with horizon}

	To more precisely quantify the contribution of stimulus-driven deterministic noise, we turned to model fitting. We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate deterministic noise from and random noise. In particular, we assume that in repeated games, the value of stimulus-driven deterministic noise is frozen whereas random noise is drawn independently both times.

	\subsubsection*{Overview of model}
	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits. To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
	\begin{equation}
	\Delta Q= \Delta R+A \Delta I+b+n_{det}+n_{ran}
	\end{equation}
	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial. For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0. $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively which are assumed to come from logistic distributions with mean 0.

	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.

	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the deterministic noise, $n_{det}$, should be the same while the random noise, $n_{ran}$ may be different. This is exactly how we distinguish deterministic noise from random noise. In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.

	\subsubsection*{Model fitting}
	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}). In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon. Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.

	\subsubsection*{Model fitting results}

	Posterior distributions over the group-level means of the deterministic and random noise variance are shown in Figure \ref{fig:mb1}. Consistent with our model-free results, we see that both random and deterministic noise variances are non-zero and that random noise is about 2-3 times larger than the deterministic noise. In addition, we find that both random and deterministic noise increase with horizon.  This increase was larger for random noise (M = 4.55, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (M = 1.78, 98.12\% of samples showed an increase in deterministic noise with horizon). But intriguingly the relative increase in both types of noise was the same (Figure \ref{fig:ratio}). That is when we compute the relative increase in deterministic noise with horizon, $\sigma_{det, horizon6}/\sigma_{det, horizon1}$, it is almost identical to the relative increase in random noise with horizon $\sigma_{ran, horizon6}/\sigma_{ran, horizon1}$. 
	%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO

	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise. Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).  However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
			%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
			\label{fig:mb1}
		\end{center}
	\end{figure}

	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=0.6\textwidth]{figures/RanDetNoise_ratio1.png}
			\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly. The ratio in the standard deviations of noise between horizon 6 and horizon 1 is similar for random and deterministic noise.}
			\label{fig:ratio}
		\end{center}
	\end{figure}

	%\begin{figure}[H]
	%	\begin{center}
%			\includegraphics[width=1\textwidth]{figures/pinconsistent_4panel_age_acchance_2noisemodelA_2noisemodelA.png}
%			\caption{}
%			\label{fig:mb2}
%		\end{center}
%	\end{figure}

	\subsubsection*{Posterior predictive checks}
	\begin{figure}[hp]
		\begin{center}
			\includegraphics[width=1\textwidth]{figures/pinconsistent_6panel_age_acchance.png}
			\caption{
			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.  I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ... COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
			Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
			\label{fig:mb3}
		\end{center}
	\end{figure}

	In addition to fitting the model to behavior, it is also important to check whether the model captures the qualitative patterns of the data \citep{Wilson2019} --- specifically how p(high info), p(low mean) and p(inconsistent) change with horizon.

	To perform this `posterior predictive check,' we created a set of simulated data by taking the subject-level parameters from the hierarchical Bayesian fits and having the model play the same sequence of games as seen by the subjects. We then applied the same model-free analysis as described in the previous sections to this simulated data set and compared the model's behavior to that of participants. As shown in Figure  \ref{fig:mb3}, the model can account for all qualitative patterns in the data --- the increase in p(high info), p(low mean), and p(inconsistent) with horizon, and that p(inconsistent) is in between pure random and pure deterministic noise.  The quantitative agreement is almost perfect for p(high info), but the model seems to systematically underestimate p(low mean) and p(inconsistent), although the discrepancy is small (underestimating p(low mean) by 0.047 or 27.30\%, and p(inconsistent) by 0.027 or 14.27\%).

	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.  These models varied whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon. As shown in Supplementary Figure S5, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of responding (Note model F does not capture that  p(inconsistent) is between pure random and deterministic?). However, the quantitative fit to the data is not as good (Supplementary Figure S5).

	%The previous section suggests that behavioral variability in random exploration can be at least partially explained by deterministic processing of the stimulus.

	%dominated by random noise. To test this more explicitly, we build a series of models in which different assumptions are made regarding the presence and absence of both types of noise and whether each type of noise if exists is horizon dependent (See Table \ref{tab:models}). In model A-D, we assumed the existence of both random and deterministic noise, in model A and B, random noise is assumed to be horizon-dependent, whereas in model A and C, deterministic noise is assumed to be horizon dependent. In model E, we assumed no random noise. In model F, we assumed no deterministic noise.

	%To evaluate and compare between models, we simulated choice behavior by taking the subject-level parameters from the Hierarchical Bayesian fits. The same model-free analysis as described in the previous session is applied to all 6 sets of simulated data for the 6 models respectively. (See Figure  \ref{fig:mb3}).

	%The original measure of random exploration, p(low mean), as used in \cite{wilson2014} can be explained by having deterministic noise alone (Figure \ref{fig:mb3}, Panel E2) or having random noise alone (Figure \ref{fig:mb3}, Panel F2). That participants qualitatively exploit the high-mean option less and choose the low-mean option more in horizon 6, can be explained by having both pure deterministic noise and pure random noise, as long as noise is horizon dependent. If both deterministic and random noise are assumed to be the same for both horizons (Figure \ref{fig:mb3}, Panel D), p(low mean) becomes completely flat and no horizon dependent random exploration is observed.

	%On the other hand, by looking at the percentage of inconsistent choices in the repeated pair of game, p(inconsistent), external noise alone can not account for behavior any more (Figure \ref{fig:mb3}, Panel E3, E4). Moreover, model C and D are disqualified that the increase of choice inconsistency with horizon can only be qualitatively accounted for when random noise is horizon-dependent (Figure \ref{fig:mb3}, Panel A, B, F).

	%Among the models A, B and F, where random noise is horizon dependent, model A provides the best quantitative fit. If there is no deterministic noise (Model F), then we overestimate the level of choice inconsistency in both horizons by a constant. In addition, horizon dependent deterministic noise gives slightly better model fits than if deterministic noise is assumed to be the same in both horizons. Overall, these model simulations confirmed that the horizon dependence of random noise is the main source of random exploration.

	\section*{Discussion}

	% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY

	In this paper, we investigated whether random exploration is really random. That is, to what extent is it driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for XXX\% of the overall variability in people's behavior and increased with horizon - a hallmark of an exploratory process. This suggests that at least some of the apparent randomness in random exploration is not random at all.

	%SO WHERE DOES THIS LEAVE RANDOMNESS IN RANDOM EXPLORATION?  WELL, THE REMAINING 75\% OF THE NOISE COULD BE RANDOM OR IT COULD BE DETERMINISTIC BECAUSE WE CAN'T CONTROL EVERYTHING. 
	So where does this leave randomness in random exploration? Well, the remaining XXX \% of the noise could be random or it could be deterministic because we can't control everything. In particular, while we controlled many aspects of the stimulus across repeated games (e.g. the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose. Thus conceptually, our estimate of deterministic noise is a lower bound. Conversely, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model. In addition, our estimation method also systematically underestimate deterministic noise (part of the deterministic noise will show up as random noise in model fitting, see Supplementary figure SX), so methodologically our model also provides a lower bound of deterministic noise and an upper bound on random noise. Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration, our results suggest that at least some of the randomness is driven by a deterministic process instead.

	%Given that there is still a considerable window for truly stochastic processes in the brain to be driving random exploration, 
	%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}. In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.

	%Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise. Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.  In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}

	%More generally, our finding that random noise dominates behavioral variability over deterministic noise, is consistent with findings of \cite{drugowitsch16}. In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection. This suggests that most noise in behavior is generated randomly and that this may arise from  computational errors in computing the correct strategy. In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.

	%USE TEXT FROM THIS PARAGRAPH IN THE `WHERE DOES THIS LEAVE RANDOMNESS PARAGRAPH' 
	%Perhaps the main limitation of this work is that we cannot say whether the remaining XXX\% of the noise is deterministic or is actually random.

	Regardless of whether the remaining XXX\% is deterministic or random, the fact that thee horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration, Specifically, a reduction in the strength with which reward drives the choice. In particular, in our analysis we excluded a scalar on $Delta R$, if we include that then we get

	%REGARDLESS OF WHETHER THE REMAINING 75\% IS RANDOM OR NOT, THE FACT THAT THE HORIZON CHANGE IN THE TWO NOISES ARE PROPORTIONAL TO EACH OTHER SUGGESTS A POSSIBLE MECHANISM FOR RANDOM EXPLORATION.  SPECIFICALLY, A REDUCTION IN THE STRENGTH WITH WHICH REWARD DRIVES THE CHOICE.  IN PARTICULAR, IN OUR ANALYSIS WE EXCLUDED A SCALAR ON $\Delta R$.  IF WE INCLUDE THAT THEN WE GET
	\begin{equation}
	    \Delta Q = \beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
	\end{equation}

	A decrease in $beta$ here would be equivalent to an increase in variance of both deterministic and random noise. Such reduced coding of reward would be consistent with \cite{ebitz17} (NEED CHECK).
	%A DECREASE IN $\beta$ HERE WOULD BE EQUIVALENT TO AN INCREASE IN VARIANCE OF BOTH DETERMINISTIC AND RANDOM NOISE.  MAKES THE PREDICTION THAT SPATIAL BIAS SHOULD ALSO CHANGE IN PROPORTION (CHECK THIS!).

	%SUCH REDUCED CODING OF REWARD WOULD BE CONSISTENT WITH EBITZ (I THINK).

	%Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between random and deterministic noise that we found in Figure \ref{fig:mb1}, where random noise is 2-3 times the size of deterministic noise.

	\bibliographystyle{plainnat}
	%\bibliographystyle{unsrt}
	\bibliography{Refs/refs}

	% add the Bibliography to the Table of Contents
	\cleardoublepage
	\ifdefined\phantomsection
	\phantomsection  % makes hyperref recognize this section properly for pdf link
	\else
	\fi
	\addcontentsline{toc}{chapter}{Bibliography}

\end{document}

### 2022-12-12 manuscript/2022 draft - Plos Comp Bio/OlderVersion/08-14-20/Internal-External Noise.tex
updates

--- 2022-12-12 manuscript/2022 draft - Plos Comp Bio/OlderVersion/08-05-20/Internal-External Noise.tex
@@
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-In particular, we find evidence that at least part (IT SOUNDS HARD TO ME TO ARGUE THAT THIS NUMBER IS NOT TASK SPECIFIC.
-- BOB SAYS: CAN WE PUT A NUMBER ON THIS E.G.
-25\%?) of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus.
+In particular, we find evidence that at least part (27.05\%) of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus.
 This still leaves open the possibility that much of random exploration is truly `random,' but narrows the window of opportunity for stochastic processes to explain this behavior.
@@
 
-	Results of this parameter recovery procedure are shown in Supplementary Figure S4.
-As is clear from this figure, parameter recovery is good in terms of correlation between the true vs fitted parameters for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Supplementary Figure S4 Panel C).
+	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}.
+Parameter recovery is good (in terms of correlation between the true vs fitted parameters) for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C).
 The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
-This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6.
-In addition we see better recovery for random noise than deterministic noise because we effectively have half as many trials for deterministic noise since we are only generating one sample of deterministic noise for each repeated game pair.
+This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
+In addition we see better recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
 
-	One limitation of our model is that we observed a systematic underestimation of deterministic noise (despite the strong correlation).
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/precover.png}
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
+			}
+			\label{fig:pararecover1}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RanDetNoise_pure_ran_det_simu_hyperprior.png}
+			\caption{Parameter recovery over the posterior of random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for purely random noise (left column) and purely deterministic noise (right column) games.
+			}
+			\label{fig:pararecover2}
+		\end{center}
+	\end{figure}
+
+	Finally, despite the strong correlation between simulated and fit values of deterministic noise, we observed a systematic underestimation of $\sigma_{det}$.
 This is further illustrated in Supplementary Figure S5.
-In thee simulation with 0 random noise and full deterministic noise, our model perfectly recovered both random and deterministic noise, however in the simulation with full random noise and 0 deterministic noise, our model recovered only XXX \% of deterministic noise, and the remaining shows up as random noise instead (See Supplementary Figure S5).
+In the simulation with 0 random noise and fully deterministic noise, our model perfectly recovered both random and deterministic noise, however in the simulation with full random noise and 0 deterministic noise, our model recovered only XXX \% of deterministic noise, and the remaining shows up as random noise instead (Figure \ref{fig:pararecover2}).
 Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
 
-	Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.
+	%Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.
 
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/pinconsistent_6panel_age_acchance.png}
+			\includegraphics[width=1\textwidth]{figures/RanDetNoise_2noise_modelA.png}
 			\caption{
@@
 
-	In this paper, we investigated whether random exploration is really random.
-That is, to what extent is it driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
-Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for XXX\% of the overall variability in people's behavior and increased with horizon - a hallmark of an exploratory process.
-This suggests that at least some of the apparent randomness in random exploration is not random at all.
+	In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
+Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 27.05\% of the overall variability in people's behavior.
+%and increased with horizon - a hallmark of an exploratory process.
 
-	%SO WHERE DOES THIS LEAVE RANDOMNESS IN RANDOM EXPLORATION?  WELL, THE REMAINING 75\% OF THE NOISE COULD BE RANDOM OR IT COULD BE DETERMINISTIC BECAUSE WE CAN'T CONTROL EVERYTHING. 
 	So where does this leave randomness in random exploration?
-Well, the remaining XXX \% of the noise could be random or it could be deterministic because we can't control everything.
+Well, the remaining 72.95 \% of the noise could be random or it could be deterministic because we can't control everything.
 In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
@@
 Conversely, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
-In addition, our estimation method also systematically underestimate deterministic noise (part of the deterministic noise will show up as random noise in model fitting, see Supplementary figure SX), so methodologically our model also provides a lower bound of deterministic noise and an upper bound on random noise.
-Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration, our results suggest that at least some of the randomness is driven by a deterministic process instead.
+In addition, our estimation method also systematically underestimate deterministic noise (part of the deterministic noise will show up as random noise in model fitting, see Figure \ref{fig:pararecover2}), so methodologically our model also provides a lower bound of deterministic noise and an upper bound on random noise.
+Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 72.95\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
+			%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
+
+	Randomness in exploratory behavior in random exploration can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).
+The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
+In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
+In our study, most of the decision noise seen can still be random noise (up to 72.95\%), and this is consistent with findings of \cite{drugowitsch16}.
+In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+This suggests that a large proportion of noise in behavior is generated randomly and that this may arise from computational errors in computing the correct strategy.
+% In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
+Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
+In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}.
+
+	On the other hand, our result suggests that at least a quarter of the decision noise in random exploration arises from a deterministic process.
+This is in line with works of \citep{Bair1996} in which MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli.
+In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
+Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+
+	%One way in which this may be achieved is if random exploration is modulated, not by increasing noise, but by decreasing reward sensitivity.
+For example, modifying equation XXX to include a reward sensitivity term we get ...
 
 	%Given that there is still a considerable window for truly stochastic processes in the brain to be driving random exploration,
-	%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
-In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
 
-	%Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
-Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
-In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}
-
-	%More generally, our finding that random noise dominates behavioral variability over deterministic noise, is consistent with findings of \cite{drugowitsch16}.
-In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
-This suggests that most noise in behavior is generated randomly and that this may arise from  computational errors in computing the correct strategy.
-In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.
-
-	%USE TEXT FROM THIS PARAGRAPH IN THE `WHERE DOES THIS LEAVE RANDOMNESS PARAGRAPH' 
-	%Perhaps the main limitation of this work is that we cannot say whether the remaining XXX\% of the noise is deterministic or is actually random.
-
-	Regardless of whether the remaining XXX\% is deterministic or random, the fact that thee horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration, Specifically, a reduction in the strength with which reward drives the choice.
-In particular, in our analysis we excluded a scalar on $Delta R$, if we include that then we get
+	Regardless of whether the remaining 72.95 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration.
+Specifically, a reduction in the strength with which reward drives the choice.
+In particular, in our analysis we excluded a scalar on $\Delta R$, if we include that then we get
 
@@
 
-	A decrease in $beta$ here would be equivalent to an increase in variance of both deterministic and random noise. Such reduced coding of reward would be consistent with \cite{ebitz17} (NEED CHECK).
+	A decrease in $beta$ here would be equivalent to an increase in variance of both deterministic and random noise. Such reduced coding of reward would be consistent with \cite{ebitz17}.
 	%A DECREASE IN $\beta$ HERE WOULD BE EQUIVALENT TO AN INCREASE IN VARIANCE OF BOTH DETERMINISTIC AND RANDOM NOISE.

### 2022-12-12 manuscript/2022 draft - Plos Comp Bio/Supplementary/DetRanNoise_Supplementary.tex
updates

--- 2022-12-12 manuscript/2020 rejected - JEP General/SupplementaryFigures_RanDetNoise/DetRanNoise_Supplementary.tex
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
@@
 
+       \maketitle
+
 	\newpage
-
 	\begin{figure}[H]
@@
 
+	\newpage
+	\begin{table}[h]
+		\small
+		\centering
+		\begin{tabular}{|c|c|c|}
+			\hline
+			Model & Deterministic noise & Random noise \\
+			\hline
+			A & Horizon dependent & Horizon dependent\\
+			\hline
+			B & Fixed & Horizon dependent\\
+			\hline
+			C & Horizon dependent & Fixed\\
+			\hline
+			D & Fixed & Fixed\\
+			\hline
+			E & Horizon dependent & None\\
+			\hline
+			F & None & Horizon dependent\\
+			\hline
+		\end{tabular}
+		\caption{Model description for figure S4.}%CHECK UPDATE}
+		\label{tab:models}	
+	\end{table}
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/pinconsistent_6panel_age_acchance.png}
+			\caption{
+			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
+			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
+I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
+COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
+			Model comparison: model A - both deterministic and random noise are horizon dependent, model B - only random noise is horizon dependent, model C - only deterministic noise is horizon dependent, model D - neither random nor deterministic noise is horizon dependent, model E - only deterministic noise is assumed to be present, model F - only random noise is assumed to be present.}
+			\label{fig:mb3}
+		\end{center}
+	\end{figure}
 	%\bibliographystyle{plainnat}

### 2022-12-12 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
updates

--- 2022-12-12 manuscript/2022 draft - Plos Comp Bio/OlderVersion/08-14-20/Internal-External Noise.tex
@@
 
+%\usepackage[final]{changes}
+\usepackage{changes}
+\definechangesauthor[name={Siyu Wang}, color=red]{siyu}
+
 \title{The nature of decision noise in random exploration}
 
-\author[1]{Siyu Wang}
-\author[1,2]{Robert C. Wilson}
+\author[1,\textcurrency]{Siyu Wang}
+\author[1,2,3]{Robert C. Wilson}
 
-\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
-\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}
+\affil[1]{Department of Psychology, University of Arizona, Tucson AZ, USA}
+\affil[2]{Neuroscience and Physiological Sciences Graduate Interdisciplinary Program, University of
+	Arizona, Tucson AZ, USA}
+\affil[3]{Cognitive Science Program, University of Arizona, Tucson AZ, USA}
+\affil[ \textcurrency]{\added[id=siyu]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}}
 
@@
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-In particular, we find evidence that at least part (27.05\%) of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus.
+In particular, we find evidence that at least part \deleted[id=siyu]{27$\%$} of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus.
 This still leaves open the possibility that much of random exploration is truly `random,' but narrows the window of opportunity for stochastic processes to explain this behavior.
@@
 
-		\section*{Methods}
+	\section*{Materials and Methods}
 	\subsection*{Participants}

### 2022-12-12 manuscript/Older drafts/Backup2/Internal-External Noise/Internal-External Noise.tex
updates

--- 2022-11-21 draft/Older drafts/Siyu#final/Internal-external noise Siyu#final/Internal-External Noise.tex
@@
 
-\title{What is the nature of decision noise in random exploration?}
+\title{Any way the brain blows?  The nature of decision noise in random exploration}
 
@@
 	\begin{abstract}
-	[150 words for Nature Human Behavior, currently 162]
-	Human decision making appears to contains components that are inherently random.
-While this behavioral variability has traditionally been seen as noise, recent work suggests that random choices may actually be adaptive.
-An example arises when deciding between exploring unknown options or exploiting options we know well.
-Theory shows that a little randomness in explore-exploit decisions is remarkably effective, leading to greater rewards overall.
-Meanwhile, experiments show that people use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore.
-Despite this progress, the nature of adaptive decision noise is unknown and it is unclear whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world?
-Here we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise.
-This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control.
+
+Human decision making is inherently variable.
+While this variability is often seen as a sign of suboptimality in human behavior, recent work suggests that randomness can actually be adaptive.
+An example  arises when we must choose between exploring unknown options or exploiting options we know well.
+A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore.
+Moreover, people actually use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore.
+Despite this progress, the nature of adaptive `decision noise’ for exploration is unknown -- specifically whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world.
+Here we show that, while both internal and external noise drive variability in behavior, the noise driving random exploration is predominantly internal.
+This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.
+
+%Human decision making is inherently variable.
+While this variability was originally seen as suboptimal, recent work suggests that randomness can actually be adaptive.
+An example arises when we must between exploring unknown options or exploiting options we know well.
+A little randomness in these `explore-exploit' decisions is remarkably effective, as it encourages to explore options we might otherwise ignore.
+Moreover, people actually use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore.
+Despite this progress, it is unknown whether variability for random exploration is generated externally, from stochastic stimuli in the world, or internally, from stochastic processes in the brain.
+That is do we explore by going any way the wind blows or any way our brains blow?
+Here we show that, while both types of noise are present that it is noise in our brains that dominates random exploration.
+This suggests that decision noise for exploration is subject to cognitive control.
 
@@
 	Imagine trying to decide where to go to dinner with a friend.
-You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing.
-Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.
+You can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try the new restaurant that you know nothing about.
+Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.
 From a theoretical perspective, making optimal explore-exploit choices, i.e.
-choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}.
+choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}.
 In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.
@@
 	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}.
-In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.
+In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
+In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
 
-	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.
-The key manipulation in the task is the horizon condition, i.e.
-the number of decisions remaining to make.
+	Using a simple gambling task, known as the Horizon Task, we have recently shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.
+The key manipulation in the Horizon Task is the horizon condition, i.e.
+the number of decisions remaining for the participant to make.
 Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards.
-For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new place.
-Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.
+For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant.
+Using such a horizon manipulation we found that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma.
 
-	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown.
-Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain.
+	A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.
+In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain.
+	%internally, from random processes in the brain, or externally, in stochastic input from the world.
+In the restaurant case, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
+	%$Of these, internal noise is the simplest to 
+	%externally, in the input from the world, or is generated internally, within the brain. 
 	In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants.
+Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in.
+Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again.
+	%If random exploration is driven by external noise, this implies that people adapt their susceptibility to random external with horizon. Thus
+	%Exploration driven by external noise, 
 	Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
-Previous work makes a strong case for both types of noise being relevant to behavior.
+Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.
+
+	Previous work makes a strong case for both external and internal noise being relevant to behavior.
 For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}.
@@
 
-	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \citep{wilson2014}.
+	In this paper, we investigate which source of noise, external vs internal, drives random exploration in humans in a modified version of the Horizon Task.
 To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice.
-If decision noise is purely externally driven, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times.
+If decision noise is purely external, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times.
 Meanwhile, if noise is internally driven their choices should be inconsistent, since the internal noise is different both times.
-By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the contribution of internal noise to random exploration far exceeds that contributed by the stimulus.
+By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the variability related to random exploration is dominated by internal noise.
 
@@
 
-	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit, in the form of four example plays distributed either unequally between bandits (i.e.
-1 play of one bandit, 3 plays of the other) or equally (2 plays each).
+	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
+1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
 These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
@@
 
-	\begin{figure}[h]
+	\begin{figure}[hp]
 		\begin{center}
 			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
-			\caption{A key additional manipulation here is repeated games.
-Each pair of repeated games with identical example trials will appear twice during the experiment.
-We setup the repeated games such that they are at least 5 games apart from each other.
-A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
+			\caption{BOB: ADD A PANEL A ABOVE THIS PANEL, SHOWING THE DYNAMICS FOR ONE GAME - LIKE IN HASHEM'S PAPER.  
+			Schematic of the experiment.
+(A) Dynamics of an example horizon 6 game.
+Here the first four trials are forced trials in which participants are instructed which option to play.
+After the forced trials, participants are free to choose between the two options for the remainder of the game.
+(B) Different possible states of the game after the first free choice over the course of the experiment.
+Overall participants play 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
+In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.
+These repeated games allow us to compute the relative contribution of external and internal noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}%BOB: UP THE FONT SIZE FOR TEXT TO MATCH FONT SIZE FOR NUMBERS}
 			\label{fig:taskfig}
@@
 			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
-			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
+			\caption{BOB: DO THE YTICKS ON THE 0.1S NOT 0.15S!  MORE WHITE SPACE BETWEEN PANELS A AND B.
+			Replication of previous findings.
+Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.
+}%BOB: SWITCH A AND B TO HAVE RANDOM FIRST}
 			\label{fig:modelfree}
@@
 	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven.
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
+Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (For [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; For [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that not all of the noise was stimulus driven.
+In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98 BOB DID YOU FIX THIS?
+DIDNT USE REPEATED ANOVA, WILL CHANGE, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
 
@@
 			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
-			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
-For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
-Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently.
-}
+			\caption{Model-free analysis suggests that both external and internal noise contribute to the choice variability in random exploration.
+For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
+However, the extent to which their choices are inconsistent lies between what is predicted by purely external and internal noise, suggesting that both noise sources influence the decision.}
 			\label{fig:mf2}
@@
 			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
-			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
-Both internal and external noises are nonzero, and internal noise has a much greater magnitude.
-Posterior distributions over the group-level means of the change of external and internal noise standard deviation.
-Internal noise increases significantly with horizon, external noise increases with horizon as well.}
+			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  internal and external noise.
+Both internal (A, B) and external (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, internal noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than external noise.}
 			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
@@
 	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment.
-Using a version of the Horizon Task with repeated games, we found that evidence for both types of noise in explore-exploit decisions.
+Using a version of the Horizon Task with repeated games, we found evidence for both types of noise in explore-exploit decisions.
 In addition, we see that both internal and external noise increase with horizon, but that the horizon effect is much larger for internal noise.
@@
 	Taken at face value, the horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
+This is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
 In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.
@@
 	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
-Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
+Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
 In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}
@@
 
-	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment.
+	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment.
 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
 This left 60 for the main analysis.
-Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
+Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
 
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
-			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
-			\label{fig:modelfree2}
-		\end{center}
-	\end{figure}
+	%SIYU: SORRY, WHAT DO YOU MEAN BY POPING THESE INTO A SEPARATE DOCUMENT, DO YOU MEAN I START A NEW LATEX DOC AND PUT THE SUPPLEMENTARY GRAPHS THERE, DO I STILL KEEP THE FIGURES HERE
+	%BOB: SUPPLEMENTARY FIGURES SHOULD START NUMBERING FROM 1.  SO Supplementary Figure 1 etc ... Pop these into a separate document for sharing.  To do refs just hard code numbers.
 
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
-			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
-For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
-Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently.
-}
-			\label{fig:mf22}
-		\end{center}
-	\end{figure}
-	\newpage
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
-			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
-Both internal and external noises are nonzero, and internal noise has a much greater magnitude.
-Posterior distributions over the group-level means of the change of external and internal noise standard deviation.
-Internal noise increases significantly with horizon, external noise increases with horizon as well.}
-			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
-			\label{fig:mb12}
-		\end{center}
-	\end{figure}
 	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities.
@@
 I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
-	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME
+	%THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME
 
@@
 
+	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
+We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
+The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
+After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfigb}.
+
 	\begin{figure}[H]
@@
 			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
-			\caption{Time line of a horizon 6 game.
+			\caption{BOB: DONT NEED THIS FIGURE IF YOU UPDATE EARLIER TASKFIGURE.
+Time line of a horizon 6 game.
 The game begins with four forced choice trials in which participants are instructed which option to play.
 After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
+			%BOB: THIS FONT SIZE IS GOOD - MAYBE A LITTLE BIGGER FOR TEXT
 			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
 			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
-			I WILL SEND YOU THE POWERPOINTS TOO}
+			%I WILL SEND YOU THE POWERPOINTS TOO
+			}
 			\label{fig:taskfigb}
@@
 	\end{figure}
-
-	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
-The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
-(See Figure \ref{fig:taskfigb})
-
 %	\begin{figure}[H]
@@
 			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
-			\caption{Hierarchical Bayesian model}
+			\caption{Schematic of the hierarchical Bayesian model using notation of BOB CITE WAGNERMAKERS AND LEE.}
 			\label{fig:model}
@@
 			\includegraphics[width=1\textwidth]{figures/precover.png}
-			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, internal noise variance, $\sigma_{int}$, and external noise variance, $\sigma_{ext}$, for horizon 1 (left column) and horizon 6 (right column) games.
+%for horizon 1 games
 			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.
@@
 			%USE ADDABCS FUNCTION TO ADD ABCs!
+			%BOB: BIT MORE SPACE BETWEEN ROWS. OTHERWISE GREAT!
 			}
@@
 
-	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
+%	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
+%BOB: I THINK THIS IS OK.
 	%\begin{figure}[H]-

### 2022-12-12 manuscript/Older drafts/Internal-external noise Supplementary Figures/Internal-External Noise.tex
updates

--- 2022-11-21 draft/Older drafts/Siyu#final/Internal-external noise Siyu#final/Internal-External Noise.tex
@@
 
-\title{What is the nature of decision noise in random exploration?}
+\title{Any way the brain blows?  What is the nature of decision noise in random exploration?}
 
@@
 	\newpage
-	\begin{abstract}
-	[150 words for Nature Human Behavior, currently 162]
-	Human decision making appears to contains components that are inherently random.
-While this behavioral variability has traditionally been seen as noise, recent work suggests that random choices may actually be adaptive.
-An example arises when deciding between exploring unknown options or exploiting options we know well.
-Theory shows that a little randomness in explore-exploit decisions is remarkably effective, leading to greater rewards overall.
-Meanwhile, experiments show that people use such `random exploration' in practice,  increasing their behavioral variability when it is more valuable to explore.
-Despite this progress, the nature of adaptive decision noise is unknown and it is unclear whether it is generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world?
-Here we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise.
-This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control.
-
-	%From everyday decisions, like deciding where to dine, to major life milestones, like deciding whom to marry, many decisions involve a tradeoff between exploring options that are unknown and exploiting options we know well.
-
-	%Recently we have shown that humans make simple explore-exploit choices using two distinct strategies. One of these strategies, known as random exploration, involves choice randomization
-
-	%While making optimal explore-exploit choices is a hard computational problem, we have recently shown that humans make effective explore-exploit choices using heuristic strategies.
-One particularly simple strategy is `random exploration'.
-In this strategy, the decision process is noisy meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-
-	%Despite this progress, questions remain about the nature of random exploration in particular, is the noise in behavior generated internally, from stochastic processes in the brain, or externally, from stochastic stimuli in the world?
-
-	%as there are a number of ways in which seemingly stochastic choices could be generated.
-In one strategy, the `external noise strategy,' participants could rely on stochasticity in the world and allow irrelevant features of the stimulus to drive choice.
-In another strategy called `internal noise strategy,' people could rely on stochastic processes within their own brains.
-
-	%In this work, we modified our recently published `Horizon Task' in such a way as to distinguish these two strategies.
-Using both a model-free and model-based analysis of human behavior we show that, while both types of noise are present in explore-exploit decisions, random exploration is dominated by internal noise.
-This suggests that random exploration depends on adaptive noise processes in the brain which are subject to (perhaps unconscious) cognitive control.
-	\end{abstract}
-	\newpage
-
-	%In theory, such `random exploration', can be surprisingly effective in explore-exploit problems and, if implemented correctly, can come close to optimal performance.
-	%Moreover, recent work suggests that humans actually use random exploration in simple explore-exploit tasks.
-
-	\section*{Introduction}
-
-	Imagine trying to decide where to go to dinner with a friend.
-You can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened and about which you know nothing.
-Such decisions, in which we must choose between a well-known exploit option and a lesser known explore option, are known as explore-exploit decisions.
-From a theoretical perspective, making optimal explore-exploit choices, i.e.
-choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}.
-In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.
-
-	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}.
-In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.
-
-	Recently we have shown that humans appear to actually use random exploration and actively adapt their decision noise to solve simple explore-exploit problems \citep{wilson2014}.
-The key manipulation in the task is the horizon condition, i.e.
-the number of decisions remaining to make.
-Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards.
-For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new place.
-Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.
-
-	However, a key limitation of this work was that the source of the decision noise used for exploration is unknown.
-Of particular interest is whether the adaptive decision noise that is linked to exploration arises externally, in the input from the world, or is generated internally, within the brain.
-In the restaurant case, an example of external noise would be if you happen to spot an old friend walking in to one of the restaurants.
-Conversely, internal noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
-Previous work makes a strong case for both types of noise being relevant to behavior.
-For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}.
-Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.
-
-	In this paper, we investigate which source of noise, internal vs external, drives random exploration in humans in a simple explore-exploit task adapted from our previous work \citep{wilson2014}.
-To distinguish between the two types of noise, we had people make the exact same explore-exploit decision twice.
-If decision noise is purely externally driven, then people's choices should be identical both times, that is their choices should be consistent, since the stimulus is the same both times.
-Meanwhile, if noise is internally driven their choices should be inconsistent, since the internal noise is different both times.
-By analyzing behavior on this task in both a model-free and model-based manner, we show that, while both types of noise are present in explore-exploit decisions, the contribution of internal noise to random exploration far exceeds that contributed by the stimulus.
-
-	%\begin{figure}[H]
-	%	\begin{center}
-	%		\includegraphics[width=0.7\textwidth]{figs/bayes_s.png}
-	%			\caption[Subject level estimated mean of external and internal noise variance.]{Subject level estimated mean of external and internal noise variance.
-Both internal and external noise increase with horizon, internal noise is dominant.}
-	%			\label{fig:mb3}
-	%		\end{center}
-	%	\end{figure}
-
-	\section*{Results}
-
-	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of internal vs external noise on people's decisions (Figure \ref{fig:taskfig}).
-In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
-Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
-
-	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit, in the form of four example plays distributed either unequally between bandits (i.e.
-1 play of one bandit, 3 plays of the other) or equally (2 plays each).
-These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
-Thus, by giving people the exact same example plays twice in two separate games (separated by several minutes in time so as to avoid detection), the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.
-
-	These `repeated games' are the key manipulation in this paper and allow us to distinguish between external and internal sources of noise.
-Specifically, if noise is externally driven, then choices on repeated games should be consistent.
-Conversely if noise is internally driven, then choices on repeated games should be inconsistent.
-
-	\begin{figure}[h]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
-			\caption{A key additional manipulation here is repeated games.
-Each pair of repeated games with identical example trials will appear twice during the experiment.
-We setup the repeated games such that they are at least 5 games apart from each other.
-A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
-			\label{fig:taskfig}
-		\end{center}
-	\end{figure}
-
-	\subsection*{Both behavioral variability and information seeking increase with horizon}
-
-	%This further allowed us to have people make the exact same
-
-	%The key manipulation was to use repeated games to let people make the same decision twice.
-In the restaurant example, if your decision is mainly driven by external noise, then a few months later when you see the friend walking in front of you into the restaurant, you are very likely to make the same decision and follow him into that restaurant to say hi.
-However, if your decision is mainly driven by internal noise, then next time you make a split-second decision about where to go, you are equally likely to go to the other restaurant.
-
-	%More specifically, we look at the contribution of external and internal noise by providing with participants the same decision problem twice in the task.
-In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
-To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.
-
-	%In each game they made multiple decisions between two options.
-Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution.
-The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game.
-One of the bandit will always have a higher mean than the other.
-Participants were instructed to maximize the points earned over the entire task.
-
-	%The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
-The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
-In unequal uncertainty condition, people are more likely to choose the option that they know less about - the more informative option - to explicitly explore that option more.
-This type of information driven exploration is known as directed exploration.
-
-	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}.
-As in our previous work, we find evidence for two types of exploration in the Horizon Task.
-Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.
-
-	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
-This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]).
-Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition.
-Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(59)=6.88, p $<$ 0.001).
-
-	%These conditions allow us to measure directed and random exploration in a model-free way.
-Directed exploration is measured as the probability of choosing the more informative option in [1 3] condition whereas random exploration is measured as the probability of choosing the low mean option in [2 2] condition.
-In line with previous results \citep{wilson2014}, we showed that both directed and random exploration increase with horizon.
-Both direct (t(59)=6.88, p $<$ 0.001) and random exploration (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]) increases with horizon.
-(Figure \ref{fig:modelfree}, A,B).
-Directed exploration is considered to be driven by information bias and random exploration is considered to be driven by decision noise, in this work, we are investigating which source of decision noise, external vs internal, drives random exploration.
-
-	%the influence of directed exploration is measured as the probability of choosing the more informative option, i.e.
-the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.
 
@@
 			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
-			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
-			\label{fig:modelfree}
-		\end{center}
-	\end{figure}
-
-	%Finally, the crucial additional manipulation in this task is repeated games (Figure \ref{fig:taskfig}).
-In each pair of repeated games, the four forced-choice trials were yoked, meaning that on the first free choice trial participants were faced with identical stimuli.
-After the first free choice trial, the outcomes on the repeated games were not yoked and the outcomes were sampled independently from the underlying Gaussian distribution.
-Not yoking the later trials made it harder for participants to detect repeated games.
-In addition, the presentation of repeated games was controlled so that each repeated pair was at least five games away from each other.
-
-	\subsection*{Model-free analysis shows that random exploration may involve both internal and external noise}
-
-	%In this section we use both model-free and model-based analyses to show that both internal and external noise contribute to the behavioral variability in random exploration.
-Using the model-based hierarchical Bayesian analysis, we also show that the effect of internal noise is the main source of noise in random exploration.
-
-	%\subsubsection*{Model-free analysis}
-	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.
-The idea behind this measure is that purely external noise should lead to consistent choices as the external stimulus is identical both times.
-Conversely, purely internal noise should lead to independent choices, and hence more inconsistent choices both times.
-
-	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven.
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
-
-	%As is clear from the plots, in both the unequal ([1 3], panel A) and equal ([2 2], panel B) conditions,
-
-	%In addition, whether participants make consistent choices in repeated games is used as a model-free measure of internal noise, since in repeated trials, external noise should be identical on both trials.
-So only internal noise can differ and drive the choice inconsistency.
-The degree to which people make consistent choices in repeated trials can reflect the internal noise.
-Since choice inconsistency in both [1 3](t(59) = 5.10, P([1 3]) $<$ 0.001) and [2 2](t(59) = 6.08, P([2 2])$<$0.001) condition increase with horizon (Figure \ref{fig:modelfree} C), this is a behavioral evidence that internal noise increases with horizon.
-
-	%Choice inconsistency between repeated games was non-zero in both horizon conditions, suggesting that not all of the noise was stimulus driven.
-In addition, choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (Figure \ref{fig:modelfree}), suggesting that at least some of the horizon dependent noise is internal.
-
-	\begin{figure}[h]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
-			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
-For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
-Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently.
-}
-			\label{fig:mf2}
-		\end{center}
-	\end{figure}
-
-	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely external and purely internal noise cases.
-For purely external noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case.
-For purely internal noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
-	\begin{equation*}
-	\begin{split}
-	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
-	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
-	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean})(1-p(\mbox{low mean}))
-	\end{split}
-	\end{equation*}
-	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(59) = 5.10, p $<$ 0.001 for horizon 1, t(59) = 5.65, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 3.49, p $<$ 0.001 for horizon 1, t(59) = 3.31, p $<$ 0.001 for horizon 6.
-Behavior is different from pure external noise prediction in both [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency.
-Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.
-
-	\subsection*{Model-based analysis shows that random exploration is dominated by internal noise}
-
-	To more precisely quantify internal and external noise, we turned to model fitting.
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noises.
-In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.
-
-	\subsubsection*{Overview of model}
-	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the information bias between the two bandits.
-To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
-	\begin{equation}
-	\Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
-	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.
-
-	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game; and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays and changes every game.
-
-	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the external noise, $n_{ext}$, should be the same while the internal noise, $n_{int}$ may be different.
-This is exactly how we distinguish external noise from internal noise.
-In symbolic terms, for repeated games $i$ and $j$,  $n_{ext}^i=n_{ext}^j$  and $n_{int}^i \neq n_{int}^j$.
-
-	\subsubsection*{Model fitting}
-	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
-In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon.
-Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
-%The mean and standard deviation of information bonus $A$ and spatial bias $B$ are sampled from a Gaussian prior and an exponential prior respectively.
-The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter $k$ and $\lambda$ for the gamma distribution are sampled from exponential priors.
-
-	%To capture the idea that external noise should be identical on repeated games, we sampled one value of the external noise, $n_{ext}$ for each pair of repeated games.
-Conversely, because internal noise is expected to change between games we sampled two values of the internal noise, $n_{int}^1$ and $n_{int}^2$, i.e.
-one for each individual game.
-
-	%The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS software \citep{jags, matjags}.
-
-	%Before presenting the results of the model-based analysis we begin with a brief overview of the most salient points of the model.
-A full description of the model can be found in the Methods and code to implement the model can be found in the Supplementary Material.
-
-	%Conceptually, the model breaks the explore-exploit choice down into two components: a learning component, in which participants estimate the mean payoff of each option from the rewards they see, and a decision component, in which participants use this estimated payoff to guide their choice.
-The learning component assumes that participants compute an estimate of the average payoff for each slot machine, $R^i_t$, using a simple delta rule update equation (based on a Kalman filter \cite{Kalman1960-uo}, see Methods):
-	%\begin{equation}
-	%R^i_{t+1}= R^i_t + \alpha_t^i (r_t - R^i_t)
-	%\end{equation}
-	%where $r_t$ is the reward on trial $t$ and $\alpha_t^i$ is the time-varying learning rate that determines the extent to which the prediction error, $(r_t - R^i_t)$, updates the estimate of the mean of bandit $i$.
-The learning process is described by three free parameters: the initial value of the estimated payoff, $R_0$, and two learning rates, the initial learning rate, $\alpha_1$, and the asymptotic learning rate, $\alpha_{\inf}$, which together describe the evolution of the actual learning rate, $\alpha_t$, over time.
-For simplicity, we assume that these parameters are independent of horizon and uncertainty condition (Table 1).
-
-	%The decision component of the model assumes that participants choose between the two options (left and right) probabilistically according to 
-	%\begin{equation}
-	%	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
-	%\end{equation}
-	%where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition).
-The decision process is described by three free parameters: the information bonus $A$, the spatial bias $B$, and the decision noise $\sigma$.
-We estimate separate values of the decision parameters for each horizon and (since the information bonus is only used in the [1 3] condition) separate values of only the bias and decision noise for each uncertainty condition.
-
-	%Overall, subject's behavior in each session (vertex vs RFPC stimulation) is described by 13 free parameters (Table \ref{table1}): three describing learning ($R_0$, $\alpha_1$ and $\alpha_{\infty}$) and 10 describing the decision process ($A$ in the two horizon conditions, $B$ and $\sigma$ in the four horizon-x-uncertainty conditions).
-These 13 parameters were fit to each subject in each stimulation condition using a hierarchical Bayesian approach \cite{Lee2014-yh} (see Methods).
-
-	\subsubsection*{Model fitting results}
-
-	%Group-level estimates for the variances of internal and external noise are shown in Figure \ref{fig:modelbased}A.
-While both variances are non-zero, this shows that the internal noise is much larger than external noise.
-This horizon-based change is probed further in Figure XXXB in which we plot the posterior distributions over the change in internal and external noise with horizon.
-This clearly shows that only internal noise varies with horizon (STATS XXX\% of the samples above zero for internal noise, XXX\% of samples above zero for external noise).
-
-	Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:mb1}.
-Consistent with our model-free results, we see that both internal and external noise variances are non-zero and that internal noise is about 2-3 times larger than the external noise.
-In addition, we find that internal noise increases dramatically with horizon (M = 4.55, p $<$ 0.001) whereas there is only a slight increase of external noise (M = 1.78, p = 0.009).
-Taken together, these results suggest that random exploration is dominated by internal noise.
-
-	%internal noise changes
-
-	%appear to increase with horizon (Figure \ref{fig:mb2}, 100\% of the samples above zero for internal noise, 96.33\% of samples above zero for external noise), the change in internal noise is much larger (STATS).
-
-	%With regard to the change in noise between horizons,
-
-	%While both variances are non-zero, this shows that the internal noise is much larger than external noise.
-This horizon-based change is probed further in Figure \ref{fig:mb2} in which we plot the posterior distributions over the change in internal and external noise with horizon.
-As is clear from this plot, internal noise increases dramatically with horizon whereas there is a slight increase of external noise with horizon as well (0 just falls in the critical region at level of 5\% for external noise but lies right at the border, although significant, it's not as strong as the internal noise).
-This clearly shows that internal noise dominates and significantly varies with horizon  (100\% of the samples above zero for internal noise, 96.33\% of samples above zero for external noise).
-
-	%Furthermore, we plotted the subject level estimates of both internal and external noises in a similar fashion of the model-free plot of choice-inconsistency for comparison (Figure \ref{fig:mb3}).
-We showed again that both internal and external noises increase significantly with horizon.
-But in terms of magnitude, internal noise is what' really dominant here (See also Figure \ref{fig:rcurvepd6} in Appendix \ref{ch:appendicies:reward}).
-Subject-level estimates also allow us to look at the relationship between internal and external noises, and we showed that internal noise correlates with external noise significantly in both horizon conditions ($R^2 = 0.32, p < 0.001$ in horizon 1, $R^2 = 0.24, p < 0.001$ in horizon 6, see Figure \ref{fig:corrintext} in Appendix \ref{ch:appendix:corrintext}).
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
-			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
-Both internal and external noises are nonzero, and internal noise has a much greater magnitude.
-Posterior distributions over the group-level means of the change of external and internal noise standard deviation.
-Internal noise increases significantly with horizon, external noise increases with horizon as well.}
-			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
-			\label{fig:mb1}
-		\end{center}
-	\end{figure}
-
-	%LEFT EACH TYPE OF NOISE IN EACH HORIZON
-	%RIGHT IS DISTRIBUTION OVER CHANGE IN NOISE WITH HORIZON
-
-	%Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:modelbased}.
-As is clear from this plot, only internal noise increases  with horizon.
-	%In both horizon 1 and horizon 6, internal noises are significantly higher than internal noises (for both horizon 1 and horizon 6, P $<$ 0.001).
-
-%	\begin{figure}[H]
-%		\begin{center}
-%			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
-%			\caption[]{}			
-%			\label{fig:mb2}
-%		\end{center}
-%	\end{figure}
-
-	\section*{Discussion}
-	In this paper, we investigated whether random exploration is driven by internal noise, putatively arising in the brain, or external noise, arising from the environment.
-Using a version of the Horizon Task with repeated games, we found that evidence for both types of noise in explore-exploit decisions.
-In addition, we see that both internal and external noise increase with horizon, but that the horizon effect is much larger for internal noise.
-Taken together our results suggest that random exploration, i.e.
-the use and adaptation of decision noise to drive exploration, is driven primarily by internal noise.
-
-	Perhaps the main limitation of this work is in the interpretation of the different types of noise as being internal and external.
-In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
-the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose.
-Thus, our estimate of external noise is likely a lower bound.
-Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of external noise would be interpreted as internal noise in our model.
-Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb1}, where internal noise is 2-3 times the size of external noise.
-
-	Taken at face value, the horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
-In addition, this work is consistent with a recent report from \cite{ebitz17} in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.
-
-	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
-Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
-In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}
-
-	More generally, our finding that internal noise dominates behavioral variability over external noise, is consistent with findings of \cite{drugowitsch16}.
-In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
-This suggests that most noise in behavior is generated internally and that this may arise from  computational errors in computing the correct strategy.
-In the context of the Horizon Task, such computational errors would likely be larger  in the long horizon condition as the correct course of action in these cases is much harder to compute.
-	%MORE GENERALLY, OUT FINDING THAT INTERNAL DOMINATES BEHAVIORAL VARIABILITY OVER EXTERNAL NOISE, IS CONSISTENT WITH FINDINGS OF DRUGOWITSCH ET AL.  IN PARTICULAR THESE AUTHORS SHOW THAT 
-	%	LINK IN THE Jan Drugowitsch stuff on variability.  ALEX POUGET STUFF. INTERNAL NOISE COMES FROM COMPUTATIONAL ERRORS IN COMPUTING THE CORRECT STRATEGY IN LONG  HORIZON CONDITIONS.
-
-	%THREE IDEAS
-	%NOISE ADDED TO THE PROCESS
-	%ATTENTION = DECREASE SIGNAL VS INCREASE NOISE
-	%NOT NOISY BUT WRONG
-	%PLANNING WITH FEW SAMPLES DEEP EXPLORATION
-
-	%There are many ways that internal noise can be implemented to cause the computational error and ultimately drive different behaviors.
-As far as behavior, choice variability would go up as long as the signal-noise ratio goes up.
-But neurally there are two ways that signal-noise ratio can go up.
-It can be that people are just paying less attention in a long horizon game and devaluing the signal, it can also be that it's really the noise that goes up.
-We can not tell these two apart easily just with behavior.
-One way that may get at it at a behavioral level is to fit both the choices and the reaction times using drift-diffusion model, then we are able to get whether it's the drift rate that goes down (signal), or it's the noise that goes up (noise).
-It would also be interesting to see how external and internal noises are coded in the DDM model.
-Another way to get at this question is to do an EEG study which directly looks at how noise is represented in the electrical signals and see how it changes with horizon.
-
-	%Another possible source of noise is the imperfectness of our model.
-It's not that noise is added to the process, it's simply that the model is wrong and only accounted for part of behavior.
-In our case, the stimulus people get has two components - reward, and the order of rewards (in which order two bandits are played), our model only accounts for the difference in mean rewards between the two bandit, the pattern of choices and the individual reward could have an influence on people's behavior.
-But this only explains for external noise, but not internal.
-Since in repeated games, people are facing the same stimuli, even if the model is suboptimal, it's suboptimal in the same way for repeated pairs, so this idea of 'not noisy but wrong' only explains possible sources for external noise but not internal.
-
-	%Also, the theory of deep exploration is a possible mechanism through which decision noise can be generated when planning with fewer samples.
-It also has a potential to explain how noise is adapted to horizon.
-If implemented well with a well-designed task, two hypothesis under this model can be tested: Does the model quantitatively explain the change of decision noise across horizons?
-Do the number of samples each participants use correlate with their individual reaction times since in theory the more sampling you do, the more planning you do, the longer it should take you to make the decision.
-
-	%In summary, current works shows that both external and internal noises are guilty for the behavioral variability, but internal noise dominates in driving random exploration.
-Further work needs to be done to understand how noises are generated or implemented in our brains.
-
-	\section*{Methods}
-	\subsection*{Participants}
-
-	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment.
-20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-This left 60 for the main analysis.
-Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
 			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
+			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
 			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
+			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
 			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
@@
 	\end{figure}
-	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities.
-(\ref{appendix:participants})
 
-	\subsection*{Task}
-	The task was a modified version of the Horizon Task \citep{wilson2014}.
-In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
-In each game they made multiple decisions between two options.
-Each option paid out a random reward between 1 and 100 points sampled from a Gaussian distribution.
-The means of the underlying Gaussian were different for the two bandit options, remained the same within a game, but changed with each new game.
-One of the bandits always had a higher mean than the other.
-Participants were instructed to maximize the points earned over the entire task.
-To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.
-
-	The number of games participants played depended on how well they performed, which acted as the primary incentive for performing the task.
-Thus, the better participants performed, the sooner they got to leave the experiment.
-On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?
-I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
-	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME
-
-	As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points.
-Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20.
-One of the means was always equal to either 40 or 60 and the second was set accordingly.
-Participants were informed that in every game one of the bandits always has a higher mean reward than the other.
-The order of games was randomized.
-Mean sizes and order of presentation were counterbalanced.
-
-	Each game consisted of 5 or 10 choices.
-Every game started with a fixation cross, then a bar of boxes will show up indicating the horizon for that game.
-For the first 4 games - the instructed games, we highlight the box on one of the bandits to instruct the participant to choose that option, they have to press the corresponding key to reveal the outcome.
-From the $5^{th}$ trial, boxes on both bandits will be highlighted and they are free to make their own decision.
-There was no time limit for decisions.
-During free choices they could press either the left arrow key or right arrow key to indicate their choice of left or right bandit.
-The score feedback was presented for 300ms.
-The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.
-(See Figure \ref{fig:taskfigb})
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
-			\caption{Time line of a horizon 6 game.
-The game begins with four forced choice trials in which participants are instructed which option to play.
-After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
-			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
-			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
-			I WILL SEND YOU THE POWERPOINTS TOO}
-			\label{fig:taskfigb}
-		\end{center}
-	\end{figure}
-
-	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
-The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
-(See Figure \ref{fig:taskfigb})
-
-%	\begin{figure}[H]
-%		\begin{center}
-%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
-%			\caption{Task conditions}
-%			\label{fig:taskfig2}
-%		\end{center}
-%	\end{figure}
-
-	%Finally, the random seeds were not perfectly controlled between subjects.
-The first 16 subjects ran the task with identical random seeds and thus all 16 saw the same sequence of forced-choice trials in both vertex and RFPC sessions.
-For the remaining subjects the random seed was unique for each subject and each session, thus these subjects had unique a series of forced-choice trials for each session.
-Despite this limitation we saw no evidence of different behavior across the two groups.
-
-	\subsection*{Data and code}
-	%Behavioral data as well as Matlab code to recreate the main figures from this paper can be found on the Dataverse website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.
-
-	\subsection*{Model-based analysis}
-	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate internal and external noise.
-In particular, we assume that in repeated games, external noise remains the same whereas internal noise can change.
-
-	\subsubsection*{Hierarchical Bayesian Model}
-
-	To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
-	\begin{equation}
-	\Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
-	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.
-
-	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game; and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays and changes every game.
-
-	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the external noise, $\sigma_{ext}$, and the standard deviation of the internal noise, $\sigma_{int}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}).
-Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.
-In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure.
-The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.
-
-	The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}.
-For example, we assume that the information bonus, $A^{is}$, for each horizon condition $i$ and for each participant $s$, is sampled from a Gaussian prior with mean $\mu^{A}_{i}$ and standard deviation $\sigma_{i}^A$.
-These prior parameters are sampled in turn from their respective hyperpriors: $\mu_{i}^{A}$, from a Gaussian distribution with mean 0 and standard deviation 10, and $\sigma_{i}^A$ from an Exponential distribution with parameters 0.1.
-	%\begin{table}[h]
-	%	\small
-	%	\begin{center}
-	%		\begin{tabular}{|c|c|c|c|}
-	%			\hline
-	%			Parameter &
-	%			Horizon dependent? &
-	%			Uncertainty dependent? &
-	%			Repeated game dependent? \\
-	%			\hline
-	%			information bonus, $A$ &
-	%			yes &
-	%			n/a &
-	%			no \\
-	%			\hline
-	%			spatial bias, $B$ &
-	%			yes &
-	%			no &
-	%			no \\
-	%			\hline
-	%			external decision noise, $\sigma_{ext}$ &
-	%			yes & 
-	%			yes &
-	%			no \\
-	%			\hline
-	%			internal decision noise, $\sigma_{int}$ &
-	%			yes & 
-	%			yes &
-	%			yes \\
-	%			\hline
-	%		\end{tabular}
-	%	\end{center}
-	%	\caption{Model parameters and their dependence on horizon, uncertainty condition and repeated games }
-	%	\label{para1}
-	%\end{table}%
-
-	\begin{table}[h]
-		\small
-		\begin{tabular}{|c|c|c|c|}
-			\hline
-			Parameter & Prior & Hyperparameters & Hyperpriors \\
-			\hline
-			information bonus, $A_{is}$ 
-			& $A_{is} \sim $  Gaussian($\mu_i^{A}$, $\sigma_i^{A}$) 
-			& $\theta_{i}^{A} = (\mu_i^{A}, \sigma_i^{A}) $
-			& \specialcell{
-				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
-			\hline
-			spatial bias, $b_{is}$ 
-			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
-			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
-			& \specialcell{
-				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
-			\hline
-			deviation of external noise, $\sigma^{ext}_{isg}$ 
-			& $\sigma_{i} \sim $  Gamma($k_i^{ext}$, $\lambda_{i}^{ext}$) 
-			& $\theta_{i}^{ext} = (k_i^{ext}, \lambda_i^{ext}) $
-			& \specialcell{
-				$k_i^{ext} \sim $ Exponential(0.01) \\ 
-				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
-			\hline
-			deviation of internal noise, $\sigma^{int}_{isgr}$ 
-			& $\sigma_{is} \sim $  Gamma($k_i^{int}$, $\lambda_{i}^{int}$) 
-			& $\theta_{i}^{int} = (k_i^{int}, \lambda_i^{int}) $
-			& \specialcell{
-				$k_i^{int} \sim $ Exponential(0.01) \\ 
-				$\lambda_i^{int} \sim $ Exponential(10)}		\\
-			\hline
-		\end{tabular}
-		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
-		\label{tab:pars2}	
-	\end{table}
-
-	\subsubsection*{Model fitting using MCMC}
-	The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \citep{jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags).
-This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.
-
-	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).
-Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
-Convergence of the Markov chains was confirmed {\it post hoc} by eye.
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
-			\caption{Hierarchical Bayesian model}
-			\label{fig:model}
-		\end{center}
-	\end{figure}
-
-	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}
-
-	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.
-In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated the choices to see whether we can recover the parameters.
-
-	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}.
-As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C, D).
-The recovery for the noise parameters, $\sigma_{ext}$ and $\sigma_{int}$, is slightly better for horizon 1 than horizon 6.
-This is because it requires more trials to recover bigger noises, so with the same number of choices it is harder to recover overall bigger noises in horizon 6.
-In addition we see better recovery for internal noise than external noise because we effectively have half as many trials for external noise since we are only generating one sample of external noise for each repeated game pair.
-Overall, we are able to recover both external and internal noises using our model to a satisfactory extent.
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/precover.png}
-			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
-			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.
-REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.
-DON'T USE LEGEND, USE TEXT FUNCTION TO PUT TEXT ON PLOTS.
-RED LINES CHANGE TO DOTTED BLACK LINES.
-			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
-			%USE ADDABCS FUNCTION TO ADD ABCs!
-			}
-			\label{fig:pararecover1}
-		\end{center}
-	\end{figure}
-
-	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
-	%\begin{figure}[H]-
-	%	\begin{center}
-	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
-	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
-	%		\label{fig:pararecover2}
-	%	\end{center}
-	%\end{figure}
-	%References
-	%Plummer, M.
-(2003, March).
-JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling.
-In Proceedings of the 3rd international workshop on distributed statistical computing (Vol.
-124, p.
-125).
-	%Rescorla, R.
-A., & Wagner, A.
-R.
-(1972).
-A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement.
-Classical conditioning II: Current research and theory, 2, 64-99.
-
-	\bibliographystyle{plainnat}
+	%\bibliographystyle{plainnat}
 	%\bibliographystyle{unsrt}
-	\bibliography{Refs/refs}
+	%\bibliography{Refs/refs}
 
 	% add the Bibliography to the Table of Contents
-	\cleardoublepage
-	\ifdefined\phantomsection
-	\phantomsection  % makes hyperref recognize this section properly for pdf link
-	\else
-	\fi
-	\addcontentsline{toc}{chapter}{Bibliography}
+	%\cleardoublepage
+	%\ifdefined\phantomsection
+	%\phantomsection  % makes hyperref recognize this section properly for pdf link
+	%\else
+	%\fi
+	%\addcontentsline{toc}{chapter}{Bibliography}
 

### 2022-12-12 manuscript/Older drafts/Older version/V1/Refs/ZajkowskiEtAl_r2.tex
updates

\documentclass[12pt]{article}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
%\usepackage[maxbibnames=99]{biblatex}
%\usepackage{setspace}
%\usepackage{geometry}
\usepackage[font={sf,small}]{caption}
%\usepackage{setspace}
%\usepackage{geometry}
%\usepackage{hyperref}
%\hypersetup{
%    colorlinks,
%    citecolor=black,
%    filecolor=black,
%    linkcolor=black,
%    urlcolor=black
%}
%\geometry{letterpaper}

\usepackage{amssymb}
%\usepackage{epstopdf}
%\usepackage{float}
%\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\title{A causal role for right frontopolar cortex in directed, but not random, exploration}

\author[1]{Wojciech Zajkowski}
\author[2,3]{Malgorzata Kossut}
\author[3,4]{Robert C. Wilson}

\affil[1]{\small University of Social Sciences and Humanities, Warsaw, Poland}
\affil[2]{Departament of Psychology, University of Social Sciences and Humanities, Warsaw, Poland}
\affil[3]{Nencki Institute, Warsaw, Poland}
\affil[4]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[5]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
\maketitle

\newpage
\begin{abstract}
The explore-exploit dilemma occurs anytime we must choose between exploring unknown options for information and exploiting known resources for reward. Previous work suggests that people use two different strategies to solve the explore-exploit dilemma: directed exploration, driven by information seeking, and random exploration, driven by decision noise. Here, we show that these two strategies rely on different neural systems. Using transcranial magnetic stimulation to inhibit the right frontopolar cortex, we were able to selectively inhibit directed exploration while leaving random exploration intact. This suggests a causal role for right frontopolar cortex in directed, but not random, exploration and that directed and random exploration rely on (at least partially) dissociable neural systems.
\end{abstract}
\newpage

\section*{Introduction}
In an uncertain world adaptive behavior requires us to carefully balance the exploration of new opportunities with the exploitation of known resources.  Finding the optimal balance between exploration and exploitation is a hard computational problem and there is considerable interest in understanding how humans and animals strike this balance in practice \cite{Badre2012-nt,Cavanagh2011-tb,Cohen2007-po,Daw2006-vw,Frank2009-ni,Hills2015-mi,Mehlhorn2015-wl,Wilson2014-ob}.  Recent work has suggested that humans use two distinct strategies to solve the explore-exploit dilemma: directed exploration, based on information seeking, and random exploration, based on decision noise \cite{Wilson2014-ob}.  Even though both of these strategies serve the same purpose, i.e. balancing exploration and exploitation, it is likely they rely on different cognitive mechanisms.  Directed exploration is driven by information and is thought to be computationally complex \cite{Gittins1979-gy,Auer2002-lz,Gittins1974-vc}. On the other hand, random exploration can be implemented in a simpler fashion by using neural or environmental noise to randomize choice \cite{Thompson1933-cl}.

A key question is whether these dissociable behavioral strategies rely on dissociable neural systems. Of particular interest is the frontopolar cortex (FPC) -- an area that has been associated with a number of functions, such as tracking pending and/or alternate options \cite{koechlin2007anterior,Boorman2009-ns}, strategies \cite{Domenech2015-nz} and goals \cite{Pollmann2016-lc} and that has been implicated in exploration itself \cite{Badre2012-nt,Cavanagh2011-tb,Daw2006-vw}. Importantly, however, the exact role that FPC plays in exploration is unknown as how exploration is defined varies from paper to paper. In one line of work, exploration is defined as information seeking. Understood this way, exploration correlates with RFPC activity measured via fMRI \cite{Badre2012-nt} and a frontal theta component in EEG \cite{Cavanagh2011-tb}, suggesting a role for RFPC in directed exploration.  However, in another line of work, exploration is operationalized differently, as choosing the low value option, not the most informative.  Such a measure of exploration is more consistent with random exploration where decision noise drives the sampling of low value options by chance. Defined in this way, exploratory choice correlates with lateral FPC activation \cite{Daw2006-vw} and stimulation and inhibition of RFPC with direct current (tDCS) can increase and decrease the frequency with which such exploratory choices occur \cite{Raja_Beharelle2015-xg}.

Taken together, these two sets of findings suggest that RFPC plays a crucial role in both directed and random exploration. However, we believe that such a conclusion is premature because of a subtle confound that arises between reward and information in most explore-exploit tasks. This confound arises because participants only gain information from the options they choose, yet are incentivized to choose more rewarding options. Thus, over many trials, participants gain more information about more rewarding options and the two ways of defining exploration, i.e. choosing high information or low reward options, become confounded \cite{Wilson2014-ob}.  This makes it impossible to tell whether the link between RFPC and exploration is specific to either directed or random exploration, or whether it is general to both.

To distinguish these interpretations and investigate the causal role of FPC in directed and random exploration, we used continuous theta-burst TMS \cite{Huang2005-mm} to selectively inhibit right frontopolar cortex (RFPC) in participants performing the `Horizon Task', an explore-exploit task specifically designed to separate directed and random exploration \cite{Wilson2014-ob}.  Using this task we find evidence that inhibition of RFPC selectively inhibits directed exploration while leaving random exploration intact.

\section*{Results}
We used our previously published `Horizon Task' (Figure \ref{fig:task}) to measure the effects of TMS stimulation of RFPC on directed and random exploration.  In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions. To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.

\begin{figure}[t]
\begin{center}
		\includegraphics[width=0.8\textwidth]{figures/Figure_1.pdf}
\caption{The Horizon Task. Participants make a series of decisions between two one-armed bandits that pay out probabilistic rewards with unknown means. At the start of each game, `forced-choice' trials give participants partial information about the mean of each option.  We use the forced-choice trials to set up one of two information conditions: (A) an unequal (or [1 3]) condition in which participants see 1 play from one option and 3 plays from the other and (B) an equal (or [2 2]) condition in which participants see 2 plays from both options. A model-free measure of directed exploration is then defined as the change in information seeking with horizon in the unequal condition (A). Likewise a model-free measure of random exploration is defined as the change choosing the low mean option in the equal condition (B).}
\label{fig:task}
\end{center}
\end{figure}

The Horizon Task has two key manipulations that allow us to measure directed and random exploration.  The first manipulation is the horizon itself, i.e. the number of decisions remaining in each game.  The idea behind this manipulation is that when the horizon is long (6 trials), participants should explore more frequently, because any information they acquire from exploring can be used to make better choices later on. In contrast, when the horizon is short (1 trial), participants should exploit the option they believe to be best. Thus, this task allows us to quantify directed and random exploration as changes in information seeking and behavioral variability that occur with horizon.

The second manipulation is the amount of information participants have about each option {\em before} making their first choice.  This information manipulation is achieved by using four forced-choice trials, in which participants are told which option to pick, at the start of each game. We use these forced-choice trials to setup one of two information conditions: an unequal, or [1 3], condition, in which participants see 1 play from one option and 3 plays from the other option, and an unequal, or [2 2], condition, in which participants see two outcomes from both options. By varying the amount of information participants have about each option {\em independent} of the mean payout of that option, this information manipulation allows us to remove the reward-information confound, at least on the first free-choice trial (Figure \ref{fig:confound}).  After the first free-choice trial, however, participants tend to choose more rewarding options more frequently and reward and information are rapidly confounded.  For this reason the bulk of our analyses are focussed on the first free-choice trial where the confound has been removed.

\begin{figure}[t]
\begin{center}
		\includegraphics[width=0.6\textwidth]{figures/Figure_2.pdf}
\caption{The reward-information confound. The y-axis corresponds to the correlation between the sign of the difference in mean ($\mbox{sgn}(\mu_{left} - \mu_{right})$) between options and the sign of difference in the number of times each option has been played ($\mbox{sgn}(n_{left} - n_{right})$). The forced trials are chosen such that the the correlation is approximately zero on the first free-choice trial.  After the first trial, however, a positive correlation quickly emerges as participants choose the more rewarding options more frequently.  This strong confound between reward and information makes it difficult to dissociate directed and random exploration on later trials.
}
\label{fig:confound}
\end{center}
\end{figure}

\subsection*{RFPC stimulation selectively inhibits directed exploration on the first free-choice}

In this section we analyze behavior on the first free-choice trial in each game. This way we are able to remove any effect of the reward-information confound and fairly compare behavior between horizon conditions. We analyze the data with both a model-free approach, using simple statistics of the data to quantify directed and random exploration, as well as a model-based approach, using a cognitive model of the behavior to draw more precise conclusions. Both analyses point to the same conclusion that RFPC stimulation selectively inhibits directed, but not random, exploration.

\subsubsection*{Model-free analysis}

The two information conditions in the Horizon Task allow us to quantify directed and random exploration in a model-free way.  In particular, directed exploration, which involves information seeking, can be quantified as the probability of choosing the high information option, $p(\mbox{high info})$ in the [1 3] condition, while random exploration, which involves decision noise, can be quantified as the probability of making a mistake, or choosing the low mean reward option, $p(\mbox{low mean})$, in the [2 2] condition.

%Information seeking is quantified as the probability of choosing the more uncertain option in the [1 3] condition (i.e. the option played once in the forced-choice trials), $p(\mbox{high info})$.  Behavioral variability is quantified as the number of mistakes, i.e. choosing the low value option, in the [2 2] condition, $p(\mbox{low mean})$. To remove the reward-information confound, both of these measures are computed on the first free-choice trial in each game, i.e. before participants' choices have a chance to confound information and reward.

Using these measures of exploration, we found that inhibiting the RFPC had a significant effect on directed exploration but not random exploration (Figure \ref{fig:mf}A, B).  In particular, for directed exploration, a repeated measures ANOVA with horizon, TMS condition and order as factors revealed a significant interaction between stimulation condition and horizon ($F(1, 24) = 4.96$, $p = 0.036$). Conversely, a similar analysis for random exploration revealed no effects of stimulation condition (main effect of stimulation condition, $F(1, 24) = 0.88$, $p = 0.36$; interaction of stimulation condition with horizon, $F(1, 24) = 1.24$, $p = 0.28$).  Post hoc analyses revealed that the change in directed exploration was driven by changes in information seeking in horizon 6 (one-sided t-test, $t(24) = 2.62$, $p = 0.008$) and not in horizon 1 (two-sided t-test, $t(24) = -0.30$; $p = 0.77$).

\begin{figure}[t]
\begin{center}
		\includegraphics[width=\textwidth]{figures/Figure_3.pdf}
\caption{Model-free analysis of the first free-choice trial shows that RPFC stimulation affects directed, but not random, exploration. (A) In the control (vertex) condition, information seeking increases with horizon, consistent with directed exploration. When RFPC is stimulated, directed exploration is reduced, an effect that is entirely driven by changes in horizon 6 (* denotes $p < 0.02$ and ** denotes $p < 0.005$; error bars are $\pm$ s.e.m.). (B) Random exploration increases with horizon but is not affected by RFPC stimulation. }
\label{fig:mf}
\end{center}
\end{figure}

\subsubsection*{Model-based analysis}
While the model-free analyses are intuitive, the model-free statistics, $p(\mbox{high info})$ and $p(\mbox{low mean})$, are not pure reflections of information seeking and behavioral variability and could be influenced by other factors such as spatial bias and learning.  To account for these possibilities we performed a model-based analysis using a model that extends our earlier work \cite{Wilson2014-ob,Somerville2017-fq,kruegerEtAl2017} see Methods for a complete description.  In this model, the level of directed and random exploration is captured by two parameters: an information bonus for directed exploration, and decision noise for random exploration. In addition the model includes terms for the spatial bias and to describe learning.

\paragraph{Overview of model}
Before presenting the results of the model-based analysis we begin with a brief overview of the most salient points of the model.  A full description of the model can be found in the Methods and code to implement the model can be found in the Supplementary Material.

Conceptually, the model breaks the explore-exploit choice down into two components: a learning component, in which participants estimate the mean payoff of each option from the rewards they see, and a decision component, in which participants use this estimated payoff to guide their choice.  The learning component assumes that participants compute an estimate of the average payoff for each slot machine, $R^i_t$, using a simple delta rule update equation (based on a Kalman filter \cite{Kalman1960-uo}, see Methods):
\begin{equation}
R^i_{t+1}= R^i_t + \alpha_t^i (r_t - R^i_t)
\end{equation}
where $r_t$ is the reward on trial $t$ and $\alpha_t^i$ is the time-varying learning rate that determines the extent to which the prediction error, $(r_t - R^i_t)$, updates the estimate of the mean of bandit $i$. The learning process is described by three free parameters: the initial value of the estimated payoff, $R_0$, and two learning rates, the initial learning rate, $\alpha_1$, and the asymptotic learning rate, $\alpha_{\inf}$, which together describe the evolution of the actual learning rate, $\alpha_t$, over time. For simplicity, we assume that these parameters are independent of horizon and uncertainty condition (Table 1).

The decision component of the model assumes that participants choose between the two options (left and right) probabilistically according to 
\begin{equation}
	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
\end{equation}
where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition). The decision process is described by three free parameters: the information bonus $A$, the spatial bias $B$, and the decision noise $\sigma$.  We estimate separate values of the decision parameters for each horizon and (since the information bonus is only used in the [1 3] condition) separate values of only the bias and decision noise for each uncertainty condition.

\begin{table}[t]
\small
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
Parameter &
Horizon dependent? &
Uncertainty dependent? &
TMS dependent? \\
\hline
prior mean, $R_0$ &
no &
no &
yes \\
\hline
initial learning rate, $\alpha_1$ &
no &
no &
yes \\
\hline
asymptotic learning rate, $\alpha_{\infty}$ &
no &
no &
yes \\
\hline
information bonus, $A$ &
yes &
n/a &
yes \\
\hline
spatial bias, $B$ &
yes &
yes &
yes \\
\hline
decision noise, $\sigma$ &
yes & 
yes &
yes \\
\hline
\end{tabular}
\end{center}
\caption{Model parameters. Subject's behavior on the first free choice of each session is described by 13 free parameters.  Three of these parameters ($R_0$, $\alpha_1$ and $\alpha_{\infty}$) describe the learning process and do not vary with horizon or uncertainty condition. Ten of these parameters ($A$, $B$ and $\sigma$ in the different horizon and information conditions) describe the decision process.  All parameters are estimated for each subject in each stimulation condition and the key analysis asks whether parameters change between vertex and RFPC stimulation.}
\label{table1}
\end{table}%

Overall, subject's behavior in each session (vertex vs RFPC stimulation) is described by 13 free parameters (Table \ref{table1}): three describing learning ($R_0$, $\alpha_1$ and $\alpha_{\infty}$) and 10 describing the decision process ($A$ in the two horizon conditions, $B$ and $\sigma$ in the four horizon-x-uncertainty conditions). These 13 parameters were fit to each subject in each stimulation condition using a hierarchical Bayesian approach \cite{Lee2014-yh} (see Methods).

\paragraph{Model fitting results} Posterior distributions over the group-level means are shown in the left column of Figure \ref{fig:mb1}, while posteriors over the TMS-related change in parameters are shown in the right column.  Both columns suggest a selective effect of RFPC stimulation on the information bonus in horizon 6.
\begin{figure}[thp]
\begin{center}
		\includegraphics[width=\textwidth]{figures/Figure_4.pdf}
\caption{Model-based analysis of the first free-choice trial showing the effect of RFPC stimulation on each of the 13 parameters. Left column: Posterior distributions over each parameter value for RFPC and vertex stimulation condition. Right column: posterior distributions over the change in each parameter between stimulation conditions.  Note that, because information bonus, decision noise and spatial bias are all in units of points, we plot them on the same scale to facilitate comparison of effect size.}
\label{fig:mb1}
\end{center}
\end{figure}

Focussing on the left column first, overall the parameter values seem reasonable.  The prior mean is close to the generative mean of 50 used in the actual experiment, and the decision parameters are comparable to those found in our previous work \cite{Wilson2014-ob}.  The learning rate parameters, $\alpha_1$ and $\alpha_{\infty}$, were not included in our previous models and are worth discussing in more detail.  As expected for Bayesian learning \cite{Kalman1960-uo,Nassar2010-fg}, the initial learning rate is higher than the asymptotic learning rate (95\% of samples in the vertex condition, 94\% in the RFPC condition).  However, the actual values of the learning rates are quite far from their `optimal' settings of $\alpha_1 = 1$ and $\alpha_{\infty} = 0$ that would correspond to perfectly computing the mean reward.  This suggests a greater than optimal reliance on the prior ($\alpha_1 < 1$) and a pronounced recency bias ($\alpha_{\infty} > 0$) such that the most recent rewards are weighted more heavily in the computation of expected reward, $R^i_{t}$.  Both of these findings are likely due to the fact that the version of the task we employed did not keep the outcomes of the forced trials on screen and instead relied on people's memories to compute the expected value. %Indeed, when we fit this same model to the data from our original experiment \cite{Wilson2014-ob} we find that $\alpha_1$ and $\alpha_{\infty}$ are much closer to 1 and 0 respectively (see Figure SXXX).

Turning to the right hand column of Figure \ref{fig:mb1}, we can see that the model-based analysis yields similar result to the model-free analysis.  In particular we see a reduction (of about 4.8 points) in the information bonus in horizon 6 (with 99\% of samples showing a reduced information bonus in the RFPC stimulation condition) and no effect on decision noise in either horizon in either the [2 2] or [1 3] uncertainty conditions (with between 40\% and 63\% of samples below zero).

In addition to the effect on the information bonus in horizon 6, there is also a hint of an effect on the information bonus in horizon 1 (85\% samples less than zero) {\color{red} and on the prior mean $R_0$ (88\% samples above zero). While these results may suggest that RFPC stimulation affects more than just information bonus in horizon 6, they more likely reflect an inherent tradeoff between prior mean and information bonus that is peculiar to this task.  In particular, because the prior mean has a stronger effect on the more uncertain option, an increase in $R_0$ increases the value of the more informative option in much the same way as an information bonus.  Thus, when applied to this task, the model has a built in tradeoff between prior mean and information bonus that can muddy the interpretation of both.  Note that this tradeoff is not a general feature of the model and could be removed with a different task design that employed more forced choice trials and hence more time for the effects of the prior to be removed.

Figure \ref{fig:mb2} exposes the tradeoff between $R_0$ and $A$ in more detail.  Panels A and B plot samples from the posterior over the TMS-related change in information bonus, $A(\mbox{vertex}) - A(\mbox{RFPC})$, against the TMS-related change in prior mean, $R_0(\mbox{vertex}) - R_0(\mbox{RFPC})$.  For both horizon conditions we see a strong negative correlation such that increasing $R_0$ decreases $A$.  This negative correlation is especially problematic for the interpretation of the horizon 1 change in information bonus where a sizable fraction of the posterior centers on no change in either variable.  In contrast the negative correlation between $A$ and $R_0$ does not affect our interpretation of the horizon 6 result where the TMS-related change in $A$ is negative regardless of of the change in $R_0$.

Finally  we asked whether the horizon-dependent change in information seeking, i.e. $\Delta A = A(h = 6) - A(h = 1)$, was different in each TMS condition.  As shown in Figure \ref{fig:mb2}C, the TMS-related change in $\Delta A$ is about -3.1 points (94\% samples below 0) and is uncorrelated with the TMS-related change in $R_0$.  Taken together, this suggests that we can be fairly confident in our claim that RFPC stimulation has a selective effect on directed exploration.
}

\begin{figure}[thp]
\begin{center}
		\includegraphics[width=\textwidth]{figures/Figure_5.pdf}
\caption{\color{red} Correlation between TMS-induced changes in information bonus, $A$, and TMS-induced changes in the prior mean, $R_0$.  (A, B) Samples from the posterior distributions over the TMS-related changes in prior mean, $R_0$, and TMS-related change in information bonus in horizon 1 (A) and horizon 6 (B).  In both cases we see a negative correlation between the change in $R_0$ and the change in $A$ consistent with a tradeoff between these variables in the model.  (C)  Samples from the posterior over the effect of TMS stimulation on the horizon-related change in information bonus, $\Delta A = A(h = 6) - A(h = 1)$ plotted against samples from the TMS-related change in prior mean.  Here we see no correlation between variables and the majority of $\Delta A(\mbox{vertex}) - \Delta A(\mbox{RFPC})$ samples below zero consistent with an effect of RFPC stimulation on directed exploration.}
\label{fig:mb2}
\end{center}
\end{figure}

\subsection*{The effect of RFPC stimulation on later trials}
Our analyses so far have focussed on just the first free choice and have ignored the remaining five choices in the horizon 6 games. The reason for this is the reward-information confound, illustrated in Figure \ref{fig:confound}, which makes interpretation of the later trials more difficult. Despite this difficulty, we note that in Figure \ref{fig:confound} the size of the confound is almost identical in the two stimulation conditions and so we proceed, with caution, to present a model-free analysis of the later trials below.

\begin{figure}[th]
\begin{center}
		\includegraphics[width=0.8\textwidth]{figures/Figure_6.pdf}
\caption{Model-free analysis of all trials. (A, B) Model-free measures of directed (A) and random (B) exploration as a function of trial number suggests a reduction in both directed and random exploration over the course of the game.  (C, D) TMS-induced change in measures of directed and random exploration as a function of trial number. This suggests that the reduction in directed exploration on the first free-choice trial, persists into the second trial of the game.}
\label{fig:later}
\end{center}
\end{figure}

In Figure \ref{fig:later} we plot the model-free measures, $p(\mbox{high info})$ and $p(\mbox{low mean})$, as a function of trial number. Both measures show a decrease over the course of the horizon 6 games although, because of the confound, it is difficult to say whether these changes reflect a reduction in directed exploration, random exploration, or both. Interestingly, the differences in $p(\mbox{high info})$ between vertex and RFPC conditions on the first free-choice trial appear to persist into the second, a result that becomes more apparent when we plot the TMS-related change, i.e. $p(\mbox{high info, RFPC}) - p(\mbox{high info, vertex})$ (Figure \ref{fig:later}C, D).  More formally a repeated measures ANOVA with trial number, TMS condition as factors reveals a significant main effect of trial number ($F(5,120) = 126$, $p < 10^{-45}$), no main effect of TMS condition ($F(1, 120) = 1.17$, $p = 0.29$) and a near significant interaction between trial number and TMS condition ($F(5, 120) = 2.26, p = 0.053$).  A post hoc, one-sided t-test on the second trial reveals a marginally significant reduction in $p(\mbox{high info})$ on the second trial ($t(24) = 1.61$; $p = 0.06$).  In contrast, a similar analysis for random exploration shows no evidence for any effect of TMS condition on $p(\mbox{low mean})$ (main effect of TMS, $F(1, 120) = 0.16$, $p = 0.69$; TMS x trial number, $F(5, 120) = 0.69$, $p = 0.63$) although the main effect of trial number persists ($F(5,120) = 13.7$, $p < 10^{-9}$).  Thus, the analysis of later trials provides additional, albeit modest, support for the idea that RFPC stimulation selectively disrupts directed but not random exploration at long horizons.

\section*{Discussion}
In this work we used continuous theta-burst transcranial magnetic stimulation (cTBS) to investigate whether right frontopolar cortex (RFPC) is causally involved in directed and random exploration.  Using a task that is able to behaviorally dissociate these two types of exploration, we found that inhibition of RFPC caused a selective reduction in directed, but not random exploration.  To the best of our knowledge, this finding represents the first causal evidence that directed and random exploration rely on dissociable neural systems and is consistent with our recent findings showing that directed and random exploration have different developmental profiles \cite{Somerville2017-fq}.  This suggests that, contrary to the assumption underlying many contemporary studies \cite{Daw2006-vw,Badre2012-nt}, exploration is not a unitary process, but a dual process in which the distinct strategies of information seeking and choice randomization are implemented via distinct neural systems.

Such a dual-process view of exploration is consistent with the classical idea that there are multiple types of exploration \cite{Berlyne1966-fg}.  In particular Berlyne's constructs of `specific exploration', involving a drive for information and `diversive exploration', involving a drive for variety, bear a striking resemblance to our definitions of directed and random exploration. Despite the importance of Berlyne's work, more modern views of exploration tend not to make the distinction between different types of exploration, considering instead a single exploratory state or exploratory drive that controls information seeking across a wide range of tasks \cite{Berlyne1966-fg,Aston-Jones2005-hx,Hills2015-mi,Kidd2015-so}. At face value, such unitary accounts seem at odds with a dual-process view of exploration. However, these two viewpoints can be reconciled if we allow for the possibility that, while directed and random exploration are implemented by different systems, their levels are set by a common exploratory drive.

Intriguingly, individual differences in behavior on the Horizon Task provide some support for the idea that directed and random exploration are driven by a common source. In particular, in a large behavioral data set of 277 people performing the Horizon Task, we find a positive correlation between the levels of directed and random exploration such that people with high levels of directed exploration also tend to have high levels random exploration ($r(275) = 0.29$, $p < 10^{-5}$), Figure \ref{fig:dirRandCorr}.  This is consistent with the idea that the levels of directed and random exploration are set by the strength of an exploratory drive that varies as an individual difference between people.

\begin{figure}[t]
\begin{center}
		\includegraphics[width=0.6\textwidth]{figures/Figure_7.pdf}
\caption{Correlation between individual differences in the levels of directed and random exploration in a sample of 277 people performing the Horizon Task.}
\label{fig:dirRandCorr}
\end{center}
\end{figure}

While the present study does allow us to conclude that directed and random exploration rely on different neural systems, the limited spatial specificity of TMS limits our ability to say exactly what those systems are. In particular, because the spatial extent of TMS is quite large, stimulation aimed at frontal pole may directly affect activity in nearby areas such as ventromedial prefrontal cortex (vmPFC) and orbitofrontal cortex (OFC), both areas that have been implicated in exploratory decision making and that may be contributing to our effect \cite{Daw2006-vw}. In addition to such direct effects of TMS on nearby regions, indirect changes in areas that are connected to the frontal pole could also be driving our effect.  For example, cTBS of left frontal pole has been associated with changes in blood perfusion in areas such as amygdala, fusiform gyrus and posterior parietal cortex \cite{Volman2011-fz} and by changes in the fMRI BOLD signal in OFC, insula and striatum \cite{hanlon2017left}. In addition \cite{Volman2011-fz} showed that unilateral cTBS of left frontal pole is associated with changes in blood perfusion to the right frontal pole. Indeed, such a bilateral effect of cTBS may explain why our intervention was effective at all given that a number of neuroimaging studies have shown bilateral activation of the frontal pole associated with exploration \cite{Daw2006-vw,Badre2012-nt}. Future work combining cTBS with neuroimaging will be necessary to shed light on these issues.

With the above caveats that our results may not be entirely due to disruption of frontal pole, the interpretation that RFPC plays a role in directed, but not random, exploration is consistent with a number of previous findings.  For example, frontal pole has been associated with tracking the value of the best unchosen option \cite{Boorman2009-ns}, inferring the reliability of alternate strategies \cite{Boorman2009-ns,Domenech2015-nz}, arbitrating between old and new strategies \cite{Donoso2014-at,Mansouri2015-xa}, and reallocating cognitive resources among potential goals in underspecified situations \cite{Pollmann2016-lc}.  Taken together, these findings suggest a role for frontal pole in model-based decisions \cite{Daw2006-vw} that involve long-term planning and the consideration of alternative actions. From this perspective, it is perhaps not surprising that directed exploration relies on RFPC, since computing an information bonus relies heavily on an internal model of the world. It is also perhaps not surprising that random exploration is independent of RFPC, as this simpler strategy could be implemented without reference to an internal model. Indeed, the ability to explore effectively in a model-free manner, may be an important function of random exploration as it allows us to explore even when our model of the world is wrong.

More generally, it is unlikely that frontal pole is the only area involved in directed exploration, and more work will be needed to map out the areas involved in directed and random exploration and expose their causal relationship to explore-exploit behavior.

\section*{Methods}
\subsection*{Participants}
31 healthy right-handed, adult volunteers (19 female, 12 male; ages 19-32). An initial sample size of 16 was chosen based on two studies using a very similar cTBS design that stimulated lateral FPC (Costa et. al. 2011; 2013) and this was augmented to 31 on the basis of feedback from reviewers. 5 participants (5 female, 0 male) were excluded from the analysis due to chance-level performance in both experimental sessions. One (female) participant failed to return for the second (vertex stimulation condition) session and is excluded from the model-free analyses but not the model-based analyses as that can handle missing data more gracefully. Thus our final data set consisted of 25 participants (13 female, 12 male, ages 19-32) with complete data and one participant (female, aged 20) with data from the RFPC session only.

All participants were informed about potential risks connected to TMS and signed a written consent. The study was approved by University of Social Sciences and Humanities ethics committee.

\subsection*{Procedure}
There were two experimental TMS sessions and a preceding MRI session.  On the first session T1 structural images were acquired using a 3T Siemens TRIO scanner. The scanning session lasted up to 10 minutes. Before the first two sessions, participants filled in standard safety questionnaires regarding MRI scanning and TMS. During the experimental sessions, prior to the stimulation participants went through 16 training games to get accustomed to the task. Afterwards, resting motor thresholds were obtained and the stimulation took place. Participants began the main task immediately after stimulation. The two experimental sessions were performed with an intersession interval of at least 5 days. The order of stimulation conditions was counterbalanced across subjects. All sessions took place at Nencki Institute of Experimental Biology in Warsaw.

\subsection*{Stimulation site}

The RFPC peak was defined as [x,y,z]= [35,50,15] in MNI (Montreal Neurological Institute) space. The coordinates were based on a number of fMRI findings that indicated RFPC involvement in exploration \cite{Badre2012-nt,Boorman2009-ns,Daw2006-vw} and constrained by the plausibility of stimulation (e.g. defining `z' coordinate lower would result in the coil being placed uncomfortably close to the eyes). Vertex corresponded to the Cz position of the 10-20 EEG system. In order to locate the stimulation sites we used a frameless neuronavigation system (Brainsight software, Rogue Research, Montreal, Canada) with a Polaris Vicra infrared camera (Northern Digital, Waterloo, Ontario, Canada).

\subsection*{TMS protocol}

We used continuous theta burst stimulation (cTBS) (Huang et al, 2005). cTBS requires 50Hz stimulation at 80\% resting motor threshold. 40 second stimulation is equivalent to 600 pulses and can decrease cortical excitability for up to 50 minutes \cite{Wischnewski2015-rd}.

Individual resting motor thresholds were assessed by stimulating the right motor knob and inspecting if the stimulation caused an involuntary hand twitch in 50\% of the cases.  We used a MagPro X100 stimulator (MagVenture, Hueckelhoven, Germany) with a 70mm figure-eight coil. The TMS was delivered in line with established safety guidelines \cite{Rossi2009-sf}.

\subsection*{Limitations}

Defining stimulation target by peak coordinates based on findings from previous studies did not allow to account for individual differences in either brain anatomy or the impact of TMS on brain networks \cite{Gratton2013-wz}. However, a study by Volman and colleagues \cite{Volman2011-fz} that used the same theta-burst protocol on the left frontopolar cortex has shown bilateral inhibitory effects on blood perfusion in the frontal pole. This suggests that both right and left parts of the frontopolar cortex might have been inhibited in our experiment, which is consistent with imaging results indicating bilateral involvement of the frontal pole in exploratory decisions.

\subsection*{Task}
The task was a modified version of the Horizon Task \cite{Wilson2014-ob}. As in the original paper, the distributions of payoffs tied to bandits were independent between games and drawn from a Gaussian distribution with variable means and fixed standard deviation of 8 points. Participants were informed that in every game one of the bandits was objectively `better' (has a higher payoff mean). Differences between the mean payouts of the two slot machines were set to either 4, 8, 12 or 20. One of the means was always equal to either 40 or 60 and the second was set accordingly. The order of games was randomized. Mean sizes and order of presentation were counterbalanced. Participants played 160 games and the whole task lasted between 39 and 50 minutes (mean 43.4 minutes).

Each game consisted of 5 or 10 choices. Every game started with a screen saying `New game' and information about whether it was a long or short horizon, followed by sequentially presented choices. Every choice was presented on a separate screen, so that participants had to keep previous the scores in memory. There was no time limit for decisions. During forced choices participants had to press the prompted key to move to the next choice. During free choices they could press either `z' or `m' to indicate their choice of left or right bandit. The decision could not be made in a time shorter than 200ms, preventing participants from accidentally responding too soon. The score feedback was presented for 500ms. A counter at the bottom of the screen indicated the number of choices left in a given game. The task was programmed using PsychoPy software v1.86 \cite{Peirce2007-gr}.

Participants were rewarded based on points scored in two sessions. The payoff bounds were set between 50 and 80 zl (equivalent to approximately 12 and 19 euro). Participants were informed about their score and monetary reward after the second session.

Finally, the random seeds were not perfectly controlled between subjects.  The first 16 subjects ran the task with identical random seeds and thus all 16 saw the same sequence of forced-choice trials in both vertex and RFPC sessions. For the remaining subjects the random seed was unique for each subject and each session, thus these subjects had unique a series of forced-choice trials for each session.  Despite this limitation we saw no evidence of different behavior across the two groups.

\subsection*{Data and code}
Behavioral data as well as Matlab code to recreate the main figures from this paper can be found on the Dataverse website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.

\subsection*{Model-based analysis}
We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \cite{Wilson2014-ob} that was modified to include a learning component. In particular, we assume that participants use the outcomes of the forced-choice trials to learn an estimate of the mean reward of each option, before inputting that mean reward into a decision function that includes terms for directed and random exploration.  This model naturally decomposes into a learning component and a decision component and we consider each of these components in turn.

\subsubsection*{Learning component}
The learning component of the model assumes that participants use a Kalman filter to learn a value for the mean reward of each option. The Kalman filter \cite{Kalman1960-uo} has been used to model learning in other explore-exploit tasks \cite{Daw2006-vw} and is a popular model of Bayesian learning as it is both analytically tractable and easily relatable to the delta-rule update equations of reinforcement learning.

More specifically, the Kalman filter assumes a generative model in which the rewards from each bandit, $r_t$, are generated from Gaussian distribution with a fixed standard deviation, $\sigma_r$, and a mean, $m^i_t$, that is different for each bandit and can vary over time. The time dependence of the mean is determined by a Gaussian random walk with mean 0 and standard deviation $\sigma_d$. Note that this generative model, assumed by the Kalman filter, is slightly different to the true generative model used in the Horizon Task, which assumes that the mean of each bandit is constant over time, i.e. $\sigma_d = 0.$  This mismatch between the assumed and actual generative models, is quite deliberate and allows us to account for the suboptimal learning of the subjects. In particular, this mismatch, introduces the possibility of a recency bias (when $\sigma_d > 0$) whereby more recent rewards are over-weighted in the computation of $R^i_t$.

The actual equations of the Kalman filter model are straightforward. The model keeps track of an estimate of both the mean reward, $R^i_t$, of each option, $i$, and the uncertainty in that estimate, $\sigma^i_t$.  When option $i$ is played on trial $t$, these two parameters update according to

\begin{equation}
\label{eq:1}
	R^i_{t+1} = R^i_t + \frac{(\sigma^{i}_{t+1})^2}{\sigma_r^2} (r_t - R^i_t)
\end{equation}
and
\begin{equation}
\label{eq:2}
	\frac{1}{(\sigma^i_{t+1})^2} = \frac{1}{(\sigma^i_t)^2 + \sigma_d^2} + \frac{1}{\sigma_r^2}
\end{equation}
When option i is not played on trial t we assume that the estimate of the mean stays the same, but that the uncertainty in this estimate grows as the generative model assumes the mean drifts over time.  Thus for unchosen option $j$ we have
\begin{equation}
R^j_{t+1} = R^j_t \quad \quad \mbox{and} \quad \quad (\sigma^j_{t+1})^2 = (\sigma^j_{t})^2 + \sigma_d^2
\end{equation}
When the option is played, the update equation \ref{eq:1} for $R^i_t$ is essentially just a `delta rule' \cite{rescorla1972theory,Schultz1997-oz}, with the estimate of the mean being updated in proportion to the prediction error, $r_t - R^i_t$.  This relationship to the reinforcement learning literature is made more clear by rewriting the learning equations in terms of the time varying learning rate, 
\begin{equation}
	\alpha^i_{t} = \frac{(\sigma^i_{t+1})^2}{\sigma_r^2 }
\end{equation}
Written in terms of this learning rate, equations \ref{eq:1} and \ref{eq:2} become
\begin{equation}
	R^i_{t+1} = R^i_t + \alpha^i_{t} (r_t - R^i_t) \\
\end{equation}
and
\begin{equation}
	\frac{1}{\alpha^i_{t}} = \frac{1}{\alpha^i_{t-1} + \alpha_d} + 1
\end{equation}
where 
\begin{equation}
	\alpha_{d} = \frac{\sigma_{d}^2}{\sigma_r^2}
\end{equation}
The learning model has four free parameters, the noise variance, $\sigma_r^2$, the drift variance, $\sigma_d^2$, and the initial values of the estimated reward, $R_0$, and uncertainty in that variance estimate, $\sigma_0^2$.  In practice, only three of these parameters are identifiable from behavioral data, and we will find it useful to reparameterize the learning model in terms of $R_0$ and an initial, $\alpha_1$, and asymptotic, $\alpha_{\infty}$, learning rate.  In particular, the initial value of the learning rate relates to $\sigma_0$, $\sigma_r$ and $\sigma_d$ as
\begin{equation}
\alpha_1 = \frac{\sigma_0^2 + \sigma_d^2}{\sigma_r^2}
\end{equation}
While the asymptotic value of the learning rate, which corresponds to the steady state value of $\alpha^i_t$ if option $i$ is played forever, relates to $\alpha_d$ (and hence $\sigma_d$ and $\sigma_r$) as
\begin{equation}
\alpha_\infty = \frac{1}{2} \left( -\alpha_d + \sqrt{\alpha_d^2 + 4 \alpha_d} \right)
\end{equation}
While this choice to parameterize the learning equations in terms of $\alpha_1$ and $\alpha_{\infty}$ is somewhat arbitrary, we feel that the learning rate parameterization has the advantage of being slightly more intuitive and leads to parameter values between 0 and 1 which are (at least for us) easier to interpret.

\subsubsection*{Decision component}
Once the payoffs of each option, $R^i_t$, have been estimated from the outcomes of the forced-choice trials, the model makes a decision using a simple logistic choice rule:
\begin{equation}
	p(\mbox{choose right}) = \frac{1}{ 1 + \exp \left( \frac{\Delta R + A \Delta I + B}{\sigma}\right) }
\end{equation}
where $\Delta R$ ( $= R^{left}_t - R^{right}_t$ ) is the difference in expected reward between left and right options and $\Delta I$ is the difference in information between left and right options (which we define as +1 when left is more informative, -1 when right is more informative, and 0 when both options convey equal information in the [2 2] condition). The three free parameters of the decision process are: the information bonus, $A$, the spatial bias, $B$, and the decision noise $\sigma$.  We assume that these three decision parameters can take on different values in the different horizon and uncertainty conditions (with the proviso that $A$ is undefined in the [2 2] information condition since $\Delta I = 0$). Thus the decision component of the model has 10 free parameters ($A$ in the two horizon conditions, and $B$ and $\sigma$ in the 4 horizon x uncertainty conditions).  Directed exploration is then quantified as the change in information bonus with horizon, while random exploration is quantified as the change in decision noise with horizon.

\subsection*{Model Fitting}
\subsubsection*{Hierarchical Bayesian Model}
Between the learning and decision components of the model, each subject's behavior is described by 13 free parameters, all of which are allowed to vary between TMS conditions.  These parameters are: the initial mean, R$_0$, the initial learning rate, $\alpha_1$, the asymptotic learning rate,$\alpha_{\infty}$, the information bonus, $A$, in both horizon conditions, the spatial bias, $B$, in the four horizon x uncertainty conditions, and the decision noise, $\sigma$, in the four horizon x uncertainty conditions (Table \ref{tab:pars2}, Figure \ref{fig:graphicalModel}).

Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \cite{Lee2014-yh}.  In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure. The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.  For notational convenience, we refer to the hyperparameters that define the prior for variable X as $\theta^X$.  In addition we use subscripts to refer to the dependence of both parameters and hyperparameters on TMS stimulation condition, $\tau$, horizon condition, $h$, uncertainty condition, $u$, subject, $s$, and game, $g$.

The particular priors and hyperpriors for each parameter are shown in Table \ref{tab:pars2}. For example, we assume that the prior mean, $R_0^{\tau s}$, for each stimulation condition $\tau$ and horizon condition $h$, is sampled from a Gaussian prior with mean $\mu_{R_0}^{\tau}$ and standard deviation $\sigma_{R_0}^\tau$. These prior parameters are sampled in turn from  their respective hyperpriors: $\mu_{R_0}^{\tau}$, from a Gaussian distribution with mean 50 and standard deviation 14, $\sigma_{R_0}^\tau$ from a Gamma distribution with shape parameter 1 and rate parameter 0.001.

\begin{table}[htdp]
\small
\begin{tabular}{|c|c|c|c|}
	\hline
	Parameter & Prior & Hyperparameters & Hyperpriors \\
	\hline
	prior mean, $R_0^{\tau s}$ 
		& $R_0^{\tau s} \sim$ Gaussian($\mu_{R_0}^{\tau }$, $\sigma_{R_0}^{\tau }$) 
		&  $\theta^{\tau}_{R_0} = (\mu_{R_0}^{\tau }, \sigma_{R_0}^{\tau })$
		&\specialcell{
			$\mu_{R_0}^{\tau } \sim $ Gaussian( 50, 14 ) \\ 
			$\sigma_{R_0}^{\tau } \sim $ Gamma( 1, 0.001 )}
		\\
	\hline
	initial learning rate, $\alpha_1^{\tau s}$ 
		& $\alpha_1^{\tau s} \sim$ Beta($a_{\alpha_1}^{\tau }$, $b_{\alpha_1}^{\tau }$) 
		& $\theta^{\tau}_{\alpha_1} = (a_{\alpha_1}^{\tau }, b_{\alpha_1}^{\tau })$
		& \specialcell{
			$a_{\alpha_1}^{\tau } \sim $ Uniform( 0.1, 10 ) \\ 
			$b_{\alpha_1}^{\tau } \sim $ Uniform( 0.5, 10 )}
		\\
	\hline
	asymptotic learning rate, $\alpha_{\infty}^{\tau s}$ 
		& $\alpha_{\infty}^{\tau s} \sim $ Beta($a_{\alpha_{\infty}}^{\tau }$, $b_{\alpha_{\infty}}^{\tau }$) 
		& $\theta^{\tau}_{\alpha_\infty} = (a_{\alpha_\infty}^{\tau }, b_{\alpha_\infty}^{\tau })$
		& \specialcell{
			$a_{\alpha_{\infty}}^{\tau } \sim $ Uniform( 0.1, 10 ) \\ 
			$b_{\alpha_{\infty}}^{\tau } \sim $ Uniform( 0.1, 10 )}
		\\
	\hline
	information bonus, $A^{\tau s h u}$ 
		& $A^{\tau s h u} \sim $  Gaussian($\mu_A^{\tau h u}$, $\sigma_A^{\tau h u}$) 
		& $\theta_A^{\tau h u} = (\mu_A^{\tau h u}, \sigma_A^{\tau h u}) $
		& \specialcell{
			$\mu_A^{\tau h u} \sim $ Gaussian( 0, 100 ) \\ 
			$\sigma_A^{\tau h u} \sim $ Gamma( 1, 0.001 )}		\\
	\hline
	spatial bias, $B^{\tau s h u}$ 
		& $B^{\tau s h u} \sim $ Gaussian($\mu_B^{\tau h u}$, $\sigma_B^{\tau h u}$) 
		& $\theta_B^{\tau h u} = (\mu_B^{\tau h u}, \sigma_B^{\tau h u}) $
		& \specialcell{
			$\mu_B^{\tau h u} \sim $ Gaussian( 0, 100) \\ 
			$\sigma_B^{\tau h u} \sim $ Gamma( 1, 0.001 )}		\\
	\hline
	decision noise, $\sigma^{\tau s h u}$ 
		& $\sigma^{\tau s h u} \sim $ Gamma($k_\sigma^{\tau h u}$, $\lambda_\sigma^{\tau h u}$) 
		& $\theta_\sigma^{\tau h u} = (k_\sigma^{\tau h u}, \lambda_\sigma^{\tau h u}) $
		& \specialcell{
			$k_{\sigma}^{\tau h u} \sim $ Exp(0.1) \\ 
			$\lambda_{\sigma}^{\tau h u} \sim $ Exp( 10 )}		\\
	\hline
\end{tabular}
\caption{Model parameters, priors, hyperparameters and hyperpriors.}
\label{tab:pars2}	
\end{table}%

%horizonTask_TMS_KFcond_v2.pdf

\begin{figure}[tp]
\begin{center}
		\includegraphics[width=0.8\textwidth]{figures/Figure_8.pdf}
\caption{Graphical representation of the model.  Each variable is represented by a node, with edges denoting the dependence between variables.  Shaded nodes correspond to observed variables, i.e. the free choices $c^{\tau shug}$, forced-trial rewards, ${\bf r}^{\tau shug}$ and forced-trial choices ${\bf a}^{\tau shug}$. Unshaded nodes correspond to unobserved variables whose values are inferred by the model. }
\label{fig:graphicalModel}
\end{center}
\end{figure}

\subsubsection*{Model fitting using MCMC}
The model was fit to the data using Markov Chain Monte Carlo approach implemented in the JAGS package \cite{plummer2003jags} via the MATJAGS interface (psiexp.ss.uci.edu/research/programs\_data/jags/).  This package approximates the posterior distribution over model parameters by generating samples from this posterior distribution given the observed behavioral data.

In particular we used 4 independent Markov chains to generate 4000 samples from the posterior distribution over parameters (1000 samples per chain).  Each chain had a burn in period of 500 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.  Convergence of the Markov chains was confirmed post hoc by eye. Code and data to replicate our analysis and reproduce our Figures is provided as part of the Supplementary Materials.
%References
%Plummer, M. (2003, March). JAGS: A program for analysis of Bayesian graphical models using Gibbs sampling. In Proceedings of the 3rd international workshop on distributed statistical computing (Vol. 124, p. 125).
%Rescorla, R. A., & Wagner, A. R. (1972). A theory of Pavlovian conditioning: Variations in the effectiveness of reinforcement and nonreinforcement. Classical conditioning II: Current research and theory, 2, 64-99.

\bibliographystyle{apalike}
\bibliography{references}

\end{document}

### 2022-12-12 manuscript/Older drafts/Older version/V2/Internal-External Noise.tex
updates

--- 2022-11-21 draft/Older drafts/Older version/V1/Internal-External Noise.tex
@@
 
-	Imagine  trying to decide where to go to dinner with a friend ...
-EXPLORE-EXPLOIT EXAMPLE.
-More generally, such decisions are known as explore-exploit decisions ...
-there's considerable interest in how humans and animals solve it CITE 10 PAPERS INCLUDING WILSON ET AL.
-2014...
-
-	%The explore-exploit tradeoff is a fundamental behavioral dilemma faced by all adaptive organisms.
-Should we explore new options in the hopes of finding a better meal, a better house or a better mate, or should we exploit the options we currently believe to be best?
-Striking the right balance between exploration and exploitation is hard computational problem and there is significant interest in how humans and other animals make explore-exploit decisions in practice.
+	Imagine trying to decide where to go to dinner with a friend, you can go to your favorite restaurant that you both really enjoy and always go to, or you can try the new restaurant that just opened a few days ago right across the street to the other restaurant which may end up being your newly favorite.
+More generally, such decisions are known as explore-exploit decisions.
+The explore-exploit decision refers to deciding between exploiting the best known option so far, like going to your old favorite restaurant, and exploring other options for potential better decisions in the future, like trying the new restaurant.
+There's considerable interest in how humans and animals solve it.
+(CITE 10 PAPERS INCLUDING WILSON ET AL.
+2014...)
 
@@
 In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-IN OUR RESTAURANT EXAMPLE ...
+In our restaurant example, your restaurant decision does not always depend on the quality of the food, you are very likely to go to the new restaurant if you happen to see another old friend walking right in, or you wait until the last moment to make a split second decision about where to go as if you flipped a mental coin in your head and decide to go to your old favorite if it's heads up.
 In theory, such “random exploration”, is surprisingly effective and, if implemented correctly, can come close to optimal performance \cite{}.
@@
 the number of decisions remaining to make.
-THE IDEA BEHIND THIS MANIPULATION IS THAT PEOPLE SHOULD EXPLORE MORE IN THE LONG HORIZON CONDITION (LINK BACK TO EXAMPLE).
-USING SUCH A HORIZON MANIPULATION WE FOUND THAT  people have greater decision noise in the long versus the short horizon condition.
+The idea behind this manipulation is that people should explore more in the long horizon condition.
+If you are leaving town tomorrow for vacation, you probably want to go to your old favorite to guarantee a good last meal, but if you are not going anywhere, you would be more likely to try the new restaurant.
+Using such a horizon manipulation we found that people have greater decision noise in the long versus the short horizon condition.
 
@@
 Of particular interest is whether the adaptive decision noise that is linked to exploration is generated internally, within the brain, or arises externally, in the input from the world.
+In the restaurant example, an old friend walking by would be a source of external noise, but flipping a mental coin would be an internal noise.
 Previous work makes a strong case for both types of noise being relevant to behavior.
@@
 	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of our previously published `Horizon Task' (Figure \ref{fig:task}) to show the influence of internal vs external noise on people's decisions (CITE).
-THE KEY MANIPULATION WAS REPEATED GAMES SAME DECISION TWICE - LINK INTO EXAMPLE
+	We used a modified version of our previously published `Horizon Task' (Figure \ref{fig:task}) to show the influence of internal vs external noise on people's decisions \cite{}.
+The key manipulation was to use repeated games to let people make the same decision twice.
+In the restaurant example, if your decision is mainly driven by external noise, then a few months later when you see the friend walking in front of you into the restaurant, you are very likely to make the same decision and follow him into that restaurant and says hi.
+However, if your decision is mainly driven by internal noise, then next time you make a split-second decision about where to go, you are equally likely to go to the other restaurant.
 
-	MORE SPECIFICALLY, 
+	More specifically, we look at the contribution of external and internal noise by providing with participants the same decision problem twice in the task.
 In this task, participants play a set of games in which they make choices between two slot machines (one-armed bandits) that pay out rewards from different Gaussian distributions.
@@
 After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
+In unequal uncertainty condition, people are more likely to choose the option that they know less about - the more informative option - to explicitly explore that option more.
+This type of information driven exploration is known as directed exploration.
 
-	THESE CONDITIONS ALLOW US TO MEASURED DIRECTED AND RNADOM EXPLORATION IN A MODEL-FREE WAY.  REPLICATION OF PREVIOUS RESULTS
+	These conditions allow us to measure directed and random exploration in a model-free way.
+Directed exploration is measured as the probability of choosing the more informative option in [1 3] condition whereas random exploration is measured as the probability of choosing the low mean option in [2 2] condition.
+In line with previous results, we showed that both directed and random exploration increase with horizon.
+Directed exploration is considered to be driven by information bias and random exploration is considered to be driven by decision noise, in this work, we are investigating which source of decision noise, external vs internal, drives random exploration.
+
+	%MOVE THIS PARAGRAPH UP As in our previous paper \cite{}, the influence of directed exploration is measured as the probability of choosing the more informative option, i.e.
+the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfig.PNG}
+			\includegraphics[width=\textwidth]{figures/task.PNG}
 			\caption{A key additional manipulation here is repeated games.
@@
 We setup the repeated games such that they are at least 5 games apart from each other.
-A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of disagreement for repeated games.}
+A model-free measure of choice inconsistency which reflex the underlying decision noise is defined as the proportion of inconsistent choices for repeated games.}
 			\label{fig:task2}
@@
 The idea behind this measure is that purely external noise should lead to consistent choices as the external stimulus is identical both time.
-Conversely, internal noise should lead to inconsistent choices both times.
-SPELL OUT MORE
+Conversely, internal noise should lead to independent choices, and hence possible inconsistent choices both times.
+More specifically, we look at the proportion of times that a participant make inconsistent decisions in repeated game.
 
 	%used choice inconsistency in the repeated games as a measure of internal noise.  
-
-	%GOT TO INTRODUCE DIRECTED EXPLORATION - MAYBE BEFORE THIS
-
-	%MOVE THIS PARAGRAPH UP As in our previous paper \cite{}, the influence of directed exploration is measured as the probability of choosing the more informative option, i.e.
-the bandit with only 1 outcome in [1 3] condition; decision noise is measured as the probability of choosing the low mean option in [2 2] condition.
-
 	In addition, whether participants make consistent choices in repeated games is used as a model-free measure of internal noise, since in repeated trials, external noise should be identical on both trials.
@@
 The degree to which people make consistent choices in repeated trials can reflect the internal noise.
+Since choice inconsistency increases with horizon, this is a behavioral evidence that internal noise increases with horizon.
 
@@
 			\includegraphics[width=\textwidth]{figures/Modelfree.png}
-			\caption{Both direct (P $<$ 0.001) and random exploration (P $<$ 0.001 for both [1 3] and [2 2]) increases with horizon.(A,B) Choice consistency in both [1 3] and [2 2] condition increases with horizon(P([1 3]) $<$ 0.001, P([2 2])$<$0.001).(C)}
+			\caption{Both direct (t(406)=-6.33, p $<$ 0.001) and random exploration (t(406) = -5.30, p $<$ 0.001 for [1 3], t(406) = -7.71, p $<$ 0.001 for [2 2]) increases with horizon.(A,B) Choice consistency in both [1 3] and [2 2] condition increases with horizon(P([1 3]) $<$ 0.001, P([2 2])$<$0.001).(C)}
 			\label{fig:mf}
@@
 
-	%In line with previous findings, both directed and random exploration increase with horizon (Figure \ref{fig:mf}). In addition, 
 	Choice inconsistency between repeated games was non-zero in both horizon conditions, suggesting that not all of the noise was stimulus driven.
@@
 	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the information bias between the two bandits.
-To model participants’ choices on this first free-choice trial, we assume that they make decisions by computing the difference of value $\Delta Q$ between the two options:
+To model participants’ choices on this first free-choice trial, we assume that they make decisions by computing the difference of value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
+Specifically, we write
 
-	$$\Delta Q= \Delta R+A \Delta I+b+n_{ext}+n_{int}$$
+	\begin{equation}
+	    \Delta Q= \Delta R+A \Delta    I+b+n_{ext}+n_{int}
+	\end{equation}
 
-	Where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial (for simplicity we define $\Delta I$ to be +1, -1 or 0: +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0).
+	Where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
+For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0: +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0).
 $n_{ext}$ and $n_{int}$ are external noise and internal noise respectively.
 
-	The subject-and-condition-specific parameters are: the spatial bias, b, which determines the extent to which participants prefer the option on the right, the information bonus A, which controls the level of directed exploration, ϵ denotes the external, external noise, which is identical on the repeat versions of each game, and n denotes internal noise, which is uncorrelated between repeat plays.
-For repeated games, the set of forced-choice trials are exact the same, so the external noise should be the same, but the free decision noise may be different.
+	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right, the information bonus $A$, which controls the level of directed exploration, $n_{ext}$ denotes the external, external noise, which is identical on the repeat versions of each game, and $n_{int}$ denotes internal noise, which is uncorrelated between repeat plays.
+
+	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the external noise, $n_{ext}$, should be the same while the internal noise, $n_{int}$ may be different.
 This is exactly how we distinguish external noise from internal noise.
-In symbolic terms, for repeated games i and j,  $\sigma_{ext}^i=\sigma_{ext}^j$  and $\sigma_{int}^i \neq \sigma_{int}^j$.
-	The decision model assumes that participants choose right if $\Delta Q > 0$, otherwise the participant will choose the left option.
+In symbolic terms, for repeated games i and j,  $n_{ext}^i=n_{ext}^j$  and $n_{int}^i \neq n_{int}^j$.
 
-	We used hierarchical Bayesian analysis to fit the parameters (see Figure \ref{fig:model} for an graphical representation of the model in the style of [5]).
-In particular, we fit values of the information bonus A, spatial bias B, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon and uncertainty condition.
-The mean and standard deviation of information bonus A and spatial bias B are sampled from a Gaussian prior and an exponential prior respectively.
-The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter k and λ for the gamma distribution are sampled from exponential priors.
-At the game level, we sampled one stimulus driven noise for each pair of repeated games and we sampled two free decision noises for each individual game in the pair.
-The model in Figure \ref{fig:model} is fitted using the MATJAGS and JAGS \cite{}.
+\paragraph{Model fitting}
+	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of CITE).
+In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of internal noise $\sigma_{int}^2$, and variance of external noise, $\sigma_{ext}^2$ for each participant in each horizon and uncertainty condition.
+The mean and standard deviation of information bonus $A$ and spatial bias $B$ are sampled from a Gaussian prior and an exponential prior respectively.
+The variance for both type of noises were sampled from a gamma distribution, and the group-level parameter $k$ and $\lambda$ for the gamma distribution are sampled from exponential priors.
+
+	To capture the idea that external noise should be identical on repeated games, we sampled one value of the external noise, $n_{ext}$ for each pair of repeated games.
+Conversely, because internal noise is expected to change between games we sampled two values of the internal noise, $n_{int}^1$ and $n_{int}^2$, i.e.
+one for each individual game.
+
+	The model in Figure \ref{fig:model} is fitted using the MATJAGS  and JAGS software \cite{}.
 
@@
 	\paragraph{Model fitting results}
-	%Posterior distributions over the group-level means are shown in the left column of Figure \ref{fig:mb1}, while posteriors over the TMS-related change in parameters are shown in the right column.
-Both columns suggest a selective effect of RFPC stimulation on the information bonus in horizon 6.
 
-	Only internal noise increases  with horizon (Figure \ref{fig:modelbased}). 
+	Group-level estimates for the variances of internal and external noise are shown in Figure \ref{fig:modelbased}A.
+While both variances are non-zero, this shows that the internal noise is much larger than external noise.
+This horizon-based change is probed further in Figure XXXB in which we plot the posterior distributions over the change in internal and external noise with horizon.
+This clearly shows that only internal noise varies with horizon (STATS XXX\% of the samples above zero for internal noise, XXX\% of samples above zero for external noise).
+
+	%LEFT EACH TYPE OF NOISE IN EACH HORIZON
+	%RIGHT IS DISTRIBUTION OVER CHANGE IN NOISE WITH HORIZON
+
+	%Posterior distributions over the group-level means of the external and internal noise variance are shown in Figure \ref{fig:modelbased}.
+As is clear from this plot, only internal noise increases  with horizon.
 	%In both horizon 1 and horizon 6, internal noises are significantly higher than internal noises (for both horizon 1 and horizon 6, P $<$ 0.001).
@@
 			\includegraphics[width=0.5\textwidth]{figures/bayes.png}
-			\caption{The change of external and internal decision noises between horizons. The errorbars are confidence intervals at $95\%$. Only internal noise increases significantly with horizon.}
+			\caption{Both internal and external noises increase significantly with horizon.}
 			\label{fig:modelbased}
@@
 
-	Figure 3 also demonstrates the existence of both external noise and internal noise.
-Hierarchical Bayesian analysis provides us a way to quantify the magnitude of external and internal noise respectively, from the result, internal noise is the main source of random exploration.
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.5\textwidth]{figures/bayesdist.png}
+			\caption{The posterior distribution of internal and external decision noise.}
+			\label{fig:modelbased}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.5\textwidth]{figures/s1.png}
+			\caption{Both internal and external noise contribute to random exploration, internal noise is dominant.}
+			\label{fig:modelbased}
+		\end{center}
+	\end{figure}
 
@@
 Thus, our estimate of external noise is likely a lower bound.
-Likewise, our estimate of internal noise is likely an upper bound as these “missing” sources of stimulus driven noise would be interpreted as internal noise in our model.
-Despite this, the sheer magnitude of the difference between internal and external noise (internal noise is 2-3 times the size of external noise, Figure 3), suggests that our interpretation may be safe as an awful lot of external noise would have to be explained by variables not relevant to the task.
+Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of stimulus driven noise would be interpreted as internal noise in our model.
+Despite this, the sheer magnitude of the difference between internal and external noise (internal noise is 2-3 times the size of external noise, Figure XXX), suggests that our interpretation may be safe as an awful lot of external noise would have to be explained by variables not relevant to the task.
+
 	The horizon-dependent increase in internal noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
 This is in line with work in the bird song literature in which song variability during song learning has been tied to neural variability arising from specific areas of the brain.
-In addition, this work is consistent with a recent report from Ebitz et al [7] in which behavioral variability in an “explore” state was also tied to internal rather than external sources of noise.
+In addition, this work is consistent with a recent report from Ebitz et al [7] in which the behavioral variability of monkeys in an `explore' state was also tied to internal rather than external sources of noise.
+
 	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory [4] suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
@@
 
+	MORE GENERALLY, OUT FINDING THAT INTERNAL DOMINATES BEHAVIORAL VARIABILITY OVER EXTERNAL NOISE, IS CONSISTENT WITH FINDINGS OF DRUGOWITSCH ET AL.  IN PARTICULAR THESE AUTHORS SHOW THAT 
+	LINK IN THE Jan Drugowitsch stuff on variability.  ALEX POUGET STUFF. INTERNAL NOISE COMES FROM COMPUTATIONAL ERRORS IN COMPUTING THE CORRECT STRATEGY IN LONG  HORIZON CONDITIONS.
+
+	THREE IDEAS
+	    NOISE ADDED TO THE PROCESS
+	    ATTENTION = DECREASE SIGNAL VS INCREASE NOISE
+	    NOT NOISY BUT WRONG
+	    PLANNING WITH FEW SAMPLES DEEP EXPLORATION
+
 	\section*{Methods}
 	\subsection*{Participants}
-	78 Participants (37 male, 41 female, ages 17-25) from the UA subject pool, played the task.
-27 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 51 for the analysis.
+	A total of 204 participants from the UA subject pool did the task in 4 different experiments. \\
+	 017\\ 
+	 49 out of 148 participants are included for analysis.
+59 were excluded because of the passive condition.
+59 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 49 for the analysis.\\
+	075\\ 
+	 58 out of 78 participants are included for analysis.
+3 were excluded because of young age, 17 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 49 for the analysis.\\
+	personality \\
+	 65 out of 91 participants are included for analysis. 26 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 65 for the analysis.\\
+     test-retest \\
+	 33 out of 45 participants are included for analysis. 12 were excluded on the basis of performance (using the same exclusion criterion as in [2]) leaving 33 for the analysis.\\
+
+	 There are no differences of p(low mean) or p(inconsistent) across studies. No differences were found between ages, races and ethnicities.
+
 	%31 healthy right-handed, adult volunteers (19 female, 12 male; ages 19-32).

### 2022-12-12 codes/Figure_gridsimu_all.m

@@ -1,10 +1,9 @@
 function Figure_gridsimu_all(plt, rdsp)
 %% all levels
 W.library('tools_violin');
-plt.figure(11,2);
+plt.figure(4,6);
 for i = 1:2
     for j = 1:11
-        plt.ax(j,i);
         plt.setfig_ax('ylim', [j-3 j+1], 'ytick', j-1,...
             'xtick',[1:11], 'xticklabel', 0:10);
         if j == 6
@@ -27,7 +26,9 @@ for i = 1:2
         hold on;
         plt.setfig_ax('color', 'AZsand');
         plt.lineplot(ones(1,13)* (j-1), [],0:12,'-');
+        plt.new;
     end
+    plt.new;
 end
 plt.update('parameterrecovery_gridsimu_all');
 end

### 2022-12-15 manuscript/2022 draft - Plos Comp Bio/Supplementary/DetRanNoise_Supplementary.tex
u

@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree_all.png}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree__all.jpg}
 			\caption{Replication of previous findings with data from all participants (i.e.
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_2noise_all.png}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_pinconsistent__all.jpg}
 			\caption{Model-free analysis with data from all participants (i.e.
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_hyperprior_all.png}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_hyperprior__all.jpg}
 			\caption{Model based analysis with data from all participants (i.e.
@@
 			\label{fig:mb12}
+		\end{center}
+	\end{figure}
+
+	\newpage
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet_R20.jpg}
+			\caption{}
+		\end{center}
+	\end{figure}
+
+	\newpage
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_gridsimu_all.jpg}
+			\caption{}
+		\end{center}
+	\end{figure}
+
+	\newpage
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_gridsimu_all_R20.jpg}
+			\caption{}
 		\end{center}

### 2022-12-15 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
u

@@
 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved driven by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p > 0.001$ were excluded in the analysis.
+In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved driven by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
 This left 66 for the main analysis.
@@
 (B) Different possible states of the game after the first free choice over the course of the experiment.
-Overall participants play 160 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
+On average, participants play more than 150 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
 In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.
@@
 Thus, the better participants performed, the sooner they got to leave the experiment.
-On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).
-Participants played an average of 67.31 repeated pairs of games (minimum = 25 repeated pairs, maximum =  100 repeated pairs).
+On average, participants played 153.7 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.37 and 32.15 minutes (mean 22.78 minutes).
+Participants played an average of 65.3 repeated pairs of games (minimum = 30 repeated pairs, maximum =  79 repeated pairs).
 	%BOB SAYS: ADD HOW MANY REPEATED GAMES THEY PLAYED - GIVE A MEAN AND RANGE FOR THIS NUMBER
@@
 	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of the rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
+	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of the rewards shown on the forced trials, and $\Delta I$, the difference of information available for playing the two options on the first free-choice trial.
 For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.
 
-	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.
+	The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.
+$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.
 
-	Each subject's behavior in each horizon condition is described by 4 free parameters: the information bonus $A$, the spatial bias, $b$, the standard deviation of the deterministic noise, $\sigma_{det}$, and the standard deviation of the random noise, $\sigma_{ran}$ (Table \ref{tab:pars2}, Figure \ref{fig:model}).
+	Each subject's behavior in each horizon condition is described by 4 free parameters (Table \ref{tab:pars2}): the information bonus $A$, the spatial bias, $b$, the standard deviation of the deterministic noise, $\sigma_{det}$, and the standard deviation of the random noise, $\sigma_{ran}$.
 Each of the free parameters is fit to the behavior of each subject using a hierarchical Bayesian approach \citep{hbm1}.
-In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure.
+In this approach to model fitting, each parameter for each subject is assumed to be sampled from a group-level prior distribution whose parameters, the so-called `hyperparameters', are estimated using a Markov Chain Monte Carlo (MCMC) sampling procedure (Figure \ref{fig:model}).
 The hyper-parameters themselves are assumed to be sampled from `hyperprior' distributions whose parameters are defined such that these hyperpriors are broad.
@@
 			deviation of deterministic noise, $\sigma^{det}_{isg}$ 
-			& $\sigma_{i} \sim $  Gamma($k_i^{det}$, $\lambda_{i}^{det}$) 
+			& $\sigma^{det}_{is} \sim $  Gamma($k_i^{det}$, $\lambda_{i}^{det}$) 
 			& $\theta_{i}^{det} = (k_i^{det}, \lambda_i^{det}) $
@@
 			deviation of random noise, $\sigma^{ran}_{isgr}$ 
-			& $\sigma_{is} \sim $  Gamma($k_i^{ran}$, $\lambda_{i}^{ran}$) 
+			& $\sigma^{ran}_{is} \sim $  Gamma($k_i^{ran}$, $\lambda_{i}^{ran}$) 
 			& $\theta_{i}^{ran} = (k_i^{ran}, \lambda_i^{ran}) $
@@
 
-	In particular we used 4 independent Markov chains to generate 16000 samples from the posterior distribution over parameters (4000 samples per chain).
-Each chain had a burn in period of 2000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
+	In particular we used 10 independent Markov chains to generate 50000 samples from the posterior distribution over parameters (5000 samples per chain).
+Each chain had a burn in period of 5000 samples, which were discarded to reduce the effects of initial conditions, and posterior samples were acquired at a thin rate of 1.
 Convergence of the Markov chains was confirmed {\it post hoc} by eye.
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
+			\includegraphics[width=1\textwidth]{figures/EEHorizon_2sigma.pdf}
     			\caption{Schematic of the hierarchical Bayesian model using notation of \cite{lee_wagenmakers_2014}}
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_pure_ran_det_simu_hyperprior.png}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet.jpg}
 			\caption{Parameter recovery over the posterior of random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for purely random noise (left column) and purely deterministic noise (right column) games.
@@
 	\subsection*{Data and code}
-	Behavioral data as well as Matlab code to recreate the main figures from this paper will be made available on the Dataverse website upon publication.
+	Behavioral data as well as MATLAB codes to recreate the main figures from this paper will be made available upon publication.
 %website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
 			\caption{
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_pinconsistent.jpg}
 			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration.
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_hyperprior.jpg}
 			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.

### 2022-12-15 codes/mainFigures_paramrecover_hyperprior.m

@@ -1,0 +1,74 @@
+outputdir = '../bayesoutput/all/';
+sp = load(fullfile(outputdir, ['HBI_DetRanNoise_samples.mat'])).samples;
+%% compute confidence interval
+ci = struct;
+ci.dNoiseRan = quantile(reshape(sp.dNoiseRan,1,[]), [0.025, 0.975]);
+ci.dNoiseDet = quantile(reshape(sp.dNoiseDet,1,[]), [0.025, 0.975]);
+for i = 1:2
+    ci.NoiseRan(i,:) = quantile(reshape(sp.NoiseRan(:,:,i),1,[]), [0.025, 0.975]);
+    ci.NoiseDet(i,:) = quantile(reshape(sp.NoiseDet(:,:,i),1,[]), [0.025, 0.975]);
+end
+%% compute coverage
+simudir = '../bayesoutput/simurepeat';
+coverage = struct;
+recovered = struct;
+for repi = 1:200
+    tstat = load(fullfile(simudir, sprintf('HBI_DetRanNoise_simu_%d_stat.mat', repi)));
+    recovered.dNoiseRan(repi) = tstat.stats.mean.dNoiseRan;
+    recovered.dNoiseDet(repi) = tstat.stats.mean.dNoiseDet;
+    coverage.dNoiseRan(repi) = W.is_between(tstat.stats.mean.dNoiseRan, ci.dNoiseRan);
+    coverage.dNoiseDet(repi) = W.is_between(tstat.stats.mean.dNoiseDet, ci.dNoiseDet);
+    for i = 1:2
+        coverage.NoiseRan(repi, i) = W.is_between(tstat.stats.mean.NoiseRan(i), ci.NoiseRan(i,:));
+        coverage.NoiseDet(repi, i) = W.is_between(tstat.stats.mean.NoiseDet(i), ci.NoiseDet(i,:));
+    end
+    recovered.NoiseRan(repi,:) = tstat.stats.mean.NoiseRan;
+    recovered.NoiseDet(repi,:) = tstat.stats.mean.NoiseDet;
+end
+%% hyper prior recovery
+plt = W_plt('savedir', '../figures', 'savepfx', 'RDBayes', 'isshow', true, ...
+    'issave', true);
+plt.figure(2,3, 'is_title', 1);
+names = ["NoiseRan", "NoiseDet"];
+xbins = -10:0.02:50;
+color = {{'AZred','AZred'},{'AZblue','AZblue'}};
[…]

### 2022-12-15 codes/main_1_behavior.m

@@ -4,6 +4,7 @@ savedir = '../data/all';
 files = dir(fullfile(rawdatadir,'Imported_EE_blinkCuriosity_2016S.csv'));
 filename = fullfile(files.folder, files.name);
 %% basic analysis
+W.library_wang('Wang_EEHorizon')
 opt_sub = {'subjectID', 'date', 'time'};
 opt_preprocess = 'EEpreprocess_game_basic';
 opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};

### 2022-12-19 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
draft to bob

@@
 \renewcommand{\baselinestretch}{1.5}
+\usepackage{float}
 
+% control the fontsize
+\PassOptionsToPackage{textsize=scriptsize}{todonotes}
 %\usepackage[final]{changes}
@@
 
-\title{The nature of decision noise in random exploration}
+\usepackage{xpatch}
 
+%\usepackage{unravel}
+%\unravelsetup{max-action=1000, max-input=1000, max-output=1000}
+\providecommand\unravel[1]{#1}
+
+% auto insert "comment"
+\makeatletter
+\xpatchcmd\replaced
+{\setkeys{Changes@replaced}{#1}}
+{\setkeys{Changes@replaced}{comment={(was): \protect\truncateto{10}{#3}}, #1}}
+{}{}
+\xpatchcmd\deleted
+{\setkeys{Changes@deleted}{#1}}
+{\setkeys{Changes@deleted}{comment={(deleted): \protect\truncateto{10}{#2}}, #1}}
+{}{}
+
+% fix \listofchanges
+\xpatchcmd\ChangesListline
+{0px}
+{0pt}
+{}{}
+
+\newcounter{truncate}
+\newcounter{truncate@max}
+
+% truncate #2 to leave at most first #1 words
+\def\truncateto#1#2{%
+	\setcounter{truncate}{0}%
+	\setcounter{truncate@max}{#1}%
+	\truncate@loop{#2}%
+}
+
+\def\truncate@loop#1{%
+	\ifnum\c@truncate=\c@truncate@max
+	\def\next{[...]}% tail
+	\else
+	\in@{ }{#1}%
+	\ifin@
+	\stepcounter{truncate}%
+	\truncate@split#1\@nil
+	\else
+	\def\next{#1}%
+	\fi
+	\fi
+	\next
+}
+
+\def\truncate@split#1 #2\@nil{%
+	\def\next{#1 \truncate@loop{#2}}%
+}
+\makeatother
+
+\makeatletter
+\setdeletedmarkup{\@gobble{#1}}
+\makeatother
+
+\title{\replaced[id=siyu]{Separating random and deterministic sources of computational noises in explore-exploit decisions}{The nature of decision noise in random exploration}}
 \author[1,\textcurrency]{Siyu Wang}
@@
 \affil[3]{Cognitive Science Program, University of Arizona, Tucson AZ, USA}
-\affil[ \textcurrency]{\added[id=siyu]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}}
+\affil[ \textcurrency]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}
 
@@
 	\begin{abstract}
-
 Human decision making is inherently variable.
@@
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-In particular, we find evidence that at least part \deleted[id=siyu]{27$\%$} of the variability in `random' exploration can be accounted for by deterministic processing of the stimulus.
-This still leaves open the possibility that much of random exploration is truly `random,' but narrows the window of opportunity for stochastic processes to explain this behavior.
+\added[id=siyu]{By a novel bayesian analysis, we are able to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.} In particular, we find evidence that at least 20$\%$ of the variability in random exploration can \added[id=siyu]{indeed} be accounted for by deterministic processing of the stimulus.
+This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
+\added[id=siyu]{Our work sheds light on how random exploration may be implemented by reducing the filtering of task-irrelevant stimuli in the decision process.}
+
 %suggests that at least some `random' exploration is not random, but leaves open the possibility that the 
@@
 	%BOB SAYS: IN GENERAL, NEED TO GO THROUGH THE PAPER AND MAKE SURE THAT `RANDOM' NOISE IS REPLACES WITH `NOT A DETERMINISTIC FUNCTION OF THE STIMULUS' OR SOMETHING LIKE THAT
+	%SIYU Says: We can define "random noise" as "not a deterministic funciton of the stimulus" at the beginning, and use the term "random noise" in the results section.
+It feels a little lengthy to use "not a deterministic function of the stimulus" in the results section each time.
+I changed a few of them back to "random noise" in the results section, in places that I think random noise is better.
+See if you like the changes.
 
 	%(BOB SAYS: TAKE OR LEAVE THIS CHANGE TO DATE, BUT SETTING UP AS A DATE ALLOWS YOU TO TALK ABOUT AVOID YOUR EX LATER ON, WHICH MAY BE A BETTER EXAMPLE OF A DETERMINISTIC PROCESS!).
+	% Siyu Say: I kept this.
+
 	Imagine trying to decide where to go to dinner on a date, you can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try a new restaurant that you know nothing about.
@@
 choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}.
-In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.
+In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \added[id=siyu]{\citep{Mehlhorn15, SCHULZ20197,WILSON202149}}.
+%\citep{eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}
 
-	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}.
-In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise,' meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
-
-	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.
-In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.
+	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90} \added[id=siyu]{, also known as random exploration}.
+In this strategy, \deleted[id=siyu]{the decision process between exploration and exploitation is corrupted by `decision noise', meaning that }high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
+\added[id=siyu]{In modeling terms, random exploration works by adding 'decision noise' to the value of the options such that sub-optimal exploratory options can sometimes have a higher total score (i.e., value + noise) than the exploit option and get chosen.} \replaced[id=siyu]{Such}{In theory, such} random exploration, is surprisingly effective and, if implemented correctly, can come close to optimal performance \added[id=siyu]{in theory}\citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
 
@@
 For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant.
-Using such a horizon manipulation we found that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.
+Using such a horizon manipulation it has been shown that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.
 
-	One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random.
-That is, whether behavioral variability is due to intrinsic stochastic processes in the brain or whether it is due to deterministic processes that we simply have not observed.
-For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
+	\replaced[id=siyu]{One limitation, however, is that it is difficult to tell whether what we have called 'decision noise' in previous research is truly a random noise.}{One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random.} That is, whether behavioral variability is due to intrinsic stochastic processes in the brain or whether it is due to deterministic processes that we \replaced[id=siyu]{failed to observe}{simply have not observed}.
+\added[id=siyu]{Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model.
+A missing deterministic component from the model could give rise to variability in behavior that might be mistakenly taken as a random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
 Avoiding an ex is a deterministic process, but if we fail to take ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
 
-	%This difficulty arises in how noise is defined in these previous tasks. For example, in the Horizon Task, variability is defined as the inverse temperature parameter of a softmax choice function.
-
-	%In particular, in most cases noise is defined by fitting a softmax decision rule to behavior.
-
 	%because noise in previous work is defined as the component of the decision process that cannot be accounted for by a model. 
-	%, what we have called 'noise' in previous research could actually just be some missing deterministic components from the model.
-Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model.
+	%, 
 	%For example, in the restaurant case above, if you happen to spot an old friend walking in to one of the restaurants, you may be more likely to follow them in.
@@
 
-	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by such deterministic processing of the stimulus.
-In particular, we modify the Horizon Task to have people face the exact same explore-exploit choice twice.
-If the decision is a purely deterministic function of the stimulus, then people's choices should be identical both times.
+	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by such \added[id=siyu]{unobserved} deterministic processing of the stimulus \added[id=siyu]{(will refer to as deterministic noise)}.
+\replaced[id=siyu]{To distinguish between stimulus-driven `deterministic noise' and non-stimulus-driven `random noise'}{In particular}, we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
+If the decision is a purely deterministic function of the stimulus \added[id=siyu]{(i.e., decision noise is purely deterministic noise)}, then people's choices should be identical both times.
 That is, their choices should be consistent, since the stimulus is the same both times.
-Conversely, the more their decision is driven by other processes, including both stochastic and unobserved deterministic processes, the less consistent their behavior should be.
+Conversely, the more their decision is driven by \replaced[id=siyu]{random noise}{other processes, including both stochastic and unobserved deterministic processes}, the less consistent their behavior should be.
 By analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
-This does not prove that random exploration is entirely deterministic or stochastic, but provides a lower bound on how much deterministic processes contribute to random exploration.
-This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.
+\replaced[id=siyu]{Our work}{This does not prove that random exploration is entirely deterministic or stochastic, but} provides a lower bound on how much deterministic processes contribute to random exploration\added[id=siyu]{, and an upper bound on how much true random noises contribute to random exploration}.
+\deleted[id=siyu]{This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.}
+
+	% SIYU says: "Conversely, the more their decision is driven by other processes, including both stochastic and unobserved deterministic processes, the less consistent their behavior should be.
+" is not correct.
+Unobserved deterministic processes will be the same for repeated games and will not contribute to choice inconsistency.
 
@@
 
+	\section*{Results}
+
+	\subsection*{The Repeated-Games Horizon Task}
+	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus-driven `random noise' \replaced[id=siyu]{in explore-exploit}{on people's} decisions (Figure \ref{fig:taskfig}).
+In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
+Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
+			\caption{ 
+				Schematic of the experiment.
+(A) Dynamics of an example horizon 6 game.
+Here the first four trials are forced trials in which participants are instructed which option to play.
+After the forced trials, participants are free to choose between the two options for the remainder of the game.
+(B) Example repeated games over the course of the experiment.
+				% Different possible states of the game after the first free choice over the course of the experiment. 
+				On average, participants play more than 150 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
+In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.
+These repeated games allow us to compute the relative contribution of deterministic and random noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}
+			\label{fig:taskfig}
+		\end{center}
+	\end{figure}
+
+	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
+1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
+These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
+
+	Relative to the original Horizon Task, the key modification here is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games (separated by several minutes in time so as to avoid detection).
+By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
+Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their \replaced[id=siyu]{choice}{behavior} on the two versions of each game will be consistent.
+Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games \replaced[id=siyu]{could}{should} be inconsistent.
+\replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which their behavior was \replaced[id=siyu]{driven by a deterministic noise process vs a random noise process}{a deterministic function of the stimulus or not}.
+
+	%, the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.
+
+	%These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.
+Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent.
+Conversely if noise is randomly driven, then choices on repeated games should be inconsistent.
+
+	\subsection*{Both behavioral variability and information seeking increase with horizon}
+
+	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}.
+As in our previous work, we find evidence for two types of exploration in the Horizon Task: Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information (Figure \ref{fig:modelfree}, Supplementary Figure S1).
+
+	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
+This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(64) = 6.55, p $<$ 0.001 for [1 3], t(64) = 7.99, p $<$ 0.001 for [2 2]).
+Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info})$ in the unequal, or [1 3], condition.
+Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(64) = 6.92, p $<$ 0.001).
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
+			\caption{
+				Replication of previous findings.
+Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.
+}
+			\label{fig:modelfree}
+		\end{center}
+	\end{figure}
+
+	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}
+
+	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.
+The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times.
+\deleted[id=siyu]{Conversely, if choice is not a deterministic function of the stimulus, participants' choices should be independent, and hence more inconsistent across the repetitions of the game.}\added[id=siyu]{Conversely, if choice is not entirely driven by a deterministic process and is also driven by random noise, participants' choices should be more inconsistent across the repetitions of the game.
+Moreover, if decision noise is purely random noise, meaning there is no unobserved deterministic process, we will show that we can actually predict the expected level of choice inconsistencies across repetitions of games by accounting for the known deterministic processes and assuming that the random noise process is independent in repetitions of the game.}
+
+	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}, Supplementary Figure S2).
+Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was stimulus driven (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions\deleted[id=siyu]{For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6}).
+In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (For [1 3] condition, t(64) = 5.41, p $<$ 0.001; for [2 2] condition, t(64) = 6.26, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is not a deterministic function of the stimulus\added[id=siyu]{, but rather random noise}.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_pinconsistent.jpg}
+			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration.
+For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
+However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
+			\label{fig:mf2}
+		\end{center}
+	\end{figure}
+
+	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely deterministic and purely random noise cases.
+For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case.
+For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
+	\begin{equation*}
+		\begin{split}
+			p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
+			&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
+			\mbox{hence},\quad p(\mbox{inconsistent}) &=  
+			1 - p(\mbox{consistent}) = 
+			2 p(\mbox{low mean})(1-p(\mbox{low mean}))
+		\end{split}
+	\end{equation*}
+	\added[id=siyu]{Furthermore, to account for that $p(\mbox{low mean})$ is a function of reward difference $\Delta R$ between the two bandits and the information condition $I$, we estimated the conditional probability:
+		$$p(\mbox{inconsistent}|\Delta R, I) = 2 p(\mbox{low mean}|\Delta R, I)(1-p(\mbox{low mean}|\Delta R, I))$$
+	Then based on the likelihood that each condition ($\Delta R$ vs $I$) occurs in the task $\rho(\Delta R, I)$, we have
+		$$p(\mbox{inconsistent}) = \sum_{\Delta R, I}\rho(\Delta R, I)p(\mbox{inconsistent}|\Delta R, I)$$}
+
+	As shown in Figure \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction.
+Specifically, behavior is different from pure random noise prediction in the both the [1 3] condition (t(64) = 4.83, p $<$ 0.001 for horizon 1, t(64) = 3.12 p $=$ 0.003 for horizon 6) and the [2 2] condition (t(64) = 3.92, p $<$ 0.001 for horizon 1, t(64) = 3.71, p $<$ 0.001 for horizon 6).
+Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition (t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6).
+\added[id=siyu]{As a negative control of our method for estimating $p(\mbox{inconsistent})$ for purely random noise, we simulated choices using a decision model that only includes random noise (for details, see the results section on model-based analysis, and methods), and found that $p(\mbox{inconsistent})$ in this simulated data is not different from our pure random noise prediction in all horizon and uncertainty conditions (p $>$ 0.05, Supplementary Figure S3).} \replaced[id=siyu]{Together, our results suggest that both random noise and deterministic noise contribute to the choice variability in random exploration.}{This suggests that at least some of the `noise' is a deterministic function of the stimulus, } Although from this analysis it is not clear whether this deterministic noise increases with horizon or not.
+
+	\subsection*{Model-based analysis \replaced[id=siyu]{provides a lower-bound estimate of deterministic noise and an upper-bound estimate of random noise}{shows deterministic noise changes with horizon}}
+
+	To more precisely quantify the contribution of deterministic noise \added[id=siyu]{and random noise}, we turned to model fitting.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate \replaced[id=siyu]{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise')}{deterministic noise from random noise}.
+In particular, we assume that in repeated games, the value of stimulus-driven deterministic noise is frozen whereas random noise is drawn independently both times.
+
+	\subsubsection*{Overview of model}
+	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
+To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
+Specifically, we write
+	\begin{equation}
+		\Delta Q= \Delta R+A \Delta I+b+n_{det}+n_{ran}
+	\end{equation}
+	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference \added[id=siyu]{of} information available for playing the two options on the first free-choice trial.
+For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
+\added[id=siyu]{$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeated plays and changes every game.} $n_{det}$ and $n_{ran}$ are assumed to come from logistic distributions with mean 0\added[id=siyu]{, and standard deviations $\sigma_{det}$ and $\sigma_{ran}$}.
+The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; \added[id=siyu]{$\sigma_{det}$ and $\sigma_{ran}$, which control the respective level of deterministic noise and random noise in random exploration.}
+
+	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the deterministic noise, $n_{det}$, should be the same while the random noise, $n_{ran}$ may be different.
+This is exactly how we distinguish deterministic noise from random noise.
+In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.
+
+	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
+In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon.
+Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
+
+	\subsubsection*{\replaced[id=siyu]{Model validation}{Parameter recovery}\label{ch:appendix:bayesrecovery}}
+	To be sure that our fit parameter values were meaningful \added[id=siyu]{and to understand the limits of our model, we evaluated our model in several ways.
+Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
+In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition on explore-exploit decisions.
+$\Delta Q, \Delta R, n_{det}, n_{ran}$ below represent the same variables as in the full model.
+	$$\Delta Q= \Delta R+n_{det}+n_{ran}$$
+	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
+By comparing the	fitted posterior distributions over the group-level means of the deterministic and random noise parameters $\sigma_{det}$ and $\sigma_{ran}$, as expected, we observed an increase in deterministic noise and no change in random noise between the reduced and the full model (Figure \ref{fig:reducedmodel}).
+	}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_reduced_model.jpg}
+			\caption{Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
+In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
+Random noise remains unchanged between the two models.}
+			\label{fig:reducedmodel}
+		\end{center}
+	\end{figure}
+
+	\added[id=siyu]{Secondly, we evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
+In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
+The simulation and re-fitting was repeated for 200 times.
+Then we counted out of the 200 repetitions, whether the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval for more than 95\% of the times (this ratio will be referred to as the coverage rate).
+For random noise, the coverage rate is 100\% for both horizon 1, horizon 6 and noise differences between horizons.
+For deterministic noise, the coverage rate is 66\% for horizon 1, 69\% for horizon 6, and 97\% for noise differences between horizons.
+By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of re-fitted parameters to the simulated data, it is clear that our model systematically underestimates deterministic noise.
+For random noise, our model fitting procedure yields a faithful recovery.
+However, because our random noise is modeled as non-stimulus-driven noise which includes both truly stochastic neural noise and possible deterministic noises irrelevant to the stimuli, our random noise estimate provides an upper bound of `random noise' induced by intrinsic stochastic processes in the brain.
+Because the underestimation of deterministic noise is partially canceled out when taking the difference, our model fitting procedure passed the coverage test in estimating the horizon differences in both random and deterministic noise.}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior_v2.jpg}
+			\caption{Parameter recovery over the posterior distribution of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices.
+Lighter color shades represent the re-fitted posterior to the simulated choices.
+Our model fitting procedure faithfully recovers the non-stimulus-driven random noise (A, B), but systematically underestimates deterministic noise in both horizons (D, E).
+The horizon differences in random noise is also faithfully recovered (C).
+The horizon differences in deterministic noise is also underestimated but not significant (F).}
+			\label{fig:coverage2}
+		\end{center}
+	\end{figure}
+
+	\added[id=siyu]{Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Supplementary Figure S5 and S6).
+The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).
+Despite the strong correlation between simulated and fit values of deterministic noise (R $>$ 0.8), we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
+
+	%Parameter recovery is good (in terms of) for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C).
+The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
+
+	\added[id=siyu]{Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
+All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ are simulated.
+In a special case, we evaluated how our model performs when there is only random noise or only deterministic noise (Figure \ref{fig:puredetran}).
+In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with full random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
+However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Supplementary Figures S7).
+Overall, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
+
+	%This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
+In addition we see better recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet.jpg}
+			\caption{Parameter recovery over the posterior of random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for purely random noise (top row) and purely deterministic noise (bottom row) games.
+			}
+			\label{fig:puredetran}
+		\end{center}
+	\end{figure} 
+	% Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
+
+	\added[id=siyu]{Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
+We see better parameter recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+	}
+	\subsubsection*{Model-based results}
+
+	Posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ are shown in Figure \ref{fig:mb1} and Supplementary Figure S8.
+Consistent with our model-free results, we see that both random and deterministic noise are non-zero.
+\replaced[id=siyu]{Numerically,}{and that} random noise is about 2-3 times larger than the deterministic noise.
+\added[id=siyu]{By computing the 5\% quantile of the posterior distribution of $\sigma_{det}/(\sigma_{det}+\sigma_{ran})$, our data suggests that at least 20.05\% of the variability in random exploration can be accounted for by deterministic noise.} In addition, we find that both random and deterministic noise increase with horizon.
+This increase was larger for random noise (M = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (M = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon).
+But intriguingly the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}).
+That is when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
+	%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
+			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
+			%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
+			\label{fig:mb1}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.5\textwidth]{figures/RDBayes_ratio__paironly.jpg}
+			\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
+The ratio in the standard deviations of noise between horizon 6 and horizon 1 is similar for random and deterministic noise.}
+			\label{fig:ratio}
+		\end{center}
+	\end{figure}
+
+\subsubsection*{Posterior predictive checks}
+In addition to fitting the model to behavior, it is also important to check whether the model captures the qualitative patterns of the data \citep{Wilson2019} --- specifically how p(high info), p(low mean) and p(inconsistent) change with horizon.
+
+To perform this `posterior predictive check,' we created a set of simulated data by taking the subject-level parameters from the hierarchical Bayesian fits and having the model play the same sequence of games as seen by the subjects.
+We then applied the same model-free analysis as described in the previous sections to this simulated data set and compared the model's behavior to that of participants.
+As shown in Figure  \ref{fig:mb3}, the model can account for all qualitative patterns in the data --- the increase in p(high info), p(low mean), and p(inconsistent) with horizon, and that p(inconsistent) is in between pure random and pure deterministic noise.
+The quantitative agreement is almost perfect for p(high info) and for p(inconsistent) in the [1 3] condition, but the model seems to systematically overestimate p(low mean) and p(inconsistent) in [2 2] conditions, although the discrepancy is relatively small (overestimating p(low mean) by 0.054 or 31.37\%, and p(inconsistent) by 0.049 or 27.83\% in [2 2] condition).
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[width=1\textwidth]{figures/RDBayes_2noise_modelA.jpg}
+		\caption{
+			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
+			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
+I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
+COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
+			Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
+		\label{fig:mb3}
+	\end{center}
+\end{figure}
+
+\subsubsection*{Comparison with alternative models}
+To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
+These models varied \added[id=siyu]{in} whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+\added[id=siyu]{Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).}
+
+		\begin{table}[h]
+				\small
+				\centering
+				\begin{tabular}{|c|c|c|}
+						\hline
+						Model & Deterministic noise & Random noise \\
+						\hline
+						$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
+						\hline
+						$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
+						\hline
+						$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
+						\hline
+						$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
+						\hline
+						$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
+						\hline
+						$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
+						\hline
+					\end{tabular}
+				\caption{Variants of the model.}
+			\label{tab:models}	
+		\end{table}
+\added[id=siyu]{
+	The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Supplementary Figure S9.
+We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
+}As shown in Supplementary Figure S10, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of \replaced[id=siyu]{behavior}{responding}.
+However, the quantitative fit to the data is not as good (Supplementary Figure S10).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
+
+\added[id=siyu]{
+	Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model.
+The simulation was repeated 50 times for each model.
+Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}, S).
+This phenomenon was also examined and discussed in the section "Model validation" above.
+}
+
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
+		\caption{
+			Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+		\label{fig:6model}
+	\end{center}
+\end{figure}
+
+%The previous section suggests that behavioral variability in random exploration can be at least partially explained by deterministic processing of the stimulus.
+
+%dominated by random noise.
+To test this more explicitly, we build a series of models in which different assumptions are made regarding the presence and absence of both types of noise and whether each type of noise if exists is horizon dependent (See Table \ref{tab:models}).
+In model A-D, we assumed the existence of both random and deterministic noise, in model A and B, random noise is assumed to be horizon-dependent, whereas in model A and C, deterministic noise is assumed to be horizon dependent.
+In model E, we assumed no random noise.
+In model F, we assumed no deterministic noise.
+
+%To evaluate and compare between models, we simulated choice behavior by taking the subject-level parameters from the Hierarchical Bayesian fits.
+The same model-free analysis as described in the previous session is applied to all 6 sets of simulated data for the 6 models respectively.
+(See Figure  \ref{fig:mb3}).
+
+%The original measure of random exploration, p(low mean), as used in \cite{wilson2014} can be explained by having deterministic noise alone (Figure \ref{fig:mb3}, Panel E2) or having random noise alone (Figure \ref{fig:mb3}, Panel F2).
+That participants qualitatively exploit the high-mean option less and choose the low-mean option more in horizon 6, can be explained by having both pure deterministic noise and pure random noise, as long as noise is horizon dependent.
+If both deterministic and random noise are assumed to be the same for both horizons (Figure \ref{fig:mb3}, Panel D), p(low mean) becomes completely flat and no horizon dependent random exploration is observed.
+
+%On the other hand, by looking at the percentage of inconsistent choices in the repeated pair of game, p(inconsistent), external noise alone can not account for behavior any more (Figure \ref{fig:mb3}, Panel E3, E4).
+Moreover, model C and D are disqualified that the increase of choice inconsistency with horizon can only be qualitatively accounted for when random noise is horizon-dependent (Figure \ref{fig:mb3}, Panel A, B, F).
+
+%Among the models A, B and F, where random noise is horizon dependent, model A provides the best quantitative fit.
+If there is no deterministic noise (Model F), then we overestimate the level of choice inconsistency in both horizons by a constant.
+In addition, horizon dependent deterministic noise gives slightly better model fits than if deterministic noise is assumed to be the same in both horizons.
+Overall, these model simulations confirmed that the horizon dependence of random noise is the main source of random exploration.
+
+\section*{Discussion}
+
+	% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY
+	\added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).} In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
+Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 20.05\% of the overall variability in people's behavior.
+
+	%	\added[id=siyu]{We quantified the contribution of both deterministic noise (unobserved deterministic processes of the stimulus) and random noise (intrinsic stochasticity in decision making, or stimulus irrelevant processes) contribute to behavioral variability in random exploration,
+
+	So where does this leave randomness in random exploration?
+\replaced[id=siyu]{In our model, random noise is modeled as non-stimulus-driven variability, which may include truly random noise (intrinsic stochasticity in decision making) or stimulus irrelevant deterministic processes that we could not control in the lab.
+As a result,}{Well, } the remaining 79.95 \% of the decision noise could be random or it could be deterministic \replaced[id=siyu]{processes that is not bound to the stimulus and that we can not control}{because we can't control everything}.
+In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
+the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were fidgeting or scratching their nose \citep{Musall2019}.
+Thus conceptually, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
+Conversely, our estimate of deterministic noise is a lower bound.
+\added[id=siyu]{Future work is needed to identify these additional sources of deterministic noise that are not controlled in our work, for example by tracking people's pose, body movements and eye movements during the experiment.}
+
+	In addition to the conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also systematically underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
+From both a conceptual and methodological perspective, our model provides a lower bound of deterministic noise and an upper bound of random noise.
+Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 79.95\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
+
+	\added[id=siyu]{This is in line with works of \citep{Bair1996} in which MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli.
+In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
+A more recent work showed that neural variability in the visual areas reflect statistics in the stimuli images \citep{Festa2021}.
+Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+}
+
+	\added[id=siyu]{Other research are in line with random exploration \citep{drugowitsch16}.
+In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+Although there is a distinction between suboptimal inference and random noise \citep{Pouget12}, they both show up as random noise in our definition.
+This suggests that a large proportion of variability in behavior may arise from computational errors in computing the correct strategy.
+In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+	}
+	%this is consistent with findings of \cite{drugowitsch16}.
+In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+This suggests that a large proportion of noise in behavior is generated randomly and that this may arise from computational errors in computing the correct strategy.
+% In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+
+	%In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+
+	%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. 
+	\added[id=siyu]{
+	Regardless of whether the remaining 79.95 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
+Specifically, a reduction in the strength with which reward drives the choice.
+In our decision model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, thus, the absolute value of different terms in the model do not matter, $\Delta Q$ is only affected by the relative magnitude of reward, information, bias and noise.
+Mathematically, our model is equivalent to
+	}
+	\begin{equation*}
+		\begin{split}
+			\Delta Q' &= \beta \Delta Q \\
+			&= \beta \Delta R + \beta A \Delta I + \beta b + \beta n_{det} + \beta n_{ran}\\
+			&= \beta \Delta R + A' \Delta I + b' + n_{det}' + n_{ran}'
+		\end{split}
+	\end{equation*}
+	\added[id=siyu]{
+	In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
+From a model fitting perspective, these two forms are equivalent.
+A decrease in $\beta$ here would be equivalent to an increase in variance of both deterministic and random noise.
+From a psychological perspective, however, they are quite different.
+An increase in both random and deterministic noise at a similar ratio, as observed in our experiment when horizon increases, suggests that either $\beta$ is reduced or $n_{det}'$ and $n_{ran}'$ are increased simultaneously at the same ratio.
+Neurally, reducing the $\beta$ corresponds to a reduction in reward sensitivity and reward coding, whereas simultaneously increasing $n_{det}'$ and $n_{ran}'$ corresponds to a reduction in some common noise filtering mechanism that reduces both random and deterministic noises.
+}
+
+	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
+Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
+In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}.
+
+	%On the other hand, our result suggests that at least one fifth of the decision noise in random exploration arises from a deterministic process.
+
+	%SIYU says the song bird literature supports that behavioral variability is reflected in neural variability.
+But it can't tell whether the neural variability arises from deterministic processes of external stimuli or random noise.
+I didn't read/remember the songbird paper, this is only based on the text below.
+			%"The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}"
+
+	%SIYU says the Ebitz paper sounds like choice signal is reduced during exploration, and p(switch|reward outcome) increases behaviorally in exploration state compared to exploitation?
+I am not sure it supported reward reduction?
+			%Such reduced coding of reward would be consistent with \cite{ebitz17}.%In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
+
+			%In our study, most of the decision noise seen can still be random noise (up to 72.95\%), and this is consistent with findings of \cite{drugowitsch16}.
+In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+This suggests that a large proportion of noise in behavior is generated randomly and that this may arise from computational errors in computing the correct strategy.
+% In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+
+	%And you can change the relative contribution of noise by changing $\beta$.
+
+	%In particular, in our analysis we excluded a scalar on $\Delta R$, if we include that then we get
+
+	%To see this, we need to return to our equation for the computation of value
+	%\begin{equation}
+	%	\Delta Q = \Delta R + A \Delta I + b + n_{det} + n_{ran}
+	%\end{equation}
+	%In this version of the equation, the difference in value is a weighted of sum of reward, information, bias, deterministic noise,  and random noise.
+Crucially, we assumed that the weighting given to the difference in reward is 1 --- that is the multiplier on $\Delta R$ is 1, unlike the variable $A$ which modulates the effect of information.
+	%A more general form of this equation would also include a scalar on $\Delta R$
+
+	%REGARDLESS OF WHETHER THE REMAINING 75\% IS RANDOM OR NOT, THE FACT THAT THE HORIZON CHANGE IN THE TWO NOISES ARE PROPORTIONAL TO EACH OTHER SUGGESTS A POSSIBLE MECHANISM FOR RANDOM EXPLORATION.
+SPECIFICALLY, A REDUCTION IN THE STRENGTH WITH WHICH REWARD DRIVES THE CHOICE.
+IN PARTICULAR, IN OUR ANALYSIS WE EXCLUDED A SCALAR ON $\Delta R$.
+IF WE INCLUDE THAT THEN WE GET
+	%\begin{equation}
+	%	\Delta Q = \beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
+	%\end{equation}
+	%In this equation, $\beta$ determines the weight given to the reward in the decision.
+From a model fitting perspective, these two forms are EQUIVALENT.
+From a psychological perspective they are quite different.
+And you can change the relative contribution of noise by changing $\beta$.
+
+%	A decrease in $beta$ here would be equivalent to an increase in variance of both deterministic and random noise. Such reduced coding of reward would be consistent with \cite{ebitz17}.
+	%A DECREASE IN $\beta$ HERE WOULD BE EQUIVALENT TO AN INCREASE IN VARIANCE OF BOTH DETERMINISTIC AND RANDOM NOISE.
+MAKES THE PREDICTION THAT SPATIAL BIAS SHOULD ALSO CHANGE IN PROPORTION (CHECK THIS!).
+
+	%SUCH REDUCED CODING OF REWARD WOULD BE CONSISTENT WITH EBITZ (I THINK).
+
+	%Randomness in exploratory behavior in random exploration can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).
+The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
+In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
+In our study, most of the decision noise seen can still be random noise (up to 72.95\%), and this is consistent with findings of \cite{drugowitsch16}.
+In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+This suggests that a large proportion of noise in behavior is generated randomly and that this may arise from computational errors in computing the correct strategy.
+% In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+
+	%Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
+Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
+In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}.
+
+	%On the other hand, our result suggests that at least one fifth of the decision noise in random exploration arises from a deterministic process.
+This is in line with works of \citep{Bair1996} in which MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli.
+In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
+Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+
+	%One way in which this may be achieved is if random exploration is modulated, not by increasing noise, but by decreasing reward sensitivity.
+For example, modifying equation XXX to include a reward sensitivity term we get ...
+
+	%Given that there is still a considerable window for truly stochastic processes in the brain to be driving random exploration,
+
+	%Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between random and deterministic noise that we found in Figure \ref{fig:mb1}, where random noise is 2-3 times the size of deterministic noise.
+
 	\section*{Materials and Methods}
@@
 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved driven by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
-This left 66 for the main analysis.
-Note that including the 14 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
+In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved \deleted[id=siyu]{driven}by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
+This left 65 for the main analysis.
+Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
 
@@
 To maximize their rewards in each game, participants need to exploit the slot machine with the highest mean, but they cannot identify this best option without exploring both options first.
-
-		\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
-			\caption{ 
-			Schematic of the experiment.
-(A) Dynamics of an example horizon 6 game.
-Here the first four trials are forced trials in which participants are instructed which option to play.
-After the forced trials, participants are free to choose between the two options for the remainder of the game.
-(B) Different possible states of the game after the first free choice over the course of the experiment.
-On average, participants play more than 150 such games, with varying horizon (1 vs 6), uncertainty condition ([1 3] vs [2 2]) and observed rewards.
-In addition, all games are repeated (as Game 18 and 100 are here) such that participants will be faced with the exact same pattern of forced trials and exact same outcomes from those forced trials twice within each experiment.
-These repeated games allow us to compute the relative contribution of deterministic and random noise by analyzing the extent to which choices are {\em consistent} across the repeated games.}
-			\label{fig:taskfig}
-		\end{center}
-	\end{figure}
 
@@
 	The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
+We used these forced-choice instructed trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
 The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
@@
 	\subsection*{Model-based analysis}
-	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').
-Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.
+	We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate \replaced[id=siyu]{deterministic noise from random noise.}{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').}Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.
 
@@
 For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1, -1 or 0, +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-
 The other variables are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively.
@@
 
-	\subsubsection*{Parameter recovery\label{ch:appendix:bayesrecovery}}
-
-	To be sure that our fit parameter values were meaningful, we tested the ability of our model fitting procedure to recovery parameters from simulated data.
-In particular, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated choices to see whether we can recover the parameters.
-
-	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}.
-Parameter recovery is good (in terms of correlation between the true vs fitted parameters) for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C).
-The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
-This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
-In addition we see better recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/precover.png}
-			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
-			}
-			\label{fig:pararecover1}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet.jpg}
-			\caption{Parameter recovery over the posterior of random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for purely random noise (left column) and purely deterministic noise (right column) games.
-			}
-			\label{fig:pararecover2}
-		\end{center}
-	\end{figure}
-
-	Finally, despite the strong correlation between simulated and fit values of deterministic noise, we observed a systematic underestimation of $\sigma_{det}$.
-This is further illustrated in Supplementary Figure S5.
-In the simulation with 0 random noise and fully deterministic noise, our model perfectly recovered both random and deterministic noise, however in the simulation with full random noise and 0 deterministic noise, our model recovered only XXX \% of deterministic noise, and the remaining shows up as random noise instead (Figure \ref{fig:pararecover2}).
-Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
-
-	%Overall, we are able to recover both deterministic and random noises using our model to a satisfactory extent.
-
 	\subsection*{Data and code}
@@
 %website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.
-
-	\section*{Results}
-
-	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus driven `random noise' on people's decisions (Figure \ref{fig:taskfig}).
-In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
-Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
-
-	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
-1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
-These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
-
-	Relative to the original Horizon Task, the key modification here is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games (separated by several minutes in time so as to avoid detection).
-By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
-Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their behavior on the two versions of each game will be consistent.
-Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games should be inconsistent.
-
-	On average participants played 67.31 repeated games each allowing us to quantify the extent to which their behavior was a deterministic function of the stimulus or not.
-
-	%, the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.
-
-	%These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.
-Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent.
-Conversely if noise is randomly driven, then choices on repeated games should be inconsistent.
-
-	\subsection*{Both behavioral variability and information seeking increase with horizon}
-
-	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}.
-As in our previous work, we find evidence for two types of exploration in the Horizon Task.
-Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.
-
-	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
-This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(65) = 6.60, p $<$ 0.001 for [1 3], t(65) = 8.14, p $<$ 0.001 for [2 2]).
-Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition.
-Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(65) = 7.00, p $<$ 0.001).
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
-			\caption{
-			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
-			\label{fig:modelfree}
-		\end{center}
-	\end{figure}
-
-	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}
-
-	Next we asked whether participants' choices were consistent or inconsistent in the two repetitions of each game.
-The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times.
-Conversely, if choice is not a deterministic function of the stimulus, participants' choices should be independent, and hence more inconsistent across the repetitions of the game.
-
-	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was stimulus driven (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions.
-For [1 3] condition, t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6; For [2 2] condition, t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6).
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 196) = 61.19, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is not a deterministic function of the stimulus.
-
-	\begin{figure}[h]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_pinconsistent.jpg}
-			\caption{Model-free analysis suggests that both deterministic and random noise contribute to the choice variability in random exploration.
-For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
-However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
-			\label{fig:mf2}
-		\end{center}
-	\end{figure}
-
-	To gain more quantitative insight into these results, we computed theoretical values for the choice inconsistency for the purely deterministic and purely random noise cases.
-For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case.
-For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
-	\begin{equation*}
-	\begin{split}
-	p(\mbox{consistent}) &= p(\mbox{low mean})^2 + p(\mbox{high mean})^2\\
-	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
-	\mbox{hence},\quad p(\mbox{inconsistent}) &=  
-	1 - p(\mbox{consistent}) = 
-	2 p(\mbox{low mean})(1-p(\mbox{low mean}))
-	\end{split}
-	\end{equation*}
-	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction Specifically, behavior is different from pure random noise prediction in the both the [1 3] condition (t(65) = 5.60, p $<$ 0.001 for horizon 1, t(65) = 5.62 p $<$ 0.001 for horizon 6) and [2 2] condition (t(65) = 3.86, p $<$ 0.001 for horizon 1, t(65) = 3.42, p $<$ 0.001 for horizon 6).
-Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition ( t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6).
-This suggests that at least some of the `noise' is a deterministic function of the stimulus, although from this analysis it is not clear whether this deterministic noise increases with horizon or not.
-
-	\subsection*{Model-based analysis shows deterministic noise changes with horizon}
-
-	To more precisely quantify the contribution of stimulus-driven deterministic noise, we turned to model fitting.
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate deterministic noise from and random noise.
-In particular, we assume that in repeated games, the value of stimulus-driven deterministic noise is frozen whereas random noise is drawn independently both times.
-
-	\subsubsection*{Overview of model}
-	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
-To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
-	\begin{equation}
-	\Delta Q= \Delta R+A \Delta I+b+n_{det}+n_{ran}
-	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{det}$ and $n_{ran}$ are deterministic noise and random noise respectively which are assumed to come from logistic distributions with mean 0.
-
-	The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; $n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeat plays and changes every game.
-
-	For each pair of repeated games, the set of forced-choice trials are exactly the same, so the deterministic noise, $n_{det}$, should be the same while the random noise, $n_{ran}$ may be different.
-This is exactly how we distinguish deterministic noise from random noise.
-In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.
-
-	\subsubsection*{Model fitting}
-	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
-In particular, we fit values of the information bonus $A$, spatial bias $B$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon.
-Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
-
-	\subsubsection*{Model fitting results}
-
-	Posterior distributions over the group-level means of the deterministic and random noise variance are shown in Figure \ref{fig:mb1}.
-Consistent with our model-free results, we see that both random and deterministic noise variances are non-zero and that random noise is about 2-3 times larger than the deterministic noise.
-In addition, we find that both random and deterministic noise increase with horizon.
-This increase was larger for random noise (M = 4.55, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (M = 1.78, 98.12\% of samples showed an increase in deterministic noise with horizon).
-But intriguingly the relative increase in both types of noise was the same (Figure \ref{fig:ratio}).
-That is when we compute the relative increase in deterministic noise with horizon, $\sigma_{det, horizon6}/\sigma_{det, horizon1}$, it is almost identical to the relative increase in random noise with horizon $\sigma_{ran, horizon6}/\sigma_{ran, horizon1}$.
-	%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RDBayes_hyperprior.jpg}
-			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
-Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
-However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
-			%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
-			\label{fig:mb1}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{figures/RanDetNoise_ratio1.png}
-			\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
-The ratio in the standard deviations of noise between horizon 6 and horizon 1 is similar for random and deterministic noise.}
-			\label{fig:ratio}
-		\end{center}
-	\end{figure}
-
-	%\begin{figure}[H]
-	%	\begin{center}
-%			\includegraphics[width=1\textwidth]{figures/pinconsistent_4panel_age_acchance_2noisemodelA_2noisemodelA.png}
-%			\caption{}
-%			\label{fig:mb2}
-%		\end{center}
-%	\end{figure}
-
-	\subsubsection*{Posterior predictive checks}
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_2noise_modelA.png}
-			\caption{
-			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
-			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
-I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
-COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
-			Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
-			\label{fig:mb3}
-		\end{center}
-	\end{figure}
-
-	In addition to fitting the model to behavior, it is also important to check whether the model captures the qualitative patterns of the data \citep{Wilson2019} --- specifically how p(high info), p(low mean) and p(inconsistent) change with horizon.
-
-	To perform this `posterior predictive check,' we created a set of simulated data by taking the subject-level parameters from the hierarchical Bayesian fits and having the model play the same sequence of games as seen by the subjects.
-We then applied the same model-free analysis as described in the previous sections to this simulated data set and compared the model's behavior to that of participants.
-As shown in Figure  \ref{fig:mb3}, the model can account for all qualitative patterns in the data --- the increase in p(high info), p(low mean), and p(inconsistent) with horizon, and that p(inconsistent) is in between pure random and pure deterministic noise.
-The quantitative agreement is almost perfect for p(high info), but the model seems to systematically underestimate p(low mean) and p(inconsistent), although the discrepancy is small (underestimating p(low mean) by 0.047 or 27.30\%, and p(inconsistent) by 0.027 or 14.27\%).
-
-	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
-These models varied whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
-As shown in Supplementary Figure S5, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of responding (Note model F does not capture that  p(inconsistent) is between pure random and deterministic?).
-However, the quantitative fit to the data is not as good (Supplementary Figure S5).
-
-	%The previous section suggests that behavioral variability in random exploration can be at least partially explained by deterministic processing of the stimulus.
-
-	%dominated by random noise.
-To test this more explicitly, we build a series of models in which different assumptions are made regarding the presence and absence of both types of noise and whether each type of noise if exists is horizon dependent (See Table \ref{tab:models}).
-In model A-D, we assumed the existence of both random and deterministic noise, in model A and B, random noise is assumed to be horizon-dependent, whereas in model A and C, deterministic noise is assumed to be horizon dependent.
-In model E, we assumed no random noise.
-In model F, we assumed no deterministic noise.
-
-	%To evaluate and compare between models, we simulated choice behavior by taking the subject-level parameters from the Hierarchical Bayesian fits.
-The same model-free analysis as described in the previous session is applied to all 6 sets of simulated data for the 6 models respectively.
-(See Figure  \ref{fig:mb3}).
-
-	%The original measure of random exploration, p(low mean), as used in \cite{wilson2014} can be explained by having deterministic noise alone (Figure \ref{fig:mb3}, Panel E2) or having random noise alone (Figure \ref{fig:mb3}, Panel F2).
-That participants qualitatively exploit the high-mean option less and choose the low-mean option more in horizon 6, can be explained by having both pure deterministic noise and pure random noise, as long as noise is horizon dependent.
-If both deterministic and random noise are assumed to be the same for both horizons (Figure \ref{fig:mb3}, Panel D), p(low mean) becomes completely flat and no horizon dependent random exploration is observed.
-
-	%On the other hand, by looking at the percentage of inconsistent choices in the repeated pair of game, p(inconsistent), external noise alone can not account for behavior any more (Figure \ref{fig:mb3}, Panel E3, E4).
-Moreover, model C and D are disqualified that the increase of choice inconsistency with horizon can only be qualitatively accounted for when random noise is horizon-dependent (Figure \ref{fig:mb3}, Panel A, B, F).
-
-	%Among the models A, B and F, where random noise is horizon dependent, model A provides the best quantitative fit.
-If there is no deterministic noise (Model F), then we overestimate the level of choice inconsistency in both horizons by a constant.
-In addition, horizon dependent deterministic noise gives slightly better model fits than if deterministic noise is assumed to be the same in both horizons.
-Overall, these model simulations confirmed that the horizon dependence of random noise is the main source of random exploration.
-
-	\section*{Discussion}
-
-	% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY
-
-	In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
-Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 27.05\% of the overall variability in people's behavior.
-%and increased with horizon - a hallmark of an exploratory process.
-
-	So where does this leave randomness in random exploration?
-Well, the remaining 72.95 \% of the noise could be random or it could be deterministic because we can't control everything.
-In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
-the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose.
-Thus conceptually, our estimate of deterministic noise is a lower bound.
-Conversely, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
-In addition, our estimation method also systematically underestimate deterministic noise (part of the deterministic noise will show up as random noise in model fitting, see Figure \ref{fig:pararecover2}), so methodologically our model also provides a lower bound of deterministic noise and an upper bound on random noise.
-Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 72.95\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
-			%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain.
-
-	Randomness in exploratory behavior in random exploration can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).
-The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
-In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
-In our study, most of the decision noise seen can still be random noise (up to 72.95\%), and this is consistent with findings of \cite{drugowitsch16}.
-In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
-This suggests that a large proportion of noise in behavior is generated randomly and that this may arise from computational errors in computing the correct strategy.
-% In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
-Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
-Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
-In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}.
-
-	On the other hand, our result suggests that at least a quarter of the decision noise in random exploration arises from a deterministic process.
-This is in line with works of \citep{Bair1996} in which MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli.
-In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
-Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
-
-	%One way in which this may be achieved is if random exploration is modulated, not by increasing noise, but by decreasing reward sensitivity.
-For example, modifying equation XXX to include a reward sensitivity term we get ...
-
-	%Given that there is still a considerable window for truly stochastic processes in the brain to be driving random exploration,
-
-	Regardless of whether the remaining 72.95 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration.
-Specifically, a reduction in the strength with which reward drives the choice.
-In particular, in our analysis we excluded a scalar on $\Delta R$, if we include that then we get
-
-	%REGARDLESS OF WHETHER THE REMAINING 75\% IS RANDOM OR NOT, THE FACT THAT THE HORIZON CHANGE IN THE TWO NOISES ARE PROPORTIONAL TO EACH OTHER SUGGESTS A POSSIBLE MECHANISM FOR RANDOM EXPLORATION.
-SPECIFICALLY, A REDUCTION IN THE STRENGTH WITH WHICH REWARD DRIVES THE CHOICE.
-IN PARTICULAR, IN OUR ANALYSIS WE EXCLUDED A SCALAR ON $\Delta R$.
-IF WE INCLUDE THAT THEN WE GET
-	\begin{equation}
-	    \Delta Q = \beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
-	\end{equation}
-
-	A decrease in $beta$ here would be equivalent to an increase in variance of both deterministic and random noise. Such reduced coding of reward would be consistent with \cite{ebitz17}.
-	%A DECREASE IN $\beta$ HERE WOULD BE EQUIVALENT TO AN INCREASE IN VARIANCE OF BOTH DETERMINISTIC AND RANDOM NOISE.
-MAKES THE PREDICTION THAT SPATIAL BIAS SHOULD ALSO CHANGE IN PROPORTION (CHECK THIS!).
-
-	%SUCH REDUCED CODING OF REWARD WOULD BE CONSISTENT WITH EBITZ (I THINK).
-
-	%Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between random and deterministic noise that we found in Figure \ref{fig:mb1}, where random noise is 2-3 times the size of deterministic noise.
 
@@
 
+	%\listofchanges
+
 \end{document}

### 2022-12-19 codes/EEplot_2noise_hyperprior_6model_recovery.m

@@ -1,7 +1,6 @@
 function EEplot_2noise_hyperprior_6model_recovery(plt, sp)
-    plt.figure(6,4, 'fontsize_label', 12, 'fontsize_leg',10, ...
-        'fontsize_axes',12, 'fontsize_title', 15, ...
-        'gapH_custom', [10 0 0 0 0 0 20]);
+    plt.figure(6,4, 'fontsize_label', 18, 'fontsize_leg',15, 'fontsize_axes',18, 'fontsize_title', 20, ...
+        'gapH_custom', [0 0 0 0 0 0 10]);
     names = {'Random noise','\Delta Random noise',...
         'Deterministic noise','\Delta Deterministic noise'};
     plt.setfig_all('ytick', []);
@@ -14,6 +13,9 @@ function EEplot_2noise_hyperprior_6model_recovery(plt, sp)
     plt.setfig(21:24, 'xlabel', {'\sigma_{ran}','\Delta \sigma_{ran}','\sigma_{det}','\Delta \sigma_{det}'});
     plt.setfig(1:4, 'title', names);
     fns = {'NoiseRan', 'NoiseDet'};
+    plt.setfig('ylabel', [{'\sigma^{ran}_{horizon}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{horizon}, \sigma^{det}_{}'} repmat({''},1,3), ...
+        {'\sigma^{ran}_{}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{}, \sigma^{det}_{}'} repmat({''},1,3), ...
+        {'\sigma^{ran}_{horizon}'} repmat({''},1,3), {'\sigma^{det}_{horizon}'} repmat({''},1,3)]);
     for mi = 1:6
         for i = 1:length(fns)
             fn = fns{i};
@@ -25,12 +27,12 @@ function EEplot_2noise_hyperprior_6model_recovery(plt, sp)
             end
             ymax = max(tl* 1.05,[],'all');
             plt.setfig_ax('ylim', [0 ymax], 'xtick', 0:5:40,'xlim', [0 20], ...
-                'legend', legs{i*2 - 1}, 'ylabel', ylabs{i*2-1});
-            plt.plot(tm, tl, [], 'line', 'color', color{i*2 -1});
+                'legend', legs{i*2 - 1});
[…]

### 2022-12-19 codes/EEplot_2noise_modelcomparison.m

@@ -2,9 +2,9 @@ function EEplot_2noise_modelcomparison(plt, gp, simugp)
     plt.figure(6,4,'fontsize_label', 18, 'fontsize_leg',15, 'fontsize_axes',18, 'fontsize_title', 20, ...
         'gapH_custom', [10 0 0 0 0 0 10], 'gapW_custom', [10 0 0 0 0]);   
     plt.setfig('xlabel', [repmat({''},1,20), repmat({'horizon'},1,4)], ...
-        'ylabel', [{'model A'} repmat({''},1,3), {'model B'} repmat({''},1,3), ...
-        {'model C'} repmat({''},1,3), {'model D'} repmat({''},1,3), ...
-        {'model E'} repmat({''},1,3), {'model F'} repmat({''},1,3)], ...
+        'ylabel', [{'\sigma^{ran}_{horizon}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{horizon}, \sigma^{det}_{}'} repmat({''},1,3), ...
+        {'\sigma^{ran}_{}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{}, \sigma^{det}_{}'} repmat({''},1,3), ...
+        {'\sigma^{ran}_{horizon}'} repmat({''},1,3), {'\sigma^{det}_{horizon}'} repmat({''},1,3)], ...
         'title', [{'p(high info)', 'p(low mean)', 'p(inconsistent), [1 3]', 'p(inconsistent), [2 2]', ...
         } repmat({''}, 1, 20)], 'xlim', [0.5 2.5], 'xtick', [1 2], 'xticklabel', [1 6], ...
         'ylim', repmat({[0.35 0.65],[0 0.35],[0 0.45],[0 0.45]},1,6), ...
@@ -28,6 +28,7 @@ function EEplot_2noise_modelcomparison(plt, gp, simugp)
         plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13, tgp.GPav_GPste_p_inconsistent13,'line', 'color', 'AZblue');
         plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13_randomtheory_byR, tgp.GPav_GPste_p_inconsistent13_randomtheory_byR,'line', 'color', 'AZblue50','linestyle','--');
         plt.setfig_ax('legend',{'data','model','\sigma^{ran} only'});
+%         plt.setfig_ax('legend',{'data','model'});
         
[…]

### 2022-12-19 codes/EEplot_2noise_modelcomparison1.m

@@ -2,7 +2,6 @@ function EEplot_2noise_modelcomparison1(plt, gp, simugp)
     plt.figure(1,4,'is_title',1);
     plt.setfig('xlabel', [repmat({'horizon'},1,4)], ...
         'ylabel', [{'p(high info)', 'p(low mean)', 'p(inconsistent), [1 3]', 'p(inconsistent), [2 2]'}], ...
-        'legend', [repmat({{'data', 'model'}},1,4)], ...
         'ylim', repmat({[0.35 0.65],[0 0.35],[0 0.45],[0 0.45]},1,1), ...
         'ytick', repmat({[0.4:0.1:0.6],[0:.1:0.3],[0:.1:0.5],[0:.1:0.5]},1,1));
     plt.setfig_all('xlim', [0.5 2.5], 'xtick', [1 2], 'xticklabel', [1 6], ...
@@ -12,22 +11,26 @@ function EEplot_2noise_modelcomparison1(plt, gp, simugp)
         plt.ax(mi,1);
         plt.plot(1:2,gp.GPav_p_hi13, gp.GPste_p_hi13,'line','color','AZred');
         plt.plot(1:2,tgp.GPav_GPav_p_hi13, tgp.GPav_GPste_p_hi13,'line','color','AZblue');
-        
+        plt.setfig_ax('legend',{'data','model'});
         plt.ax(mi,2);
         plt.plot(1:2,gp.GPav_p_lm22, gp.GPste_p_lm22,'line','color','AZred');
         plt.plot(1:2,tgp.GPav_GPav_p_lm22, tgp.GPav_GPste_p_lm22,'line','color','AZblue');
-        
+        plt.setfig_ax('legend',{'data','model'});
         plt.ax(mi,3);
         plt.plot(1:2,gp.GPav_p_inconsistent13, gp.GPste_p_inconsistent13,'line', 'color', 'AZred');
-%         plt.plot(1:2,gp.GPav_p_inconsistent13_randomtheory, gp.GPste_p_inconsistent13_randomtheory,'line', 'color', 'AZred50','linestyle','--');
-%         plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13_randomtheory, tgp.GPste_GPav_p_inconsistent13_randomtheory,'line', 'color', 'AZblue50','linestyle','--');
[…]

### 2022-12-19 codes/EEplot_2noise_parameter_recovery.m

@@ -15,7 +15,7 @@ function EEplot_2noise_parameter_recovery(plt, p1, p2)
             d1 = [d1{:}];
             d2 = [d2{:}];
             str{i*2-2+hi} = plt.scatter(d1', d2', 'diag');
-            lsline
+%             lsline
         end
     end
     plt.setfig('legend', str);

### 2022-12-19 codes/EEplot_2noise_pinconsistent.m

@@ -1,24 +1,36 @@
-function plt = EEplot_2noise_pinconsistent(plt, gp, ver)
-    if ~exist('ver', 'var') || isempty(ver)
-        ver = '';
+function plt = EEplot_2noise_pinconsistent(plt, gp, ver13, ver22, strpfx)
+    if ~exist('ver13', 'var') || isempty(ver13)
+        ver13 = '';
+    end
+    if ~exist('ver22', 'var') || isempty(ver22)
+        ver22 = '';
+    end
+    if ~exist('strpfx', 'var') || isempty(strpfx)
+        strpfx = '';
     end
     leg = {'behavioral data','deterministic noise only','random noise only'};
     plt.setfig('xlabel', {'horizon','horizon'}, 'ylabel', {'p(inconsistent)','p(inconsistent)'}, ...
         'title', {'[1 3] condition','[2 2] condition'}, 'legloc', {'SW','SW'}, ...
         'legend', {leg,leg}, ...
         'xlim', {[0.5 2.5],[0.5 2.5]}, 'xtick', {[1 2],[1 2]}, 'xticklabel', {[1 6],[1 6]});
-    plt.plot(1:2,gp.GPav_p_inconsistent13, gp.GPste_p_inconsistent13, 'line', ...
+    plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent13']), gp.([strpfx 'GPste_p_inconsistent13']), 'line', ...
         'color', 'AZcactus');
     plt.plot(1:2,[0 0], [0 0],'line', 'color', 'AZblue');
-    plt.plot(1:2,gp.(['GPav_p_inconsistent13_randomtheory' ver]), ...
-        gp.(['GPste_p_inconsistent13_randomtheory' ver]),'line', 'color', 'AZred');
-    plt.sigstar_y(gp.GPav_p_inconsistent13, 1:2, gp.GPpvalue_p_inconsistent13, 'right');
+    plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent13_randomtheory' ver13]), ...
+        gp.([strpfx 'GPste_p_inconsistent13_randomtheory' ver13]),'line', 'color', 'AZred');
+    plt.sigstar_y(gp.([strpfx 'GPav_p_inconsistent13']), 1:2, gp.([strpfx 'GPpvalue_p_inconsistent13']), 'right');
     plt.new;
[…]

### 2022-12-19 codes/mainFigures_paramrecover_coverageanalysis.m

@@ -1,13 +1,13 @@
 outputdir = '../bayesoutput/all/';
 sp = load(fullfile(outputdir, ['HBI_DetRanNoise_samples.mat'])).samples;
-%% compute confidence interval
-ci = struct;
-ci.dNoiseRan = quantile(reshape(sp.dNoiseRan,1,[]), [0.025, 0.975]);
-ci.dNoiseDet = quantile(reshape(sp.dNoiseDet,1,[]), [0.025, 0.975]);
-for i = 1:2
-    ci.NoiseRan(i,:) = quantile(reshape(sp.NoiseRan(:,:,i),1,[]), [0.025, 0.975]);
-    ci.NoiseDet(i,:) = quantile(reshape(sp.NoiseDet(:,:,i),1,[]), [0.025, 0.975]);
-end
+% %% compute confidence interval
+% ci = struct;
+% ci.dNoiseRan = quantile(reshape(sp.dNoiseRan,1,[]), [0.025, 0.975]);
+% ci.dNoiseDet = quantile(reshape(sp.dNoiseDet,1,[]), [0.025, 0.975]);
+% for i = 1:2
+%     ci.NoiseRan(i,:) = quantile(reshape(sp.NoiseRan(:,:,i),1,[]), [0.025, 0.975]);
+%     ci.NoiseDet(i,:) = quantile(reshape(sp.NoiseDet(:,:,i),1,[]), [0.025, 0.975]);
+% end
 %% compute coverage
 simudir = '../bayesoutput/simurepeat';
 coverage = struct;
@@ -16,11 +16,11 @@ for repi = 1:200
     tstat = load(fullfile(simudir, sprintf('HBI_DetRanNoise_simu_%d_stat.mat', repi)));
     recovered.dNoiseRan(repi) = tstat.stats.mean.dNoiseRan;
     recovered.dNoiseDet(repi) = tstat.stats.mean.dNoiseDet;
-    coverage.dNoiseRan(repi) = W.is_between(tstat.stats.mean.dNoiseRan, ci.dNoiseRan);
-    coverage.dNoiseDet(repi) = W.is_between(tstat.stats.mean.dNoiseDet, ci.dNoiseDet);
+    coverage.dNoiseRan(repi) = W.is_between(mean(sp.dNoiseRan,'all'), [tstat.stats.ci_low.dNoiseRan tstat.stats.ci_high.dNoiseRan]);
[…]

### 2022-12-19 codes/main_1_behavior.m

@@ -20,7 +20,7 @@ save(fullfile(savedir, 'idxsub_exclude.mat'), 'id', 'idall','idxsub');
 %% analyze group
 clear gp;
 additionalcompare = {{'p_inconsistent13'},{'p_inconsistent22'},...
-    {'p_inconsistent13','p_inconsistent13_randomtheory'},{'p_inconsistent22','p_inconsistent22_randomtheory'}};
+    {'p_inconsistent13','p_inconsistent13_randomtheory_byR'},{'p_inconsistent22','p_inconsistent22_randomtheory_byR'}};
 gp{1} = W.analysis_1group(sub, true, additionalcompare);
 gp{2} = W.analysis_1group(suball, true, additionalcompare);
 suffix = {'','_all'};
@@ -34,13 +34,13 @@ for gi = 1:2
     plt.figure(1,2,'is_title',1);
     plt.setfig(1:2, 'ylim', {[0, 0.3] + gi*0.05,[0.35 0.65]}, ...
         'ytick', {0.1:.1:0.4, 0.4:.1:0.6}, ...
-        'legend', {{'[1 3]'},{'[2 2]','[1 3]'}});
+        'legend', {{'[2 2]','[1 3]'},{'[1 3]'}}, 'legloc',{'SE','SE'});
     plt = EEplot_modelfree(plt, gp{gi},1,1);
     %% 2 noise figure
     plt.figure(1,2,'is_title',1);
     plt.setfig(1:2, 'ylim', {[0 0.45],[0 0.45]}, ...
         'ytick', 0:0.1:0.4);
-    plt = EEplot_2noise_pinconsistent(plt, gp{gi}, '_byR');
+    plt = EEplot_2noise_pinconsistent(plt, gp{gi}, '_byR', '_byR');
 end
 %% save bayes data
 load('../data/all/idxsub_exclude.mat');

### 2022-12-21 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
siyu final draft

@@
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-\added[id=siyu]{By a novel bayesian analysis, we are able to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.} In particular, we find evidence that at least 20$\%$ of the variability in random exploration can \added[id=siyu]{indeed} be accounted for by deterministic processing of the stimulus.
+\added[id=siyu]{We built a novel computational model which could detect and separate previously unobserved deterministic processes from random noise, without having to make a priori assumptions about the nature of the deterministic process.
+} In particular, we find evidence that around 15$\%$ of the variability in random exploration can be accounted for by deterministic processing of the stimulus.
 This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
-\added[id=siyu]{Our work sheds light on how random exploration may be implemented by reducing the filtering of task-irrelevant stimuli in the decision process.}
+\added[id=siyu]{Our work takes us one step closer to understanding of the nature of decision noise in random exploration.}
 
@@
 \added[id=siyu]{Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model.
-A missing deterministic component from the model could give rise to variability in behavior that might be mistakenly taken as a random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
+A missing deterministic component from the model could give rise to variability in behavior that might appear to be a random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
 Avoiding an ex is a deterministic process, but if we fail to take ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
@@
 Conversely, the more their decision is driven by \replaced[id=siyu]{random noise}{other processes, including both stochastic and unobserved deterministic processes}, the less consistent their behavior should be.
-By analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
+\added[id=siyu]{We built a computational model to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.
+Our model could detect the existence of previously unobserved deterministic processes without making a priori assumptions about the deterministic process that we want to observe.} By analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
 \replaced[id=siyu]{Our work}{This does not prove that random exploration is entirely deterministic or stochastic, but} provides a lower bound on how much deterministic processes contribute to random exploration\added[id=siyu]{, and an upper bound on how much true random noises contribute to random exploration}.
@@
 By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
-Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their \replaced[id=siyu]{choice}{behavior} on the two versions of each game will be consistent.
+Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their \replaced[id=siyu]{choices}{behavior} on the two versions of each game will be consistent.
 Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games \replaced[id=siyu]{could}{should} be inconsistent.
-\replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which their behavior was \replaced[id=siyu]{driven by a deterministic noise process vs a random noise process}{a deterministic function of the stimulus or not}.
+\replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which \added[id=siyu]{the variability in} their behavior was \replaced[id=siyu]{driven by a deterministic process vs a random noise process}{a deterministic function of the stimulus or not}.
 
@@
 Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
+We test this by leaving out one known deterministic process from the decision model, and ask if our method could recover that known deterministic process as deterministic noise.
 In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition on explore-exploit decisions.
-$\Delta Q, \Delta R, n_{det}, n_{ran}$ below represent the same variables as in the full model.
 	$$\Delta Q= \Delta R+n_{det}+n_{ran}$$
-	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
+	Here, $\Delta Q, \Delta R, n_{det}, n_{ran}$ represent the same variables as in the full model.	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
 By comparing the	fitted posterior distributions over the group-level means of the deterministic and random noise parameters $\sigma_{det}$ and $\sigma_{ran}$, as expected, we observed an increase in deterministic noise and no change in random noise between the reduced and the full model (Figure \ref{fig:reducedmodel}).
+This suggests that our model is capable of detecting missing deterministic processes.
 	}
@@
 	\added[id=siyu]{Secondly, we evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
-In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and the re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
+In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
 The simulation and re-fitting was repeated for 200 times.
-Then we counted out of the 200 repetitions, whether the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval for more than 95\% of the times (this ratio will be referred to as the coverage rate).
-For random noise, the coverage rate is 100\% for both horizon 1, horizon 6 and noise differences between horizons.
-For deterministic noise, the coverage rate is 66\% for horizon 1, 69\% for horizon 6, and 97\% for noise differences between horizons.
-By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of re-fitted parameters to the simulated data, it is clear that our model systematically underestimates deterministic noise.
+Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval.
+If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations (this ratio will be referred to as the coverage rate).
+For random noise, the coverage rate is 100\% for both horizon 1, horizon 6, and the horizon difference.
+For deterministic noise, the coverage rate is 66\% for horizon 1 and 69\% for horizon 6.
+By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of recovered parameters, it is clear that our model systematically underestimates deterministic noise (Figure \ref{fig:coverage2}).
+Despite the underestimation of deterministic noise in both horizons, we could still reliably detect the horizon changes of deterministic noise (coverage rate is 97\%).
+This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizons.
 For random noise, our model fitting procedure yields a faithful recovery.
-However, because our random noise is modeled as non-stimulus-driven noise which includes both truly stochastic neural noise and possible deterministic noises irrelevant to the stimuli, our random noise estimate provides an upper bound of `random noise' induced by intrinsic stochastic processes in the brain.
-Because the underestimation of deterministic noise is partially canceled out when taking the difference, our model fitting procedure passed the coverage test in estimating the horizon differences in both random and deterministic noise.}
+However, there is a conceptual limitation.
+Because random noise is modeled as non-stimulus-driven noise, it can include both true stochastic neural noise and possible deterministic noises which do not depend on the stimuli.
+Because of this, our random noise estimate provides an upper bound of true `random noise' induced by intrinsic stochastic processes in the brain.}
 
@@
 The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).
-Despite the strong correlation between simulated and fit values of deterministic noise (R $>$ 0.8), we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
+The strength of correlation between simulated and fit values are strong for both deterministic noise (R $>$ 0.8) and random noise (R $>$ 0.9).
+Despite the strong inter-subject correlations, we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
 
@@
 	\added[id=siyu]{Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
-All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ are simulated.
+All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ were tested.
 In a special case, we evaluated how our model performs when there is only random noise or only deterministic noise (Figure \ref{fig:puredetran}).
-In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with full random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
+In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with fully random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
 However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Supplementary Figures S7).
-Overall, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
+Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
+
+	%Overall, we see better recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
 
 	%This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
-In addition we see better recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
 
@@
 	\added[id=siyu]{Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
-We see better parameter recovery for random noise than deterministic noise.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+In addition, We see better parameter recovery for random noise than deterministic noise.
 This is likely because we effectively have half as many trials for deterministic noise.
 In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-Our model provides a lower bound for deterministic noise and an upper bound for random noise.
 	}
@@
 \replaced[id=siyu]{Numerically,}{and that} random noise is about 2-3 times larger than the deterministic noise.
-\added[id=siyu]{By computing the 5\% quantile of the posterior distribution of $\sigma_{det}/(\sigma_{det}+\sigma_{ran})$, our data suggests that at least 20.05\% of the variability in random exploration can be accounted for by deterministic noise.} In addition, we find that both random and deterministic noise increase with horizon.
-This increase was larger for random noise (M = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (M = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon).
-But intriguingly the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}).
+\added[id=siyu]{By computing the posterior distribution of $\sigma^2_{det}/(\sigma^2_{det}+\sigma^2_{ran})$, our data suggests that 14.25\% of the variability in random exploration is accounted for by deterministic noise ([4.90\%, 28.81\%], 95\% CI).} In addition, we find that both random and deterministic noise increase with horizon.
+This increase was larger for random noise (mean = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (mean = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon).
+But intriguingly, the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}).
 That is when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
@@
 	\added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).} In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
-Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 20.05\% of the overall variability in people's behavior.
+Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.
 
@@
 \replaced[id=siyu]{In our model, random noise is modeled as non-stimulus-driven variability, which may include truly random noise (intrinsic stochasticity in decision making) or stimulus irrelevant deterministic processes that we could not control in the lab.
-As a result,}{Well, } the remaining 79.95 \% of the decision noise could be random or it could be deterministic \replaced[id=siyu]{processes that is not bound to the stimulus and that we can not control}{because we can't control everything}.
+As a result,}{Well, } the remaining 85\% of the decision noise could be random or it could be deterministic \replaced[id=siyu]{processes that is not bound to the stimulus and that we can not control}{because we can't control everything}.
 In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
@@
 From both a conceptual and methodological perspective, our model provides a lower bound of deterministic noise and an upper bound of random noise.
-Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 79.95\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
+Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 85\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
 
-	\added[id=siyu]{This is in line with works of \citep{Bair1996} in which MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli.
+	\added[id=siyu]{The deterministic noise hypothesis is in line with works in which neural variability can be accounted for by fluctuations in sensory inputs.
+For example, }MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli \citep{Bair1996}.
 In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
-A more recent work showed that neural variability in the visual areas reflect statistics in the stimuli images \citep{Festa2021}.
 Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
-}
 
-	\added[id=siyu]{Other research are in line with random exploration \citep{drugowitsch16}.
-In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
-Although there is a distinction between suboptimal inference and random noise \citep{Pouget12}, they both show up as random noise in our definition.
+	\added[id=siyu]{The random noise hypothesis, on the other hand, is consistent with findings of \cite{drugowitsch16}.} In particular these authors show that randomness in behavior could arise from imperfections in mental inference, which happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
 This suggests that a large proportion of variability in behavior may arise from computational errors in computing the correct strategy.
-In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
-	}
+\added[id=siyu]{Although suboptimal inference is different from simply adding random noise to a neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not deterministically determined by the stimuli, it is a form of random noise in our definition.
+}In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
+	%SIYU says the song bird literature supports that behavioral variability is reflected in neural variability.
+But it can't tell whether the neural variability arises from deterministic processes of external stimuli or random noise.
+I didn't read/remember the songbird paper, this is only based on the text below.
+		%"The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}"
+
 	%this is consistent with findings of \cite{drugowitsch16}.
@@
 	\added[id=siyu]{
-	Regardless of whether the remaining 79.95 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
+	Regardless of whether the remaining 85\% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
 Specifically, a reduction in the strength with which reward drives the choice.
@@
 	\end{equation*}
-	\added[id=siyu]{
-	In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
+	\added[id=siyu]{In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
 From a model fitting perspective, these two forms are equivalent.
@@
 From a psychological perspective, however, they are quite different.
-An increase in both random and deterministic noise at a similar ratio, as observed in our experiment when horizon increases, suggests that either $\beta$ is reduced or $n_{det}'$ and $n_{ran}'$ are increased simultaneously at the same ratio.
-Neurally, reducing the $\beta$ corresponds to a reduction in reward sensitivity and reward coding, whereas simultaneously increasing $n_{det}'$ and $n_{ran}'$ corresponds to a reduction in some common noise filtering mechanism that reduces both random and deterministic noises.
+An increase in both random and deterministic noise at a similar ratio relative to the reward, as observed in our experiment when horizon increases, suggests that either $\beta$ is reduced or $n_{det}'$ and $n_{ran}'$ are increased simultaneously at the same ratio.
+Neurally, reducing the $\beta$ corresponds to a reduction in reward sensitivity and reward coding, whereas simultaneously increasing $n_{det}'$ and $n_{ran}'$ suggests a common noise gating mechanism in which the noise filtering neural circuit is inhibited so that both random and deterministic noises are amplified.
 }
+	%SIYU says the Ebitz paper sounds like choice signal is reduced during exploration, and p(switch|reward outcome) increases behaviorally in exploration state compared to exploitation?
+I am not sure it supported reward reduction?
+		%Such reduced coding of reward would be consistent with \cite{ebitz17}.%In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
 
@@
 	%On the other hand, our result suggests that at least one fifth of the decision noise in random exploration arises from a deterministic process.
-
-	%SIYU says the song bird literature supports that behavioral variability is reflected in neural variability.
-But it can't tell whether the neural variability arises from deterministic processes of external stimuli or random noise.
-I didn't read/remember the songbird paper, this is only based on the text below.
-			%"The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}"
-
-	%SIYU says the Ebitz paper sounds like choice signal is reduced during exploration, and p(switch|reward outcome) increases behaviorally in exploration state compared to exploitation?
-I am not sure it supported reward reduction?
-			%Such reduced coding of reward would be consistent with \cite{ebitz17}.%In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
 

### 2022-12-21 codes/EEplot_2noise_modelcomparison.m

@@ -13,29 +13,27 @@ function EEplot_2noise_modelcomparison(plt, gp, simugp)
     for mi = 1:6
         tgp = simugp.(['model' char(64+mi)]);
         plt.ax(mi,1);
-        plt.plot(1:2,gp.GPav_p_hi13, gp.GPste_p_hi13,'line','color','AZred');
         plt.plot(1:2,tgp.GPav_GPav_p_hi13, tgp.GPav_GPste_p_hi13,'line','color','AZblue');
-        plt.setfig_ax('legend',{'data','model'});
-
+        plt.plot(1:2,gp.GPav_p_hi13, gp.GPste_p_hi13,'line','color','AZred');
+        plt.setfig_ax('legend',{'model','data'});
         plt.ax(mi,2);
-        plt.plot(1:2,gp.GPav_p_lm22, gp.GPste_p_lm22,'line','color','AZred');
         plt.plot(1:2,tgp.GPav_GPav_p_lm22, tgp.GPav_GPste_p_lm22,'line','color','AZblue');
-        plt.setfig_ax('legend',{'data','model'});
-        
+        plt.plot(1:2,gp.GPav_p_lm22, gp.GPste_p_lm22,'line','color','AZred');
+        plt.setfig_ax('legend',{'model','data'});
         plt.ax(mi,3);
+        plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13_randomtheory_byR, tgp.GPav_GPste_p_inconsistent13_randomtheory_byR,'line', 'color', 'AZblue50','linestyle','--');
+        plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13, tgp.GPav_GPste_p_inconsistent13,'line', 'color', 'AZblue');
         plt.plot(1:2,gp.GPav_p_inconsistent13, gp.GPste_p_inconsistent13,'line', 'color', 'AZred');
 %         plt.plot(1:2,gp.GPav_p_inconsistent13_randomtheory_byR, gp.GPste_p_inconsistent13_randomtheory_byR,'line', 'color', 'AZred50','linestyle','--');
-        plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13, tgp.GPav_GPste_p_inconsistent13,'line', 'color', 'AZblue');
-        plt.plot(1:2,tgp.GPav_GPav_p_inconsistent13_randomtheory_byR, tgp.GPav_GPste_p_inconsistent13_randomtheory_byR,'line', 'color', 'AZblue50','linestyle','--');
[…]

### 2022-12-21 codes/EEplot_2noise_pinconsistent.m

@@ -8,23 +8,23 @@ function plt = EEplot_2noise_pinconsistent(plt, gp, ver13, ver22, strpfx)
     if ~exist('strpfx', 'var') || isempty(strpfx)
         strpfx = '';
     end
-    leg = {'behavioral data','deterministic noise only','random noise only'};
+    leg = {'random noise only','behavioral data','deterministic noise only'};
     plt.setfig('xlabel', {'horizon','horizon'}, 'ylabel', {'p(inconsistent)','p(inconsistent)'}, ...
         'title', {'[1 3] condition','[2 2] condition'}, 'legloc', {'SW','SW'}, ...
         'legend', {leg,leg}, ...
         'xlim', {[0.5 2.5],[0.5 2.5]}, 'xtick', {[1 2],[1 2]}, 'xticklabel', {[1 6],[1 6]});
+    plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent13_randomtheory' ver13]), ...
+        gp.([strpfx 'GPste_p_inconsistent13_randomtheory' ver13]),'line', 'color', 'AZred');
     plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent13']), gp.([strpfx 'GPste_p_inconsistent13']), 'line', ...
         'color', 'AZcactus');
     plt.plot(1:2,[0 0], [0 0],'line', 'color', 'AZblue');
-    plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent13_randomtheory' ver13]), ...
-        gp.([strpfx 'GPste_p_inconsistent13_randomtheory' ver13]),'line', 'color', 'AZred');
     plt.sigstar_y(gp.([strpfx 'GPav_p_inconsistent13']), 1:2, gp.([strpfx 'GPpvalue_p_inconsistent13']), 'right');
     plt.new;
+    plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent22_randomtheory' ver22]), ...
+        gp.([strpfx 'GPste_p_inconsistent22_randomtheory' ver22]),'line', 'color', 'AZred');
     plt.plot(1:2,gp.([strpfx 'GPav_p_inconsistent22']), gp.([strpfx 'GPste_p_inconsistent22']),'line', ...
         'color', 'AZsand');
     plt.plot(1:2,[0 0], [0 0],'line', 'color', 'AZblue');
[…]

### 2022-12-21 codes/EEplot_2noise_reduced.m

@@ -6,14 +6,14 @@ function plt = EEplot_2noise_reduced(plt, sp)
     ymax = 1;
     plt.setfig('ylim', {[0 ymax],[0 ymax],[0 0.7],[0 0.7]});
     plt.setfig_all('ytick', [0:0.2:ymax]);
-    color = {{'AZred50','AZred'},{'AZblue50','AZblue'},...
-        {'AZred50','AZred'},{'AZblue50','AZblue'}};
+    color = {{'AZred','AZred50'},{'AZblue','AZblue50'},...
+        {'AZred','AZred50'},{'AZblue','AZblue50'}};
     fns = {'NoiseRan', 'NoiseDet'};
     stepsize = 0.02;
     xbins = [-10:stepsize:30];
     plt.setfig(1:2, 'title', names);
     plt.setfig([-1 0] + length(sp)*2, 'xlabel', {'noise standard deviation [points]', 'noise standard deviation [points]'});
-    plt.setfig_all('legend', {'reduced model','full model'}, 'legloc', 'NE');
+    plt.setfig_all('legend', {'full model', 'reduced model'}, 'legloc', 'NE');
     for hi = 1:2
         for i = 1:2
             fn = fns{i};

### 2022-12-21 codes/FigureLoad_gridsimu.m

@@ -6,6 +6,10 @@ for ii = 1:length(rans)
     for j = 0:10
         W.print('%d,%d', i,j);
         tfilename = sprintf('HBI_DetRanNoise_simugrid_ran%d_det%d_samples.mat', i, j);
+        if ~exist(fullfile(griddir, tfilename), 'file')
+            W.print('file not exists: ran %d, det %d', i, j);
+            continue;
+        end
         tsp = load(fullfile(griddir, tfilename)).samples;
         randetsp{1}{ii,j+1} = mean(tsp.NoiseRan, 3);
         randetsp{2}{ii,j+1} = mean(tsp.NoiseDet, 3);

### 2022-12-21 codes/mainFigures_bayes.m

@@ -1,6 +1,6 @@
 %% bayesian plots
 plt = W_plt('savedir', '../figures', 'savepfx', 'RDBayes', 'isshow', true, ...
-    'issave', true);
+    'issave', true, 'extension',{'svg', 'jpg'});
 outputdir = '../bayesoutput/all';
 %% figure - posterior 
 suffix = {'','_all','_paironly'};
@@ -21,10 +21,27 @@ for gii = 1:3
 end
 %% reduced models
 sp = {};
-sp{2} = paramsub{1};
-sp{1} = load(fullfile(outputdir, ['HBI_DetRanNoise_dRonly_samples.mat'])).samples;
+sp{1} = paramsub{1};
+sp{2} = load(fullfile(outputdir, ['HBI_DetRanNoise_dRonly_samples.mat'])).samples;
 %
 EEplot_2noise_reduced(plt, sp);
+%% deterministic noise ratio
+sp = load(fullfile(outputdir, ['HBI_DetRanNoise_samples.mat'])).samples;
+%%
+det = []; ran = [];
+vardet = [];
+for i = 1:2
+    det(i,:) = reshape(sp.NoiseDet(:,:,i),1,[]);
+    ran(i,:) = reshape(sp.NoiseRan(:,:,i),1,[]);
+    vardet(i,:) = det(i,:).^2./(det(i,:).^2 + ran(i,:).^2);
+end
+ddet = reshape(sp.dNoiseDet,1,[]);
+dran = reshape(sp.dNoiseRan,1,[]);
+vard = (ddet.^2./(ddet.^2+dran.^2));
+W.print('explained variance: %.2f', mean(vardet, 'all')*100)
+W.print('explained variance 5%%: %.2f', quantile(reshape(vardet,1,[]),[0.025 0.975])*100)
+
+
 %% ratio of deterministic vs random noise
 %% ratio plot (h = 6/h = 1, for ran and det, A and b)
 sp = paramsub{1};

### 2022-12-21 codes/mainFigures_paramrecover_coverageanalysis.m

@@ -27,7 +27,7 @@ for repi = 1:200
 end
 %% hyper prior recovery
 plt = W_plt('savedir', '../figures', 'savepfx', 'RDBayes', 'isshow', true, ...
-    'issave', true);
+    'issave', true, 'extension',{'svg', 'jpg'});
 plt.figure(2,3, 'is_title', 1);
 names = ["NoiseRan", "NoiseDet"];
 xbins = -10:0.02:50;

### 2022-12-21 codes/mainFigures_paramrecover_hyperprior.m

@@ -16,7 +16,7 @@ W.save('./Temp/param_recovery.mat','recovered', recovered);
 load('./Temp/param_recovery.mat')
 %% hyper prior recovery
 plt = W_plt('savedir', '../figures', 'savepfx', 'RDBayes', 'isshow', true, ...
-    'issave', true);
+    'issave', true, 'extension',{'svg', 'jpg'});
 plt.figure(2,3, 'is_title', 1);
 names = ["NoiseRan", "NoiseDet"];
 xbins = -10:0.02:50;

### 2022-12-21 codes/main_1_behavior.m

@@ -26,7 +26,7 @@ gp{2} = W.analysis_1group(suball, true, additionalcompare);
 suffix = {'','_all'};
 %% behavioral plots
 plt = W_plt('savedir', '../figures', 'savepfx', 'RanDetNoise', 'isshow', true, ...
-    'issave', true);
+    'issave', true, 'extension',{'svg', 'jpg'});
 %%
 for gi = 1:2
     plt.set_pltsetting('savesfx', suffix{gi});
@@ -34,7 +34,7 @@ for gi = 1:2
     plt.figure(1,2,'is_title',1);
     plt.setfig(1:2, 'ylim', {[0, 0.3] + gi*0.05,[0.35 0.65]}, ...
         'ytick', {0.1:.1:0.4, 0.4:.1:0.6}, ...
-        'legend', {{'[2 2]','[1 3]'},{'[1 3]'}}, 'legloc',{'SE','SE'});
+        'legend', {{'[1 3]','[2 2]'},{'[1 3]'}}, 'legloc',{'SE','SE'});
     plt = EEplot_modelfree(plt, gp{gi},1,1);
     %% 2 noise figure
     plt.figure(1,2,'is_title',1);

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/GRAPHICALMODEL/EEHorizon_2sigma.tex
add older version

--- 2022-11-21 draft/2018 rejected - Nat Human behavior/GRAPHICALMODEL/Siyu_EEHorizon_2sigma_final.tex
@@
 \rput(6,8.5){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$b_{is}$}
-\rput(8,8.5){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma^n_{is}$}
-\rput(10,8.5){\GM@node[nodeSize=12mm]{epsilon}}\GM@label[offset=-2mm]{epsilon}{$\sigma^\epsilon_{is}$}
+\rput(8,8.5){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma^{ran}_{is}$}
+\rput(10,8.5){\GM@node[nodeSize=12mm]{epsilon}}\GM@label[offset=-2mm]{epsilon}{$\sigma^{det}_{is}$}
 
@@
 
-\rput(7.28,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^n_{i}$}
-\rput(8.35,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^n_{i}$}
+\rput(7.28,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^{ran}_{i}$}
+\rput(8.35,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^{ran}_{i}$}
 
-\rput(9.42,10){\GM@node[nodeSize=12mm]{kEpsilon}}\GM@label[offset=-2mm]{kEpsilon}{$k^\epsilon_{i}$}
-\rput(10.49,10){\GM@node[nodeSize=12mm]{lambdaEpsilon}}\GM@label[offset=-2mm]{lambdaEpsilon}{$\lambda^\epsilon_{i}$}
+\rput(9.42,10){\GM@node[nodeSize=12mm]{kEpsilon}}\GM@label[offset=-2mm]{kEpsilon}{$k^{det}_{i}$}
+\rput(10.49,10){\GM@node[nodeSize=12mm]{lambdaEpsilon}}\GM@label[offset=-2mm]{lambdaEpsilon}{$\lambda^{det}_{i}$}
 
-\rput(8,6){\GM@node[nodeSize=12mm]{sigmar}}\GM@label[offset=-4mm]{sigmar}{$n_{isgr}$}
-\rput(10,4.5){\GM@node[nodeSize=12mm]{epsilonr}}\GM@label[offset=-3mm]{epsilonr}{$\epsilon_{isg}$}
+\rput(8,6){\GM@node[nodeSize=12mm]{sigmar}}\GM@label[offset=-4mm]{sigmar}{$n^{ran}_{isgr}$}
+\rput(10,4.5){\GM@node[nodeSize=12mm]{epsilonr}}\GM@label[offset=-3mm]{epsilonr}{$n^{det}_{isg}$}
 
@@
 \rput(12.75,9.5){\pnode{mub}}\GM@label{mub}{$\mu^b_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^b_{i} \sim \mbox{Exponential}(0.01)$}
-\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^n_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^n_{i} \sim \mbox{Exponential}(10)$}
-\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^\epsilon_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^\epsilon_{i} \sim \mbox{Exponential}(10)$}
+\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^{ran}_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^{ran}_{i} \sim \mbox{Exponential}(10)$}
+\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^{det}_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^{det}_{i} \sim \mbox{Exponential}(10)$}
 
@@
 \rput(12.75,7){\pnode{b}}\GM@label{b}{$B_{is} \sim \mbox{Gaussian}(\mu^B_{i}, \sigma^B_{i})$}
-\rput(12.75,6.5){\pnode{sigma}}\GM@label{sigma}{$\sigma^n_{is} \sim \mbox{Gamma}(k^n_{i}, \lambda^n_{i})$}
-\rput(12.75,6){\pnode{epsilon}}\GM@label{epsilon}{$\sigma^\epsilon_{is} \sim \mbox{Gamma}(k^\epsilon_{i}, \lambda^\epsilon_{i})$}
+\rput(12.75,6.5){\pnode{sigma}}\GM@label{sigma}{$\sigma^{ran}_{is} \sim \mbox{Gamma}(k^{ran}_{i}, \lambda^{ran}_{i})$}
+\rput(12.75,6){\pnode{epsilon}}\GM@label{epsilon}{$\sigma^{det}_{is} \sim \mbox{Gamma}(k^{det}_{i}, \lambda^{det}_{i})$}
 
 % game level params
-\rput(12.75,5){\pnode{game}}\GM@label{game}{Stimulus driven noise for repeated game}
-\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$\epsilon_{isg} \sim \mbox{Logistic}(0, \sigma^\epsilon_{is})$}
+\rput(12.75,5){\pnode{game}}\GM@label{game}{Deterministic noise for repeated game}
+\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$n^{det}_{isg} \sim \mbox{Logistic}(0, \sigma^{det}_{is})$}
 
 % repeated game level params
-\rput(12.75,4){\pnode{repeat}}\GM@label{repeat}{Free noise for each game}
-\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n_{isgr} \sim \mbox{Logistic}(0, \sigma^n_{is})$}
+\rput(12.75,4){\pnode{repeat}}\GM@label{repeat}{Random noise for each game}
+\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n^{ran}_{isgr} \sim \mbox{Logistic}(0, \sigma^{ran}_{is})$}
 
@@
 \GM@label{q}{$\Delta Q_{isgr} \leftarrow
-\Delta R_{isg}+A_{is}\Delta I_{isg} + B_{is} + n_{isgr} + \epsilon_{isg}$}
+\Delta R_{isg}+A_{is}\Delta I_{isg} + b_{is} + n^{ran}_{isgr} + n^{det}_{isg}$}
 \rput(12.75,1.5){\pnode{c}}

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/GRAPHICALMODEL/OlderFiles/Siyu_EEHorizon_2sigma.tex
add older version

--- 2022-11-21 draft/2018 rejected - Nat Human behavior/GRAPHICALMODEL/Siyu_EEHorizon_2sigma_final.tex
@@
 \ifpdf
-\pdfpaperheight=10.5in
-\pdfpaperwidth=5.5in
+\pdfpaperheight=14in
+\pdfpaperwidth=9in
 \else
-\special{papersize=10.5in,5.5in}
+\special{papersize=14in,9in}
 \fi
@@
 
-\begin{pspicture}(4,2)(19,8)
+\begin{pspicture}(1,1)(18,11)
 %\showgrid
@@
 \rput(4,8.5){\GM@node[nodeSize=12mm]{A}}\GM@label[offset=-3mm]{A}{$A_{is}$}
-\rput(6,8.5){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$b_{is}$}
+\rput(6,8.5){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$B_{is}$}
 \rput(8,8.5){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma^n_{is}$}
@@
 
-\rput(5.14,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^b_{i}$}
-\rput(6.21,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^b_{i}$}
+\rput(5.14,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^B_{i}$}
+\rput(6.21,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^B_{i}$}
 
@@
 \rput(12.75,10.5){\pnode{priors}}\GM@label{priors}{Priors}
-\rput(12.75,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^A_{i} \sim \mbox{Exponential}(0.01)$}
-\rput(12.75,9.5){\pnode{mub}}\GM@label{mub}{$\mu^b_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^b_{i} \sim \mbox{Exponential}(0.01)$}
-\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^n_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^n_{i} \sim \mbox{Exponential}(10)$}
-\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^\epsilon_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^\epsilon_{i} \sim \mbox{Exponential}(10)$}
+\rput(12.75,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^A_{i} \sim \mbox{Exponential}(0.1)$}
+\rput(12.75,9.5){\pnode{mub}}\GM@label{mub}{$\mu^B_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^B_{i} \sim \mbox{Exponential}(0.1)$}
+\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^n_{i} \sim \mbox{Exponential}(0.1)$, $\lambda^n_{i} \sim \mbox{Exponential}(10)$}
+\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^\epsilon_{i} \sim \mbox{Exponential}(0.1)$, $\lambda^\epsilon_{i} \sim \mbox{Exponential}(10)$}
 
@@
 \rput(12.75,5){\pnode{game}}\GM@label{game}{Stimulus driven noise for repeated game}
-\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$\epsilon_{isg} \sim \mbox{Logistic}(0, \sigma^\epsilon_{is})$}
+\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$\epsilon_{isg} \sim \mbox{Gaussian}(0, \sigma^\epsilon_{is})$}
 
@@
 \rput(12.75,4){\pnode{repeat}}\GM@label{repeat}{Free noise for each game}
-\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n_{isgr} \sim \mbox{Logistic}(0, \sigma^n_{is})$}
+\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n_{isgr} \sim \mbox{Gaussian}(0, \sigma^n_{is})$}
 

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/draft/Older drafts/Bob#1/Internal-external noise Bob#1/Internal-External Noise.tex
add older version

--- 2022-11-21 draft/Older drafts/Siyu#final/Internal-external noise Siyu#final/Internal-External Noise.tex
@@
 From a theoretical perspective, making optimal explore-exploit choices, i.e.
-choices that maximize long-term reward, is incredibly hard \citep{eegittins74, eeBasu18}.
+choices that maximize long-term reward, is incredibly hard \citep{eegittins74} (CITE BELMAN).
 In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \citep{, eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}.
@@
 In this strategy, the decision process between exploration and exploitation is corrupted by `decision noise', meaning that high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11} ADDED TWO NEW PAPERS ON THOMPSON SAMPLING, NOT SURE WHETHER THEY ARE OKAY.
+In theory, such `random exploration,' is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eebridle90} CITE THOMPSON AND NEW PAPERS ON THOMPSON SAMPLING.
 
@@
 For instance, external, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than internal noise \citep{eeBrunton13}.
-Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures \citep{songbird2}.
+Conversely internal, neural noise is thought to drive exploratory singing behavior in song birds  \citep{songbird2} and the generation and control of this internal noise has been linked to specific neural structures CITES.
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
+			\includegraphics[width=\textwidth]{figures/taskfig.PNG}
 			\caption{A key additional manipulation here is repeated games.
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
-			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
+			\includegraphics[width=\textwidth]{figures/modelfree.png}
+			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions. TAKE OUT PANEL C }
 			\label{fig:modelfree}
@@
 Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (STATS), suggesting that not all of the noise was stimulus driven.
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
+In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (STATS ANOVA WITH HORIZON AND UNCERTAINTY CONDITIONS AS FACTORS), suggesting that at least some of the horizon dependent noise is internal.
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/theory_da_info.png}
+			\includegraphics[width=\textwidth]{figures/modelfree_intext.png}
 			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
@@
 	&= p(\mbox{low mean})^2 + (1-p(\mbox{low mean}))^2\\ 
-	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean})(1-p(\mbox{low mean}))
+	\mbox{hence},\quad p(\mbox{inconsistent}) &=  2 p(\mbox{low mean}) CHECK????
 	\end{split}
 	\end{equation*}
-	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(59) = 5.10, p $<$ 0.001 for horizon 1, t(59) = 5.65, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 3.49, p $<$ 0.001 for horizon 1, t(59) = 3.31, p $<$ 0.001 for horizon 6.
-Behavior is different from pure external noise prediction in both [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency.
+	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (STATS T-TEST RELATIVE TO BOTH), suggesting that both external and internal noise are present in driving this choice inconsistency.
 Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.
@@
 Consistent with our model-free results, we see that both internal and external noise variances are non-zero and that internal noise is about 2-3 times larger than the external noise.
-In addition, we find that internal noise increases dramatically with horizon (M = 4.55, p $<$ 0.001) whereas there is only a slight increase of external noise (M = 1.78, p = 0.009).
+In addition, we find that internal noise increases dramatically with horizon (STATS) whereas there is only a slight increase of external noise (STATS).
 Taken together, these results suggest that random exploration is dominated by internal noise.
@@
 
-	\begin{figure}[H]
+	\begin{figure}[h]
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/hyperprior.png}
-			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
+			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoiseh.png}
+			\caption[Posterior distributions over the group-level means of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise variance.
 Both internal and external noises are nonzero, and internal noise has a much greater magnitude.
-Posterior distributions over the group-level means of the change of external and internal noise standard deviation.
-Internal noise increases significantly with horizon, external noise increases with horizon as well.}
-			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
+COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'}
 			\label{fig:mb1}
@@
 
-%	\begin{figure}[H]
-%		\begin{center}
-%			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
-%			\caption[]{}			
-%			\label{fig:mb2}
-%		\end{center}
-%	\end{figure}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/dist_hypernoise.png}
+			\caption[Posterior distributions over the group-level means of the change of external and internal noise variance.]{Posterior distributions over the group-level means of the change of external and internal noise variance.
+Internal noise increases significantly with horizon, external noise increases with horizon as well.}
+			\label{fig:mb2}
+		\end{center}
+	\end{figure}
 
@@
 Likewise, our estimate of internal noise is likely an upper bound as these `missing' sources of external noise would be interpreted as internal noise in our model.
-Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb1}, where internal noise is 2-3 times the size of external noise.
+Despite this, it seems hard to imagine that these additional noise sources could be enough to account for the large differences between internal and external noise that we found in Figure \ref{fig:mb2}, where internal noise is 2-3 times the size of external noise.
 
@@
 	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating internal levels of noise.
-Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.
-In addition there is some evidence that NE plays a direct role in random exploration \citep{eeWarren17}, although this finding is complicated by other work showing no effect of NE drugs on exploration \citep{jepma2012, nieuwenhuis05}
+Indeed, manipulation of the NE system has been found to change behavioral variability in both humans and other animals in a variety of tasks (CITES KARPOVA, PAPER I JUST SENT YOU AND REFS WITHIN, ALSO JANE'S BIOARXIV PAPER).
+In addition there is some evidence that NE plays a direct role in random exploration (CITE WARREN ET AL.
+2017), although this finding is complicated by other work showing no effect of NE drugs on exploration CITE NIEWENHUIS AND JEPMA DRUG STUDY.
 
@@
 
-	80 participants (ages 18-25, 43 male, 37 female) from the University of Arizona undergraduate subject pool participated in the experiment.
+	80 participants (ages 18-XXX, XXX male, XXX female, XXX no response) from the University of Arizona undergraduate subject pool participated in the experiment.
 20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
 This left 60 for the main analysis.
-Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures \ref{fig:modelfree2}-\ref{fig:mb12})% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
+Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures XXX-YYY).
+XXX MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
 
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figall/line_modelfree.png}
-			\caption{Both directed and random exploration increase with horizon. Choice inconsistency also increases with horizon for both [1 3] and [2 2] conditions.}
-			\label{fig:modelfree2}
-		\end{center}
-	\end{figure}
-
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figall/theory_da_info.png}
-			\caption[Both external and internal noise contribute to the choice variability in random exploration]{Both external and internal noise contribute to the choice variability in random exploration.
-For both [1 3] and [2 2] condition, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only external noise exists where people should behave identically in repeated games.
-Also, there is a significant difference between people's behavior and predicted choice inconsistency assuming that only internal noise exists where people treat repeated games independently.
-}
-			\label{fig:mf22}
-		\end{center}
-	\end{figure}
-	\newpage
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figall/hyperprior.png}
-			\caption[Posterior distributions over the group-level noise standard deviation of the external and internal noise variance.]{Posterior distributions over the group-level means of the external and internal noise standard deviation.
-Both internal and external noises are nonzero, and internal noise has a much greater magnitude.
-Posterior distributions over the group-level means of the change of external and internal noise standard deviation.
-Internal noise increases significantly with horizon, external noise increases with horizon as well.}
-			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
-			\label{fig:mb12}
-		\end{center}
-	\end{figure}
 	%No differences of task-related behavior (accuracy, directed-exploration, random-exploration and choice inconsistency) were found between ages, genders, races and ethnicities.
@@
 Thus, the better participants performed, the sooner they got to leave the experiment.
-On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 12.34 and 32.74 minutes (mean 23.31 minutes).%  IS THIS NUMBER CORRECT?
+On average, participants played 151.6 games (minimum = 90 games, maximum = 192 games) and the whole task lasted between 10.6 and 29.2 minutes (mean 19.7 minutes).
+IS THIS NUMBER CORRECT?
 I THOUGHT IT WAS MORE LIKE AN HOUR?%(\ref{appendixb})
-	THIS IS ONLY THE TASK TIME, NOT INCLUDING BASELINE PUPIL OR THE TWO CURIOSITY SURVEY OR RA SETUP TIME
 
@@
 The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.
-(See Figure \ref{fig:taskfigb})
+(See Figure \ref{fig:taskfig1})
 
@@
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
+			\includegraphics[width=\textwidth]{figures/taskfig1.PNG}
 			\caption{Time line of a horizon 6 game.
@@
 After these forced trials they are free to choose between the two options for the rest of the game (6 trials in this case).
-			%COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
-			%CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
-			I WILL SEND YOU THE POWERPOINTS TOO}
-			\label{fig:taskfigb}
+			COMBINE THIS WITH FIGURE 7 FOR TASK CONDITIONS
+			CAN YOU MAKE THESE FIGURES LOOK NICER I.E. MORE LIKE MINE?  REDUCE LINE WIDTHS ETC ... IF NOT SEND ME POWER POINTS AND I WILL MAKE INTO PDFS}
+			\label{fig:taskfig1}
 		\end{center}
@@
 After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
-(See Figure \ref{fig:taskfigb})
+(See Figure \ref{fig:taskfig2})
 
-%	\begin{figure}[H]
-%		\begin{center}
-%			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
-%			\caption{Task conditions}
-%			\label{fig:taskfig2}
-%		\end{center}
-%	\end{figure}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/taskfig2.PNG}
+			\caption{Task conditions}
+			\label{fig:taskfig2}
+		\end{center}
+	\end{figure}
 
@@
 				$\mu_i^{A} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{A} \sim $ Exponential(0.01)}		\\
+				$\sigma_i^{A} \sim $ Exponential(0.1)}		\\
 			\hline
-			spatial bias, $b_{is}$ 
-			& $b_{is} \sim $  Gaussian($\mu_i^{b}$, $\sigma_i^{b}$) 
-			& $\theta_{i}^{b} = (\mu_i^{b}, \sigma_i^{b}) $
+			spatial bias, $B_{is}$ 
+			& $B_{is} \sim $  Gaussian($\mu_i^{B}$, $\sigma_i^{B}$) 
+			& $\theta_{i}^{B} = (\mu_i^{B}, \sigma_i^{B}) $
 			& \specialcell{
-				$\mu_i^{b} \sim $ Gaussian( 0, 100 ) \\ 
-				$\sigma_i^{b} \sim $ Exponential(0.01)}		\\
+				$\mu_i^{B} \sim $ Gaussian( 0, 100 ) \\ 
+				$\sigma_i^{B} \sim $ Exponential(0.1)}		\\
 			\hline
-			deviation of external noise, $\sigma^{ext}_{isg}$ 
+			deviation of external noise, $\epsilon_{isg}$ 
 			& $\sigma_{i} \sim $  Gamma($k_i^{ext}$, $\lambda_{i}^{ext}$) 
@@
 			& \specialcell{
-				$k_i^{ext} \sim $ Exponential(0.01) \\ 
+				$k_i^{ext} \sim $ Gaussian(0.1) \\ 
 				$\lambda_i^{ext} \sim $ Exponential(10)}		\\
 			\hline
-			deviation of internal noise, $\sigma^{int}_{isgr}$ 
+			deviation of internal noise, $\sigma_{isgr}$ 
 			& $\sigma_{is} \sim $  Gamma($k_i^{int}$, $\lambda_{i}^{int}$) 
@@
 			& \specialcell{
-				$k_i^{int} \sim $ Exponential(0.01) \\ 
+				$k_i^{int} \sim $ Gaussian(0.1) \\ 
 				$\lambda_i^{int} \sim $ Exponential(10)}		\\
@@
 		\end{tabular}
-		\caption{Model parameters, priors, hyperparameters and hyperpriors. }%CHECK UPDATE}
+		\caption{Model parameters, priors, hyperparameters and hyperpriors. CHECK UPDATE}
 		\label{tab:pars2}	
@@
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
+			\includegraphics[width=\textwidth]{figures/hbm.pdf}
 			\caption{Hierarchical Bayesian model}
@@
 	Results of this parameter recovery procedure are shown in Figure \ref{fig:pararecover1}.
-As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C, D).
+As is clear from this figure, parameter recovery is good for all parameters apart from the bias, which is likely due to this parameter being so close to zero (Figure XXX PANEL NUMBER).
 The recovery for the noise parameters, $\sigma_{ext}$ and $\sigma_{int}$, is slightly better for horizon 1 than horizon 6.
@@
 
-	\begin{figure}[H]
+	\begin{figure}[h]
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/precover.png}
-			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) %for horizon 1 games
-			%COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.
+			\includegraphics[width=1\textwidth]{figures/paramrecover_1.png}
+			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 1 games
+			COMBINE THESE TWO, HAVE TWO COLUMNS, LEFT IS HORIZON 1, RIGHT IS HORIZON 6.
 REDUCE NUMBER OF SIGNIFICANT FIGURES IN p VALUES.
@@
 RED LINES CHANGE TO DOTTED BLACK LINES.
-			%XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
-			%USE ADDABCS FUNCTION TO ADD ABCs!
+			XLABELS - SIMULATED $A$, YLABELS FIT $A$, I.E. MAKE SPECIFIC FOR EACH PARAMETER
+			USE ADDABCS FUNCTION TO ADD ABCs!
 			}
@@
 
-	THIS PLOT CHANGES EVERY TIME I RUN IT. THE $R^2$ VALUE FOR INFO BONUS AND INTERNAL NOISE ARE SIMILAR ALL THE TIME, BUT FOR EXTERNAL NOISE IT RANGES FROM 0.2 TO 0.7, FOR BIAS, FROM 0 TO 0.3
-	%\begin{figure}[H]-
-	%	\begin{center}
-	%		\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
-	%		\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
-	%		\label{fig:pararecover2}
-	%	\end{center}
-	%\end{figure}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/paramrecover_2.png}
+			\caption{Parameter recovery over the subject-level means of information bonus($A$), spatial bias($b$), internal noise variance($\sigma_{int}$) and external noise variance($\sigma_{ext}$) for horizon 6 games}
+			\label{fig:pararecover2}
+		\end{center}
+	\end{figure}
 	%References

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/draft/Older drafts/Internal-External Noise/Internal-External Noise.tex
add older version

--- 2022-12-12 manuscript/Older drafts/Backup2/Internal-External Noise/Internal-External Noise.tex
@@
 			\includegraphics[width=\textwidth]{figures/taskfiga.pdf}
-			\caption{BOB: ADD A PANEL A ABOVE THIS PANEL, SHOWING THE DYNAMICS FOR ONE GAME - LIKE IN HASHEM'S PAPER.  
+			\caption{
+			%BOB: ADD A PANEL A ABOVE THIS PANEL, SHOWING THE DYNAMICS FOR ONE GAME - LIKE IN HASHEM'S PAPER.  
 			Schematic of the experiment.
@@
 	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
-This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(59) = 6.17, p $<$ 0.001 for [1 3], t(59) = 7.26, p $<$ 0.001 for [2 2]).
+This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(65) = 6.60, p $<$ 0.001 for [1 3], t(65) = 8.14, p $<$ 0.001 for [2 2]).
 Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info}$ in the unequal, or [1 3], condition.
-Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(59)=6.88, p $<$ 0.001).
+Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(65) = 7.00, p $<$ 0.001).
 
@@
 			\includegraphics[width=\textwidth]{figures/line_modelfree.png}
-			\caption{BOB: DO THE YTICKS ON THE 0.1S NOT 0.15S!  MORE WHITE SPACE BETWEEN PANELS A AND B.
+			\caption{
+			%BOB: DO THE YTICKS ON THE 0.1S NOT 0.15S!  MORE WHITE SPACE BETWEEN PANELS A AND B.
 			Replication of previous findings.
@@
 	To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (For [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; For [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that not all of the noise was stimulus driven.
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 236) = 37.98 BOB DID YOU FIX THIS?
-DIDNT USE REPEATED ANOVA, WILL CHANGE, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
+Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions (For [1 3] condition, t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6; For [2 2] condition, t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6.), suggesting that not all of the noise was stimulus driven.
+In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (F(1, 196) = 61.19, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is internal.
 
@@
 	\end{equation*}
-	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(59) = 5.10, p $<$ 0.001 for horizon 1, t(59) = 5.65, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 3.49, p $<$ 0.001 for horizon 1, t(59) = 3.31, p $<$ 0.001 for horizon 6.
-Behavior is different from pure external noise prediction in both [1 3] condition, t(59) = 11.78, p $<$ 0.001 for horizon 1, t(59) = 15.41, p $<$ 0.001 for horizon 6; and [2 2] condition, t(59) = 8.76, p $<$ 0.001 for horizon 1, t(59) = 17.02, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency.
+	As shown in Figure  \ref{fig:mf2}, people's behavior falls in between the pure external noise prediction and the pure internal noise prediction (Behavior is different from pure internal noise prediction in both [1 3] condition, t(65) = 5.60, p $<$ 0.001 for horizon 1, t(65) = 5.62 p $<$ 0.001 for horizon 6; and [2 2] condition, t(65) = 3.86, p $<$ 0.001 for horizon 1, t(65) = 3.42, p $<$ 0.001 for horizon 6.
+Behavior is different from pure external noise prediction in both [1 3] condition, t(65) = 12.92, p $<$ 0.001 for horizon 1, t(65) = 16.76, p $<$ 0.001 for horizon 6; and [2 2] condition, t(65) = 9.67, p $<$ 0.001 for horizon 1, t(65) = 17.74, p $<$ 0.001 for horizon 6.), suggesting that both external and internal noise are present in driving this choice inconsistency.
 Since choice inconsistency only reflects internal noise, Figure \ref{fig:mf2} suggests that internal noise increases with horizon.
@@
 	80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment.
-20 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-This left 60 for the main analysis.
-Note that including the 20 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
+14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
+This left 66 for the main analysis.
+Note that including the 14 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
 
@@
 The task was programmed using Psychtoolbox in MATLAB \citep{psychtoolbox1, psychtoolbox2}.
-(See Figure \ref{fig:taskfigb})
+(See Figure \ref{fig:taskfig})
 
@@
 The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfigb}.
+After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfig}.
 
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=\textwidth]{figures/taskfigb.pdf}
-			\caption{BOB: DONT NEED THIS FIGURE IF YOU UPDATE EARLIER TASKFIGURE.
+%	\begin{figure}[H]
+%		\begin{center}
+%			\includegraphics[width=\textwidth]{figures/taskfigb_final.pdf}
+%			\caption{BOB: DONT NEED THIS FIGURE IF YOU UPDATE EARLIER TASKFIGURE.
 Time line of a horizon 6 game.
@@
 			%I WILL SEND YOU THE POWERPOINTS TOO
-			}
-			\label{fig:taskfigb}
-		\end{center}
-	\end{figure}
+%			}
+%			\label{fig:taskfigb}
+%		\end{center}
+%	\end{figure}
 %	\begin{figure}[H]
@@
 			\includegraphics[width=1\textwidth]{figures/Siyu_EEHorizon_2sigma_final.pdf}
-			\caption{Schematic of the hierarchical Bayesian model using notation of BOB CITE WAGNERMAKERS AND LEE.}
+    			\caption{Schematic of the hierarchical Bayesian model using notation of \cite{lee2014}}
+			%BOB CITE WAGNERMAKERS AND LEE.}
 			\label{fig:model}

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEplot_2noise_hyperprior_6modelori.m

@@ -1,0 +1,71 @@
+function EEplot_2noise_hyperprior_6modelori(plt, sp)
+    plt.figure(6,4);
+    plt.setup_pltparams('fontsize_face', 20, 'fontsize_leg', 10, 'linewidth', 2);
+    names = {'Random noise','\Delta Random noise',...
+        'Deterministic noise','\Delta Deterministic noise'};
+    plt.setfig('ytick', []);
+    color = {{'AZred50','AZred'},{'black'},{'AZblue50','AZblue'},{'black'}};
+    fns = {'NoiseRan', 'NoiseDet'};
+    legs = {{'h = 1','h = 6'},{'change'},{'h = 1','h = 6'},{'change'}};
+    ylabs = {'density','','',''};
+    stepsize = 0.02;
+    xbins = [-10:stepsize:30];
+    plt.param_fig.colordelete = 1;
+    plt.setfig(21:24, 'xlabel', {'\sigma_{ran}','\Delta \sigma_{ran}','\sigma_{det}','\Delta \sigma_{det}'});
+    plt.setfig(1:4, 'title', names);
+    fns = {'NoiseRan', 'NoiseDet'};
+    for mi = 1:6
+        for i = 1:length(fns)
+            fn = fns{i};
+            plt.new;
+            tl = [];
+
+            if ~isfield(sp{mi}, fn)
+                tl(1,:) = zeros(1,length(xbins));
+                tl(1,dsearchn(xbins', 0)) = 1;
+                tl(2,:) = tl(1,:); 
+            else
+                sz = size(sp{mi}.(fn));
+                if sz(end) == 2
+                    for hi = 1:2
+                        td = sp{mi}.(fn);
+                        td = squeeze(td(:,:,hi));
+                        td = reshape(td, 1, []);
+                        tl(hi,:) = hist(td, xbins)/(length(td)*stepsize);
+                    end
+                else
+                    td = sp{mi}.(fn);
+                    td = squeeze(td(:,:));
+                    td = reshape(td, 1, []);
+                    tl(1,:) = hist(td, xbins)/(length(td)*stepsize);
+                    tl(2,:) = tl(1,:);
+                end
+            end
+            ymax = max(tl* 1.05,[],'all');
+            plt.setfig_ax('ylim', [0 ymax], 'xtick', 0:4:20,'xlim', [0 20], 'color', color{i*2 -1}, ...
+                'legend', legs{i*2 - 1}, 'ylabel', ylabs{i*2-1});
[…]

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEplot_2noise_modelcomparison.m

@@ -1,0 +1,43 @@
+function EEplot_2noise_modelcomparison(plt, gp, simugp)
+    plt.figure(6,4);
+    plt.setup_pltparams('fontsize_face', 20, 'fontsize_leg', 10, 'linewidth', 2);
+    
+    plt.setfig('xlabel', [repmat({''},1,20), repmat({'horizon'},1,4)], ...
+        'ylabel', [{'model A'} repmat({''},1,3), {'model B'} repmat({''},1,3), ...
+        {'model C'} repmat({''},1,3), {'model D'} repmat({''},1,3), ...
+        {'model E'} repmat({''},1,3), {'model F'} repmat({''},1,3)], ...
+        'title', [{'p(high info)', 'p(low mean)', 'p(inconsistent), [1 3]', 'p(inconsistent), [2 2]', ...
+        } repmat({''}, 1, 20)], 'xlim', [0.5 2.5], 'xtick', [1 2], 'xticklabel', [1 6], ...
+        'legend', [repmat({{'',''}},1,23), {{'data', 'pure ran', 'model'}}], ...
+        'legloc', 'SouthEast', ...
+        'ylim', repmat({[0.35 0.65],[0 0.35],[0 0.45],[0 0.45]},1,6), ...
+        'ytick', repmat({[0.4:0.1:0.6],[0:.1:0.3],[0:.1:0.5],[0:.1:0.5]},1,6));
+    for mi = 1:6
+        tgp = simugp.(['model' char(64+mi)]);
+        plt.new;
+        plt.setfig_ax('color', {'AZred', 'AZblue'});
+        plt.lineplot(gp.av_p_hi13, gp.ste_p_hi13);
+        plt.lineplot(tgp.av_p_hi13, tgp.ste_p_hi13);
+        
+        plt.new;
+        plt.setfig_ax('color', {'AZred', 'AZblue'});
+        plt.lineplot(gp.av_p_lm22, gp.ste_p_lm22);
+        plt.lineplot(tgp.av_p_lm22, tgp.ste_p_lm22);
+        
+        plt.new;
+        plt.setfig_ax('color', {'AZred', 'AZred50', 'AZblue50', 'AZblue'});
+        plt.lineplot(gp.av_p_inconsistent13, gp.ste_p_inconsistent13);
+        plt.lineplot(gp.av_p_inconsistent13_randomtheory, gp.ste_p_inconsistent13_randomtheory, [], '--');
[…]

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/mainALL_ANALYSIS_FIGURE_RanDetNoise.m

@@ -1,0 +1,60 @@
+%% main analysis 2016
+rawdatadir = '../data';
+savedir = '../data/all';
+files = dir(fullfile(rawdatadir,'Imported_EE_blinkCuriosity_2016S.csv'));
+filename = fullfile(files.folder, files.name);
+%% basic analysis
+opt_sub = {'subjectID', 'date', 'time'}; 
+opt_preprocess = 'EEpreprocess_game_basic';
+opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};
+opt_analysis = {'EEanalysis_sub_basic'};
+[sub0, game, idxsub] = W_sub.analysis_pipeline(filename, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, savedir);
+W.writetable(game, fullfile(savedir,'data_all.csv'));
+%% exclude participants
+sub0 = readtable('../data/all/SUB_EE_blinkCuriosity_2016S.csv', 'Delimiter', ',');
+load('../data/all/idxsub_EE_blinkCuriosity_2016S.mat');
+[suball, idall] = W_sub.excludesubject(sub0, 'age', [18 NaN]);
+[sub, id] = W_sub.excludesubject(sub0, 'age', [18 NaN], 'pvalue_ac', [NaN 0.05]);
+save(fullfile(savedir, 'idxsub_exclude.mat'), 'id', 'idall','idxsub');
+%% analyze group
+clear gp;
+additionalcompare = {{'p_inconsistent13'},{'p_inconsistent22'},...
+    {'p_inconsistent13','p_inconsistent13_randomtheory'},{'p_inconsistent22','p_inconsistent22_randomtheory'}};
+gp{1} = W_sub.analysis_1group(sub, true, additionalcompare);
+gp{2} = W_sub.analysis_1group(suball, true, additionalcompare);
+suffix = {'','_all'};
+%%
+plt = W_plt('isshow', 1);
+for gi = 1:2
+%% model-free plots
+    plt.setup_W_plt('fig_dir', '../figures/all','fig_suffix', suffix{gi},'fig_projectname', 'RanDetNoise');
+    plt.param_fig.islocked = true;
+    plt.setfig(2, 'color', {{'AZsand','AZcactus'}, {'AZcactus'}}, ...
+        'ylim', {[0, 0.3] + gi*0.05,[0.35 0.65]}, ...
+        'ytick', {0.1:.1:0.4, 0.4:.1:0.6});
+    EEplot_modelfree(plt, gp{gi},1,1);
[…]

### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/main_bayes_40chains.m

@@ -1,0 +1,198 @@
+%% get version
+ver = 'all';
+versave = 'all_50chains';
+% ver = 'paironly';
+%%
+cd(W.funcout('fileparts', 1, matlab.desktop.editor.getActiveFilename));
+outputdir = fullfile('../bayesoutput', versave);
+outputdir = GetFullPath(outputdir);
+wj = W_JAGS;
+%% set up chains
+nchains = 50; % How Many Chains?
+nburnin = 2000; % How Many Burn-in Samples?
+nsamples = 1000; % How Many Recorded Samples?
+HBItest = '';
+wj.setup_params(nchains, nburnin, nsamples);
+wj.openpool;
+%% test
+% HBItest = 'test_';
+% wj.setup_params;
+%%
+nchains = wj.bayes_params.nchains;
+%% main bayes
+files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
+suffix = {'','_all'};
+for fi = 1:2
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise' suffix{fi}]);
+    wj.run;
+end
+%% 6 model comparison (rest of the 5 models)
+data = load(fullfile(fullfile('../data',ver), 'bayes_2noise.mat')).(['bayes_data']);
+for mi = 2:6
+    data.modelname = ['2noisemodel' char(64 + mi)];
+    wj.setup_data_dir(data, outputdir);
+    wjinfo = EEbayes_analysis(data, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise_' char(64 + mi)]);
+    wj.run;
+end
+%% reduced model
+files = {'bayes_2noise_dRonly.mat', 'bayes_2noise_0model.mat', 'bayes_2noise_dIonly.mat'};
+suffix = {'_dRonly','_0model','_dIonly'};
+for fi = 1:1 % just the first one
+    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
[…]

### 2022-12-22 OLD/oldversions/Internal-External%20Noise/Internal-External Noise.tex
old version

--- 2022-12-12 manuscript/2022 draft - Plos Comp Bio/OlderVersion/08-14-20/Internal-External Noise.tex
@@
 	In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
-Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 27.05\% of the overall variability in people's behavior.
+Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for at least 27\% of the overall variability in people's behavior.
 %and increased with horizon - a hallmark of an exploratory process.
@@
 	So where does this leave randomness in random exploration?
-Well, the remaining 72.95 \% of the noise could be random or it could be deterministic because we can't control everything.
+Is the remaining 73 \% of the noise that we cannot attribute to deterministic noise really random?
+Or is this noise also deterministic, perhaps driven by factors we could not control in the lab?
 In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
-the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were scratching their nose.
+the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were fidgeting or scratching their nose CITE ANNE CHURCHLAND FIGETING (https://escholarship.org/content/qt1v23n1j7/qt1v23n1j7.pdf).
 Thus conceptually, our estimate of deterministic noise is a lower bound.
 Conversely, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
-In addition, our estimation method also systematically underestimate deterministic noise (part of the deterministic noise will show up as random noise in model fitting, see Figure \ref{fig:pararecover2}), so methodologically our model also provides a lower bound of deterministic noise and an upper bound on random noise.
-Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 72.95\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
+Clearly, future work should try to control for these additional sources of deterministic noise as much as possible, for example by tracking people's pose, body movements and eye movements during the experiment.
+
+	In addition to this conceptual limitation in measuring deterministic noise, parametere recovery simulations of our own experiment  suggest that our estimation method also systematically underestimate deterministic noise in this task (BY A FACTOR OF WHAT) (part of the deterministic noise will show up as random noise in model fitting, see Figure \ref{fig:pararecover2}).
+
+	Thus, both conceptually and methodologically our model provides a lower bound of deterministic noise and an upper bound on random noise.
+Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 73\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
 			%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. 
-
-	Randomness in exploratory behavior in random exploration can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).
+CUT FROM HERE
+    If random exploration is random 
+	Apparent randomness in random exploration can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).
 The idea that random exploration is driven by truly stochastic processes in the brain is in line with work in the bird song literature in which variability during song learning has been tied to neural variability arising from specific areas of the brain \citep{songbird1, songbird2}.
 In a recent report from \cite{ebitz17}, the behavioral variability of monkeys in an `explore' state was also tied to random rather than deterministic sources of noise.
-In our study, most of the decision noise seen can still be random noise (up to 72.95\%), and this is consistent with findings of \cite{drugowitsch16}.
+In our study, most of the decision noise seen can still be random noise (up to 73\%), and this is consistent with findings of \cite{drugowitsch16}.
 In particular these authors show that randomness in behavior arises from imperfections in mental inference, that happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
@@
 Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+END CUT HERE
 
@@
 
-	Regardless of whether the remaining 72.95 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration.
+	Regardless of whether the remaining 73 \% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other suggests a possible mechanism for random exploration.
 Specifically, a reduction in the strength with which reward drives the choice.
-In particular, in our analysis we excluded a scalar on $\Delta R$, if we include that then we get
+
+	To see this, we need to return to our equation for the computation of value
+		\begin{equation}
+	    \Delta Q = \Delta R + A \Delta I + b + n_{det} + n_{ran}
+	\end{equation}
+	In this version of the equation, the difference in value is a weighted of sum of reward, information, bias, deterministic noise,  and random noise.
+Crucially, we assumed that the weighting given to the difference in reward is 1 --- that is the multiplier on $\Delta R$ is 1, unlike the variable $A$ which modulates the effect of information.
+
+	A more general form of this equation would also include a scalar on $\Delta R$
 
@@
 	\end{equation}
+	In this equation, $\beta$ determines the weight given to the reward in the decision.
+From a model fitting perspective, these two forms are EQUIVALENT.
+From a psychological perspective they are quite different.
+And you can change the relative contribution of noise by changing $\beta$.
 

### 2022-12-22 OLD/oldversions/supplementary/Internal-External Noise.tex
old version

--- 2022-12-12 manuscript/2022 draft - Plos Comp Bio/Supplementary/DetRanNoise_Supplementary.tex
@@
 
-       \maketitle
+	\newpage
 
-	\newpage
 	\begin{figure}[H]
@@
 			\label{fig:mb12}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/precover.png}
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
+			}
+			\label{fig:pararecover1}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RanDetNoise_pure_ran_det_simu_hyperprior.png}
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise variance, $\sigma_{ran}$, and deterministic noise variance, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
+			}
+			\label{fig:pararecover1}
 		\end{center}
@@
 	\end{figure}
+
 	%\bibliographystyle{plainnat}

### 2023-12-21 manuscript/2022 unused - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
u

--- 2022-12-21 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
@@
 This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
-\added[id=siyu]{Our work takes us one step closer to understanding of the nature of decision noise in random exploration.}
+\added[id=siyu]{Moreover, our results suggest that both deterministic noise and random noise change proportionally to each other as the planning horizon changes.
+This sheds light on a common noise gating mechanism that is at play in random exploration.
+Our work takes us one step closer to understanding of the nature of decision noise in explore-exploit decisions.}
 

### 2023-12-21 manuscript/Older drafts/2023 unused - overleaf/manuscript_DeterministicRandomNoise.tex
u

--- 2022-12-21 manuscript/2022 draft - Plos Comp Bio/manuscript_DeterministicRandomNoise.tex
@@
 \definechangesauthor[name={Siyu Wang}, color=red]{siyu}
+\definechangesauthor[name={Bob}, color=blue]{bob}
 
@@
 
-\title{\replaced[id=siyu]{Separating random and deterministic sources of computational noises in explore-exploit decisions}{The nature of decision noise in random exploration}}
+\title{\replaced[id=siyu]{Separating random and deterministic sources of computational noise in explore-exploit decisions}{The nature of decision noise in random exploration}}
 \author[1,\textcurrency]{Siyu Wang}
@@
 Human decision making is inherently variable.
-While this variability is often seen as a sign of suboptimality behavior, recent work suggests that variability can actually be adaptive.
+While this variability is often seen as a sign of suboptimal behavior, recent work suggests that variability can actually be adaptive.
 An example arises when we must choose between exploring unknown options or exploiting options we know well.
-A little randomness in these `explore-exploit' decisions is remarkably effective as it encourages us to explore options we might otherwise ignore.
-Recent work suggests that people may actually use such `random exploration' in practice, increasing their behavioral variability when it is more valuable to explore.
-%Other work, however, suggests that variability in human behavior is not always random, and is instead driven by deterministic factors (such as repeating past actions) 
-A key question is whether the variability in random exploration is actually random.
-That is, is random exploration driven by stochastic processes in the brain or by some unobserved deterministic process that we have failed to account for when measuring behavioral variability?
+\added[id=bob]{A little randomness in these `explore-exploit' decisions is remarkably effective as it can encourage us to explore options we might otherwise ignore.
+In line with this idea, several studies have found evidence that people increase their behavioral variability when it is valuable to explore.
+A key question, however, is whether this variability in so-called `random exploration' is actually random.} That is, is random exploration driven by stochastic processes in the brain or by some unobserved deterministic process that we have failed to account for when measuring behavioral variability?
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-\added[id=siyu]{We built a novel computational model which could detect and separate previously unobserved deterministic processes from random noise, without having to make a priori assumptions about the nature of the deterministic process.
-} In particular, we find evidence that around 15$\%$ of the variability in random exploration can be accounted for by deterministic processing of the stimulus.
-This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
-\added[id=siyu]{Our work takes us one step closer to understanding of the nature of decision noise in random exploration.}
+\added[id=bob]{By modeling behavior in this task, we were able to estimate a lower bound on the amount of variability that is deterministically driven by the stimulus and an upper bound on the amount of variability that is random.} Using this approach, we find evidence that \added[id=siyu]{at least 15$\%$} of the variability in random exploration can be accounted for by deterministic processing of the stimulus.
+\added[id=bob]{Conversely, this suggests that up to 85$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.}
+% This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
+\added[id=bob]{Finally, our results suggest that both deterministic and random sources of variability change proportionally to each other as the value of exploration increases, suggesting that a common noise gating mechanism may be at play in random exploration.}
 
@@
 
-	Imagine trying to decide where to go to dinner on a date, you can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try a new restaurant that you know nothing about.
+	Imagine trying to decide where to go to dinner on a date.
+You can go to your favorite restaurant, the one you both really enjoy and always go to, or you can try a new restaurant that you know nothing about.
 Such decisions, in which we must choose between a well-known `exploit' option and a lesser known `explore' option, are known as explore-exploit decisions.
@@
 choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}.
-In part because of this difficulty, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \added[id=siyu]{\citep{Mehlhorn15, SCHULZ20197,WILSON202149}}.
+In part because of this computational complexity, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \added[id=siyu]{\citep{Mehlhorn15, SCHULZ20197,WILSON202149}}.
 %\citep{eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}
@@
 In this strategy, \deleted[id=siyu]{the decision process between exploration and exploitation is corrupted by `decision noise', meaning that }high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-\added[id=siyu]{In modeling terms, random exploration works by adding 'decision noise' to the value of the options such that sub-optimal exploratory options can sometimes have a higher total score (i.e., value + noise) than the exploit option and get chosen.} \replaced[id=siyu]{Such}{In theory, such} random exploration, is surprisingly effective and, if implemented correctly, can come close to optimal performance \added[id=siyu]{in theory}\citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
+\added[id=siyu]{In modeling terms, random exploration works by adding `decision noise' to the value of the options such that sub-optimal exploratory options can sometimes have a higher total score (i.e., value + noise) than the exploit option and get chosen.} \replaced[id=siyu]{Such}{In theory, such} random exploration, is surprisingly effective and, if implemented correctly, can come close to optimal performance \added[id=siyu]{in some cases}\citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
 
-	It has recently been shown that humans appear to use random exploration and can increase such decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014}.
-In one of these tasks, known as the Horizon Task, the key manipulation is the horizon condition, i.e.
+	It has recently been shown that humans appear to use random exploration and can increase  decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014} \added[id=bob]{ALSO CITE FINDLING ET AL.}.
+In one of these tasks, known as the Horizon Task \citep{wilson2014}, the key manipulation is the horizon condition, i.e.
 the number of decisions remaining for the participant to make.
 Increasing the horizon makes exploration more valuable as there is more time to use the information gained by exploration to maximize future rewards.
-For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you would be more likely to explore the new restaurant.
+For example, if you are leaving town tomorrow (short horizon), you will probably exploit the restaurant you know and love, but if you are in town for a while (long horizon), you will be more likely to explore the new restaurant.
 Using such a horizon manipulation it has been shown that people's behavior is more variable in long horizons than short horizons, suggesting that they use adaptive decision noise to solve the explore-exploit dilemma \citep{wilson2014}.
 
-	\replaced[id=siyu]{One limitation, however, is that it is difficult to tell whether what we have called 'decision noise' in previous research is truly a random noise.}{One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random.} That is, whether behavioral variability is due to intrinsic stochastic processes in the brain or whether it is due to deterministic processes that we \replaced[id=siyu]{failed to observe}{simply have not observed}.
-\added[id=siyu]{Decision noise as defined in previous researches are more or less a quantification of what's not predictable by the model.
-A missing deterministic component from the model could give rise to variability in behavior that might appear to be a random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
-Avoiding an ex is a deterministic process, but if we fail to take ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
+	\replaced[id=siyu]{One limitation of this previous research, however, is that it is difficult to tell whether what we have called `decision noise' comes from a truly random process.}{One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random.} That is, whether behavioral variability is due to genuinely stochastic processes in the brain or whether it is due to deterministic processes that we \replaced[id=siyu]{failed to observe}{simply have not observed}.
+\added[id=bob]{Decision noise as defined in previous research essentially quantifies the extent to which behavior cannot be predicted by a computational model.
+A missing deterministic component from the model could give rise to variability in behavior that might appear to be random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
+Avoiding an ex is a deterministic process, but if we fail to take the ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
 
@@
 
-	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by such \added[id=siyu]{unobserved} deterministic processing of the stimulus \added[id=siyu]{(will refer to as deterministic noise)}.
-\replaced[id=siyu]{To distinguish between stimulus-driven `deterministic noise' and non-stimulus-driven `random noise'}{In particular}, we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
-If the decision is a purely deterministic function of the stimulus \added[id=siyu]{(i.e., decision noise is purely deterministic noise)}, then people's choices should be identical both times.
-That is, their choices should be consistent, since the stimulus is the same both times.
-Conversely, the more their decision is driven by \replaced[id=siyu]{random noise}{other processes, including both stochastic and unobserved deterministic processes}, the less consistent their behavior should be.
-\added[id=siyu]{We built a computational model to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.
-Our model could detect the existence of previously unobserved deterministic processes without making a priori assumptions about the deterministic process that we want to observe.} By analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
-\replaced[id=siyu]{Our work}{This does not prove that random exploration is entirely deterministic or stochastic, but} provides a lower bound on how much deterministic processes contribute to random exploration\added[id=siyu]{, and an upper bound on how much true random noises contribute to random exploration}.
-\deleted[id=siyu]{This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.}
+	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by deterministic processing of the stimulus \added[id=bob]{(which we refer to as `deterministic noise') vs other processes, including deterministic processing of unobserved stimuli as well as truly stochastic processes (which we refer to as `random noise')}.
+\added[id=siyu]{To distinguish between these two types of noise} we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
+If the decision is a purely deterministic function of the stimulus \added[id=bob]{(i.e., decision noise is purely deterministic noise), then people's choices should be identical for both decisions, since the stimulus is the same both times.
+Conversely, if the decision is a purely random function of the stimulus (i.e., decision noise is purely random noise), then people's choices will be different 50$\%$ of the time, since the random noise is different each time.
+In between these two extremes of purely deterministic and purely random drivers of behavioral variability, the extent to which people's decisions are consistent between the two decisions can be used to estimate the amount of deterministic and random noise.} %\added[id=siyu]{We built a computational model to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.
+Our model could detect the existence of previously unobserved deterministic processes that arise from the stimuli without explicitly knowing what the deterministic process is.}
+
+    \added[id=bob]{In the following, we analyze behavior on the repeated decisions version of the Horizon Task in both a model-free and model-based manner.
+Our model-free analysis estimates the extent to which people's behavior is consistent across repeated versions of the same decision.
+By measuring how this choice consistency changes as a function of horizon, this model-free analysis offers qualitative insight into the extent to which behavioral variability is driven by deterministic vs random noise.
+Our model-based analysis uses a computational model of the explore-exploit decision in the Horizon Task that incorporates both noise processes.
+By fitting this model to the behavioral data, this model-based analysis allows us to quantify the relative size of the two sources of noise and how they change in the service of exploration.}
+
+    % to the behavioral data.
+
+    % of behavior that incorporates the two noise processes.
+
+    % Our model-based analysis uses a
+
+    % Thus, by analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
+This does not prove that random exploration is entirely deterministic or stochastic, but provides a lower bound on how much deterministic processes contribute to random exploration\added[id=siyu]{, and an upper bound on how much true random noises contribute to random exploration.
+Our work sheds light on how random exploration may be partially implemented by amplifying the effects of task-irrelevant stimuli on choice.} \deleted[id=siyu]{This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.}
 
@@
 	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus-driven `random noise' \replaced[id=siyu]{in explore-exploit}{on people's} decisions (Figure \ref{fig:taskfig}).
-In this task, participants make repeated choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
+In this task, participants make a series of \added[id=bob]{BE CAREFUL USING THE WORD REPEATED.
+ONLY USE IT TO REFER TO THE REPEATED GAMES - HERE I REPLACED REPEATED WITH `SERIES OF'} choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
 Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
@@
 
-	Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
+    \added[id=bob]{THERE NEEDED TO BE MORE INFORMATION ABOUT THE HORIZON TASK HERE - REMEMBER THIS COUDL BE THE FIRST TIME PEOPLE HAVE SEEN THE TASK.
+The Horizon Task has two key features that together allow it to quantify explore-exploit behavior.
+The first of these features is  the time horizon --- the number of decisions participants will make in the future.
+By changing this horizon from short (1 trial) to long (6 trials), the Horizon Task allows us to control the relative value of exploration and exploitation.
+Just like the restaurant example in the introduction, when the horizon is short, participants should be more likely to exploit the option they believe to be best, because this leads to the highest payoff in the short term.
+Conversely, when the horizon is long, participants should be more likely to explore at first, because this allows them to gather information to make better choices later on.
+By contrasting behavior between short and long horizon conditions {\it on the very first trial}, when all else is equal, the Horizon Task allows us to quantify how behavior changes when it is valuable to explore.
+
+    The second key feature of the Horizon Task are the four `example plays' at the start of each game that allow us to control exactly what participants know about the two bandits before they make their choice.
+In these example plays, participants are instructed which of the bandits to play allowing us to control how much information they have about each of the options.
+The instructed trials are used to set up one of two information conditions and `unequal information' or [1 3] condition, in which participants play one option once and the other three times, and an `equal information' or [2 2] condition, in which participants play both options twice.}
+
+	% Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
 1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
@@
 
-	Relative to the original Horizon Task, the key modification here is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games (separated by several minutes in time so as to avoid detection).
+	Relative to the original Horizon Task, the key modification in this paper is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games separated by several minutes in time so as to avoid detection.
 By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
-Thus, if their behavior is a deterministic function of the stimuli, they will behave identically on the two games, that is their \replaced[id=siyu]{choices}{behavior} on the two versions of each game will be consistent.
-Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games \replaced[id=siyu]{could}{should} be inconsistent.
-\replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which \added[id=siyu]{the variability in} their behavior was \replaced[id=siyu]{driven by a deterministic process vs a random noise process}{a deterministic function of the stimulus or not}.
+\added[id=bob]{Thus, if their behavior is a deterministic function of the stimuli, then they will make the same decision in both games and their choices will be consistent.
+Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games will be inconsistent some fraction of the time.} \replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which \added[id=siyu]{the variability in} their behavior was \replaced[id=siyu]{driven by a deterministic process vs a random noise process}{a deterministic function of the stimulus or not}.
 
@@
 
-	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results \citep{wilson2014}.
-As in our previous work, we find evidence for two types of exploration in the Horizon Task: Random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information (Figure \ref{fig:modelfree}, Supplementary Figure S1).
+    \added[id=bob]{GENERAL COMMENT - WE NEED TO WALK PEOPLE THROUGH THE BASIC RESULTS MUCH MORE SLOWLY, OTHERWISE WILL BE CONFUSING FOR PEOPLE NEW TO THE HORIZON TASK.}
 
-	Random exploration is quantified in a model-free way as the probability of choosing the low mean option, $p(\mbox{low mean})$ in the equal, or [2 2], condition.
-This value increases with horizon, consistent with the idea that behavior is more random in horizon 6 (t(64) = 6.55, p $<$ 0.001 for [1 3], t(64) = 7.99, p $<$ 0.001 for [2 2]).
-Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info})$ in the unequal, or [1 3], condition.
-Again this measure increases with horizon, showing that people are more information seeking in horizon 6 (t(64) = 6.92, p $<$ 0.001).
+	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results using both a model-free and model-based approach \citep{wilson2014}.
+%As in our previous work, we find evidence for two types of exploration in the Horizon Task: random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.
+    \added[id=bob]{In both analyses, we focus on just the first free-choice trial in each game, where the only thing that differs between the horizon conditions is the number of choices that participants will make in the future.} %(Figure \ref{fig:modelfree}, Supplementary Figure S1).
+    % As in our previous work, we find evidence for two types of exploration in the Horizon Task: random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information (Figure \ref{fig:modelfree}, Supplementary Figure S1).
+
+    \subsubsection*{Model-free analysis}
+    \added[id=bob]{
+    In the model-free analysis, we quantify random and directed exploration using simple choice probabilities.
+Random exploration is quantified as the probability of choosing the option that has the lower average payout in the example plays in the equal, or [2 2], condition, $p(\mbox{low mean})$.
+The idea here is that, in the equal condition, the optimal strategy is to compute the mean payout for each bandit from the example plays and then always choose the option with the highest mean.
+When participants do not choose the option with the higher mean, the assumption is that this is due to some kind of `decision noise,' making the probability of choosing the low mean option a measure of behavioral variability.
+In this view, random exploration corresponds to an increase in $p(\mbox{low mean})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}A; t(64) = 7.99, p $<$ 0.001 for [2 2]).}
+
+    \added[id=bob]{
+    Directed exploration is quantified as the probability of choosing the more informative option $p(\mbox{high info})$ in the unequal, or [1 3], condition.
+The more informative option is the option played once during the example plays as choosing this option gives relatively more information (doubling the number of samples from 1 to 2) than choosing the option played three times (only increasing the number of sample by a third, from 3 to 4).
+In this view, directed exporation corresponds to an increase in $p(\mbox{high info})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}B; t(64) = 6.92, p $<$ 0.001).}
+
+    \subsubsection*{Model-based analysis}
+
+    \added[id=bob]{I'M ADDING THIS SECTION, BECAUSE IT ALLOWS US TO INTRODUCE THE MODEL EARLY. THAT WAY WHEN THE MODEL COMES IN LATER WE CAN SEE HOW IT'S CHANGED}
+
+    \added[id=bob]{Another approach to understanding behavior in the Horizon Task is to use a computational model.
+In this case, we model  participants' choices on the first free-choice trial by assuming they make decisions by computing the difference in value (or utility) $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
+Specifically, we write
+	\begin{equation}
+    \label{eq:origmodel}
+		\Delta Q= \Delta R+A \Delta I+b+n
+	\end{equation}
+    where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference in information available for playing the two options on the first free-choice trial.
+For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
+$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n$ denotes decision noise, which, in this version of the model is a combination of deterministic and random noise.
+$n$ is assumed to come from a logistic distributions with mean 0 and standard deviations $\sigma$.
+}
+
+    \added[id=bob]{The free parameters of this model are: the information bonus $A$, which controls the level of directed exploration; the noise standard deviation, $\sigma$, which controls the level of random exploration, and the spatial bias, $b$, which determines the extent to which participants prefer the option on the right.
+These free parameters are fit separately for each participant in each horizon and information condition, allowing us to test whether directed and random exploration increase with horizon.
+Consistent with previous research, we find that this is indeed the case (Figure \ref{fig:modelfree}C, D; STATS FOR CHANGE IN INFO BONUS AND DECISION NOISE).
+}
+
+    \added[id=bob]{Taken together our model-free and model-based analyses agree with previous findings showing increased behavioral variability and increased information seeking in the long horizon condition consistent with humans using random and directed exploration.
+However, for random exploration, this previous analysis cannot distinguish between deterministic and random sources of noise.
+For this we analyze the extent to which people's choices are consistent on the repeated games.}
+
+	%  In this analysis, random exploration is quantified in a model-free way as the probability of choosing the option that has the lower average payout in the example plays in the equal, or [2 2], condition, $p(\mbox{low mean})$.
+The idea here is that in the equal condition, the optimal strategy is to compute the average payout for each bandit from the example plays and then always choose the option with the highest average.
+When participants do not choose the higher average option, the assumption is that this is due to some kind of `decision noise.' In this view, random exploration corresponds to an increase in $p(\mbox{low mean})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}A; t(64) = 6.55, p $<$ 0.001 for [1 3], t(64) = 7.99, p $<$ 0.001 for [2 2]).}
+
+ % Directed exploration, is measured as the probability of choosing the more informative option $p(\mbox{high info})$ in the unequal, or [1 3], condition.
+Again this measure increases with horizon, showing that people are more information seeking in horizon 6 than horizon 1 (t(64) = 6.92, p $<$ 0.001).
+\added[id=bob]{Together, these results are consistent with the idea that people use both random exploration and directed exploration and increase both behavioral variability and information seeking when the time horizon is long.}
 
@@
 			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
-			\caption{
+			\caption{\added[id=bob]{ADD C AND D PANELS FOR MODEL-BASED RESULTS HERE}
 				Replication of previous findings.
@@
 
-	\subsection*{Model-free analysis shows that random exploration may involve both random and deterministic noise}
+	\subsection*{Model-free analysis of repeated games suggests that random exploration involves both random and deterministic noise}
 
@@
 	\end{equation*}
-	\added[id=siyu]{Furthermore, to account for that $p(\mbox{low mean})$ is a function of reward difference $\Delta R$ between the two bandits and the information condition $I$, we estimated the conditional probability:
+	\added[id=siyu]{Furthermore, to account for the fact that $p(\mbox{low mean})$ is a function of reward difference $\Delta R$ between the two bandits and the information condition $I$, we estimated the conditional probability:
 		$$p(\mbox{inconsistent}|\Delta R, I) = 2 p(\mbox{low mean}|\Delta R, I)(1-p(\mbox{low mean}|\Delta R, I))$$
@@
 	As shown in Figure \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction.
-Specifically, behavior is different from pure random noise prediction in the both the [1 3] condition (t(64) = 4.83, p $<$ 0.001 for horizon 1, t(64) = 3.12 p $=$ 0.003 for horizon 6) and the [2 2] condition (t(64) = 3.92, p $<$ 0.001 for horizon 1, t(64) = 3.71, p $<$ 0.001 for horizon 6).
+Specifically, behavior is different from the pure random noise prediction in the both the [1 3] condition (t(64) = 4.83, p $<$ 0.001 for horizon 1, t(64) = 3.12 p $=$ 0.003 for horizon 6) and the [2 2] condition (t(64) = 3.92, p $<$ 0.001 for horizon 1, t(64) = 3.71, p $<$ 0.001 for horizon 6).
 Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition (t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6).
-\added[id=siyu]{As a negative control of our method for estimating $p(\mbox{inconsistent})$ for purely random noise, we simulated choices using a decision model that only includes random noise (for details, see the results section on model-based analysis, and methods), and found that $p(\mbox{inconsistent})$ in this simulated data is not different from our pure random noise prediction in all horizon and uncertainty conditions (p $>$ 0.05, Supplementary Figure S3).} \replaced[id=siyu]{Together, our results suggest that both random noise and deterministic noise contribute to the choice variability in random exploration.}{This suggests that at least some of the `noise' is a deterministic function of the stimulus, } Although from this analysis it is not clear whether this deterministic noise increases with horizon or not.
+\added[id=bob]{As a negative control of our method for estimating $p(\mbox{inconsistent})$ for purely random noise, we simulated choices using a decision model that only includes random noise (Equation \ref{eq:origmodel}), and found that $p(\mbox{inconsistent})$ in this simulated data is not different from our pure random noise prediction in all horizon and uncertainty conditions (p $>$ 0.05, Supplementary Figure S3).} \replaced[id=siyu]{Together, our results suggest that both random noise and deterministic noise contribute to the choice variability in random exploration.}{This suggests that at least some of the `noise' is a deterministic function of the stimulus, }.
+However,  how each of these types of noise change with horizon is difficult to discern.
 
@@
 	To more precisely quantify the contribution of deterministic noise \added[id=siyu]{and random noise}, we turned to model fitting.
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate \replaced[id=siyu]{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise')}{deterministic noise from random noise}.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} (Equation \ref{eq:origmodel}) that was modified to differentiate \replaced[id=siyu]{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise')}{deterministic noise from random noise}.
 In particular, we assume that in repeated games, the value of stimulus-driven deterministic noise is frozen whereas random noise is drawn independently both times.
@@
 	\subsubsection*{Overview of model}
-	As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
-To model participants' choices on this first free-choice trial, we assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
-Specifically, we write
+	%As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
+    \added[id=bob]{To model participants' choices on the first free-choice trial, we use a modified version of Equation \ref{eq:origmodel}.
+    }
+    %assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
 	\begin{equation}
+    \label{eq:newmodel}
 		\Delta Q= \Delta R+A \Delta I+b+n_{det}+n_{ran}
 	\end{equation}
-	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference \added[id=siyu]{of} information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
+	where, as before $\Delta R$, is the the difference in mean rewards shown on the forced trials, $\Delta I$, is the difference in information, $A$ is the information bonus, and $b$ is the spatial bias.
+ %\added[id=siyu]{of} information available for playing the two options on the first free-choice trial.
+%For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
 \added[id=siyu]{$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeated plays and changes every game.} $n_{det}$ and $n_{ran}$ are assumed to come from logistic distributions with mean 0\added[id=siyu]{, and standard deviations $\sigma_{det}$ and $\sigma_{ran}$}.
-The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration; \added[id=siyu]{$\sigma_{det}$ and $\sigma_{ran}$, which control the respective level of deterministic noise and random noise in random exploration.}
+    %The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration;
+    New in Equation \ref{eq:newmodel} are the terms $n_{det}$ and $n_{ran}$.
+$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeated plays and changes every game.
+$n_{det}$ and $n_{ran}$ are assumed to come from logistic distributions with mean 0, and standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+
+    % $\sigma_{det}$ and $\sigma_{ran}$, which control the respective level of deterministic noise and random noise in random exploration.}
 
@@
 
-	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for an graphical representation of the model in the style of \cite{lee2014}).
+	We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for a graphical representation of the model in the style of \cite{lee2014}).
 In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon.
@@
 	\subsubsection*{\replaced[id=siyu]{Model validation}{Parameter recovery}\label{ch:appendix:bayesrecovery}}
-	To be sure that our fit parameter values were meaningful \added[id=siyu]{and to understand the limits of our model, we evaluated our model in several ways.
+	To be sure that our fit parameter values were meaningful \added[id=siyu]{and to understand the limits of our model, we evaluated our model extensively using simulated data.
+\added[id=bob]{This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
+Full details of this analysis are presented in the Supplementary Materials.
+
+    In this section we focus on our results for parameter recovery.
+In a parameter recover analysis, behavioral data is simulated by the model with known parameters and then this simulated behavioral data fit with the model to quantify the extent to which fit parameters match the input simulated parameters --- that is, whether the simulated parameters can be recovered.
+
+    Parameter recover in this task was good for this model, with fit values for $\sigma_{ran}$ and $\sigma_{det}$ showing strong correlations with their simulated values (INSERT STATS HERE).
+However, while the relationship was near perfect for random noise, there was a systematic bias to underestimate the level of deterministic noise by about XXX (INSERT VALUE HERE).
+Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (STATS).
+This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizon conditions.
+
+    INSERT A SIMPLE PARAMETER RECOVERY PLOT FOR SIGMARAN AND SIGMADET.
+
+    %this simulated behavioral data is fit using the same fitting procedure used to model human choices.
+
+    Thus, overall, we were able to detect both deterministic and random noises using our model.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+In addition, we see better parameter recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+	}
+
+ MOVE THE REST OF THIS SECTION TO THE SUPPLEMENTARY MATERIAL
  Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
@@
 			\includegraphics[width=0.7\textwidth]{figures/RDBayes_reduced_model.jpg}
-			\caption{Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
+			\caption{\added[id=bob]{FONT SIZE ON LEGEND IS VERY SMALL!}Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
 In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
@@
 But intriguingly, the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}).
-That is when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
+That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
 	%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO
@@
 			\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
-			\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+			\caption{\added[id=bob]{TITLE ON B AND D SHOULD SAY `CHANGE IN RANDOM NOISE'}Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
 Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
@@
 \subsubsection*{Comparison with alternative models}
+COULD POTENTIALLY MOVE THIS TO THE SUPPLEMENT
+
 To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
@@
 	% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY
-	\added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (show up in behavior as random noise) and by deterministic neural processes induced by the stimuli (show up in behavior as deterministic noise).} In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
-Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.
+	% \added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (appearing in behavior as random noise) and by deterministic neural processes induced by the stimuli (appearing in behavior as deterministic noise).}
+    In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.
 
@@
 
-	So where does this leave randomness in random exploration?
-\replaced[id=siyu]{In our model, random noise is modeled as non-stimulus-driven variability, which may include truly random noise (intrinsic stochasticity in decision making) or stimulus irrelevant deterministic processes that we could not control in the lab.
-As a result,}{Well, } the remaining 85\% of the decision noise could be random or it could be deterministic \replaced[id=siyu]{processes that is not bound to the stimulus and that we can not control}{because we can't control everything}.
-In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
+SUGGEST REFRAMING NEXT PARAGRAPH IN THE FOLLOWING WAY.
+SIMPLE READ OF RESULTS SUGGESTS RANDOM EXPLORATION IS RANDOM - THIS IS CONSISTENT WITH X.
+FOLLOWING PARA THEN GIVES LIMITATIONS AND TALKS ABOUT DETERMINISTIC INTERPRETATION.
+I'VE TRIED TO DO THIS ...
+
+    \added[id=bob]{One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random.
+Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \cite{drugowitsch16}.
+In this view, apparently random behavior is not due to sensory processing or response selection, but to suboptimal computations in the brain.
+Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition.
+Indeed, a strong interpretation of this hypothesis would suggest that randomness in explore-exploit behavior is due to imperfect inference about the correct course of action.
+In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute CITE DEEP EXPLORATION PAPER.}
+
+    While the random noise interpretation is theoretically appealing, our approach, while an improvement on previous methods, is not without limitations.
+Most important is that our measure of `random' noise is only an upper bound on the true level of randomness and that, in principle, the random decision noise could be lower.
+Specifically, in our model, what we labeled random noise was really `non-stimulus-driven variability.' While this non-stimulus-driven variability could be driven by truly random stochastic processes, it could also be driven by deterministic processing that is unrelated to the stimuli in the task.
+For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}.
+In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
+As a result, from both a conceptual and methodological perspective, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.
+
+    Like the random noise account, the deterministic noise account is also in line with previous work in which neural variability can be accounted for by fluctuations in sensory inputs.
+For example, MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli \citep{Bair1996}.
+In other words, `irrelevant' features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
+
+    % Of particular interest in our study was the fact that deterministic noise increased with horizon.
+Such a horizon increase is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+
+    Regardless of whether we interpret the noise as random or deterministic, a key finding in this paper is that both types of noise change with horizon.
+Such a horizon increase is a hallmark of an exploratory process and suggests that the modulation of deterministic and random processes may underlie random exploration.
+Moreover, the fact that the horizon change in the two types of noise are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration: a reduction in the strength with which reward drives the choice.
+
+    To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation XXX.
+In this model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, where
+    \begin{equation}
+    \label{eq:noise1}
+        \Delta Q = 
+        \Delta R + A \Delta I + b + n_{det} + n_{ran}
+    \end{equation}
+    Now imagine a case where the reward signal is scaled by a factor $\beta$.  In this case, $\Delta Q$ becomes
+    \begin{equation}
+    \label{eq:noise2}
+        \Delta Q = 
+        \beta \Delta R + A \Delta I + b + n_{det} + n_{ran}
+    \end{equation}
+    Because the choice only depends on the sign of $\Delta Q$, scaling $\Delta Q$ by a factor of $1/\beta$ will not change the behavior of the model.
+Thus, if we divide both sides of the above equation by $\beta$ we get
+    \begin{equation}
+    \label{eq:noise3}
+        \Delta Q / \beta = 
+        \Delta R + A \Delta I / \beta + b / \beta + n_{det} / \beta + n_{ran} / \beta
+    \end{equation}
+    which is equivalent to a scaling of both deterministic and random noise by the same factor $1/\beta$.
+Thus, one interpretation of our result that both deterministic and random noise change across horizons with the same ratio, is that this reflects a change in reward processing.
+That is, the reward signal is reduced in the longer horizon condition (smaller $\beta$ in horizon 6 than horizon 1).
+%In the alternate interpretation, where deterministic and random noise are varied independently it is not clear why they should change by the same scale factor.
+
+    Such a reduction in the strength of reward coding in exploration, is consistent with our recent work using a drift diffusion model (DDM) to model explore-exploit decisions CITE FENG ET AL.
+In the drift diffusion model, changes in behavioral variability can be driven by changes in the decision threshold (smaller threshold = more noise) or changes in the signal-to-noise ratio with which reward is encoded (lower SNR = more noise).
+By fitting both choices and response times, we were able to distinguish between these two accounts showing that the majority of the horizon-change in variability was driven by changes in SNR not threshold.
+However, this model could not determine whether the changes in SNR were driven by signal or noise.
+By showing that the change in deterministic and random noise have the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise.
+Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology (CITE TOMOV GERSHMAN FMRI, CITE EBITZ).
+
+    % More generally, 
+    % scaling $\Delta Q$ by a constant, $\beta$, does not change the model's choice behavior.
+That is, basing the decision on $\beta \Delta Q$ gives the same behavior as basing the behavior on $\Delta Q$
+    % \begin{equation*}
+    %     \beta \Delta Q 
+    %         = \beta \Delta R + \beta A \Delta I + \beta b + \beta n_{det} + \beta n_{ran}
+    % \end{equation*}
+    % However, scaling the noise by a constant gives just another noise
+
+ 			% % &= \beta \Delta R + A' \Delta I + b' + n_{det}' + n_{ran}'
+
+    % From a model fitting perspective, Equations \ref{eq:noise2} and \ref{eq:noise3} are equivalent as they lead to identical behavior.
+From a psychological perspective, however, they are quite different.
+In Equation \ref{eq:noise2}, how reward is processed changes across horizon with a lower $\beta$ in horizon 6 than horizon 1.
+In Equation \ref{eq:noise3}, it is the noise that changes across horizon, with both noises somehow scaled by the identical parameter $1/\beta$.
+
+    % An increase in both random and deterministic noise at a similar ratio relative to the reward, as observed in our experiment when horizon increases, suggests that either $\beta$ is reduced or $n_{det}'$ and $n_{ran}'$ are increased simultaneously at the same ratio.
+Neurally, reducing the $\beta$ corresponds to a reduction in reward sensitivity and reward coding, whereas simultaneously increasing $n_{det}'$ and $n_{ran}'$ suggests a common noise gating mechanism in which the noise filtering neural circuit is inhibited so that both random and deterministic noises are amplified.
+}
+
+    % Of particular interest in our study was the fact that deterministic noise increased with horizon.
+Such a horizon increase is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+
+    % In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
+From both a conceptual and methodological perspective, our model provides a lower bound of deterministic noise and an upper bound of random noise.
+Although there is still a considerable window for truly stochastic processes in the brain to be driving random exploration (up to 85\%), our results suggest that at least some of the apparent randomness in random exploration is not random at all.
+
+    % As a result, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.
+
+    % In particular, while we controlled many aspects of the stimulus across repeated games (e.g.
 the outcomes and the order of the forced trials), we could not perfectly control {\it all} stimuli the participant received, which would vary, for example, based on exactly what they were looking at or whether they were fidgeting or scratching their nose \citep{Musall2019}.
-Thus conceptually, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
+Thus, our estimate of random noise is an upper bound as these `missing' sources of deterministic noise would be interpreted as random noise in our model.
 Conversely, our estimate of deterministic noise is a lower bound.
@@
 
-	In addition to the conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also systematically underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
+	% In addition to the conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also systematically underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
 From both a conceptual and methodological perspective, our model provides a lower bound of deterministic noise and an upper bound of random noise.
@@
 
-	\added[id=siyu]{The deterministic noise hypothesis is in line with works in which neural variability can be accounted for by fluctuations in sensory inputs.
+	% \added[id=siyu]{The deterministic noise hypothesis is in line with works in which neural variability can be accounted for by fluctuations in sensory inputs.
 For example, }MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli \citep{Bair1996}.
-In other words, irrelevant features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
-Of particular interest was the fact that deterministic noise in our study increased with horizon, which is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
+In other words, `irrelevant' features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
+Of particular interest in our study was the fact that deterministic noise increased with horizon.
+Such a horizon increase is a hallmark of an exploratory process and suggests that the modulation of deterministic processes may underlie random exploration.
 
-	\added[id=siyu]{The random noise hypothesis, on the other hand, is consistent with findings of \cite{drugowitsch16}.} In particular these authors show that randomness in behavior could arise from imperfections in mental inference, which happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+	% \added[id=siyu]{The random noise hypothesis, on the other hand, is consistent with findings of \cite{drugowitsch16}.} In particular these authors show that randomness in behavior could arise from imperfections in mental inference, which happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
 This suggests that a large proportion of variability in behavior may arise from computational errors in computing the correct strategy.
-\added[id=siyu]{Although suboptimal inference is different from simply adding random noise to a neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not deterministically determined by the stimuli, it is a form of random noise in our definition.
+\added[id=siyu]{Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition.
 }In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
@@
 	%Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. 
-	\added[id=siyu]{
-	Regardless of whether the remaining 85\% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
+	% \added[id=siyu]{
+	% % Regardless of whether the remaining 85\% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
 Specifically, a reduction in the strength with which reward drives the choice.
@@
 Mathematically, our model is equivalent to
-	}
-	\begin{equation*}
-		\begin{split}
-			\Delta Q' &= \beta \Delta Q \\
-			&= \beta \Delta R + \beta A \Delta I + \beta b + \beta n_{det} + \beta n_{ran}\\
-			&= \beta \Delta R + A' \Delta I + b' + n_{det}' + n_{ran}'
-		\end{split}
-	\end{equation*}
-	\added[id=siyu]{In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
+	% % }
+	% \begin{equation*}
+	% 	\begin{split}
+	% 		\Delta Q' &= \beta \Delta Q \\
+	% 		&= \beta \Delta R + \beta A \Delta I + \beta b + \beta n_{det} + \beta n_{ran}\\
+	% 		&= \beta \Delta R + A' \Delta I + b' + n_{det}' + n_{ran}'
+	% 	\end{split}
+	% \end{equation*}
+	% \added[id=siyu]{In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
 From a model fitting perspective, these two forms are equivalent.
@@
 
-	Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
+	% Whether such a noise-controlling area exists in the human brain is less well established, but one candidate theory \citep{aj2005} suggests that norepinephrine (NE) from the locus coeruleus may play a role in modulating random levels of noise.
 Indeed, changes in the NE system have been associated with changes behavioral variability in both humans and other animals in a variety of tasks \citep{eeKarpova14, eeKeung18}.

### 2024-05-13 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
intro/methods

--- 2023-12-21 manuscript/Older drafts/2023 unused - overleaf/manuscript_DeterministicRandomNoise.tex
@@
 
-\title{\replaced[id=siyu]{Separating random and deterministic sources of computational noise in explore-exploit decisions}{The nature of decision noise in random exploration}}
+\title{Separating random and deterministic sources of computational noise in explore-exploit decisions}
 \author[1,\textcurrency]{Siyu Wang}
@@
 An example arises when we must choose between exploring unknown options or exploiting options we know well.
-\added[id=bob]{A little randomness in these `explore-exploit' decisions is remarkably effective as it can encourage us to explore options we might otherwise ignore.
+A little randomness in these `explore-exploit' decisions is remarkably effective as it can encourage us to explore options we might otherwise ignore.
 In line with this idea, several studies have found evidence that people increase their behavioral variability when it is valuable to explore.
-A key question, however, is whether this variability in so-called `random exploration' is actually random.} That is, is random exploration driven by stochastic processes in the brain or by some unobserved deterministic process that we have failed to account for when measuring behavioral variability?
+A key question, however, is whether this variability in so-called `random exploration' is actually random.
+That is, is random exploration driven by stochastic processes in the brain or by some unobserved deterministic process that we have failed to account for when measuring behavioral variability?
 By designing an explore-exploit task in which, unbeknownst to them, participants are presented with the exact same choice twice, we provide a partial answer to this question.
-\added[id=bob]{By modeling behavior in this task, we were able to estimate a lower bound on the amount of variability that is deterministically driven by the stimulus and an upper bound on the amount of variability that is random.} Using this approach, we find evidence that \added[id=siyu]{at least 15$\%$} of the variability in random exploration can be accounted for by deterministic processing of the stimulus.
-\added[id=bob]{Conversely, this suggests that up to 85$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.}
-% This still leaves open the possibility that much of random exploration is truly `random', but narrows the window of opportunity for \added[id=siyu]{true} stochastic processes to explain this behavior.
-\added[id=bob]{Finally, our results suggest that both deterministic and random sources of variability change proportionally to each other as the value of exploration increases, suggesting that a common noise gating mechanism may be at play in random exploration.}
-
-%suggests that at least some `random' exploration is not random, but leaves open the possibility that the 
-%However, it is unclear whether this variability is truly random
-%From a modeling perspective, behavioral variability is essentially the variance that can not be explained by a model and is modeled as the level of decision noise.
-However, what we have called "decision noise" in previous researches could actually just be missing deterministic components from the model, it is difficult to tell whether decision noise truly arises from a stochastic process.
-Here we show that, while both random and deterministic noise drive variability in behavior, the noise driving random exploration is predominantly random.
-This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.
-
+By modeling behavior in this task, we were able to estimate a lower bound on the amount of variability that is deterministically driven by the stimulus and an upper bound on the amount of variability that is random.
+Using this approach, we find evidence that at least 15$\%$ of the variability in random exploration in our studied task can be accounted for by deterministic processing of the stimulus.
+Conversely, this suggests that up to 85$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.
+Finally, our results suggest that both deterministic and random sources of variability change proportionally to each other as the value of exploration increases, suggesting that a common noise gating mechanism may be at play in random exploration.
 	\end{abstract}
@@
 	\section*{Introduction}
-
-	%BOB SAYS: IN GENERAL, NEED TO GO THROUGH THE PAPER AND MAKE SURE THAT `RANDOM' NOISE IS REPLACES WITH `NOT A DETERMINISTIC FUNCTION OF THE STIMULUS' OR SOMETHING LIKE THAT
-	%SIYU Says: We can define "random noise" as "not a deterministic funciton of the stimulus" at the beginning, and use the term "random noise" in the results section.
-It feels a little lengthy to use "not a deterministic function of the stimulus" in the results section each time.
-I changed a few of them back to "random noise" in the results section, in places that I think random noise is better.
-See if you like the changes.
-
-	%(BOB SAYS: TAKE OR LEAVE THIS CHANGE TO DATE, BUT SETTING UP AS A DATE ALLOWS YOU TO TALK ABOUT AVOID YOUR EX LATER ON, WHICH MAY BE A BETTER EXAMPLE OF A DETERMINISTIC PROCESS!).
-	% Siyu Say: I kept this.
+	%BOB SAID: IN GENERAL, NEED TO GO THROUGH THE PAPER AND MAKE SURE THAT `RANDOM' NOISE IS REPLACES WITH `NOT A DETERMINISTIC FUNCTION OF THE STIMULUS' OR SOMETHING LIKE THAT
 
@@
 choices that maximize long-term reward, is computationally intractable in most cases \citep{eegittins74, eeBasu18}.
-In part because of this computational complexity, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice \added[id=siyu]{\citep{Mehlhorn15, SCHULZ20197,WILSON202149}}.
+In part because of this computational complexity, there is considerable interest in how humans and animals solve the explore-exploit dilemma in practice {\citep{Mehlhorn15, SCHULZ20197,WILSON202149}.
 	%\citep{eeauer02,eegittins79,eekrebs78,eethompson33, eewatkins89, eebridle90,eemeyer95, eebanks97,eefrank09, eesteyvers09, eelee11, eepl12,eezhang13,eedaw06, eepl11, wilson2014}
 
-	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90} \added[id=siyu]{, also known as random exploration}.
-In this strategy, \deleted[id=siyu]{the decision process between exploration and exploitation is corrupted by `decision noise', meaning that }high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
-\added[id=siyu]{In modeling terms, random exploration works by adding `decision noise' to the value of the options such that sub-optimal exploratory options can sometimes have a higher total score (i.e., value + noise) than the exploit option and get chosen.} \replaced[id=siyu]{Such}{In theory, such} random exploration, is surprisingly effective and, if implemented correctly, can come close to optimal performance \added[id=siyu]{in some cases}\citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
-
-	It has recently been shown that humans appear to use random exploration and can increase  decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014} \added[id=bob]{ALSO CITE FINDLING ET AL.}.
+	One particularly effective strategy for solving the explore-exploit dilemma is choice randomization \citep{eethompson33, eewatkins89, eebridle90}, also known as random exploration.
+In this strategy, high value `exploit' options are not always chosen and exploratory choices are sometimes made by chance.
+From a modeling perspective, random exploration works by adding `decision noise' to the value of the options such that sub-optimal exploratory options can sometimes have a higher total score (i.e., value + noise) than the exploit option and get chosen.
+Such random exploration, is surprisingly effective and, if implemented correctly, can come close to optimal performance \citep{eethompson33, eebridle90, eeAgrawal11, eeChapelle11}.
+
+	It has recently been shown that humans appear to use random exploration and can increase  decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014, Findling19}.
 In one of these tasks, known as the Horizon Task \citep{wilson2014}, the key manipulation is the horizon condition, i.e.
@@
 
-	\replaced[id=siyu]{One limitation of this previous research, however, is that it is difficult to tell whether what we have called `decision noise' comes from a truly random process.}{One limitation of this previous work, is that it is difficult to tell whether what is measured as decision noise is truly random.} That is, whether behavioral variability is due to genuinely stochastic processes in the brain or whether it is due to deterministic processes that we \replaced[id=siyu]{failed to observe}{simply have not observed}.
-\added[id=bob]{Decision noise as defined in previous research essentially quantifies the extent to which behavior cannot be predicted by a computational model.
-A missing deterministic component from the model could give rise to variability in behavior that might appear to be random noise.} For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
+	One limitation of this previous research, however, is that it is difficult to tell whether what we have called `decision noise' actually reflect a noise process.
+From a modeling perspective, decision noise as defined in previous research essentially quantifies the extent to which behavior  cannot be explained by a computational model.
+A missing deterministic component from the model could give rise to variability in behavior that might appear to be random noise.
+For example, in the restaurant example, my usual preference for one restaurant or another may be overruled if I see an ex romantic partner going into one of them.
 Avoiding an ex is a deterministic process, but if we fail to take the ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
 
-	%because noise in previous work is defined as the component of the decision process that cannot be accounted for by a model. 
-	%, 
-	%For example, in the restaurant case above, if you happen to spot an old friend walking in to one of the restaurants, you may be more likely to follow them in.
-Such a process would be deterministic, in that it is driven by an observable stimulus, but unless a model of your behavior incorporated the presence of your friend,
-
-	%.
-If the model did not consider that the agent is favoring the behavior of following a friend in a deterministic way, the behavior of going to a less favorable restaurant because of a friend will appear to be `random' when it is really a deterministic effect.
-Hence this deterministic factor will be modeled as random decision noise.
-Crucially, however, this deterministic source of variability is very much in the stimulus and if you saw the same friend go into the same restaurant at a later date you might follow them again.
-Conversely, truly `random' noise would arise from stochastic mental processes,
-
-	%tossing a metaphorical coin in your head. Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.
-
-	%A key limitation of our previous work was that the source of the decision noise used for exploration was left unknown.
-In particular, due to the design of the task, we could not determine whether the adaptive decision noise that is linked to exploration arises deterministicly, in the input from the world, or is generated randomly, within the brain.
+	% That is, whether behavioral variability is due to genuinely stochastic processes in the brain or whether it is due to deterministic processes that we failed to observe.
+
+	% From a modeling perspective, behavioral variability is essentially the variance that can not be explained by a model and is modeled as the level of decision noise.
+However, what we have called "decision noise" in previous researches could actually just be missing deterministic components from the model, it is difficult to tell whether decision noise truly arises from a stochastic process.
+Here we show that, while both random and deterministic noise drive variability in behavior, the noise driving random exploration is predominantly random.
+This suggests that random exploration depends on adaptive noise processes in the brain which are subject to cognitive control.
 
@@
 
-    %* Difficulty with measuring random exploration, how do you know something is random, previous work model
-    %* Model all the deterministic factor (contribute to noise)
-    %* Have repeated games
-    %* Missing deterministic dependent, much smaller than non-deterministic component
-    %* Stochastic process…
-
-	%In the restaurant case, an example of deterministic noise would be if you happen to spot an old friend walking in to one of the restaurants.
-Such an event is random, in that you couldn't predict it, and has the potential to alter your behavior as you're likely to follow your friend in.
-Crucially, however, this `randomness' is very much in the world and if you saw the same friend go into the same restaurant at a later date you might follow them  again.
-
-	%Conversely, random noise would arise from stochastic neural processes tossing a metaphorical coin in your head.
-Such a process would not be influenced by the friend going into the restaurant, and if you saw the same friend again, you might make a different choice.
-
-	%Previous work makes a strong case for both deterministic and random noise being relevant to behavior.
-For instance, deterministic, stimulus-driven noise is thought to be a much greater source of choice variability in perceptual decisions than random noise \citep{eeBrunton13}.
-Conversely random, neural noise is thought to drive exploratory singing behavior in song birds and the generation and control of this random noise has been linked to specific neural structures \citep{songbird2}.
-
-	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by deterministic processing of the stimulus \added[id=bob]{(which we refer to as `deterministic noise') vs other processes, including deterministic processing of unobserved stimuli as well as truly stochastic processes (which we refer to as `random noise')}.
-\added[id=siyu]{To distinguish between these two types of noise} we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
-If the decision is a purely deterministic function of the stimulus \added[id=bob]{(i.e., decision noise is purely deterministic noise), then people's choices should be identical for both decisions, since the stimulus is the same both times.
+	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by deterministic processing of the stimulus (which we refer to as `deterministic noise') vs other processes, including deterministic processing of unobserved stimuli as well as truly stochastic processes (which we refer to as `random noise').
+To distinguish between these two types of noise, we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
+If the decision is a purely deterministic function of the stimulus (i.e., decision noise is purely deterministic noise), then people's choices should be identical for both decisions, since the stimulus is the same both times.
 Conversely, if the decision is a purely random function of the stimulus (i.e., decision noise is purely random noise), then people's choices will be different 50$\%$ of the time, since the random noise is different each time.
-In between these two extremes of purely deterministic and purely random drivers of behavioral variability, the extent to which people's decisions are consistent between the two decisions can be used to estimate the amount of deterministic and random noise.} %\added[id=siyu]{We built a computational model to quantify the relative magnitudes of stimulus-driven deterministic noise and non-stimulus-driven random noise in driving random exploration.
-Our model could detect the existence of previously unobserved deterministic processes that arise from the stimuli without explicitly knowing what the deterministic process is.}
-
-    \added[id=bob]{In the following, we analyze behavior on the repeated decisions version of the Horizon Task in both a model-free and model-based manner.
+In between these two extremes of purely deterministic and purely random drivers of behavioral variability, the extent to which people's decisions are consistent between the two decisions can be used to estimate the amount of deterministic and random noise.
+
+	% Our model could detect the existence of previously unobserved deterministic processes that arise from the stimuli without explicitly knowing what the deterministic process is.
+
+	In the following, we analyze behavior on the repeated decisions version of the Horizon Task in both a model-free and model-based manner.
 Our model-free analysis estimates the extent to which people's behavior is consistent across repeated versions of the same decision.
@@
 Our model-based analysis uses a computational model of the explore-exploit decision in the Horizon Task that incorporates both noise processes.
-By fitting this model to the behavioral data, this model-based analysis allows us to quantify the relative size of the two sources of noise and how they change in the service of exploration.}
-
-    % to the behavioral data.
-
-    % of behavior that incorporates the two noise processes.
-
-    % Our model-based analysis uses a
-
-    % Thus, by analyzing behavior on this task in both a model-free and model-based manner, we show that at least some of the `randomness' in random exploration must come from deterministic processing of the stimulus.
-This does not prove that random exploration is entirely deterministic or stochastic, but provides a lower bound on how much deterministic processes contribute to random exploration\added[id=siyu]{, and an upper bound on how much true random noises contribute to random exploration.
-Our work sheds light on how random exploration may be partially implemented by amplifying the effects of task-irrelevant stimuli on choice.} \deleted[id=siyu]{This in turn sheds light on how random exploration may be implemented by amplifying the effects of task-irrelevant stimuli on choice.}
-
-	% SIYU says: "Conversely, the more their decision is driven by other processes, including both stochastic and unobserved deterministic processes, the less consistent their behavior should be.
-" is not correct.
-Unobserved deterministic processes will be the same for repeated games and will not contribute to choice inconsistency.
+By fitting this model to the behavioral data, this model-based analysis allows us to quantify the relative size of the two sources of noise and how they change in the service of exploration.
 
@@
 	\section*{Results}	
-
 	\subsection*{The Repeated-Games Horizon Task}
-	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus-driven `random noise' \replaced[id=siyu]{in explore-exploit}{on people's} decisions (Figure \ref{fig:taskfig}).
-In this task, participants make a series of \added[id=bob]{BE CAREFUL USING THE WORD REPEATED.
-ONLY USE IT TO REFER TO THE REPEATED GAMES - HERE I REPLACED REPEATED WITH `SERIES OF'} choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
+	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus-driven `random noise' in explore-exploit decisions (Figure \ref{fig:taskfig}).
+In this task, participants make a series of choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
+They are asked to choose between the two bandits to maximize the total rewards.
+One bandit always has a higher mean payout than the other.
+Participants need to try each bandit a few times to learn about the distribution of payout from that bandit.
 Because they are initially unsure as to the mean payoff of each bandit, this task requires that participants carefully balance exploration of the lesser known bandit with exploitation of the better known bandit to maximize their overall rewards.
@@
 
-    \added[id=bob]{THERE NEEDED TO BE MORE INFORMATION ABOUT THE HORIZON TASK HERE - REMEMBER THIS COUDL BE THE FIRST TIME PEOPLE HAVE SEEN THE TASK.
+	The task is organized in games (Figure \ref{fig:taskfig}A).
+The mean payout of the two bandits are held fixed within a game and reset between games.
+Each game consists of either 5 or 10 trials.
+In the first 4 trials, participants are instructed about which bandit to choose, this allows us to manipulate what information from both bandits participants receive before they make their first free choice between the two bandits.
+From the 5th trial, participants make free choices between the two bandits.
+Participants have either 1 or 6 free choices to make.
+
 	The Horizon Task has two key features that together allow it to quantify explore-exploit behavior.
 The first of these features is  the time horizon --- the number of decisions participants will make in the future.
-By changing this horizon from short (1 trial) to long (6 trials), the Horizon Task allows us to control the relative value of exploration and exploitation.
+By changing this horizon from short (1 free-choice trial) to long (6 free-choice trials), the Horizon Task allows us to control the relative value of exploration and exploitation.
 Just like the restaurant example in the introduction, when the horizon is short, participants should be more likely to exploit the option they believe to be best, because this leads to the highest payoff in the short term.
 Conversely, when the horizon is long, participants should be more likely to explore at first, because this allows them to gather information to make better choices later on.
-By contrasting behavior between short and long horizon conditions {\it on the very first trial}, when all else is equal, the Horizon Task allows us to quantify how behavior changes when it is valuable to explore.
-
-    The second key feature of the Horizon Task are the four `example plays' at the start of each game that allow us to control exactly what participants know about the two bandits before they make their choice.
-In these example plays, participants are instructed which of the bandits to play allowing us to control how much information they have about each of the options.
-The instructed trials are used to set up one of two information conditions and `unequal information' or [1 3] condition, in which participants play one option once and the other three times, and an `equal information' or [2 2] condition, in which participants play both options twice.}
-
-	% Crucially, before people make their first choice in the Horizon Task, they are given information about the mean payoff from each bandit in the form of four example plays distributed either unequally between bandits (i.e.
-1 play of one bandit, 3 plays of the other, the [1 3] condition) or equally (2 plays each, the [2 2] condition).
-These example plays allow us to manipulate exactly what people know about each option before they make their first choice.
-
-	Relative to the original Horizon Task, the key modification in this paper is to give people `repeated games,' in which they see exact same set of example plays twice in two separate games separated by several minutes in time so as to avoid detection.
-By repeating the instructed plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
-\added[id=bob]{Thus, if their behavior is a deterministic function of the stimuli, then they will make the same decision in both games and their choices will be consistent.
-Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games will be inconsistent some fraction of the time.} \replaced[id=siyu]{The extent to which participants' choices are consistent on the repeated versions of the games allow}{On average participants played 67.31 repeated games each allowing} us to quantify the extent to which \added[id=siyu]{the variability in} their behavior was \replaced[id=siyu]{driven by a deterministic process vs a random noise process}{a deterministic function of the stimulus or not}.
-
-	%, the example plays allow us to probe how participants respond to the exact same explore-exploit choice twice.
-
-	%These `repeated games' are the key manipulation in this paper and allow us to distinguish between deterministic and random sources of noise.
-Specifically, if noise is deterministicly driven, then choices on repeated games should be consistent.
-Conversely if noise is randomly driven, then choices on repeated games should be inconsistent.
+By contrasting behavior between short and long horizon conditions {\it on the very first free-choice trial}, when all else is equal, the Horizon Task allows us to quantify how behavior changes, when it is more valuable to explore.
+
+	The second key feature of the Horizon Task are the 4 `forced-choice trials' at the start of each game that allow us to control exactly what participants know about the two bandits before they make their choice.
+In these forced-choice trials, participants are instructed which of the bandits to play allowing us to control how much information they have about each of the options.
+The forced-choice trials are used to set up one of two information conditions: an `unequal information' or [1 3] condition, in which participants play one bandit once and the other three times, and an `equal information' or [2 2] condition, in which participants play both bandits twice.
+
+	Relative to the original Horizon Task, the key modification in this paper is to give people `repeated games' (Figure \ref{fig:taskfig}B), in which they see exact same set of forced-choice plays twice in two separate games separated by several minutes in time so as to avoid detection.
+By repeating the forced-choice plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
+Thus, if their behavior is a deterministic function of the stimuli, then they will make the same decision in both games and their choices will be consistent.
+Conversely, if their behavior is not driven by a deterministic function of the stimulus, then their choices on the repeated games will be inconsistent some fraction of the time.
+The extent to which participants' choices are consistent on the repeated versions of the games allow us to quantify the extent to which the variability in their behavior was driven by a deterministic process vs a random noise process.
 
 	\subsection*{Both behavioral variability and information seeking increase with horizon}
-
-    \added[id=bob]{GENERAL COMMENT - WE NEED TO WALK PEOPLE THROUGH THE BASIC RESULTS MUCH MORE SLOWLY, OTHERWISE WILL BE CONFUSING FOR PEOPLE NEW TO THE HORIZON TASK.}
-
 	Before discussing the results for repeated games, we first confirm that the basic behavior in this task is consistent with our previously reported results using both a model-free and model-based approach \citep{wilson2014}.
-%As in our previous work, we find evidence for two types of exploration in the Horizon Task: random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information.
-    \added[id=bob]{In both analyses, we focus on just the first free-choice trial in each game, where the only thing that differs between the horizon conditions is the number of choices that participants will make in the future.} %(Figure \ref{fig:modelfree}, Supplementary Figure S1).
-    % As in our previous work, we find evidence for two types of exploration in the Horizon Task: random exploration, which is the main focus of this paper, where exploration is driven by noise, and directed exploration, where exploration is driven by information (Figure \ref{fig:modelfree}, Supplementary Figure S1).
+In both analyses, we focus on just the first free-choice trial in each game, where the only thing that differs between the horizon conditions is the number of choices that participants will make in the future.
 
 	\subsubsection*{Model-free analysis}
-    \added[id=bob]{
 	In the model-free analysis, we quantify random and directed exploration using simple choice probabilities.
-Random exploration is quantified as the probability of choosing the option that has the lower average payout in the example plays in the equal, or [2 2], condition, $p(\mbox{low mean})$.
+Random exploration is quantified as the probability of choosing the option that has the lower average payout in the forced-choice plays in the equal, or [2 2], condition, $p(\mbox{low mean})$.
 The idea here is that, in the equal condition, the optimal strategy is to compute the mean payout for each bandit from the example plays and then always choose the option with the highest mean.
-When participants do not choose the option with the higher mean, the assumption is that this is due to some kind of `decision noise,' making the probability of choosing the low mean option a measure of behavioral variability.
-In this view, random exploration corresponds to an increase in $p(\mbox{low mean})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}A; t(64) = 7.99, p $<$ 0.001 for [2 2]).}
-
-    \added[id=bob]{
+When participants do not choose the option with the higher mean, the assumption is that this is due to some kind of `decision noise', making the probability of choosing the low mean option a measure of behavioral variability.
+In this view, random exploration corresponds to an increase in $p(\mbox{low mean})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}A; t(64) = 7.99, p $<$ 0.001).
+
 	Directed exploration is quantified as the probability of choosing the more informative option $p(\mbox{high info})$ in the unequal, or [1 3], condition.
-The more informative option is the option played once during the example plays as choosing this option gives relatively more information (doubling the number of samples from 1 to 2) than choosing the option played three times (only increasing the number of sample by a third, from 3 to 4).
-In this view, directed exporation corresponds to an increase in $p(\mbox{high info})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}B; t(64) = 6.92, p $<$ 0.001).}
+The more informative option is the option played once during the forced-choice plays as choosing this option gives relatively more information (doubling the number of samples from 1 to 2) than choosing the option played three times (only increasing the number of sample by a third, from 3 to 4).
+In this view, directed exporation corresponds to an increase in $p(\mbox{high info})$ with horizon, which is exactly what we see in the data (Figure \ref{fig:modelfree}B; t(64) = 6.92, p $<$ 0.001).
 
 	\subsubsection*{Model-based analysis}	
-
-    \added[id=bob]{I'M ADDING THIS SECTION, BECAUSE IT ALLOWS US TO INTRODUCE THE MODEL EARLY. THAT WAY WHEN THE MODEL COMES IN LATER WE CAN SEE HOW IT'S CHANGED}
-
-    \added[id=bob]{Another approach to understanding behavior in the Horizon Task is to use a computational model.
+	Another approach to understanding behavior in the Horizon Task is to use a computational model.
 In this case, we model  participants' choices on the first free-choice trial by assuming they make decisions by computing the difference in value (or utility) $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
@@
 	\end{equation}
-    where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced trials, and $\Delta I$, the difference in information available for playing the two options on the first free-choice trial.
-For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n$ denotes decision noise, which, in this version of the model is a combination of deterministic and random noise.
+	where, the experimentally controlled variables are $\Delta R=R_{right}-R_{left}$, the difference between the mean of rewards shown on the forced-choice trials, and $\Delta I$, the difference in information available for playing the two options on the first free-choice trial.
+For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right in the [1 3] condition, and in [2 2] condition, $\Delta I$ is 0.
+
+	Here, $n$ denotes decision noise, which, in this version of the model is a combination of deterministic and random noise.
 $n$ is assumed to come from a logistic distributions with mean 0 and standard deviations $\sigma$.
-}
-
-    \added[id=bob]{The free parameters of this model are: the information bonus $A$, which controls the level of directed exploration; the noise standard deviation, $\sigma$, which controls the level of random exploration, and the spatial bias, $b$, which determines the extent to which participants prefer the option on the right.
+
+	The free parameters of this model are: the information bonus $A$, which controls the level of directed exploration; the noise standard deviation, $\sigma$, which controls the level of random exploration, and the spatial bias, $b$, which determines the extent to which participants prefer the option on the right.
 These free parameters are fit separately for each participant in each horizon and information condition, allowing us to test whether directed and random exploration increase with horizon.
 Consistent with previous research, we find that this is indeed the case (Figure \ref{fig:modelfree}C, D; STATS FOR CHANGE IN INFO BONUS AND DECISION NOISE).
-}
 
@@
 		\caption{\added[id=bob]{ADD C AND D PANELS FOR MODEL-BASED RESULTS HERE}
-				Replication of previous findings.
-Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.
-}
+			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
 		\label{fig:modelfree}
@@
 \subsubsection*{\replaced[id=siyu]{Model validation}{Parameter recovery}\label{ch:appendix:bayesrecovery}}
-	To be sure that our fit parameter values were meaningful \added[id=siyu]{and to understand the limits of our model, we evaluated our model extensively using simulated data.
+To be sure that our fit parameter values were meaningful and to understand the limits of our model, we evaluated our model extensively using simulated data.
 \added[id=bob]{This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
 Full details of this analysis are presented in the Supplementary Materials.
+}
 
@@
 In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-	}
 
@@
 This suggests that our model is capable of detecting missing deterministic processes.
-	}
+
 \begin{figure}[H]
@@
 % More generally, 
-    % scaling $\Delta Q$ by a constant, $\beta$, does not change the model's choice behavior.
-That is, basing the decision on $\beta \Delta Q$ gives the same behavior as basing the behavior on $\Delta Q$
+% scaling $\Delta Q$ by a constant, $\beta$, does not change the model's choice behavior. That is, basing the decision on $\beta \Delta Q$ gives the same behavior as basing the behavior on $\Delta Q$
 % \begin{equation*}

### 2024-05-13 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
intro/methods

--- 2022-12-12 manuscript/2020 rejected - JEP General/SupplementaryFigures_RanDetNoise/DetRanNoise_Supplementary.tex
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
 
-\title{Supplementary Figures: The nature of decision noise in random exploration}
+\usepackage{changes}
+\definechangesauthor[name={Siyu Wang}, color=red]{siyu}
+\title{Supplementary Figures: \replaced[id=siyu]{Separating random and deterministic sources of computational noises in explore-exploit decisions}{The nature of decision noise in random exploration}}
+\author[1,\textcurrency]{Siyu Wang}
+\author[1,2,3]{Robert C. Wilson}
 
-\author[1]{Siyu Wang}
-\author[1,2]{Robert C. Wilson}
-
-\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
-\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}
+\affil[1]{Department of Psychology, University of Arizona, Tucson AZ, USA}
+\affil[2]{Neuroscience and Physiological Sciences Graduate Interdisciplinary Program, University of
+	Arizona, Tucson AZ, USA}
+\affil[3]{Cognitive Science Program, University of Arizona, Tucson AZ, USA}
+\affil[ \textcurrency]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}
 
@@
 	\newpage
-
 	\begin{figure}[H]
 		\begin{center}
-			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree_all.png}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree__all.jpg}
 			\caption{Replication of previous findings with data from all participants (i.e.
@@
 }
-			\label{fig:modelfree2}
+			\label{fig:s1}
 		\end{center}
 	\end{figure}
+	\newpage
 
@@
 However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision.}
-			\label{fig:mf22}
+			\label{fig:s2}
+		\end{center}
+	\end{figure}
+
+	\newpage
+
+	\begin{figure}[H]
+	\begin{center}
+		\includegraphics[width=\textwidth]{figures/RDBayes_pinconsistent_validation.jpg}
+		\caption{Model-free analysis with simulated choices from a model that has only random noise validates our prediction of p(inconsistent) for pure random noise.
+The extent to which simulated choices are inconsistent completely overlaps with our pure random noise prediction(p $>$ 0.05).
+This suggests that when choice inconsistency lies below the pure random noise prediction indeed provides evidence that deterministic noise exists in random exploration (Figure 3).}
+		\label{fig:s3}
+	\end{center}
+	\end{figure}
+
+	\newpage
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior.jpg}
+			\caption{Parameter recovery over the mean estimates of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices, dashed black line is the mean of the true posterior.
+Histograms represent the mean estimates of the respective parameters in the refitting to the simulated data.
+(A) and (B) are random noise at H = 1 and H = 6, respectively.
+(C) is the random noise differences between horizons.
+(D) and (E) are deterministic noise at H = 1 and H = 6, respectively.
+(F) is the deterministic noise differences between horizons.}
+			\label{fig:s4}
+		\end{center}
+	\end{figure}
+
+\newpage
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession.jpg}
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.}
+			\label{fig:s5}
 		\end{center}
@@
 \newpage
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.6\textwidth]{figures/RDBayes_parameterrecovery_subject.jpg}
+			\caption{Same as Supplementary Figure S5, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
+			\label{fig:s6}
+		\end{center}
+	\end{figure} 
+\newpage
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_gridsimu_all.jpg}
+			\caption{Parameter recovery on arbitrary combinations of random and deterministic noises.
+A.
+Recovered posterior distributions of random noise.
+B.
+Recovered posterior distributions of deterministic noise.
+For both A and B, from the top row to the bottom row, the true noise standard deviation that is used in the simulations go from 0 to 10.
+The y limit of each panel is 4 (+/- 2 from the true value).
+Our model did a relatively good job in recovering all combinations of deterministic and random noises.}
+			\label{fig:s7}
+		\end{center}
+	\end{figure}
+
+	\newpage
 	\begin{figure}[H]
 		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RanDetNoise_hyperprior_all.png}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior__all.jpg}
 			\caption{Model based analysis with data from all participants (i.e.
@@
 However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
-			%COMBINE FIGURES 4 AND 5, USE SEPARATE PANELS FOR EACH ONE, REMOVE INCORRECT RED SHADED REGION ON TOP PANEL OF FIGURE 5, Y-AXIS CAN BE LABELED 'posterior density' X-AXIS IS 'noise standard deviation [points]'
-			\label{fig:mb12}
+			\label{fig:s8}
 		\end{center}
@@
 
-	%\bibliographystyle{plainnat}
-	%\bibliographystyle{unsrt}
-	%\bibliography{Refs/refs}
+	\newpage
+	\begin{figure}[t]
+		\begin{center}
+			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperpriors_6model.jpg}
+			\caption{Model based analysis with variations of models.
+Each row is one model.
+These models varied in whether deterministic $\sigma^{det}$ and random noise $\sigma^{ran}$ are present or not and whether either types of noise is dependent on horizon (subscript denotes the dependence on horizon).
+}
+			\label{fig:s9}
+		\end{center}
+	\end{figure}
 
-	% add the Bibliography to the Table of Contents
-	%\cleardoublepage
-	%\ifdefined\phantomsection
-	%\phantomsection  % makes hyperref recognize this section properly for pdf link
-	%\else
-	%\fi
-	%\addcontentsline{toc}{chapter}{Bibliography}
+\newpage
 
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_6modelcomparison.jpg}
+			\caption{
+			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
+			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
+I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
+COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
+			Model comparison.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+			\label{fig:s10}
+		\end{center}
+	\end{figure}
 \end{document}

### 2024-05-13 codes/main_1_behavior.m

@@ -8,7 +8,7 @@ W.library_wang('Wang_EEHorizon')
 opt_sub = {'subjectID', 'date', 'time'};
 opt_preprocess = 'EEpreprocess_game_basic';
 opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};
-opt_analysis = {'EEanalysis_sub_basic'};
+opt_analysis = {'EEanalysis_sub_basic', 'EEanalysis_sub_MLE'};
 [sub0, game, idxsub] = W.analysis_pipeline(filename, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, savedir);
 W.writetable(game, fullfile(savedir,'data_all.csv'));
 %% exclude participants

### 2024-05-14 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
added stats for fig 2CD
parameter recovery

@@
 	\subsubsection*{Model-based analysis}	
-	Another approach to understanding behavior in the Horizon Task is to use a computational model.
+	Another approach to understanding behavior in the Horizon Task is to use a computational model \citep{wilson2014}.
 In this case, we model  participants' choices on the first free-choice trial by assuming they make decisions by computing the difference in value (or utility) $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.
@@
 These free parameters are fit separately for each participant in each horizon and information condition, allowing us to test whether directed and random exploration increase with horizon.
-Consistent with previous research, we find that this is indeed the case (Figure \ref{fig:modelfree}C, D; STATS FOR CHANGE IN INFO BONUS AND DECISION NOISE).
+Consistent with previous research, we find that this is indeed the case (Figure \ref{fig:modelfree}C; t(64) = 5.35, p $<$ 0.001.
+Figure \ref{fig:modelfree}D; t(64) = 3.54, p $<$ 0.001).
 
-	\added[id=bob]{Taken together our model-free and model-based analyses agree with previous findings showing increased behavioral variability and increased information seeking in the long horizon condition consistent with humans using random and directed exploration.
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
+			\caption{Replication of previous findings that people use both random and directed exploration in this task.
+(A) model-free measure of behavioral variability, $p(\mbox{low mean})$, increases with horizon.
+(B) model-free measure of information seeking, $p(\mbox{high info})$, increases with horizon.
+(C) model-based measure of behavioral variability, decision noise $\sigma$, increases with horizon.
+(B) model-based measure of information seeking, information bonus $A$, increases with horizon.}
+			\label{fig:modelfree}
+		\end{center}
+	\end{figure}
+
+	Taken together, our model-free and model-based analyses agree with previous findings showing increased behavioral variability and increased information seeking in the long horizon condition, consistent with humans using random and directed exploration (Figure \ref{fig:modelfree}, Supplementary Figure S1).
 However, for random exploration, this previous analysis cannot distinguish between deterministic and random sources of noise.
-For this we analyze the extent to which people's choices are consistent on the repeated games.}
+For this we analyze the extent to which people's choices are consistent on the repeated games.
 
@@
 
-\begin{figure}[H]
-	\begin{center}
-		\includegraphics[width=\textwidth]{figures/RanDetNoise_modelfree.jpg}
-		\caption{\added[id=bob]{ADD C AND D PANELS FOR MODEL-BASED RESULTS HERE}
-			Replication of previous findings. Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.  }
-		\label{fig:modelfree}
-	\end{center}
-\end{figure}
-
 \subsection*{Model-free analysis of repeated games suggests that random exploration involves both random and deterministic noise}
@@
 The idea behind this measure is that purely deterministic noise should lead to consistent choices as the deterministic stimulus is identical both times.
-\deleted[id=siyu]{Conversely, if choice is not a deterministic function of the stimulus, participants' choices should be independent, and hence more inconsistent across the repetitions of the game.}\added[id=siyu]{Conversely, if choice is not entirely driven by a deterministic process and is also driven by random noise, participants' choices should be more inconsistent across the repetitions of the game.
-Moreover, if decision noise is purely random noise, meaning there is no unobserved deterministic process, we will show that we can actually predict the expected level of choice inconsistencies across repetitions of games by accounting for the known deterministic processes and assuming that the random noise process is independent in repetitions of the game.}
+Conversely, if choice is not entirely driven by a deterministic process and is also driven by random noise, participants' choices should be more inconsistent across the repetitions of the game.
+Moreover, if decision noise is purely random noise, meaning there is no unobserved deterministic process, we will show that we can actually predict the expected level of choice inconsistencies across repetitions of games by accounting for the known deterministic processes and assuming that the random noise process is independent in repetitions of the game.
 
 To quantify choice inconsistency we computed the frequency with which participants made different responses for pairs of repeated games (Figure \ref{fig:mf2}, Supplementary Figure S2).
-Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions, suggesting that not all of the noise was stimulus driven (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions\deleted[id=siyu]{For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6}).
-In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (For [1 3] condition, t(64) = 5.41, p $<$ 0.001; for [2 2] condition, t(64) = 6.26, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is not a deterministic function of the stimulus\added[id=siyu]{, but rather random noise}.
+Using this measure we found that participants made inconsistent choices in both the unequal ([1 3]) and equal ([2 2]) information conditions ($p(\mbox{inconsistent}) > 0 $), suggesting that not all of the noise was stimulus driven.
+In addition, we found that choice inconsistency was higher in horizon 6 than in horizon 1 for both [1 3] and [2 2] condition (For [1 3] condition, t(64) = 5.41, p $<$ 0.001; for [2 2] condition, t(64) = 6.26, p $<$ 0.001), suggesting that at least some of the horizon dependent noise is not a deterministic function of the stimulus, but rather random noise.
 
+% (t-test vs zero revealed that inconsistency was greater than zero for all horizon and uncertainty conditions)
+
+%\deleted[id=siyu]{For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6}
 \begin{figure}[H]
@@
 \end{equation*}
-\added[id=siyu]{Furthermore, to account for the fact that $p(\mbox{low mean})$ is a function of reward difference $\Delta R$ between the two bandits and the information condition $I$, we estimated the conditional probability:
+
+Furthermore, to account for the fact that $p(\mbox{low mean})$ is a function of reward difference $\Delta R$ between the two bandits and the information condition $I$, we estimated the conditional probability:
 	$$p(\mbox{inconsistent}|\Delta R, I) = 2 p(\mbox{low mean}|\Delta R, I)(1-p(\mbox{low mean}|\Delta R, I))$$
-	Then based on the likelihood that each condition ($\Delta R$ vs $I$) occurs in the task $\rho(\Delta R, I)$, we have
-	$$p(\mbox{inconsistent}) = \sum_{\Delta R, I}\rho(\Delta R, I)p(\mbox{inconsistent}|\Delta R, I)$$}
+	Then based on the likelihood that each condition ($\Delta R$ and $I$) occurs in the task $\rho(\Delta R, I)$, we have
+	$$p(\mbox{inconsistent}) = \sum_{\Delta R, I}\rho(\Delta R, I)p(\mbox{inconsistent}|\Delta R, I)$$
 
@@
 Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition (t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6).
-\added[id=bob]{As a negative control of our method for estimating $p(\mbox{inconsistent})$ for purely random noise, we simulated choices using a decision model that only includes random noise (Equation \ref{eq:origmodel}), and found that $p(\mbox{inconsistent})$ in this simulated data is not different from our pure random noise prediction in all horizon and uncertainty conditions (p $>$ 0.05, Supplementary Figure S3).} \replaced[id=siyu]{Together, our results suggest that both random noise and deterministic noise contribute to the choice variability in random exploration.}{This suggests that at least some of the `noise' is a deterministic function of the stimulus, }.
-However,  how each of these types of noise change with horizon is difficult to discern.
+As a negative control of our method for estimating $p(\mbox{inconsistent})$ for purely random noise, we simulated choices using a decision model that only includes random noise (Equation \ref{eq:newmodel}), and found that $p(\mbox{inconsistent})$ in this simulated data is not different from our pure random noise prediction in all horizon and uncertainty conditions (p $>$ 0.05, Supplementary Figure S3).
+Together, our results suggest that both random noise and deterministic noise contribute to the choice variability in random exploration.
+However, the relative contribution from each of these types of noise, as well as how each type of noise changes with horizon, are difficult to discern.
 
-\subsection*{Model-based analysis \replaced[id=siyu]{provides a lower-bound estimate of deterministic noise and an upper-bound estimate of random noise}{shows deterministic noise changes with horizon}}
+\subsection*{Model-based analysis provides a lower-bound estimate of deterministic noise and an upper-bound estimate of random noise}
 
-To more precisely quantify the contribution of deterministic noise \added[id=siyu]{and random noise}, we turned to model fitting.
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} (Equation \ref{eq:origmodel}) that was modified to differentiate \replaced[id=siyu]{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise')}{deterministic noise from random noise}.
+To more precisely quantify the contribution of deterministic noise and random noise, we turned to model fitting.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model (Equation \ref{eq:origmodel}) that was modified to differentiate between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').
 In particular, we assume that in repeated games, the value of stimulus-driven deterministic noise is frozen whereas random noise is drawn independently both times.
@@
 %As with our model-free analysis, the model-based analysis focuses only on the first free-choice trial since that is the only free choice when we have control over the experience participants have about two bandits.
-\added[id=bob]{To model participants' choices on the first free-choice trial, we use a modified version of Equation \ref{eq:origmodel}.
-}
+To model participants' choices on the first free-choice trial, we use a modified version of Equation \ref{eq:origmodel}.
 %assume that they make decisions by computing the difference in value $\Delta Q$ between the right and left options, choosing right when $\Delta Q > 0$ and left otherwise.  Specifically, we write
@@
 \end{equation}
-where, as before $\Delta R$, is the the difference in mean rewards shown on the forced trials, $\Delta I$, is the difference in information, $A$ is the information bonus, and $b$ is the spatial bias.
-%\added[id=siyu]{of} information available for playing the two options on the first free-choice trial.
-%For simplicity, and because information is manipulated categorically in the Horizon Task, we define $\Delta I$ to be +1 if one reward is drawn from the right option and three are drawn from the left in the [1 3] condition, -1 if one from the left and three from the right, and in [2 2] condition, $\Delta I$ is 0.
-\added[id=siyu]{$n_{det}$ denotes the deterministic noise, which is identical on the repeat versions of each game; and $n_{ran}$ denotes random noise, which is uncorrelated between repeated plays and changes every game.} $n_{det}$ and $n_{ran}$ are assumed to come from logistic distributions with mean 0\added[id=siyu]{, and standard deviations $\sigma_{det}$ and $\sigma_{ran}$}.
-%The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration;
+where, as before $\Delta R$, is the the difference in mean rewards shown on the forced-choice trials, $\Delta I$, is the difference in information, $A$ is the information bonus, and $b$ is the spatial bias.
 New in Equation \ref{eq:newmodel} are the terms $n_{det}$ and $n_{ran}$.
@@
 
-% $\sigma_{det}$ and $\sigma_{ran}$, which control the respective level of deterministic noise and random noise in random exploration.}
+%The subject-and-condition-specific parameters are: the spatial bias, $b$, which determines the extent to which participants prefer the option on the right; the information bonus $A$, which controls the level of directed exploration;
 
@@
 
-\subsubsection*{\replaced[id=siyu]{Model validation}{Parameter recovery}\label{ch:appendix:bayesrecovery}}
+\subsubsection*{Model validation \label{ch:appendix:bayesrecovery}}
 To be sure that our fit parameter values were meaningful and to understand the limits of our model, we evaluated our model extensively using simulated data.
-\added[id=bob]{This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
+This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
 Full details of this analysis are presented in the Supplementary Materials.
-}
 
-	In this section we focus on our results for parameter recovery.
-In a parameter recover analysis, behavioral data is simulated by the model with known parameters and then this simulated behavioral data fit with the model to quantify the extent to which fit parameters match the input simulated parameters --- that is, whether the simulated parameters can be recovered.
+In this section we focus on our results for parameter recovery \citep{Wilson2019}.
+In a parameter recover analysis, behavioral data is simulated by the model with known parameters and then this simulated behavioral data is fit with the model to quantify the extent to which fit parameters match the input simulated parameters --- that is, whether the simulated parameters can be recovered.
 
-	Parameter recover in this task was good for this model, with fit values for $\sigma_{ran}$ and $\sigma_{det}$ showing strong correlations with their simulated values (INSERT STATS HERE).
-However, while the relationship was near perfect for random noise, there was a systematic bias to underestimate the level of deterministic noise by about XXX (INSERT VALUE HERE).
-Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (STATS).
+\begin{figure}[H]
+	\begin{center}
+		\includegraphics[width=\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession_randet.jpg}
+		\caption{Parameter recovery analysis for random (A,B) and deterministic noise (C,D) in the two horizons.}
+		\label{fig:paramrecover_main}
+	\end{center}
+\end{figure}
+
+Parameter recover in this task was good for this model, with fit values for $\sigma_{ran}$ and $\sigma_{det}$ showing strong correlations with their simulated values in both horizon (H) conditions (For $\sigma_{ran}$, $R = 0.91 (\mbox{H} = 1) \mbox{ and } 0.84 (\mbox{H} = 6), p < 0.001$, For $\sigma_{det}$, $R = 0.60 (\mbox{H} = 1) \mbox{ and } 0.59 (\mbox{H} = 6), p < 0.001$).
+However, while the relationship was near perfect for random noise ($\frac{\mbox{\small Recovered } \sigma_{ran}}{\mbox{\small Simulated } \sigma_{ran}} = 1.01$), there was a systematic bias to underestimate the level of deterministic noise by about 32\% ($\frac{\mbox{\small Recovered } \sigma_{det}}{\mbox{\small Simulated } \sigma_{det}} = 0.68$).
+Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (Supplementary Figure XXX).
 This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizon conditions.
-
-	INSERT A SIMPLE PARAMETER RECOVERY PLOT FOR SIGMARAN AND SIGMADET.
-
-	%this simulated behavioral data is fit using the same fitting procedure used to model human choices.
+	% $\frac{\mbox{\small Recovered } \Delta \sigma_{det}}{\mbox{\small Simulated } \Delta \sigma_{det}} = 0.74$
 
@@
 
-MOVE THE REST OF THIS SECTION TO THE SUPPLEMENTARY MATERIAL
-Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
-We test this by leaving out one known deterministic process from the decision model, and ask if our method could recover that known deterministic process as deterministic noise.
-In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition on explore-exploit decisions.
-$$\Delta Q= \Delta R+n_{det}+n_{ran}$$
-Here, $\Delta Q, \Delta R, n_{det}, n_{ran}$ represent the same variables as in the full model.	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
-By comparing the	fitted posterior distributions over the group-level means of the deterministic and random noise parameters $\sigma_{det}$ and $\sigma_{ran}$, as expected, we observed an increase in deterministic noise and no change in random noise between the reduced and the full model (Figure \ref{fig:reducedmodel}).
-This suggests that our model is capable of detecting missing deterministic processes.
-
-\begin{figure}[H]
-\begin{center}
-	\includegraphics[width=0.7\textwidth]{figures/RDBayes_reduced_model.jpg}
-	\caption{\added[id=bob]{FONT SIZE ON LEGEND IS VERY SMALL!}Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
-In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
-Random noise remains unchanged between the two models.}
-	\label{fig:reducedmodel}
-\end{center}
-\end{figure}
-
-\added[id=siyu]{Secondly, we evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
-In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
-The simulation and re-fitting was repeated for 200 times.
-Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval.
-If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations (this ratio will be referred to as the coverage rate).
-For random noise, the coverage rate is 100\% for both horizon 1, horizon 6, and the horizon difference.
-For deterministic noise, the coverage rate is 66\% for horizon 1 and 69\% for horizon 6.
-By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of recovered parameters, it is clear that our model systematically underestimates deterministic noise (Figure \ref{fig:coverage2}).
-Despite the underestimation of deterministic noise in both horizons, we could still reliably detect the horizon changes of deterministic noise (coverage rate is 97\%).
-This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizons.
-For random noise, our model fitting procedure yields a faithful recovery.
-However, there is a conceptual limitation.
-Because random noise is modeled as non-stimulus-driven noise, it can include both true stochastic neural noise and possible deterministic noises which do not depend on the stimuli.
-Because of this, our random noise estimate provides an upper bound of true `random noise' induced by intrinsic stochastic processes in the brain.}
-
-\begin{figure}[H]
-\begin{center}
-	\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior_v2.jpg}
-	\caption{Parameter recovery over the posterior distribution of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
-Solid lines are true posterior used to simulate choices.
-Lighter color shades represent the re-fitted posterior to the simulated choices.
-Our model fitting procedure faithfully recovers the non-stimulus-driven random noise (A, B), but systematically underestimates deterministic noise in both horizons (D, E).
-The horizon differences in random noise is also faithfully recovered (C).
-The horizon differences in deterministic noise is also underestimated but not significant (F).}
-	\label{fig:coverage2}
-\end{center}
-\end{figure}
-
-\added[id=siyu]{Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Supplementary Figure S5 and S6).
-The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).
-The strength of correlation between simulated and fit values are strong for both deterministic noise (R $>$ 0.8) and random noise (R $>$ 0.9).
-Despite the strong inter-subject correlations, we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
-
-%Parameter recovery is good (in terms of) for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C).
-The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
-
-\added[id=siyu]{Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
-All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ were tested.
-In a special case, we evaluated how our model performs when there is only random noise or only deterministic noise (Figure \ref{fig:puredetran}).
-In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with fully random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
-However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Supplementary Figures S7).
-Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
-
-%Overall, we see better recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-
-%This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
-
-\begin{figure}[H]
-\begin{center}
-	\includegraphics[width=0.7\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet.jpg}
-	\caption{Parameter recovery over the posterior of random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for purely random noise (top row) and purely deterministic noise (bottom row) games.
-	}
-	\label{fig:puredetran}
-\end{center}
-\end{figure} 
-% Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
-
-\added[id=siyu]{Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
-Our model provides a lower bound for deterministic noise and an upper bound for random noise.
-In addition, We see better parameter recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-}
 \subsubsection*{Model-based results}

### 2024-05-14 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
parameter recovery

@@
 \definechangesauthor[name={Siyu Wang}, color=red]{siyu}
-\title{Supplementary Figures: \replaced[id=siyu]{Separating random and deterministic sources of computational noises in explore-exploit decisions}{The nature of decision noise in random exploration}}
+\definechangesauthor[name={Bob}, color=blue]{bob}
+\usepackage{blindtext}
+\usepackage{titlesec}
+
+\title{Supplementary Materials: Separating random and deterministic sources of computational noises in explore-exploit decisions}
 \author[1,\textcurrency]{Siyu Wang}
@@
 	\maketitle
+	\newpage
+	\tableofcontents
 
-	\newpage
+	\section{Additional model-free analyses}
+	\subsection{Replication of Figure 2 without excluding subjects}
 	\begin{figure}[H]
@@
 
+	\subsection{Replication of Figure 3 without excluding subjects}
 	\begin{figure}[H]
@@
 
+	\subsection{Model-free analysis with simulated choice from a pure random noise model}
 	\begin{figure}[H]
@@
 	\end{figure}
+	\newpage
+	\section{Model validation analyses}
+
+	MOVE THE REST OF THIS SECTION TO THE SUPPLEMENTARY MATERIAL
+	Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
+We test this by leaving out one known deterministic process from the decision model, and ask if our method could recover that known deterministic process as deterministic noise.
+In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition on explore-exploit decisions.
+	$$\Delta Q= \Delta R+n_{det}+n_{ran}$$
+	Here, $\Delta Q, \Delta R, n_{det}, n_{ran}$ represent the same variables as in the full model.	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
+By comparing the	fitted posterior distributions over the group-level means of the deterministic and random noise parameters $\sigma_{det}$ and $\sigma_{ran}$, as expected, we observed an increase in deterministic noise and no change in random noise between the reduced and the full model (Figure \ref{fig:reducedmodel}).
+This suggests that our model is capable of detecting missing deterministic processes.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_reduced_model.jpg}
+			\caption{\added[id=bob]{FONT SIZE ON LEGEND IS VERY SMALL!}Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
+In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
+Random noise remains unchanged between the two models.}
+			\label{fig:reducedmodel}
+		\end{center}
+	\end{figure}
+
+	\added[id=siyu]{Secondly, we evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
+In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
+The simulation and re-fitting was repeated for 200 times.
+Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval.
+If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations (this ratio will be referred to as the coverage rate).
+For random noise, the coverage rate is 100\% for both horizon 1, horizon 6, and the horizon difference.
+For deterministic noise, the coverage rate is 66\% for horizon 1 and 69\% for horizon 6.
+By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of recovered parameters, it is clear that our model systematically underestimates deterministic noise (Figure \ref{fig:coverage2}).
+Despite the underestimation of deterministic noise in both horizons, we could still reliably detect the horizon changes of deterministic noise (coverage rate is 97\%).
+This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizons.
+For random noise, our model fitting procedure yields a faithful recovery.
+However, there is a conceptual limitation.
+Because random noise is modeled as non-stimulus-driven noise, it can include both true stochastic neural noise and possible deterministic noises which do not depend on the stimuli.
+Because of this, our random noise estimate provides an upper bound of true `random noise' induced by intrinsic stochastic processes in the brain.}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior_v2.jpg}
+			\caption{Parameter recovery over the posterior distribution of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices.
+Lighter color shades represent the re-fitted posterior to the simulated choices.
+Our model fitting procedure faithfully recovers the non-stimulus-driven random noise (A, B), but systematically underestimates deterministic noise in both horizons (D, E).
+The horizon differences in random noise is also faithfully recovered (C).
+The horizon differences in deterministic noise is also underestimated but not significant (F).}
+			\label{fig:coverage2}
+		\end{center}
+	\end{figure}
+
+	\added[id=siyu]{Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Supplementary Figure S5 and S6).
+The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).
+The strength of correlation between simulated and fit values are strong for both deterministic noise (R $>$ 0.8) and random noise (R $>$ 0.9).
+Despite the strong inter-subject correlations, we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
+
+	%Parameter recovery is good (in terms of) for all parameters apart from the bias in short horizon condition, which is likely due to this parameter being so close to zero (Figure \ref{fig:pararecover1} Panel C).
+The recovery for the noise parameters, $\sigma_{det}$ and $\sigma_{ran}$, is slightly better for horizon 1 than horizon 6.
+
+	\added[id=siyu]{Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
+All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ were tested.
+In a special case, we evaluated how our model performs when there is only random noise or only deterministic noise (Figure \ref{fig:puredetran}).
+In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with fully random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
+However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Supplementary Figures S7).
+Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
+
+	%Overall, we see better recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+
+	%This is because it requires more trials to recover larger values of $\sigma_{det}$ and $\sigma_{ran}$, so with the same number of choices it is harder to recover overall larger noise variances in horizon 6.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/RDBayes_parameterrecovery_grid_pureRanDet.jpg}
+			\caption{Parameter recovery over the posterior of random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for purely random noise (top row) and purely deterministic noise (bottom row) games.
+			}
+			\label{fig:puredetran}
+		\end{center}
+	\end{figure} 
+	% Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
+
+	\added[id=siyu]{Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+In addition, We see better parameter recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+	}
 

### 2024-05-14 codes/EEplot_2noise_hyperpriors.m

@@ -2,8 +2,8 @@ function EEplot_2noise_hyperpriors(plt, sp, stepsize)
     if ~exist('stepsize', 'var')
         stepsize = [0.02, 0.02];
     end
-    names = {'Random noise - \sigma_{ran}','Random noise - \sigma_{ran}',...
-        'Deterministic noise - \sigma_{det}','Deterministic noise - \sigma_{det}'};
+    names = {'Random noise - \sigma_{ran}',{'Change in random noise - \sigma_{ran}'},...
+        'Deterministic noise - \sigma_{det}',{'Change in deterministic noise - \sigma_{det}'}};
     plt.setfig_all('ylabel', 'posterior density', ...
         'ytick', []);
     plt.setfig('xtick',{0:4:50,-3:3:15,0:4:50,-3:3:15});

### 2024-05-14 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
supplemental

@@
 	\section{Model validation analyses}
-
-	MOVE THE REST OF THIS SECTION TO THE SUPPLEMENTARY MATERIAL
-	Firstly, we checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
+	\subsection{Deterministic noise captures known deterministic processes excluded from the model}	
+	We checked if our fitted deterministic noise could indeed capture unobserved deterministic process that was not accounted for by the decision model.
 We test this by leaving out one known deterministic process from the decision model, and ask if our method could recover that known deterministic process as deterministic noise.
@@
 		\begin{center}
-			\includegraphics[width=0.7\textwidth]{figures/RDBayes_reduced_model.jpg}
-			\caption{\added[id=bob]{FONT SIZE ON LEGEND IS VERY SMALL!}Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
+			\includegraphics[width=\textwidth]{figures/RDBayes_reduced_model.jpg}
+			\caption{Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.
 In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
@@
 	\end{figure} 
-
-	\added[id=siyu]{Secondly, we evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
-In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage2}, Supplementary Figure S4).
+	\newpage
+	\subsection{Frequentist coverage analysis}	
+	We next evaluated our hierarchical Bayesian analysis procedure using the `frequentist coverage analysis'.
+In the coverage analysis, we simulated choices with the fitted parameters from the Hierarchical Bayesian analysis, and then re-fit the model to the simulated choices to see whether we can recover the parameters (Figure \ref{fig:coverage}).
 The simulation and re-fitting was repeated for 200 times.
-Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval.
+Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval (Figure \ref{fig:coverage2}).
 If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations (this ratio will be referred to as the coverage rate).
@@
 For deterministic noise, the coverage rate is 66\% for horizon 1 and 69\% for horizon 6.
-By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of recovered parameters, it is clear that our model systematically underestimates deterministic noise (Figure \ref{fig:coverage2}).
+By comparing the posterior distributions of parameters that were used to generate simulations and the posterior distribution of recovered parameters, it is clear that our model systematically underestimates deterministic noise (Figure \ref{fig:coverage},  \ref{fig:coverage2}).
 Despite the underestimation of deterministic noise in both horizons, we could still reliably detect the horizon changes of deterministic noise (coverage rate is 97\%).
@@
 However, there is a conceptual limitation.
-Because random noise is modeled as non-stimulus-driven noise, it can include both true stochastic neural noise and possible deterministic noises which do not depend on the stimuli.
-Because of this, our random noise estimate provides an upper bound of true `random noise' induced by intrinsic stochastic processes in the brain.}
+Because random noise is modeled as non-stimulus-driven noise, it can include both true stochastic random noise and possible deterministic noises which do not depend on the stimuli.
+Because of this, our random noise estimate provides an upper bound of true `random noise' induced by intrinsic stochastic processes in the brain.
 
@@
 The horizon differences in deterministic noise is also underestimated but not significant (F).}
+			\label{fig:coverage}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior.jpg}
+			\caption{Parameter recovery over the mean estimates of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices, dashed black line is the mean of the true posterior.
+Histograms represent the mean estimates of the respective parameters in the refitting to the simulated data.
+(A) and (B) are random noise at H = 1 and H = 6, respectively.
+(C) is the random noise differences between horizons.
+(D) and (E) are deterministic noise at H = 1 and H = 6, respectively.
+(F) is the deterministic noise differences between horizons.}
 			\label{fig:coverage2}
@@
 
-	\added[id=siyu]{Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Supplementary Figure S5 and S6).
+	\newpage
+	\subsection{Parameter recovery of the subject-level parameter fits}
+	Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Figure \ref{fig:paramrecover_session}, \ref{fig:paramrecover_sub}).
 The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).
-The strength of correlation between simulated and fit values are strong for both deterministic noise (R $>$ 0.8) and random noise (R $>$ 0.9).
-Despite the strong inter-subject correlations, we again observed a systematic underestimation of $\sigma_{det}$ (Supplementary Figure S5 and S6).}
+The strength of correlation between simulated and fit values are strong for both deterministic noise (R $>$ 0.8, Figure \ref{fig:paramrecover_sub}) and random noise (R $>$ 0.9, Figure \ref{fig:paramrecover_sub}).
+Despite the strong inter-subject correlations, we again observed a systematic underestimation of $\sigma_{det}$ (Figure \ref{fig:paramrecover_session}, \ref{fig:paramrecover_sub}).
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.65\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession.jpg}
+			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.}
+			\label{fig:paramrecover_session}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.65\textwidth]{figures/RDBayes_parameterrecovery_subject.jpg}
+			\caption{Same as Supplementary Figure S5, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
+			\label{fig:paramrecover_sub}
+		\end{center}
+	\end{figure}
+
+	\newpage	
+	\subsection{Parameter recovery over bigger parameter value ranges}
 
@@
 
-	\added[id=siyu]{Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
+	Lastly, in addition to testing how our model performs in parameter ranges around the actual fitted parameters, we tested the limitations of our models in arbitrary combinations of random vs deterministic noises.
 All combinations of random and deterministic noises with $0 \le \sigma_{det} \le 10$ and $0 \le \sigma_{ran} \le 10$ were tested.
@@
 In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with fully random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
-However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Supplementary Figures S7).
-Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Supplementary Figures S7).}
+However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Figure \ref{fig:gridall}).
+Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Figure \ref{fig:gridall}).
 
-	%Overall, we see better recovery for random noise than deterministic noise.
+	Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+In addition, We see better parameter recovery for random noise than deterministic noise.
 This is likely because we effectively have half as many trials for deterministic noise.
@@
 	\end{figure}
-	% Thus, our model provides a lower bound on deterministic noise and an upper bound on random noise.
-
-	\added[id=siyu]{Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
-Our model provides a lower bound for deterministic noise and an upper bound for random noise.
-In addition, We see better parameter recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-	}
-
-	\newpage
-	\begin{figure}[H]
-		\begin{center}
-			\includegraphics[width=1\textwidth]{figures/RDBayes_parameterrecovery_hyperprior.jpg}
-			\caption{Parameter recovery over the mean estimates of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
-Solid lines are true posterior used to simulate choices, dashed black line is the mean of the true posterior.
-Histograms represent the mean estimates of the respective parameters in the refitting to the simulated data.
-(A) and (B) are random noise at H = 1 and H = 6, respectively.
-(C) is the random noise differences between horizons.
-(D) and (E) are deterministic noise at H = 1 and H = 6, respectively.
-(F) is the deterministic noise differences between horizons.}
-			\label{fig:s4}
-		\end{center}
-	\end{figure}
-
-\newpage
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession.jpg}
-			\caption{Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.}
-			\label{fig:s5}
-		\end{center}
-	\end{figure} 
-\newpage
-
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=0.6\textwidth]{figures/RDBayes_parameterrecovery_subject.jpg}
-			\caption{Same as Supplementary Figure S5, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
-			\label{fig:s6}
-		\end{center}
-	\end{figure} 
-\newpage
 
@@
 Our model did a relatively good job in recovering all combinations of deterministic and random noises.}
-			\label{fig:s7}
+			\label{fig:gridall}
 		\end{center}
@@
 
-	\newpage
+	\section{Additional model-based analyses}
+	\subsection{Replication of Figure 4 without excluding subjects}
 	\begin{figure}[H]
@@
 	\end{figure}
-
 	\newpage
-	\begin{figure}[t]
+	\subsection{Fitting variations of the two-noise model to the data}
+	\begin{figure}[hp]
 		\begin{center}
@@
 
+	\subsection{Model comparison: Posterior checks using variations of the two-noise model}
 	\begin{figure}[hp]
@@
 			\caption{
-			% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
-			% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
-I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
-COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
 			Model comparison.

### 2024-05-14 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
u

@@
 \subsubsection*{Model-based results} 
-
-Posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ are shown in Figure \ref{fig:mb1} and Supplementary Figure S8.
+Posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ are shown in Figure \ref{fig:mb1} and Supplementary Figure S11.
 Consistent with our model-free results, we see that both random and deterministic noise are non-zero.
-\replaced[id=siyu]{Numerically,}{and that} random noise is about 2-3 times larger than the deterministic noise.
-\added[id=siyu]{By computing the posterior distribution of $\sigma^2_{det}/(\sigma^2_{det}+\sigma^2_{ran})$, our data suggests that 14.25\% of the variability in random exploration is accounted for by deterministic noise ([4.90\%, 28.81\%], 95\% CI).} In addition, we find that both random and deterministic noise increase with horizon.
+Numerically, random noise is about 2-3 times larger than the deterministic noise.
+By computing the posterior distribution of $\sigma^2_{det}/(\sigma^2_{det}+\sigma^2_{ran})$, our data suggests that 14.25\% of the variability in random exploration is accounted for by deterministic noise ([4.90\%, 28.81\%], 95\% CI).
+In addition, we find that both random and deterministic noise increase with horizon.
 This increase was larger for random noise (mean = 7.13, 100\% of samples showed an increase in random noise with horizon) than deterministic noise (mean = 2.59, 98.64\% of samples showed an increase in deterministic noise with horizon).
-But intriguingly, the relative increase in both types of noise was \replaced[id=siyu]{similar}{the same} (Figure \ref{fig:ratio}).
-That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is \replaced[id=siyu]{very similar}{almost identical} to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
-%BOB SAYS: CHECK NOTATION HERE AND INSERT FIGURE SHOWING RATIO
+But intriguingly, the relative increase in both types of noise was similar (Figure \ref{fig:ratio}).
+That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is very similar to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
 
@@
 	\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
-	\caption{\added[id=bob]{TITLE ON B AND D SHOULD SAY `CHANGE IN RANDOM NOISE'}Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+	\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
 Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
@@
 	\caption{
-		% BOB SAYS: KEEP THE TOP LINE OF THIS FIGURE FOR THE MAIN PAPER AND MOVE THE FULL FIGURE TO SUPPLEMENTARY MATERIAL TO FIT DESCRIPTION IN TEXT.  
-		% ALSO MAKE THE FOLLOWING CHANGES: THIS FIGURE NEEDS A LEGEND IN TOP RIGHT TO SAY WHICH LINE IS MODEL AND WHICH IS NOT.
-I WOULD ALSO USE TEXTBOXES (in Matlab annotation('textbox', [0 0 1 1], 'string', 'hi') will get you started) TO DESCRIBE THE MODELS RATHER THAN LABELLING MODEL A ETC ...
-COULD EVEN HAVE SOME FORM OF TABLE 1 ON THE LEFT HAND SIDE - SO A COLUMN FOR DETERMINISTIC NOISE AND A COLUMN FOR RANDOM NOISE.
 		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
@@
 \end{figure}
-
-\subsubsection*{Comparison with alternative models}
-COULD POTENTIALLY MOVE THIS TO THE SUPPLEMENT
-
-To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
-These models varied \added[id=siyu]{in} whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
-\added[id=siyu]{Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).}
-
-\begin{table}[h]
-\small
-\centering
-\begin{tabular}{|c|c|c|}
-	\hline
-	Model & Deterministic noise & Random noise \\
-	\hline
-	$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
-	\hline
-	$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
-	\hline
-	$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
-	\hline
-	$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
-	\hline
-	$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
-	\hline
-	$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
-	\hline
-\end{tabular}
-\caption{Variants of the model.}
-\label{tab:models}	
-\end{table}
-\added[id=siyu]{
-The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Supplementary Figure S9.
-We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
-}As shown in Supplementary Figure S10, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of \replaced[id=siyu]{behavior}{responding}.
-However, the quantitative fit to the data is not as good (Supplementary Figure S10).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
-
-\added[id=siyu]{
-Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
-To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model.
-The simulation was repeated 50 times for each model.
-Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}, S).
-This phenomenon was also examined and discussed in the section "Model validation" above.
-}
-
-\begin{figure}[H]
-\begin{center}
-	\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
-	\caption{
-		Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
-A-D.
-both deterministic and random noise are horizon dependent, E-H.
-only random noise is horizon dependent, I-L.
-only deterministic noise is horizon dependent, M-P.
-neither random nor deterministic noise is horizon dependent, Q-T.
-only deterministic noise is assumed to be present, U-X.
-only random noise is assumed to be present.}
-	\label{fig:6model}
-\end{center}
-\end{figure}
-
-%The previous section suggests that behavioral variability in random exploration can be at least partially explained by deterministic processing of the stimulus.
 
@@
 \section*{Discussion}
-
-% BOB SAYS: CAN YOU TRY TO REFRAME THIS AROUND THE DETERMINISTIC STORY
-% \added[id=siyu]{Randomness in exploratory behavior can be driven by both the intrinsic variability generated in the brain (appearing in behavior as random noise) and by deterministic neural processes induced by the stimuli (appearing in behavior as deterministic noise).}
 In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.
 
-%	\added[id=siyu]{We quantified the contribution of both deterministic noise (unobserved deterministic processes of the stimulus) and random noise (intrinsic stochasticity in decision making, or stimulus irrelevant processes) contribute to behavioral variability in random exploration,
-
-SUGGEST REFRAMING NEXT PARAGRAPH IN THE FOLLOWING WAY.
+% SUGGEST REFRAMING NEXT PARAGRAPH IN THE FOLLOWING WAY.
 SIMPLE READ OF RESULTS SUGGESTS RANDOM EXPLORATION IS RANDOM - THIS IS CONSISTENT WITH X.
@@
 
-\added[id=bob]{One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random.
+One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random.
 Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \cite{drugowitsch16}.
@@
 Indeed, a strong interpretation of this hypothesis would suggest that randomness in explore-exploit behavior is due to imperfect inference about the correct course of action.
-In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute CITE DEEP EXPLORATION PAPER.}
+In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute \citep{Wilson2020}.
 
-While the random noise interpretation is theoretically appealing, our approach, while an improvement on previous methods, is not without limitations.
+Although the random noise interpretation is theoretically appealing, our approach, while an improvement on previous methods, is not without limitations.
 Most important is that our measure of `random' noise is only an upper bound on the true level of randomness and that, in principle, the random decision noise could be lower.
-Specifically, in our model, what we labeled random noise was really `non-stimulus-driven variability.' While this non-stimulus-driven variability could be driven by truly random stochastic processes, it could also be driven by deterministic processing that is unrelated to the stimuli in the task.
+Specifically, in our model, what we labeled random noise was really `non-stimulus-driven variability'.
+While this non-stimulus-driven variability could be driven by truly random stochastic processes, it could also be driven by deterministic processing that is unrelated to the stimuli in the task.
 For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}.
-In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:coverage2}, Supplementary Figure S6).
+In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
 As a result, from both a conceptual and methodological perspective, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.
@@
 
-To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation XXX.
+To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}.
 In this model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, where
@@
 
-Such a reduction in the strength of reward coding in exploration, is consistent with our recent work using a drift diffusion model (DDM) to model explore-exploit decisions CITE FENG ET AL.
+Such a reduction in the strength of reward coding in exploration, is consistent with our recent work using a drift diffusion model (DDM) to model explore-exploit decisions \citep{Feng21}.
 In the drift diffusion model, changes in behavioral variability can be driven by changes in the decision threshold (smaller threshold = more noise) or changes in the signal-to-noise ratio with which reward is encoded (lower SNR = more noise).
@@
 By showing that the change in deterministic and random noise have the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise.
-Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology (CITE TOMOV GERSHMAN FMRI, CITE EBITZ).
+Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology \citep{Tomov20, Ebitz18,costa1,costa22}
 
@@
 14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved \deleted[id=siyu]{driven}by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
+In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
 This left 65 for the main analysis.
-Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures 1 - 3)% MAKE AND INCLUDE REPLICAS OF ALL FIGS BUT WITH ALL SUBJECTS
+Note that including the 15 badly performing subjects did not change the main results (Supplementary Figures S1, S2, S11)
 
@@
 \subsection*{Model-based analysis}
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate \replaced[id=siyu]{deterministic noise from random noise.}{between components of the noise that are deterministically driven by the stimulus (`deterministic noise') and components of the noise that are not deterministically driven by the stimulus (`random noise').}Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate deterministic noise from random noise.
+Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.
 

### 2024-05-14 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
u

@@
 	\subsection{Fitting variations of the two-noise model to the data}
+
+	\subsubsection*{Comparison with alternative models}
+	COULD POTENTIALLY MOVE THIS TO THE SUPPLEMENT
+
+	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
+These models varied \added[id=siyu]{in} whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+\added[id=siyu]{Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).}
+
+	\begin{table}[h]
+		\small
+		\centering
+		\begin{tabular}{|c|c|c|}
+			\hline
+			Model & Deterministic noise & Random noise \\
+			\hline
+			$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
+			\hline
+			$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
+			\hline
+			$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
+			\hline
+			$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
+			\hline
+			$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
+			\hline
+			$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
+			\hline
+		\end{tabular}
+		\caption{Variants of the model.}
+		\label{tab:models}	
+	\end{table}
+	\added[id=siyu]{
+		The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Supplementary Figure S9.
+We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
+	}As shown in Supplementary Figure S10, only one of these models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of \replaced[id=siyu]{behavior}{responding}.
+However, the quantitative fit to the data is not as good (Supplementary Figure S10).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
+
+	\added[id=siyu]{
+		Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model.
+The simulation was repeated 50 times for each model.
+Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}, S).
+This phenomenon was also examined and discussed in the section "Model validation" above.
+	}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
+			\caption{
+				Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+			\label{fig:6model}
+		\end{center}
+	\end{figure}
+
 	\begin{figure}[hp]

### 2024-05-16 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
noise paper

@@
 
-\date{\today}
+%\date{\today}
 
@@
 By modeling behavior in this task, we were able to estimate a lower bound on the amount of variability that is deterministically driven by the stimulus and an upper bound on the amount of variability that is random.
-Using this approach, we find evidence that at least 15$\%$ of the variability in random exploration in our studied task can be accounted for by deterministic processing of the stimulus.
-Conversely, this suggests that up to 85$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.
+Using this approach, we find evidence that at least 14$\%$ of the variability in random exploration in our studied task can be accounted for by deterministic processing of the stimulus.
+Conversely, this suggests that up to 86$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.
 Finally, our results suggest that both deterministic and random sources of variability change proportionally to each other as the value of exploration increases, suggesting that a common noise gating mechanism may be at play in random exploration.
@@
 
-	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by deterministic processing of the stimulus (which we refer to as `deterministic noise') vs other processes, including deterministic processing of unobserved stimuli as well as truly stochastic processes (which we refer to as `random noise').
+	In this paper, we investigate the extent to which the apparent randomness in random exploration can be explained by deterministic processing of the stimulus (which we refer to as `deterministic noise') vs other processes, including deterministic processing that is unrelated to the stimuli as well as truly stochastic processes (which we refer to as `random noise').
 To distinguish between these two types of noise, we modify the Horizon Task \citep{wilson2014} to have people face the exact same explore-exploit choice twice.
@@
 Each game consists of either 5 or 10 trials.
-In the first 4 trials, participants are instructed about which bandit to choose, this allows us to manipulate what information from both bandits participants receive before they make their first free choice between the two bandits.
+The first four trials of each game are `forced-choice' trials.
+In the first four trials, participants are instructed about which bandit to choose, this allows us to manipulate what information from both bandits participants receive before they make their first free choice between the two bandits.
 From the 5th trial, participants make free choices between the two bandits.
@@
 
-	The second key feature of the Horizon Task are the 4 `forced-choice trials' at the start of each game that allow us to control exactly what participants know about the two bandits before they make their choice.
+	The second key feature of the Horizon Task are the 4 forced-choice trials at the start of each game that allow us to control exactly what participants know about the two bandits before they make their choice.
 In these forced-choice trials, participants are instructed which of the bandits to play allowing us to control how much information they have about each of the options.
@@
 
-	Relative to the original Horizon Task, the key modification in this paper is to give people `repeated games' (Figure \ref{fig:taskfig}B), in which they see exact same set of forced-choice plays twice in two separate games separated by several minutes in time so as to avoid detection.
+	Relative to the original Horizon Task, the key modification in this paper is to give people `repeated games' (Figure \ref{fig:taskfig}B), in which they see the exact same set of forced-choice plays twice in two separate games separated by several minutes in time so as to avoid detection.
 By repeating the forced-choice plays for each game twice, we can set up a situation where (unbeknownst to the participants) they are faced with the exact same explore-exploit choice, with the exact same stimuli twice.
@@
 Random exploration is quantified as the probability of choosing the option that has the lower average payout in the forced-choice plays in the equal, or [2 2], condition, $p(\mbox{low mean})$.
-The idea here is that, in the equal condition, the optimal strategy is to compute the mean payout for each bandit from the example plays and then always choose the option with the highest mean.
+The idea here is that, in the equal condition, the optimal strategy is to compute the mean payout for each bandit from the forced-choice plays and then always choose the option with the highest mean.
 When participants do not choose the option with the higher mean, the assumption is that this is due to some kind of `decision noise', making the probability of choosing the low mean option a measure of behavioral variability.
@@
 	The free parameters of this model are: the information bonus $A$, which controls the level of directed exploration; the noise standard deviation, $\sigma$, which controls the level of random exploration, and the spatial bias, $b$, which determines the extent to which participants prefer the option on the right.
-These free parameters are fit separately for each participant in each horizon and information condition, allowing us to test whether directed and random exploration increase with horizon.
+These free parameters are fit separately for each participant in each horizon condition, allowing us to test whether directed and random exploration increase with horizon.
 Consistent with previous research, we find that this is indeed the case (Figure \ref{fig:modelfree}C; t(64) = 5.35, p $<$ 0.001.
@@
 We used hierarchical Bayesian analysis to fit the parameters of the model (see Figure \ref{fig:model} for a graphical representation of the model in the style of \cite{lee2014}).
-In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise, $\sigma_{det}^2$ for each participant in each horizon.
+In particular, we fit values of the information bonus $A$, spatial bias $b$, variance of random noise $\sigma_{ran}^2$, and variance of deterministic noise $\sigma_{det}^2$ for each participant in each horizon.
 Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
 
-\subsubsection*{Model validation \label{ch:appendix:bayesrecovery}}
+\subsubsection*{Model validation}
 To be sure that our fit parameter values were meaningful and to understand the limits of our model, we evaluated our model extensively using simulated data.
 This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
-Full details of this analysis are presented in the Supplementary Materials.
 
@@
 However, while the relationship was near perfect for random noise ($\frac{\mbox{\small Recovered } \sigma_{ran}}{\mbox{\small Simulated } \sigma_{ran}} = 1.01$), there was a systematic bias to underestimate the level of deterministic noise by about 32\% ($\frac{\mbox{\small Recovered } \sigma_{det}}{\mbox{\small Simulated } \sigma_{det}} = 0.68$).
-Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (Supplementary Figure XXX).
+Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (See Supplementary Materials section 2.2).
 This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizon conditions.
-	% $\frac{\mbox{\small Recovered } \Delta \sigma_{det}}{\mbox{\small Simulated } \Delta \sigma_{det}} = 0.74$
-
-Thus, overall, we were able to detect both deterministic and random noises using our model.
-Our model provides a lower bound for deterministic noise and an upper bound for random noise.
 In addition, we see better parameter recovery for random noise than deterministic noise.
@@
 In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+	% $\frac{\mbox{\small Recovered } \Delta \sigma_{det}}{\mbox{\small Simulated } \Delta \sigma_{det}} = 0.74$
+
+In addition to the conventional subject-level parameter recovery analysis presented here, we also performed parameter recovery analysis that examined how faithful the full posterior distribution of group-level parameters can be recovered in simulated data (Supplementary Figure S5, S6).
+Qualitatively, we also showed that our way of modeling deterministic noise is capable of capturing known deterministic processes intentionally omitted from the full model (Supplementary Figure S4).
+Full details of these additional analyses are presented in the Supplementary Materials.
+
+Overall, we were able to detect both deterministic and random noises using our model.
+Because random noise is modeled as non-stimulus-driven noise, it can reflect both true stochastic random noise and possible deterministic noises which do not depend on the stimuli.
+Thus conceptually our random noise estimate provides an upper bound for the true `random noise' induced by intrinsic stochastic processes in the brain.
+Thus, our model provides a lower bound for deterministic noise and an upper bound for random noise.
 
@@
 Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
-However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
-	%BOB SAYS: DON'T NEED THE NUMBERS ON THE Y-AXES. 
+}
+	%BOB SAYS: However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.
 	\label{fig:mb1}
@@
 	\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
-The ratio in the standard deviations of noise between horizon 6 and horizon 1 is similar for random and deterministic noise.}
+The ratio in the standard deviations of noises between horizon 6 and horizon 1 is similar for random and deterministic noise.}
 	\label{fig:ratio}
@@
 
-To perform this `posterior predictive check,' we created a set of simulated data by taking the subject-level parameters from the hierarchical Bayesian fits and having the model play the same sequence of games as seen by the subjects.
+To perform this `posterior predictive check', we created a set of simulated data by taking the subject-level parameters from the hierarchical Bayesian fits and having the model play the same sequence of games as seen by the subjects.
 We then applied the same model-free analysis as described in the previous sections to this simulated data set and compared the model's behavior to that of participants.
@@
 The quantitative agreement is almost perfect for p(high info) and for p(inconsistent) in the [1 3] condition, but the model seems to systematically overestimate p(low mean) and p(inconsistent) in [2 2] conditions, although the discrepancy is relatively small (overestimating p(low mean) by 0.054 or 31.37\%, and p(inconsistent) by 0.049 or 27.83\% in [2 2] condition).
+
+As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these models fail to capture all qualitative patterns (Supplementary Figure S13).
+Full details of this analysis can be found in Supplementary Materials section 3.2.
 
@@
 	\caption{
-		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and lies between the pure random and pure deterministic noise prediction.}
+		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and it lies between the pure random and pure deterministic noise prediction.}
 	\label{fig:mb3}
@@
 \section*{Discussion}
-In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 15\% of the overall variability in people's behavior.
+In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 14\% of the overall variability in people's behavior.
 
@@
 One interpretation for this low level of deterministic noise is that most of the variability in random exploration is truly random.
-Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \cite{drugowitsch16}.
+Such a random noise interpretation, would be consistent with recent work showing that variability in perceptual decisions may be driven by imperfections in mental inference \citep{drugowitsch16}.
 In this view, apparently random behavior is not due to sensory processing or response selection, but to suboptimal computations in the brain.
@@
 In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
-As a result, from both a conceptual and methodological perspective, it is possible that the remaining 85\% of the decision noise that is not stimulus-driven noise, could be deterministic.
+As a result, from both a conceptual and methodological perspective, it is possible that the remaining 86\% of the decision noise that is not stimulus-driven noise, could be deterministic.
 
@@
 However, this model could not determine whether the changes in SNR were driven by signal or noise.
-By showing that the change in deterministic and random noise have the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise.
+By showing that the change in deterministic and random noise have approximately the same ratio, the present work suggests that this SNR change is driven by changes in reward-signal processing, not noise.
 Of course, to truly see whether changes in signal or noise are driving random exploration will require more direct measurements of neural processing such as with neuroimaging and electrophysiology \citep{Tomov20, Ebitz18,costa1,costa22}
@@
 80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment.
-14 were excluded on the basis of performance, using the same exclusion criterion as in \citep{wilson2014}.
-In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved by making a completely random choice between the two options and excluded participants with a likelihood smaller than 99.999\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
+15 were excluded on the basis of performance, using the same exclusion criterion as in \cite{wilson2014}.
+In this exclusion criteria, we measured the accuracy of each participant's choices by calculating the percentage of times that a participant chose the bandit with the higher underlying mean payouts in the last choice of a long horizon game, intuitively people should figure out which bandit has a higher mean payout by the last trial and should have an accuracy measure significantly above 50\%, specifically, we computed the likelihood that the measured accuracy can be achieved by making a completely random choice between the two options and excluded participants with a likelihood greater than 0.1\%, in other words, participants who didn't show an accuracy significant above chance with $p < 0.001$ were excluded in the analysis.
 This left 65 for the main analysis.
@@
 Every game started with a fixation cross, then a bar of boxes appeared indicating the horizon for that game.
-For the first 4 trials - the instructed trials, we highlight the box on one of the bandits to instruct the participant to choose that option.
+For the first 4 trials - the instructed `forced-choice' trials, we highlight the box on one of the bandits to instruct the participant to choose that option.
 On these trials, they have to press the corresponding key to reveal the outcome.
@@
 The first four trials of each game were forced-choice trials, in which only one of the options was available for the participant to choose.
-We used these forced-choice instructed trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
+We used these forced-choice trials to manipulate the relative ambiguity of the two options, by providing the participant with different amounts of information about each bandit before their first free choice.
 The four forced-choice trials set up two uncertainty conditions: unequal uncertainty(or [1 3]) in which one option was forced to be played once and the other three times, and equal uncertainty(or [2 2]) in which each option was forced to be played twice.
-After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions),  Figure \ref{fig:taskfig}.
+After the forced-choice trials, participants made either 1 or 6 free choices (two horizon conditions).
 
 \subsection*{Model-based analysis}
-We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \citep{wilson2014} that was modified to differentiate deterministic noise from random noise.
+We modeled behavior on the first free choice of the Horizon Task using a version of the logistic choice model in \cite{wilson2014} that was modified to differentiate deterministic noise from random noise.
 Because the stimuli are identical in the repeated games, by definition, deterministic noise remains the same in repeated games, whereas random noise can change.

### 2024-05-16 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
noise paper

--- 2024-05-14 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
@@
 We test this by leaving out one known deterministic process from the decision model, and ask if our method could recover that known deterministic process as deterministic noise.
-In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition on explore-exploit decisions.
+In particular, we fit a reduced version of our model that only considers reward and ignores the influence of uncertainty condition and spatial bias on explore-exploit decisions.
 	$$\Delta Q= \Delta R+n_{det}+n_{ran}$$
-	Here, $\Delta Q, \Delta R, n_{det}, n_{ran}$ represent the same variables as in the full model.	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
+	Here, $\Delta Q, \Delta R, n_{det}, n_{ran}$ represent the same variables as in the full model (Equation 2 in the manuscript).	If deterministic noise in our model can indeed capture unobserved deterministic processes that's missed by the model, then we would expect to see a higher level of fitted deterministic noise in the reduced model compared to in the full model, whereas the level of random noise should remain unchanged.
 By comparing the	fitted posterior distributions over the group-level means of the deterministic and random noise parameters $\sigma_{det}$ and $\sigma_{ran}$, as expected, we observed an increase in deterministic noise and no change in random noise between the reduced and the full model (Figure \ref{fig:reducedmodel}).
@@
 Then we counted out of the 200 repetitions how many times the true parameter that we simulate the choices from lies in the fitted 95\% confidence interval (Figure \ref{fig:coverage2}).
-If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations (this ratio will be referred to as the coverage rate).
+This ratio will be referred to as the coverage rate.
+If our model fitting is reliable, then the fitted confidence interval should cover the true parameter for more than 95\% of the simulations.
 For random noise, the coverage rate is 100\% for both horizon 1, horizon 6, and the horizon difference.
@@
 
+	Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
+Our model provides a lower bound for deterministic noise and an upper bound for random noise.
+In addition, We see better parameter recovery for random noise than deterministic noise.
+This is likely because we effectively have half as many trials for deterministic noise.
+In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+
 	\begin{figure}[hp]
@@
 			\includegraphics[width=0.65\textwidth]{figures/RDBayes_parameterrecovery_subject.jpg}
-			\caption{Same as Supplementary Figure S5, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
+			\caption{Same as Figure \ref{fig:paramrecover_session}, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
 			\label{fig:paramrecover_sub}
@@
 	\newpage	
-	\subsection{Parameter recovery over bigger parameter value ranges}
+	\subsection{Parameter recovery over a bigger range of arbitrary combinations of deterministic and random noises \label{gridrecovery}}
 
@@
 In the simulation with fully deterministic noise and 0 random noise, our model successfully recovered both random and deterministic noise (Figure \ref{fig:puredetran} C, D), however in the simulation with fully random noise and 0 deterministic noise, although our model successfully recovered random noise, some small proportion of deterministic noise was falsely detected when they should instead be 0 (Figure \ref{fig:puredetran} A, B).
-However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this inflation of deterministic noise anymore (Figure \ref{fig:gridall}).
+However, this phenomenon only exists when the true deterministic noise is 0, once the true deterministic noise is greater than 1, we don't observe this obvious inflation of deterministic noise anymore (Figure \ref{fig:gridall}).
 Apart from this, our model did a fairly good job in recovering all combinations of random and deterministic noises (Figure \ref{fig:gridall}).
-
-	Overall, we are able to detect both deterministic and random noises using our model to a satisfactory extent.
-Our model provides a lower bound for deterministic noise and an upper bound for random noise.
-In addition, We see better parameter recovery for random noise than deterministic noise.
-This is likely because we effectively have half as many trials for deterministic noise.
-In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
 
@@
 	\section{Additional model-based analyses}
-	\subsection{Replication of Figure 4 without excluding subjects}
+	\subsection{Replication of Figure 5 without excluding subjects}
 	\begin{figure}[H]
@@
 	\newpage
-	\subsection{Fitting variations of the two-noise model to the data}
+	\subsection{Alternative variations of the two-noise model}
+	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
+These models varied in whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).
+
+	\begin{table}[h]
+		\small
+		\centering
+		\begin{tabular}{|c|c|c|}
+			\hline
+			Model & Deterministic noise & Random noise \\
+			\hline
+			$\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$  & Horizon dependent & Horizon dependent\\
+			\hline
+			$\sigma^{ran}_{horizon},\sigma^{det}_{}$ & Fixed & Horizon dependent\\
+			\hline
+			$\sigma^{ran}_{},\sigma^{det}_{horizon}$ & Horizon dependent & Fixed\\
+			\hline
+			$\sigma^{ran}_{},\sigma^{det}_{}$ & Fixed & Fixed\\
+			\hline
+			$\sigma^{ran}_{horizon}$ & Horizon dependent & None\\
+			\hline
+			$\sigma^{det}_{horizon}$ & None & Horizon dependent\\
+			\hline
+		\end{tabular}
+		\caption{Variants of the model.}
+		\label{tab:models}	
+	\end{table}
+
+	The posterior distributions over the group-level means of the deterministic and random noise standard deviation $\sigma_{det}$ and $\sigma_{ran}$ (when they exist) in these model variants are shown in Figure \ref{fig:sixmodels}.
+
+	We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
+As shown in Figure \ref{fig:posteriorcheck6models}, only one of these alternative models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of behavior.
+However, the quantitative fit to the data is not as good (Figure \ref{fig:posteriorcheck6models}).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
+
 	\begin{figure}[hp]
@@
 			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperpriors_6model.jpg}
-			\caption{Model based analysis with variations of models.
+			\caption{Model based analysis with alternative models.
 Each row is one model.
@@
 }
-			\label{fig:s9}
+			\label{fig:sixmodels}
 		\end{center}
@@
 
-	\newpage
-
-	\subsection{Model comparison: Posterior checks using variations of the two-noise model}
 	\begin{figure}[hp]
@@
 only random noise is assumed to be present.}
-			\label{fig:s10}
+			\label{fig:posteriorcheck6models}
+		\end{center}
+	\end{figure}
+
+	Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model.
+The simulation was repeated 50 times for each model.
+Indeed, we showed that our model can capture both the existence of random and deterministic noise, and whether each noise changes with horizon condition (Figure \ref{fig:6model}), with only one exception that our model falsely detected a small fraction of deterministic noise when no deterministic noise was present (Figure \ref{fig:6model}).
+This phenomenon was also examined and discussed in the section \ref{gridrecovery} above.
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_hyperprior_6model_recovery.jpg}
+			\caption{
+				Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+			\label{fig:6model}
 		\end{center}

### 2024-05-16 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
final version 240515

@@
 (C) model-based measure of behavioral variability, decision noise $\sigma$, increases with horizon.
-(B) model-based measure of information seeking, information bonus $A$, increases with horizon.}
+(D) model-based measure of information seeking, information bonus $A$, increases with horizon.}
 			\label{fig:modelfree}
@@
 This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
+Full details of this analysis is presented in the Supplementary Materials section 2.
 
@@
 		\includegraphics[width=\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession_randet.jpg}
-		\caption{Parameter recovery analysis for random (A,B) and deterministic noise (C,D) in the two horizons.}
+		\caption{Parameter recovery analysis for random (A,B) and deterministic (C,D) noises in the two horizons.}
 		\label{fig:paramrecover_main}
@@
 However, while the relationship was near perfect for random noise ($\frac{\mbox{\small Recovered } \sigma_{ran}}{\mbox{\small Simulated } \sigma_{ran}} = 1.01$), there was a systematic bias to underestimate the level of deterministic noise by about 32\% ($\frac{\mbox{\small Recovered } \sigma_{det}}{\mbox{\small Simulated } \sigma_{det}} = 0.68$).
-Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (See Supplementary Materials section 2.2).
+Despite this underestimation of deterministic noise in both horizon conditions, the difference in deterministic noise between horizons is much better captured (see Supplementary Materials section 2.2).
 This is because the underestimation of deterministic noise is partially canceled out when the difference is taken between horizon conditions.
@@
 Qualitatively, we also showed that our way of modeling deterministic noise is capable of capturing known deterministic processes intentionally omitted from the full model (Supplementary Figure S4).
-Full details of these additional analyses are presented in the Supplementary Materials.
+Full details of these additional analysis are presented in the Supplementary Materials.
 
@@
 	\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
-Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and increase with horizon (B, D).
 }
@@
 	\includegraphics[width=0.5\textwidth]{figures/RDBayes_ratio__paironly.jpg}
-	\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
-The ratio in the standard deviations of noises between horizon 6 and horizon 1 is similar for random and deterministic noise.}
+	\caption{(A) Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
+The ratios in the standard deviations of noises between horizon 6 and horizon 1 are similar for random and deterministic noise.}
 	\label{fig:ratio}
@@
 
-As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these models fail to capture all qualitative patterns (Supplementary Figure S13).
+As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these reduced models fail to capture all qualitative patterns (Supplementary Figure S13).
 Full details of this analysis can be found in Supplementary Materials section 3.2.
@@
 	\caption{
-		Our model accounts for all qualitative patterns of the data, namely, p(high info) and p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] and [2 2] conditions and it lies between the pure random and pure deterministic noise prediction.}
+		Our model accounts for all qualitative patterns of the data, namely, (A) p(high info) and (B) p(low mean) increase as a function of horizon, p(inconsistent) increases as a function of horizon for both [1 3] (C) and [2 2] (D) conditions and it lies between the pure random and pure deterministic noise prediction.}
 	\label{fig:mb3}

### 2024-05-16 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
final version 240515

@@
 no exclusions).
-Both  $p(\mbox{low mean})$ (A) and $p(\mbox{high info})$ (B) increase with horizon suggesting that people use both random and directed exploration in this task.
-}
+(A) model-free measure of behavioral variability, $p(\mbox{low mean})$, increases with horizon.
+(B) model-free measure of information seeking, $p(\mbox{high info})$, increases with horizon.
+(C) model-based measure of behavioral variability, decision noise $\sigma$, increases with horizon.
+(D) model-based measure of information seeking, information bonus $A$, increases with horizon.}
 			\label{fig:s1}
@@
 	\newpage
-	\subsection{Alternative variations of the two-noise model}
+	\subsection{Alternative models: variations of the two-noise model}
 	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.

### 2024-05-16 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
u

@@
 Using this approach, we find evidence that at least 14$\%$ of the variability in random exploration in our studied task can be accounted for by deterministic processing of the stimulus.
-Conversely, this suggests that up to 86$\%$ of the variability is truly `random,' although it is still possible that this variability is driven by deterministic factors not related to the stimulus.
+Conversely, this suggests that up to 86$\%$ of the variability is truly `random', although it is still possible that this variability is driven by deterministic factors not related to the stimulus.
 Finally, our results suggest that both deterministic and random sources of variability change proportionally to each other as the value of exploration increases, suggesting that a common noise gating mechanism may be at play in random exploration.
@@
 	We used a modified version of the `Horizon Task' \citep{wilson2014} to show the influence of stimulus-driven `deterministic noise' vs non-stimulus-driven `random noise' in explore-exploit decisions (Figure \ref{fig:taskfig}).
-In this task, participants make a series of choices between two slot machines, or `one-armed bandits,' that pay out probabilistic rewards.
+In this task, participants make a series of choices between two slot machines, or `one-armed bandits', that pay out probabilistic rewards.
 They are asked to choose between the two bandits to maximize the total rewards.
@@
 This allowed us to quantify whether deterministic and random noise can be identified under ideal conditions where the behavior is generated by the model with known parameters.
-Full details of this analysis is presented in the Supplementary Materials section 2.
+Full details are presented in the Supplementary Materials section 2.
 
@@
 	\begin{center}
-		\includegraphics[width=\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession_randet.jpg}
+		\includegraphics[width=.95\textwidth]{figures/RDBayes_parameterrecovery_subject_examplesession_randet.jpg}
 		\caption{Parameter recovery analysis for random (A,B) and deterministic (C,D) noises in the two horizons.}
@@
 \begin{center}
-	\includegraphics[width=0.7\textwidth]{figures/RDBayes_hyperprior.jpg}
+	\includegraphics[width=\textwidth]{figures/RDBayes_hyperprior.jpg}
 	\caption{Model based analysis showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
@@
 \section*{Discussion}
-In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise.'  Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise.' In particular, we found that deterministic noise accounted for around 14\% of the overall variability in people's behavior.
+In this paper, we investigated whether random exploration is really random or whether it is driven deterministically by aspects of the stimulus we have previously ignored when measuring `decision noise'.
+Using a version of the Horizon Task with repeated games, we found evidence that at least some of the noise in random exploration could be explained by such `deterministic noise'.
+In particular, we found that deterministic noise accounted for around 14\% of the overall variability in people's behavior.
 

### 2024-05-17 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
ethics statement

@@
 \begin{center}
-	\includegraphics[width=0.5\textwidth]{figures/RDBayes_ratio__paironly.jpg}
-	\caption{(A) Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
+	\includegraphics[width=0.5\textwidth]{figures/RDBayes_ratio.jpg}
+	\caption{Model based analysis showing the posterior distributions over the ratio of the group-level mean of the standard deviations of  random and deterministic noise between horizon 6 and horizon 1 respectivelly.
 The ratios in the standard deviations of noises between horizon 6 and horizon 1 are similar for random and deterministic noise.}
@@
 \section*{Materials and Methods}
+\subsection*{Ethics Statement}
+Human subject protocols were approved by the University of Arizona institutional review board (IRB \# 1411567117).
+Written informed consent was given by all participants prior to participating in the study.
+
 \subsection*{Participants}
-
 80 participants (ages 18-25, 37 male, 43 female) from the University of Arizona undergraduate subject pool participated in the experiment.

### 2024-05-17 codes/mainFigures_bayes.m

@@ -82,4 +82,4 @@ for mi = 1:2
     [tl, tm] = W.JAGS_density(td(:,:,2)./td(:,:,1), xbins);
     plt.plot(tm, tl, [], 'line', 'color', cols{mi});
 end
-plt.update('ratio');
+plt.update('ratio',' ');

### 2025-03-09 codes_original_submission/GRAPHICALMODEL/OlderFiles/horizonTask_graphicalModelBlinks_v1.tex
u

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template graphical model %
% Author: Tom Lodewyckx    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% 0. Begin document
%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{pst-all}
\usepackage{com.braju.graphicalmodels}
%\usepackage[pass,paperwidth=8.5in,paperheight=11in]{geometry}
\usepackage{ifpdf}
\ifpdf
  \pdfpaperheight=6in
  \pdfpaperwidth=9.5in
\else
  \special{papersize=9.3in,5in}
\fi
\catcode`\@=11%
\pagestyle{empty}
%\setlength{\textheight}{9.3in}
%\setlength{\textwidth}{7in}
%\setlength{\footskip}{0.5in}
%\setlength{\topmargin}{-0.5in}
%\setlength{\headheight}{0.2in}
%\setlength{\headsep}{0in}
%\setlength{\parindent}{1pc}
%\setlength{\oddsidemargin}{-0.25in}
%\setlength{\evensidemargin}{-0.25in}
\begin{document}
\TeXtoEPS
\psset{unit=14mm,arrowscale=1.5}
\SpecialCoor

% 1. Grid
%%%%%%%%%

\begin{pspicture}(0.5,2)(13,9)
\showgrid

% 2. Nodes
%%%%%%%%%%

%\rput(1,1){\GM@node[NodeOptions]{name}}
%                 \GM@label[LabelOptions]{name}{label}

% dm and dI
\rput(4,4){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{nst}$}
\rput(6,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{nst}$}

% choices
\rput(5,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{c}}\GM@label[offset=-3mm]{c}{$c_{nst}$}

% subject parameters
\rput(3,8){\GM@node[nodeSize=12mm]{A}}\GM@label[offset=-3mm]{A}{$A_{ns}$}
\rput(5,8){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$B_{ns}$}
\rput(7,8){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma_{ns}$}

% hyperparameters
\rput(0.8,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{n}$}
\rput(1.75,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{n}$}
\rput(2.7,10){\GM@node[nodeSize=12mm]{betaA}}\GM@label[offset=-2mm]{betaA}{$\beta^A_{n}$}

\rput(4.05,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^B_{n}$}
\rput(5,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^B_{n}$}
\rput(5.95,10){\GM@node[nodeSize=12mm]{betab}}\GM@label[offset=-2mm]{betab}{$\beta^B_{n}$}

\rput(7.3,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^\sigma_{n}$}
\rput(8.25,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^\sigma_{n}$}
\rput(9.2,10){\GM@node[nodeSize=12mm]{betaSigma}}\GM@label[offset=-2mm]{betaSigma}{$\beta^\sigma_{n}$}

% blink rate
\rput(-0.5,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{BR}}\GM@label[offset=-4mm]{BR}{$BR_{s}$}

% differences
%\rput(3,13){\GM@node[nodeSize=12mm]{dmuAOUT}}
%\rput(3,13){\GM@node[nodeSize=10mm]{dmuA}}\GM@label[offset=-3mm]{dmuA}{$\delta\mu^A_{u}$}
%\rput(5,13){\GM@node[nodeSize=12mm]{dkSigmaOUT}}
%\rput(5,13){\GM@node[nodeSize=10mm]{dkSigma}}\GM@label[offset=-3mm]{dkSigma}{$\delta k^\sigma_{u}$}
%\rput(7,13){\GM@node[nodeSize=12mm]{dmubOUT}}
%\rput(7,13){\GM@node[nodeSize=10mm]{dmub}}\GM@label[offset=-3mm]{dmub}{$\delta\mu^b_{u}$}

% 3. Arrows
%%%%%%%%%%%
%
\ncline[arrows=->]{dm}{c}
\ncline[arrows=->]{dI}{c}

\ncline[arrows=->]{A}{c}
\ncline[arrows=->]{sigma}{c}
\ncline[arrows=->]{b}{c}

\ncline[arrows=->]{muA}{A}
\ncline[arrows=->]{sigmaA}{A}
\ncline[arrows=->]{betaA}{A}

\ncline[arrows=->]{kSigma}{sigma}
\ncline[arrows=->]{lambdaSigma}{sigma}
\ncline[arrows=->]{betaSigma}{sigma}

\ncline[arrows=->]{mub}{b}
\ncline[arrows=->]{sigmab}{b}
\ncline[arrows=->]{betab}{b}

\ncline[arrows=->]{BR}{A}
\ncline[arrows=->]{BR}{b}
\ncline[arrows=->]{BR}{sigma}

%\ncline[arrows=->]{muA}{dmuAOUT}
%\ncline[arrows=->]{kSigma}{dkSigmaOUT}
%\ncline[arrows=->]{mub}{dmubOUT}

% 4. Plates
%%%%%%%%%%%

\rput(3.25,3){\GM@plate{3.5}{4}{game $t=1$:$T$   }}
\rput(-1.25,2.5){\GM@plate{9.5}{6.5}{subject $s=1$:$S$   }}
\rput(0.25,2){\GM@plate{9.5}{8.75}{condition $n=1$:$N$   }}
%\rput(1,1.5){\GM@plate{10}{10}{$u\in\left\{ [1\mbox{ }3], [2\mbox{ }2]\right\}$   }}

% 5. Model specification
%%%%%%%%%%%%%%%%%%%%%%%%

% priors
\rput(10,10.6){\pnode{priors}}\GM@label{priors}{Priors}
\rput(10,10.05){\pnode{muA}}\GM@label{muA}{$\mu^A_{n} \sim \mbox{Gaussian}(0, 1000)$}
\rput(10,9.6){\pnode{sigmaA}}\GM@label{sigmaA}{$\sigma^A_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
\rput(10,9.15){\pnode{betaA}}\GM@label{betaA}{$\beta^A_{n} \sim \mbox{Gaussian}(0, 1000)$}

\rput(10,8.55){\pnode{mub}}\GM@label{mub}{$\mu^B_{n} \sim \mbox{Gaussian}(0, 1000)$}
\rput(10,8.1){\pnode{sigmab}}\GM@label{sigmab}{$\sigma^B_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
\rput(10,7.65){\pnode{betaB}}\GM@label{betaB}{$\beta^B_{n} \sim \mbox{Gaussian}(0, 1000)$}

\rput(10,7.05){\pnode{kSigma}}\GM@label{kSigma}{$k^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
\rput(10,6.6){\pnode{lambdaSigma}}\GM@label{lambdaSigma}{$\lambda^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
\rput(10,6.15){\pnode{betaSigma}}\GM@label{betaSigma}{$\beta^\sigma_{n} \sim \mbox{Gaussian}(0, 1000)$}

% subject level params
\rput(10,5.5){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
\rput(10,5){\pnode{A}}\GM@label{A}{$A_{ns} \sim \mbox{Gaussian}(\mu^A_{n}, \sigma^A_{n}) + \beta_n^A BR_s$}
\rput(10,4.5){\pnode{b}}\GM@label{b}{$B_{ns} \sim \mbox{Gaussian}(\mu^B_{n}, \sigma^B_{n})+ \beta_n^B BR_s$}
\rput(10,4){\pnode{sigma}}\GM@label{sigma}{$\sigma_{ns} \sim \mbox{Gamma}(k^\sigma_{n}, \lambda^\sigma_{n})+ \beta_n^\sigma BR_s$}

% choices!
%\rput(9,5.5){\pnode{c}}
%\GM@label{c}{$p \leftarrow }  $}

\rput(10,3.4){\pnode{choices}}\GM@label{choices}{Observed choices}
\rput(10,2.9){\pnode{p}}
\GM@label{p}{$p_{nst} \leftarrow
\left[1+\exp\left( \frac{\Delta R_{nst}+A_{ns}\Delta I_{nst} + B_{ns}}{\sigma_{ns}} \right)\right]^{-1} $}
\rput(10,2.3){\pnode{c}}
\GM@label{c}{$c_{nst} \sim \mbox{Bernoulli}
\left( p_{nst} \right)$}

% 6. End document
%%%%%%%%%%%%%%%%%

\end{pspicture}

\endTeXtoEPS
\end{document}

### 2025-03-09 codes_original_submission/GRAPHICALMODEL/OlderFiles/horizonTask_graphicalModel_v3.tex
u

--- 2025-03-09 codes_original_submission/GRAPHICALMODEL/OlderFiles/horizonTask_graphicalModelBlinks_v1.tex
@@
 \else
-  \special{papersize=9.3in,5in}
+  \special{papersize=7.6in,5in}
 \fi
@@
 
-\begin{pspicture}(0.5,2)(13,9)
-\showgrid
+\begin{pspicture}(2.95,2)(13,9)
+%\showgrid
 
@@
 % dm and dI
-\rput(4,4){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{nst}$}
-\rput(6,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{nst}$}
+\rput(3,4){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{nst}$}
+\rput(5,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{nst}$}
+\rput(7,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dl}}\GM@label[offset=-5.7mm]{dl}{$\Delta l_{nst}$}
 
@@
 % hyperparameters
-\rput(0.8,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{n}$}
-\rput(1.75,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{n}$}
-\rput(2.7,10){\GM@node[nodeSize=12mm]{betaA}}\GM@label[offset=-2mm]{betaA}{$\beta^A_{n}$}
+\rput(2,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{n}$}
+\rput(3,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{n}$}
 
-\rput(4.05,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^B_{n}$}
-\rput(5,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^B_{n}$}
-\rput(5.95,10){\GM@node[nodeSize=12mm]{betab}}\GM@label[offset=-2mm]{betab}{$\beta^B_{n}$}
+\rput(4.5,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^B_{n}$}
+\rput(5.5,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^B_{n}$}
 
-\rput(7.3,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^\sigma_{n}$}
-\rput(8.25,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^\sigma_{n}$}
-\rput(9.2,10){\GM@node[nodeSize=12mm]{betaSigma}}\GM@label[offset=-2mm]{betaSigma}{$\beta^\sigma_{n}$}
-
-% blink rate
-\rput(-0.5,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{BR}}\GM@label[offset=-4mm]{BR}{$BR_{s}$}
+\rput(7,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^\sigma_{n}$}
+\rput(8,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^\sigma_{n}$}
 
@@
 \ncline[arrows=->]{dI}{c}
+\ncline[arrows=->]{dl}{c}
 
@@
 \ncline[arrows=->]{sigmaA}{A}
-\ncline[arrows=->]{betaA}{A}
 
@@
 \ncline[arrows=->]{lambdaSigma}{sigma}
-\ncline[arrows=->]{betaSigma}{sigma}
 
@@
 \ncline[arrows=->]{sigmab}{b}
-\ncline[arrows=->]{betab}{b}
-
-\ncline[arrows=->]{BR}{A}
-\ncline[arrows=->]{BR}{b}
-\ncline[arrows=->]{BR}{sigma}
 
@@
 
-\rput(3.25,3){\GM@plate{3.5}{4}{game $t=1$:$T$   }}
-\rput(-1.25,2.5){\GM@plate{9.5}{6.5}{subject $s=1$:$S$   }}
-\rput(0.25,2){\GM@plate{9.5}{8.75}{condition $n=1$:$N$   }}
+\rput(2.25,3){\GM@plate{5.5}{4}{game $g=1$:$G$   }}
+\rput(1.75,2.5){\GM@plate{6.5}{6.5}{subject $s=1$:$S$   }}
+\rput(1.25,2){\GM@plate{7.5}{8.75}{condition $n=1$:$N$   }}
 %\rput(1,1.5){\GM@plate{10}{10}{$u\in\left\{ [1\mbox{ }3], [2\mbox{ }2]\right\}$   }}
@@
 % priors
-\rput(10,10.6){\pnode{priors}}\GM@label{priors}{Priors}
-\rput(10,10.05){\pnode{muA}}\GM@label{muA}{$\mu^A_{n} \sim \mbox{Gaussian}(0, 1000)$}
-\rput(10,9.6){\pnode{sigmaA}}\GM@label{sigmaA}{$\sigma^A_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
-\rput(10,9.15){\pnode{betaA}}\GM@label{betaA}{$\beta^A_{n} \sim \mbox{Gaussian}(0, 1000)$}
+\rput(9,10.5){\pnode{priors}}\GM@label{priors}{Priors}
+\rput(9,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{n} \sim \mbox{Gaussian}(0, 1000)$}
+\rput(9,9.5){\pnode{sigmaA}}\GM@label{sigmaA}{$\sigma^A_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
 
-\rput(10,8.55){\pnode{mub}}\GM@label{mub}{$\mu^B_{n} \sim \mbox{Gaussian}(0, 1000)$}
-\rput(10,8.1){\pnode{sigmab}}\GM@label{sigmab}{$\sigma^B_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
-\rput(10,7.65){\pnode{betaB}}\GM@label{betaB}{$\beta^B_{n} \sim \mbox{Gaussian}(0, 1000)$}
+\rput(9,9){\pnode{mub}}\GM@label{mub}{$\mu^B_{n} \sim \mbox{Gaussian}(0, 1000)$}
+\rput(9,8.5){\pnode{sigmab}}\GM@label{sigmab}{$\sigma^B_{n} \sim \mbox{Gamma}(0.001, 0.001)$}
 
-\rput(10,7.05){\pnode{kSigma}}\GM@label{kSigma}{$k^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
-\rput(10,6.6){\pnode{lambdaSigma}}\GM@label{lambdaSigma}{$\lambda^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
-\rput(10,6.15){\pnode{betaSigma}}\GM@label{betaSigma}{$\beta^\sigma_{n} \sim \mbox{Gaussian}(0, 1000)$}
+\rput(9,8){\pnode{kSigma}}\GM@label{kSigma}{$k^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
+\rput(9,7.5){\pnode{lambdaSigma}}\GM@label{lambdaSigma}{$\lambda^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
 
 % subject level params
-\rput(10,5.5){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
-\rput(10,5){\pnode{A}}\GM@label{A}{$A_{ns} \sim \mbox{Gaussian}(\mu^A_{n}, \sigma^A_{n}) + \beta_n^A BR_s$}
-\rput(10,4.5){\pnode{b}}\GM@label{b}{$B_{ns} \sim \mbox{Gaussian}(\mu^B_{n}, \sigma^B_{n})+ \beta_n^B BR_s$}
-\rput(10,4){\pnode{sigma}}\GM@label{sigma}{$\sigma_{ns} \sim \mbox{Gamma}(k^\sigma_{n}, \lambda^\sigma_{n})+ \beta_n^\sigma BR_s$}
+\rput(9,6.5){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
+\rput(9,6){\pnode{A}}\GM@label{A}{$A_{ns} \sim \mbox{Gaussian}(\mu^A_{n}, \sigma^A_{n})$}
+\rput(9,5.5){\pnode{b}}\GM@label{b}{$B_{ns} \sim \mbox{Gaussian}(\mu^B_{n}, \sigma^B_{n})$}
+\rput(9,5){\pnode{sigma}}\GM@label{sigma}{$\sigma_{ns} \sim \mbox{Gamma}(k^\sigma_{n}, \lambda^\sigma_{n})$}
 
@@
 
-\rput(10,3.4){\pnode{choices}}\GM@label{choices}{Observed choices}
-\rput(10,2.9){\pnode{p}}
+\rput(9,4){\pnode{choices}}\GM@label{choices}{Observed choices}
+\rput(9,3.5){\pnode{p}}
 \GM@label{p}{$p_{nst} \leftarrow
-\left[1+\exp\left( \frac{\Delta R_{nst}+A_{ns}\Delta I_{nst} + B_{ns}}{\sigma_{ns}} \right)\right]^{-1} $}
-\rput(10,2.3){\pnode{c}}
+\left[1+\exp\left( \frac{\Delta R_{nst}+A_{ns}\Delta I_{nst} + B_{ns}\Delta l_{nst}}{\sigma_{ns}} \right)\right]^{-1} $}
+\rput(9,2.9){\pnode{c}}
 \GM@label{c}{$c_{nst} \sim \mbox{Bernoulli}

### 2025-03-09 OLD/oldversions/Noise/#1 External-Internal Noise/BAYESIAN/OLD_get_bayesiandata.m

@@ -1,74 +1,0 @@
-%% load data
-clear
-datadir = '..\';
-filename = 'BEHAVIOR_AU_075.mat';
-thres = 0.65;
-savename = sprintf('BAYES_AU_075_%d', round(thres*100));
-data = importdata(fullfile(datadir, filename));
-ind = data.sub.ac10 >= thres;
-% %% data for bayes -only repeated trials
-% savename = [savename, '_paired'];
-% bayesdata.nh = 2;
-% bayesdata.nS = length(data);
-% for si = 1:bayesdata.nS
-%     gd = data{si};
-%     nS(si) = gd.n_trial;
-% end
-% LEN = max(nS);
-% bayesdata.nForcedTrials = 4;
-% for si = 1:bayesdata.nS
-%     gd = data{si};
-%     ind = gd.gID > 0;
-%     nT = sum(ind);
-%     bayesdata.nTrial(si) = nT;
-%     bayesdata.dMu(si,:) = nanextend(gd.dM(ind),LEN);
-%     bayesdata.horizon(si,:) = nanextend(round((gd.cond_horizon(ind) + 4)/5), LEN);
-%     bayesdata.dInfo(si,:) = nanextend(gd.cond_info(ind),LEN);
-%     bayesdata.choice(si,:) = nanextend((gd.key5(ind) == 1) + 0,LEN);
-%     [~, ~, ranking] = unique(gd.gID(ind));
-%     bayesdata.ngID(si) = max(ranking);
-%     bayesdata.gID(si,:) = nanextend(ranking,LEN);
-% end
-% save(fullfile('./Model_data/',savename),'bayesdata');
-% %% control data -only repeated trials
-% clearvars -except savename, clc;
-% load(fullfile('./Model_data/',savename));
-% for si = 1:bayesdata.nS
-%     nT = sum(~isnan(bayesdata.gID(si,:)));
-%     bayesdata.gID(si,1:nT) = bayesdata.gID(si,randperm(nT));
-% end
-% save(fullfile('./Model_data/',[savename '_control']),'bayesdata');
-% %% data for bayes -only repeated trials & [2 2]
-% savename = [savename, '_paired', '_22'];
-% bayesdata.nh = 2;
-% bayesdata.nS = length(data);
-% for si = 1:bayesdata.nS
-%     gd = data{si};
-%     ind = gd.gID > 0 & gd.cond_info == 0;
-%     nT = sum(ind);
[…]

### 2025-03-09 OLD/oldversions/Noise/#1 External-Internal Noise/BAYESIAN/analysis_bayesian.m

@@ -1,101 +1,0 @@
-classdef analysis_bayesian < handle
-    properties
-        data
-        xfit
-        nchains
-        nburnin
-        nsamples
-        thin
-        modelname
-        params
-        init0
-        modelpath
-        savepath
-        poolobj
-        stats
-        samples
-    end
-    methods
-        function obj = analysis_bayesian(data, modelpath, savepath)
-            obj.poolobj = [];
-            obj.savepath = savepath;
-            obj.modelpath = modelpath;
-            obj.data = data;
-            obj.setup_chain;
-        end
-        function setup_chain(obj)
-            obj.nchains = 4; % How Many Chains?
-            obj.nburnin = 2000; % How Many Burn-in Samples?
-            obj.nsamples = 5000; % How Many Recorded Samples?
-            obj.thin = 1;
-        end
-        function analysis(obj, modelname, params)
-            obj.openpool;
-            data = obj.data;
-            obj.init0 = obj.get_bayesinit(params, obj.nchains, [data.nh, data.nS]);
-            modelfile = fullfile(obj.modelpath, [modelname '.txt']);
-            disp(['Fitting now:' modelfile]);
-            [samples, stats, structArray, tictoc] = obj.fit_matjags(modelfile, data, obj.init0, obj.nchains, obj.nburnin, obj.nsamples, params);
-            obj.stats = stats;
-            obj.samples = samples;
-            save(fullfile(obj.savepath, [modelname '_new2']),'stats','tictoc','modelname');
-            save(fullfile(obj.savepath, [modelname '_samples_new2']),'samples','-v7.3');
-            obj.closepool;
-                 end
-        function init0 = get_bayesinit(obj, params, nchains, msize)
-            if nargin < 3
-                msize = [1, 1];
-            end
-            for i=1:nchains
-                S = [];
-                for j = 1:length(params)
-                    str = params{j};
-                    if ~isempty(strfind(str,'_p'))
-                        S.(str) = ones(msize);
-                    elseif ~isempty(strfind(str,'_n'))
-                        S.(str) = zeros(msize);
-                    end
-                end
-                init0(i) = S;
-            end
-        end
-        function openpool(obj)
-            if isempty(obj.poolobj)
-                poolobj = gcp('nocreate'); % If no pool, do not create new one.
-                if isempty(poolobj)
-                    poolsize = 0;
-                else
-                    poolsize = poolobj.NumWorkers;
-                end
-                if poolsize == 0
-                    parpool;
[…]

### 2025-03-09 OLD/oldversions/Noise/#1 External-Internal Noise/select_dataset.m

@@ -1,16 +1,0 @@
-clear all, clc;
-addpath('S:\ANALYSIS\HORIZONTASK');
-explist = {'AU_2015_personality', 'AU_2015_testRetestAndPersonality', ...
-    'AU_017-Data_Fall2016','AU_075_Horizon1_Original'};
-%     explist = {'AU_075_Horizon1_Original'};
-cond_exptrial = 1;
-cond_site = 1;
-cond_repeated = 1;
-[data, summary] = select_datasets(cond_exptrial, cond_site, cond_repeated, explist);
-save('data_arizona','data','summary');
-explist = {};
-cond_exptrial = 1;
-cond_site = 2;
-cond_repeated = 1;
-[data, summary] = select_datasets(cond_exptrial, cond_site, cond_repeated, explist);
-save('data_princeton','data','summary');

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEplot_2noise_PureRanDet.m

@@ -1,48 +1,0 @@
-function plt = EEplot_2noise_PureRanDet(plt, sp, sp2)
-names = {'Random noise only - \sigma_{ran}','','Deterministic noise only - \sigma_{det}',...
-    ''};
-
-ymax = 2.5;
-plt.figure(2,2,'istitle',0);
-plt.setup_pltparams('fontsize_leg', 20, 'linewidth', 2);
-plt.setfig('xlim', {[-1 22],[-3 12], [-1 22], [-3 12]}, 'ylim', [0 ymax], ...
-    'ytick', [0:0.5:ymax], 'xtick',{0:4:20,-3:3:12,0:4:20,-3:3:12});
-color = {'AZblue','AZred'};
-fns = {'NoiseRan', 'NoiseDet'};
-stepsize = 0.02;
-xbins = [-10:stepsize:30];
-plt.setfig('color', color, 'legend', {'h = 1','h = 6'},...
-    'title', {'Random noise - \sigma_{ran}','Deterministic noise - \sigma_{det}','',''}, ...
-    'xlabel', {'','','noise standard deviation [points]', 'noise standard deviation [points]'}, ...
-    'ylabel', names, 'legloc', 'NorthEast');
-for i = 1:2
-    fn = fns{i};
-    plt.new;
-    for hi = 1:2
-        td = sp.(fn);
-        td = squeeze(td(:,:,hi));
-        td = reshape(td, 1, []);
-        tl = hist(td, xbins)/(length(td)*stepsize);
-        plt.lineplot(tl, [], xbins);
-    end
-    hold on;
-    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
-    hold off;
-end
-for i = 1:2
-    fn = fns{i};
-    plt.new;
-    for hi = 1:2
-        td = sp2.(fn);
-        td = squeeze(td(:,:,hi));
-        td = reshape(td, 1, []);
-        tl = hist(td, xbins)/(length(td)*stepsize);
-        plt.lineplot(tl, [], xbins);
-    end
-    hold on;
-    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
-    hold off;
-end
-plt.update;
-plt.save(['pure_ran_det_simu_hyperprior']);
-end

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEplot_2noise_modelcomparison.m

@@ -1,43 +1,0 @@
-function EEplot_2noise_modelcomparison(plt, gp, simugp)
-    plt.figure(6,4);
-    plt.setup_pltparams('fontsize_face', 20, 'fontsize_leg', 10, 'linewidth', 2);
-    
-    plt.setfig('xlabel', [repmat({''},1,20), repmat({'horizon'},1,4)], ...
-        'ylabel', [{'model A'} repmat({''},1,3), {'model B'} repmat({''},1,3), ...
-        {'model C'} repmat({''},1,3), {'model D'} repmat({''},1,3), ...
-        {'model E'} repmat({''},1,3), {'model F'} repmat({''},1,3)], ...
-        'title', [{'p(high info)', 'p(low mean)', 'p(inconsistent), [1 3]', 'p(inconsistent), [2 2]', ...
-        } repmat({''}, 1, 20)], 'xlim', [0.5 2.5], 'xtick', [1 2], 'xticklabel', [1 6], ...
-        'legend', [repmat({{'',''}},1,23), {{'data', 'pure ran', 'model'}}], ...
-        'legloc', 'SouthEast', ...
-        'ylim', repmat({[0.35 0.65],[0 0.35],[0 0.45],[0 0.45]},1,6), ...
-        'ytick', repmat({[0.4:0.1:0.6],[0:.1:0.3],[0:.1:0.5],[0:.1:0.5]},1,6));
-    for mi = 1:6
-        tgp = simugp.(['model' char(64+mi)]);
-        plt.new;
-        plt.setfig_ax('color', {'AZred', 'AZblue'});
-        plt.lineplot(gp.av_p_hi13, gp.ste_p_hi13);
-        plt.lineplot(tgp.av_p_hi13, tgp.ste_p_hi13);
-        
-        plt.new;
-        plt.setfig_ax('color', {'AZred', 'AZblue'});
-        plt.lineplot(gp.av_p_lm22, gp.ste_p_lm22);
-        plt.lineplot(tgp.av_p_lm22, tgp.ste_p_lm22);
-        
-        plt.new;
-        plt.setfig_ax('color', {'AZred', 'AZred50', 'AZblue50', 'AZblue'});
-        plt.lineplot(gp.av_p_inconsistent13, gp.ste_p_inconsistent13);
-        plt.lineplot(gp.av_p_inconsistent13_randomtheory, gp.ste_p_inconsistent13_randomtheory, [], '--');
[…]

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEplot_2noise_parameter_recovery.m

@@ -1,27 +1,0 @@
-function EEplot_2noise_parameter_recovery(plt, p1, p2)
-    plt.figure(4,2,'istitle');
-    plt.setup_pltparams('islegmark', false, 'dotsize', 15, 'fontsize_leg', 15, 'fontsize_face', 20, ...
-        'linewidth', 2, 'islegbox', 0);
-    fns = {'Infobonus_sub', 'bias_sub', 'NoiseRan_sub', 'NoiseDet_sub'};
-    plt.setfig('color', {{'black'},{'black'},{'black'},{'black'},{'black'},{'black'},{'black'},{'black'}},...
-        'ylabel', {'Fit A','Fit A','Fit b','Fit b', ...
-        'Fit \sigma^{ran}', 'Fit \sigma^{ran}', 'Fit \sigma^{det}', 'Fit \sigma^{det}'}, ...
-        'xlabel', {'Simulated A','Simulated A','Simulated b','Simulated b', ...
-        'Simulated \sigma^{ran}', 'Simulated \sigma^{ran}', 'Simulated \sigma^{det}', 'Simulated \sigma^{det}'}, ...
-        'title', {'H = 1', 'H = 6', '','','','','',''},...
-        'legloc', 'NorthWest');
-    for i = 1:4
-        fn = fns{i};
-        d1 = p1.(fn);
-        d2 = p2.(fn);
-        for hi = 1:2
-            plt.new;
-            str{i*2-2+hi} = plt.scatter(d1(hi,:)', d2(hi,:)', 'diag');
-        end
-    end
-    plt.setfig('legend', strcat({'A', 'A','b','b',...
-        '\sigma_{ran}','\sigma_{ran}','\sigma_{det}','\sigma_{det}'}, ...
-        {', ',', ',', ',', ',', ',', ',', ',', '}, str));
-    plt.update;
-    plt.save('parameter_recovery');
-end

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/EEsimulate_bayes_2noise.m

@@ -1,44 +1,0 @@
-function out = EEsimulate_bayes_2noise(data, As, bs, nrans, ndets)
-    disp(['Generating simulation']);
-    if size(ndets, 2) == 1
-        ndets = repmat(ndets, 1, 2);
-    end
-    if size(nrans, 2) == 1
-        nrans = repmat(nrans, 1, 2);
-    end
-    
-    for si = 1:data.nSubject
-        choice = [];
-        nT = data.nTrial(si);
-        dR = data.dR(si,1:nT)';
-        dI = data.dI(si,1:nT)';
-        repeatID = data.repeatID(si,1:nT)';
-        horizon = data.horizon(si,1:nT)';
-        n_det = NaN(1, data.nrepeatID(si));
-        for gi = 1:nT
-            if nrans(si, horizon(gi)) > 0
-                pdni = makedist('Logistic',0, nrans(si, horizon(gi)));
-                dni = random(pdni);
-            else
-                dni = 0;
-            end
-            if isnan(n_det(repeatID(gi)))
-                if ndets(si, horizon(gi)) > 0
-                    pdne = makedist('Logistic',0, ndets(si, horizon(gi)));
-                    n_det(repeatID(gi)) = random(pdne);
-                else
-                    n_det(repeatID(gi)) = 0;
-                end
-            end
-            dne = n_det(repeatID(gi));
-            dQ = dR(gi) + As(si, horizon(gi)) * dI(gi) + bs(si, horizon(gi)) + dne + dni;
-            if dQ > 0
-                choice(gi) = 1;
-            else
-                choice(gi) = 0;
-            end
-        end
-        data.choice(si, :) = W.extend_nan(choice, size(data.choice, 2));
-    end
-    out = data;
-end

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/codes_import/Import_EE_blinkCuriosity_2016S.m

@@ -1,31 +1,0 @@
-datadir = '../../data/EE_blinkCuriosity_2016S/data';
-outputdir = '../../data';
-%% import EE_16S_blinkCuriosity
-filetemplate = 'P075*horizontask*';
-version = 'timestamp_v1';
-importfunc = @(x)EEimport_game(x.game, version);
-str_subID = {'_sub','_'};
-num_subID = [1,3];
-str_datetime = {'_','.mat'}; 
-num_datetime = [3,1];
-subIDrange = [1 200];
-outputname = 'EE_blinkCuriosity_2016S';
-data = W_import.IMPORT_datafiles_subID_yyyymmddTHHMMSS(datadir, filetemplate, importfunc, ...
-    str_subID, num_subID, str_datetime, num_datetime, subIDrange);
-%%
-filetemplate = 'P075*demographicinfo*';
-importfunc = @(x)assist_import_demo(x, 'demo_siyu_v1');
-sub = W_import.IMPORT_datafiles_subID_yyyymmddTHHMMSS(datadir, filetemplate, importfunc, ...
-    str_subID, num_subID, str_datetime, num_datetime, subIDrange);
-%%
-data = W.tab_join(data, sub, {'subjectID', 'date', 'time'});
-%%
-filetemplate = 'P075*horizontask*';
-importfunc = @(x)assist_import_demo(x, 'demo_siyu_v1b');
-sub2 = W_import.IMPORT_datafiles_subID_yyyymmddTHHMMSS(datadir, filetemplate, importfunc, ...
-    str_subID, num_subID, str_datetime, num_datetime, subIDrange);
-sub2 = sub2(sub2.tasktime > 1,:);
-%%
-[data] = W.tab_join(data, sub2);
-%%
-writetable(data, fullfile(outputdir, ['Imported_', outputname, '.csv']));

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/mainALL_ANALYSIS_FIGURE_RanDetNoise.m

@@ -1,60 +1,0 @@
-%% main analysis 2016
-rawdatadir = '../data';
-savedir = '../data/all';
-files = dir(fullfile(rawdatadir,'Imported_EE_blinkCuriosity_2016S.csv'));
-filename = fullfile(files.folder, files.name);
-%% basic analysis
-opt_sub = {'subjectID', 'date', 'time'}; 
-opt_preprocess = 'EEpreprocess_game_basic';
-opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};
-opt_analysis = {'EEanalysis_sub_basic'};
-[sub0, game, idxsub] = W_sub.analysis_pipeline(filename, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, savedir);
-W.writetable(game, fullfile(savedir,'data_all.csv'));
-%% exclude participants
-sub0 = readtable('../data/all/SUB_EE_blinkCuriosity_2016S.csv', 'Delimiter', ',');
-load('../data/all/idxsub_EE_blinkCuriosity_2016S.mat');
-[suball, idall] = W_sub.excludesubject(sub0, 'age', [18 NaN]);
-[sub, id] = W_sub.excludesubject(sub0, 'age', [18 NaN], 'pvalue_ac', [NaN 0.05]);
-save(fullfile(savedir, 'idxsub_exclude.mat'), 'id', 'idall','idxsub');
-%% analyze group
-clear gp;
-additionalcompare = {{'p_inconsistent13'},{'p_inconsistent22'},...
-    {'p_inconsistent13','p_inconsistent13_randomtheory'},{'p_inconsistent22','p_inconsistent22_randomtheory'}};
-gp{1} = W_sub.analysis_1group(sub, true, additionalcompare);
-gp{2} = W_sub.analysis_1group(suball, true, additionalcompare);
-suffix = {'','_all'};
-%%
-plt = W_plt('isshow', 1);
-for gi = 1:2
-%% model-free plots
-    plt.setup_W_plt('fig_dir', '../figures/all','fig_suffix', suffix{gi},'fig_projectname', 'RanDetNoise');
-    plt.param_fig.islocked = true;
-    plt.setfig(2, 'color', {{'AZsand','AZcactus'}, {'AZcactus'}}, ...
-        'ylim', {[0, 0.3] + gi*0.05,[0.35 0.65]}, ...
-        'ytick', {0.1:.1:0.4, 0.4:.1:0.6});
-    EEplot_modelfree(plt, gp{gi},1,1);
[…]

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/mainBOTH_FIGURE_6model.m

@@ -1,64 +1,0 @@
-%% setup
-ver = 'all';
-verout = 'all_20chains';
-% ver = 'paironly';
-% verout = ver;
-outputdir = fullfile('../bayesoutput/', verout);
-figdir = '../figures';
-plt = W_plt;
-plt.setup_W_plt('fig_dir', fullfile(figdir,verout), 'fig_suffix', '','fig_projectname', 'RanDetNoise');
-%% 6-model comparison (simulation)
-%%
-% basic analysis of simulated behavior
-load(fullfile('../data',ver, 'bayes_6model_simu.mat'));
-files = dir(fullfile('../data', ver, sprintf('data_%s.csv', ver)));
-filename = fullfile(files.folder, files.name);
-game0 = readtable(filename, 'Delimiter', ',');
-idx = load(fullfile('../data', ver, 'idxsub_exclude'));
-tid = idx.idxsub(idx.id);
-game = [];
-for mi = 1:6
-    game{mi} = [];
-    for si = 1:length(tid)
-        tgame = game0(tid{si},:);
-        tgame.choice_5 = simu{mi}.choice(si,1:simu{mi}.nTrial(si))' + 1;
-        game{mi} = vertcat(game{mi} , tgame);
-    end
-    opt_sub = {'subjectID', 'date', 'time'};
-    opt_preprocess = 'EEpreprocess_game_basic';
-    opt_game_sub = W.iif(strcmp(ver, 'all'), {'EEpreprocess_game_sub_repeatedgame'},...
-        {'EEpreprocess_game_sub_repeatedpaironly'});
-    opt_analysis = {'EEanalysis_sub_basic'};
-    [simusub{mi}] = W_sub.analysis_pipeline(game{mi}, opt_sub, opt_preprocess, opt_game_sub, opt_analysis);
-    simugp.(['model' char(64 + mi)]) = W_sub.analysis_1group(simusub{mi});
-end    
-sub = readtable(fullfile('../data', ver, 'SUB_EE_blinkCuriosity_2016S.csv'),'Delimiter', ',');
-gp = W_sub.analysis_1group(sub(idx.id,:));
-%% plotting 6 panel
-EEplot_2noise_modelcomparison1(plt, gp, simugp);
-EEplot_2noise_modelcomparison(plt, gp, simugp);
-%% parameter recovery of fitted simulations (A-F)
[…]

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/mainBOTH_FIGURE_bayes2noise.m

@@ -1,133 +1,0 @@
-%% setup
-ver = 'all_20chains';
-outputdir = fullfile('../bayesoutput', ver);
-figdir = fullfile('../figures', ver);
-plt = W_plt;
-%% figure - posterior 
-suffix = {'','_all'};
-for gi = 1:2
-    plt.setup_W_plt('fig_dir', figdir,'fig_suffix', suffix{gi},'fig_projectname', 'RanDetNoise');
-    plt.setfig_loc(4, 'xlim', {[-1 10+6*gi],[-3 12], [-1 10+6*gi], [-3 12]});
-    paramsub = load(fullfile(outputdir, ['HBI_DetRanNoise' suffix{gi} '_samples.mat'])).samples;
-    EEplot_2noise_hyperpriors(plt, paramsub);
-end
-%% reduced models - dRonly, dIonly, 0model, no-bias
-plt.initialize('new', 'fig_dir', figdir,'fig_suffix', '', 'fig_projectname', 'RanDetNoise');
-sp = {};
-sp{2} = load(fullfile(outputdir, ['HBI_DetRanNoise_samples.mat'])).samples;
-% sp{2} = load(fullfile(outputdir, ['HBI_DetRanNoise_nobias_samples.mat'])).samples;
-sp{1} = load(fullfile(outputdir, ['HBI_DetRanNoise_dRonly_samples.mat'])).samples;
-% sp{4} = load(fullfile(outputdir, ['HBI_DetRanNoise_dIonly_samples.mat'])).samples;
-% sp{5} = load(fullfile(outputdir, ['HBI_DetRanNoise_0model_samples.mat'])).samples;
-plt.setfig_new;
-EEplot_2noise_reduced(plt, sp);
-
-%% ratio of deterministic vs random noise
-%% ratio plot (h = 6/h = 1, for ran and det, A and b)
-sp = load(fullfile(outputdir, ['HBI_DetRanNoise_samples.mat'])).samples;
-plt.figure(4,1,'istitle',0,'rect',[0 0 0.5 0.9],...
-    'margin',[0.1 0.11 0.02 0.02]);
-stepsize = 0.04;
-xbins = [-30:stepsize:30];
-fns = {'NoiseRan', 'NoiseDet','InfoBonus_mu_n','bias_mu_n'};
[…]

### 2025-03-09 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/main_bayes_40chains.m

@@ -1,198 +1,0 @@
-%% get version
-ver = 'all';
-versave = 'all_50chains';
-% ver = 'paironly';
-%%
-cd(W.funcout('fileparts', 1, matlab.desktop.editor.getActiveFilename));
-outputdir = fullfile('../bayesoutput', versave);
-outputdir = GetFullPath(outputdir);
-wj = W_JAGS;
-%% set up chains
-nchains = 50; % How Many Chains?
-nburnin = 2000; % How Many Burn-in Samples?
-nsamples = 1000; % How Many Recorded Samples?
-HBItest = '';
-wj.setup_params(nchains, nburnin, nsamples);
-wj.openpool;
-%% test
-% HBItest = 'test_';
-% wj.setup_params;
-%%
-nchains = wj.bayes_params.nchains;
-%% main bayes
-files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
-suffix = {'','_all'};
-for fi = 1:2
-    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
-    wj.setup_data_dir(d, outputdir);
-    wjinfo = EEbayes_analysis(d, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise' suffix{fi}]);
-    wj.run;
-end
-%% 6 model comparison (rest of the 5 models)
-data = load(fullfile(fullfile('../data',ver), 'bayes_2noise.mat')).(['bayes_data']);
-for mi = 2:6
-    data.modelname = ['2noisemodel' char(64 + mi)];
-    wj.setup_data_dir(data, outputdir);
-    wjinfo = EEbayes_analysis(data, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise_' char(64 + mi)]);
-    wj.run;
-end
-%% reduced model
-files = {'bayes_2noise_dRonly.mat', 'bayes_2noise_0model.mat', 'bayes_2noise_dIonly.mat'};
-suffix = {'_dRonly','_0model','_dIonly'};
-for fi = 1:1 % just the first one
-    d = load(fullfile(fullfile('../data',ver), files{fi})).(['bayes_data', suffix{fi}]);
-    wj.setup_data_dir(d, outputdir);
-    wjinfo = EEbayes_analysis(d, nchains);
[…]

### 2025-03-09 codes/main_3_grid_parameter_recovery.m

@@ -1,60 +1,0 @@
-JAGS_setup;
-% wj.workingdir = 'C:\Users\Siyu_Wang\Documents\TEMP\JAGS2';
-%% full-grid parameter recovery (with only random noise and only deterministic noise)
-% % simulate behavior
-% data = load('../data/all/bayes_2noise.mat').bayes_data;
-% data.modelname = '2noisemodel';
-% nsub = data.nSubject;
-% nh = data.nHorizon;
-% for numi = 0:10
-%     for numj = 0:10
-%         W.print('%d,%d', numi, numj);
-%         % zero det noise
-%         simugame = EEsimulate_bayes_2noise(data, zeros(nsub, nh), ...
-%             zeros(nsub, nh), ones(nsub, nh) * numi, ones(nsub, nh) * numj);
-%         % fit
-%         wj.setup_data_dir(simugame, '../bayesoutput/simugrid');
-%         wjinfo = EEbayes_analysis(simugame, nchains);
-%         wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoise_' sprintf('simugrid_ran%d_det%d',numi,numj)]);
-%         wj.run;
-%         W.save(sprintf('../bayesoutput/simugrid/simugame_ran%d_det%d.mat', numi, numj), ...
-%             'simugame', simugame);
-%     end
-% end
-%% same as above but use actual information and bias
-paramsub = load(fullfile(outputdir, ['HBI_DetRanNoise_stat.mat'])).stats.mean;
-data = load('../data/all/bayes_2noise.mat').bayes_data;
-data.modelname = '2noisemodel';
-nsub = data.nSubject;
-nh = data.nHorizon;
-rans = [1:10,12:2:20];
-% rans = rans(end:-1:1);
-dets = [0:10];
-% dets = dets(end:-1:1);
-for i = 1:length(rans)
-    numi = rans(i);
-    for j = 1:length(dets)
-        numj = dets(j);
-        savename = sprintf('../bayesoutput/simugrid_infobias/simugame_ran%d_det%d.mat', numi, numj);
-        W.print('%d,%d', numi, numj);
-        if exist(savename, 'file')
-            W.print('skip');
-            continue;
-        end
-        % zero det noise
-        simugame = EEsimulate_bayes_2noise(data, paramsub.Infobonus_sub', ...
[…]

### 2025-04-19 Plos Comp Bio - revision v1/manuscript_DeterministicRandomNoise.tex
reorganize codes

--- 2024-05-17 manuscript/2024 - Plot Comp Bio/manuscript_DeterministicRandomNoise.tex
@@
 \PassOptionsToPackage{textsize=scriptsize}{todonotes}
-%\usepackage[final]{changes}
+\usepackage[]{changes}
 \usepackage{changes}
@@
 		Human decision making is inherently variable.
-While this variability is often seen as a sign of suboptimal behavior, recent work suggests that variability can actually be adaptive.
+While this variability is often seen as a sign of suboptimal behavior, \added[id=siyu]{both theoretical work in machine learning and empirical human studies} suggest that variability can actually be adaptive.
 An example arises when we must choose between exploring unknown options or exploiting options we know well.
@@
 
-	It has recently been shown that humans appear to use random exploration and can increase  decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014, Findling19}.
+	It has recently been shown that humans appear to use random exploration and can increase decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014}, \added[id=siyu]{as has also been suggested in computational models of animal behavior \citep{DOYA2002495, KHAMASSI2013441}}.
 In one of these tasks, known as the Horizon Task \citep{wilson2014}, the key manipulation is the horizon condition, i.e.
@@
 Avoiding an ex is a deterministic process, but if we fail to take the ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
+\added[id=siyu]{To dissociate a missing deterministic component from a true random process, choice consistency between repeated decisions can be utilized to decompose behavioral variability into predictable deterministic components and unpredictable random components \citep{WYART2016, Wyart_2018, Findling19, Findling21}.}
 
@@
 In both analyses, we focus on just the first free-choice trial in each game, where the only thing that differs between the horizon conditions is the number of choices that participants will make in the future.
+\added[id=siyu]{Subsequent choices in Horizon 6 games were not analyzed.}
 
@@
 	Here, $n$ denotes decision noise, which, in this version of the model is a combination of deterministic and random noise.
-$n$ is assumed to come from a logistic distributions with mean 0 and standard deviations $\sigma$.
+$n$ is assumed to come from a logistic distribution with mean 0 and standard deviations $\sigma$.
 
@@
 As shown in Figure \ref{fig:mf2}, people's behavior falls in between the pure deterministic noise prediction and the pure random noise prediction.
-Specifically, behavior is different from the pure random noise prediction in the both the [1 3] condition (t(64) = 4.83, p $<$ 0.001 for horizon 1, t(64) = 3.12 p $=$ 0.003 for horizon 6) and the [2 2] condition (t(64) = 3.92, p $<$ 0.001 for horizon 1, t(64) = 3.71, p $<$ 0.001 for horizon 6).
+Specifically, behavior is different from the pure random noise prediction in both the [1 3] condition (t(64) = 4.83, p $<$ 0.001 for horizon 1, t(64) = 3.12 p $=$ 0.003 for horizon 6) and the [2 2] condition (t(64) = 3.92, p $<$ 0.001 for horizon 1, t(64) = 3.71, p $<$ 0.001 for horizon 6).
 Likewise, behavior is different from pure deterministic noise prediction in both the [1 3] condition (t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6) and the [2 2] condition (t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6).
@@
 \subsubsection*{Posterior predictive checks}
-In addition to fitting the model to behavior, it is also important to check whether the model captures the qualitative patterns of the data \citep{Wilson2019} --- specifically how p(high info), p(low mean) and p(inconsistent) change with horizon.
+In addition to fitting the model to behavior, it is also important to check whether the model captures the qualitative patterns of the data \citep{Wilson2019, Palminteri} --- specifically how p(high info), p(low mean) and p(inconsistent) change with horizon.
 
@@
 For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}.
-In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
+\added[id=siyu]{Cross-game deterministic strategies and stimuli from previous games (e.g., memory of previously seen game) were also treated as random noise in our model \citep{WYART2016, Wyart_2018}.} In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
 As a result, from both a conceptual and methodological perspective, it is possible that the remaining 86\% of the decision noise that is not stimulus-driven noise, could be deterministic.
@@
 Like the random noise account, the deterministic noise account is also in line with previous work in which neural variability can be accounted for by fluctuations in sensory inputs.
-For example, MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli \citep{Bair1996}.
+For example, MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimulus \citep{Bair1996}.
 In other words, `irrelevant' features in the stimuli are represented in a reliable way in the brain that could drive downstream choices in a predictable way.
@@
 
-To see how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}.
+\added[id=siyu]{We will show that changes in reward processing can be equivalent to changes in noise in decision models (see also \cite{Cinotti19}).}.
+To show how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}.
 In this model, choice is determined by the sign of the difference in utility $\Delta Q$ between the two options, where
@@
 In the drift diffusion model, changes in behavioral variability can be driven by changes in the decision threshold (smaller threshold = more noise) or changes in the signal-to-noise ratio with which reward is encoded (lower SNR = more noise).
-By fitting both choices and response times, we were able to distinguish between these two accounts showing that the majority of the horizon-change in variability was driven by changes in SNR not threshold.
+By fitting both choices and response times, we were able to distinguish between these two accounts showing that the majority of the horizon-change in variability was driven by changes in SNR and not threshold.
 However, this model could not determine whether the changes in SNR were driven by signal or noise.
@@
 \section*{Materials and Methods}
-\subsection*{Ethics Statement}
+\subsection*{Ethics statement}
 Human subject protocols were approved by the University of Arizona institutional review board (IRB \# 1411567117).

### 2025-04-19 Plos Comp Bio - revision v1/supplemental/DetRanNoise_Supplementary.tex
reorganize codes

--- 2024-05-16 manuscript/2024 - Plot Comp Bio/supplemental/DetRanNoise_Supplementary.tex
@@
 	\subsection{Parameter recovery of the subject-level parameter fits}
-	Next, we tested the ability of our model fitting procedure to recovery parameters from simulated data at the subject level (Figure \ref{fig:paramrecover_session}, \ref{fig:paramrecover_sub}).
+	Next, we tested the ability of our model fitting procedure to recover parameters from simulated data at the subject level (Figure \ref{fig:paramrecover_session}, \ref{fig:paramrecover_sub}).
 The correlations between the true vs fitted parameters are significant across participants for all parameters (p $<$ 0.001).

### 2025-04-19 codes/EEbayes_reformat.m

@@ -1,0 +1,362 @@
+function bayesdata = EEbayes_reformat(data, idxsub, modelname, maxtrial)
+    if ~exist('maxtrial', 'var')
+        maxtrial = Inf;
+    end
+    data = W.tab_autofieldcombine(data);
+    bayesdata.modelname = modelname;
+    switch modelname
+        case 'EE2noise'
+            bayesdata.nHorizon = 2;
+            bayesdata.nSubject = length(idxsub);
+            nT = cellfun(@(x)length(x), idxsub);
+            LEN = min(max(nT),maxtrial);
+            bayesdata.nForcedTrials = 4;
+            bayesdata.nCond = 2;
+            for si = 1:bayesdata.nSubject
+                gd = data(idxsub{si},:);
+                nT = min(size(gd,1), LEN);
+                bayesdata.nTrial(si,1) = nT;
+                bayesdata.horizon(si,:) = W.extend(ceil(gd.cond_horizon'/5), LEN); 
+                % 1 is short, 2 is long
+                bayesdata.dI(si,:) = W.extend(gd.dI',LEN); % -1, 0, 1
+                bayesdata.choice(si,:) = W.extend((gd.choice(:,5)' == 2) + 0, LEN); 
+                % 1 is right, 0 is left
+                bayesdata.dR(si,:) = W.extend(gd.dR(:,4)',LEN);
+                [~, ~, ranking] = unique(gd.repeat_id);
+                bayesdata.nrepeatID(si,1) = max(ranking);
+                bayesdata.repeatID(si,:) = W.extend(ranking',LEN);
+
+                bayesdata.infocond(si,:) = W.nan_equal(bayesdata.dI(si,:), 0) + 1;
+                % 1 - [1 3], 2 - [2 2]
+                bayesdata.rpIDHorizon(si,:) = W.extend(W.arrayfun(@(x)unique(bayesdata.horizon(si,(ranking == x))), 1:bayesdata.nrepeatID(si,1)), LEN);
+                bayesdata.rpIDCond(si,:) = W.extend(W.arrayfun(@(x)unique(bayesdata.infocond(si,(ranking == x))), 1:bayesdata.nrepeatID(si,1)), LEN);
+                
+                bayesdata.choice4(si,1:LEN,:) = W.extend((gd.choice(:, 1:4)' == 2) + 0, LEN)';
+                bayesdata.reward4(si,1:LEN,:) = W.extend(gd.reward(:, 1:4)', LEN)'; 
+                
[…]

### 2025-04-19 codes/EEsimulate_bayes_2noise_2condDIST.m

@@ -1,0 +1,86 @@
+function out = EEsimulate_bayes_2noise_2condDIST(data, As, bs, nrans, ndets)
+    disp(['Generating simulation']);
+    for si = 1:data.nSubject
+        choice = [];
+        nT = data.nTrial(si);
+        dR = data.dR(si,1:nT)';
+        dI = data.dI(si,1:nT)';
+        repeatID = data.repeatID(si,1:nT)';
+        horizon = data.horizon(si,1:nT)';
+        infocond = data.infocond(si, 1:nT)';
+        n_det = NaN(1, data.nrepeatID(si));
+        tA = nan(1,2);
+        tb = nan(1,2);
+        for h = 1:2
+            te = reshape(As(:,:, h, si),1,[]); 
+            tA(h) = mean(te);
+            te = reshape(bs(:,:, h, si),1,[]); 
+            tb(h) = mean(te);
+        end
+        nsample = size(As,1) * size(As,2);
+        if length(size(nrans)) == 5
+            tran = nan(2,2);
+            for h = 1:2
+                for c = 1:2
+                    te = reshape(nrans(:,:, h, c, si),1,[]);
+                    tid = randperm(nsample);
+                    tran(h, c) = te(tid(1));
+                end
+            end
+        else
+            tran = nan(2,2);
+            for c = 1:2
+                te = reshape(nrans(:,:, c, si),1,[]);
+                tid = randperm(nsample);
+                tran(1, c) = te(tid(1));
+                tran(2, c) = tran(1, c);
+            end
+        end
+        if length(size(ndets)) == 5
+            tdet = nan(2,2);
+            for h = 1:2
+                for c = 1:2
+                    te = reshape(ndets(:,:, h, c, si),1,[]);
+                    tid = randperm(nsample);
+                    tdet(h, c) = te(tid(1));
+                end
+            end
+        else
+            tdet = nan(2,2);
+            for c = 1:2
+                te = reshape(ndets(:,:, c, si),1,[]);
+                tid = randperm(nsample);
+                tdet(1, c) = te(tid(1));
+                tdet(2, c) = tdet(1, c);
+            end
+        end
+        for gi = 1:nT
+            ttA = tA(horizon(gi));
+            ttb = tb(horizon(gi));
+            ttran = tran(horizon(gi), infocond(gi));
+            ttdet = tdet(horizon(gi), infocond(gi));
+            if  ttran > 0
+                pdni = makedist('Logistic',0, ttran);
+                dni = random(pdni);
[…]

### 2025-04-19 codes/JAGS_setup_R1.m

@@ -1,0 +1,21 @@
+%% get version
+addpath(genpath('../codes_support/'));
+ver = 'allR1';
+%%
+W.cd();
+outputdir = fullfile('../bayesoutput', ver);
+outputdir = W.get_fullpath(W.mkdir(outputdir));
+wj = W_JAGS;
+%% set up chains
+nchains = 4; % How Many Chains?
+nburnin = 2000; % How Many Burn-in Samples?
+nsamples = 2000; % How Many Recorded Samples?
+HBItest = '';
+wj.setup_params(nchains, nburnin, nsamples);
+%% test
+% HBItest = 'test_';
+% wj.setup_params;
+%%
+nchains = wj.bayes_params.nchains;
+W.library_wang('Wang_EEHorizon', 1);
+W.library('JAGS', 1);

### 2025-04-19 codes/codedraft_revision_PlosCB/main_3_fit_Bayes_models.m

@@ -1,0 +1,63 @@
+
+%%
+plt = W_plt('savedir', '../figures_revision', 'savepfx', 'RDBayes_dIvar', 'isshow', true, ...
+    'issave', true, 'extension',{'svg', 'jpg'});
+tsp = W.load('../bayesoutput_revision/all_revision/HBI_DetRanNoiseR1_dIvar_samples');
+tst = W.load('../bayesoutput_revision/all_revision/HBI_DetRanNoiseR1_dIvar_stat');
+
+%%
+plt.set_pltsetting('savesfx', suffix{gii});
+old = plt.param_plt.fontsize_title;
+plt.param_plt.fontsize_title = 15;
+% plt.param_plt.pixel_h = 150;
+plt.figure(2,2,'is_title', 'all');
+gi = 1;
+plt.setfig('xlim', {[-1 10+10*gi],[-3 8+ gi *4], [-1 10+10*gi], [-3 8 + gi *4], [-1 10+10*gi], [-3 8 + gi *4], [-1 10+10*gi], [-3 8 + gi *4]}, ...
+    'ylim', []);
+EEplot_2noise_hyperpriors(plt, tsp, [0.02 0.02]);
+plt.param_plt.fontsize_title = old;
+
+%% check convergence
+% load('..\bayesoutput_revision\all_revision\HBI_DetRanNoiseR1_convergence_samples.mat');
+%% 
+% plt = W_plt('savedir', '../figures', 'savepfx', 'RanDetNoise', 'isshow', true, ...
+%     'issave', false, 'extension',{'svg', 'jpg'});
+% plt.figure(6,10);
+% % id = 100:10000;
+% id = 1:5000;
+% for i = 1:10
+%     plt.ax(1, i);
+%     plt.plot([], samples.InfoBonus_mu_n(i,:,1), [], 'line');
+%     plt.plot([], samples.InfoBonus_mu_n(i,:,2), [], 'line');
+%     plt.dashY(2000)
+%     plt.ax(2, i);
+%     plt.plot([], samples.bias_mu_n(i,:,1), [], 'line');
+%     plt.plot([], samples.bias_mu_n(i,:,2), [], 'line');
+%     plt.dashY(2000)
+%     plt.ax(3, i);
+%     plt.plot([], samples.NoiseRan(i,id,1,1), [], 'line');
+%     plt.plot([], samples.NoiseRan(i,id,2,1), [], 'line');
+%     plt.dashY(2000)
+%     plt.ax(4, i);
[…]

### 2025-04-19 codes/codedraft_revision_PlosCB/main_compare_hierarchical_vs_non_hierarchical.m

@@ -1,0 +1,235 @@
+JAGS_setup;
+outputdir = '../bayesoutput_revision/hierarchical';
+%% load data
+data = W.load('../data/all_revision/bayesdata');
+%%
+modelnames = {'model_2noise_2cond_nonhierarchical','model_2noise_2cond'};
+suffix = {'_2cond_nonhierarchical','_2cond'};
+for mi = 1:length(modelnames)
+    modelname = modelnames{mi};
+    %% fit models to data
+    d = data;
+    d.modelname = sprintf('2noisemodel%s', suffix{mi});
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['fit' suffix{mi}]);
+    wj.run;
+    %% load fitted data
+    st0 = W.load(fullfile(outputdir, ['HBI_fit' suffix{mi} '_stat']));
+    st0 = st0.stats.mean;
+    sp = W.load(fullfile(outputdir, ['HBI_fit' suffix{mi} '_samples']));
+    stmap = getMAP(sp);
+    save(fullfile(outputdir, ['HBI_fit' suffix{mi} '_MAP']), 'stmap');
+    %% simulate data from fitted model
+    st = st0;
+    simu = EEsimulate_bayes_2noise_2cond(data, st.Infobonus_sub, ...
+        st.bias_sub, st.NoiseRan_sub, st.NoiseDet_sub);
+    %% simulate data from shuffled model
+    nsub = size(st.Infobonus_sub, 2);
+    id1 = randperm(nsub*4);
+    id2 = randperm(nsub*4);
+    st_shuffled = st;
+    st_shuffled.NoiseRan_sub = reshape(st.NoiseRan_sub(id1),2,2,[]);
+    st_shuffled.NoiseDet_sub = reshape(st.NoiseDet_sub(id2),2,2,[]);
+    simu_shuffled = EEsimulate_bayes_2noise_2cond(data, st_shuffled.Infobonus_sub, ...
+        st_shuffled.bias_sub, st_shuffled.NoiseRan_sub, st_shuffled.NoiseDet_sub);
[…]

### 2025-04-19 codes/codedraft_revision_PlosCB/plot_parameter_recovery.m

@@ -1,0 +1,54 @@
+function plt = plot_parameter_recovery(plt, st1, st2, option, ax, tlt)
+    if option == 1
+        plt.figure(2,2, 'is_title', 'all');
+        for hi = 1:2
+            plt.ax(1, hi);
+            ylab = {'recovered'};
+                tx = squeeze(st1.NoiseRan_sub(hi,:));
+                ty = squeeze(st2.NoiseRan_sub(hi,:));
+                str = plt.scatter(tx, ty, 'diag');
+            plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+
+            plt.ax(2, hi);
+                tx = squeeze(st1.NoiseDet_sub(hi,:));
+                ty = squeeze(st2.NoiseDet_sub(hi,:));
+                str = plt.scatter(tx, ty, 'diag');
+            plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+        end
+    elseif option == 2
+        plt.figure(2,4, 'is_title', 'all');
+        cond = {'[1 3]', '[2 2]'};
+        for ci = 1:2
+            for hi = 1:2
+                plt.ax(ci, hi);
+                ylab = W.iif(hi == 1, {cond{ci}, 'recovered'}, {'recovered'});
+                tx = squeeze(st1.NoiseRan_sub(ci, hi,:));
+                ty = squeeze(st2.NoiseRan_sub(ci, hi,:));
+                str = plt.scatter(tx, ty, 'diag');
+                plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+    
+                plt.ax(ci, hi + 2);
+                tx = squeeze(st1.NoiseDet_sub(ci, hi,:));
+                ty = squeeze(st2.NoiseDet_sub(ci, hi,:));
+                str = plt.scatter(tx, ty, 'diag');
+                plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+            end
+        end
[…]

### 2025-04-19 codes/findMAP.m

@@ -1,0 +1,10 @@
+function x = findMAP(v)
+    v = reshape(v, 1, []);
+    vx = linspace(min(v), max(v), 1000);
+    dx = mean(diff(vx), 'all');
+    stv = std(v);
+    one = ones(length(v),1);
+    val = sum1d(one, v', stv, dx, min(v), max(v));
+    [~, tid] = max(val);
+    x = vx(tid);
+end

### 2025-04-19 codes/mainFigures_paramrecover_subject.m

@@ -1,0 +1,52 @@
+outputdir = '../bayesoutput/all/';
+st = load(fullfile(outputdir, ['HBI_DetRanNoise_stat.mat'])).stats.mean;
+%% compute coverage
+simudir = '../bayesoutput/simurepeat';
+stsimu = cell(1,200);
+for repi = 1:200
+    tstat = load(fullfile(simudir, sprintf('HBI_DetRanNoise_simu_%d_stat.mat', repi)));
+    stsimu{repi} = tstat.stats.mean;
+end
+stsimu = cellfun(@(x)x, stsimu);
+%%
+stsimuav = W.struct_median(stsimu);
+%%
+plt = W_plt('savedir', '../figures', 'savepfx', 'RDBayes', 'isshow', true, ...
+    'issave', true);
+mx = {[-25 25],[-25 25],[-3 3],[-3 3],[-1 20],[-1 30],[-1 5],[-1 15]};
+mxtick = {[-20:10:20],[-40:20:40],[-2:2:4],[-2:2:4],[0:10:50],[0:10:50],[0:5:30],[0:5:30]};
+plt.figure(4,2,'is_title', 1);
+plt.setfig('xtick', mxtick, ...
+    'ytick', mxtick);
+EEplot_2noise_parameter_recovery(plt, st, stsimuav);
+plt.figure(2,2,'is_title', 1);
+mx = {[-1 20],[-1 30],[-1 5],[-1 15]};
+mxtick = {[0:10:50],[0:10:50],[0:5:30],[0:5:30]};
+plt.setfig('xtick', mxtick, ...
+    'ytick', mxtick);
+EEplot_2noise_parameter_recovery(plt, st, stsimuav,2);
+%%
+plt.set_pltsetting('savesfx', 'examplesession')
+plt.figure(4,2,'is_title', 1);
+plt.setfig('xtick', mxtick, ...
+    'ytick', mxtick);
+EEplot_2noise_parameter_recovery(plt, st, stsimu(1));
+%%
+plt.set_pltsetting('savesfx', 'examplesession_randet')
+plt.figure(2,2,'is_title', 1);
+mx = {[-1 20],[-1 30],[-1 5],[-1 15]};
+mxtick = {[0:10:50],[0:10:50],[0:5:30],[0:5:30]};
+plt.setfig('xtick', mxtick, ...
+    'ytick', mxtick);
[…]

### 2025-04-19 codes/main_1b_preparedata.m

@@ -1,0 +1,26 @@
+addpath(genpath('../codes_support/'))
+%% prepare data
+load('../data/all/idxsub_exclude.mat');
+game = readtable('../data/all/data_all.csv', 'Delimiter', ',');
+g = W.tab_autofieldcombine(game);
+dI = arrayfun(@(x)getdI_onegame(g.choice(x,:), g.reward(x,:)), 1:size(g,1));
+game.dIvar = dI';
+
+bayesdata = EEbayes_reformat(game, idxsub(id), 'EE2noise');
+save('../data/all/bayesdata.mat', 'bayesdata');
+
+
+% bayes_data_2condKF = EEbayes_reformat(game, idxsub(id), '2noisemodel_2condKF');
+% save('../data/all_revision/bayes_2noise_2condKF.mat', 'bayes_data_2condKF');
+% %%
+
+% p_mu_obs = generate_likelihood();
+% g = W.tab_autofieldcombine(game);
+% dI = arrayfun(@(x)getdI_onegame(g.choice(x,:), g.reward(x,:), p_mu_obs), 1:size(g,1));
+% game.dIvar = dI';
+% bayes_data_2cond_dIvar = EEbayes_reformat(game, idxsub(id), '2noisemodel_2cond_dIvar');
+% save('../data/all_revision/bayes_2noise_2cond_dIvar.mat', 'bayes_data_2cond_dIvar');
+%%
+% 
+% bayes_data_2cond_dIvar2 = EEbayes_reformat(game, idxsub(id), '2noisemodel_2cond_dIvar');
+% save('../data/all_revision/bayes_2noise_2cond_dIvar2.mat', 'bayes_data_2cond_dIvar2');

### 2025-04-19 codes/main_7_2cond_models.m

@@ -1,0 +1,94 @@
+%% test convergence
+JAGS_setup_R1;
+nchains = 4; % How Many Chains?
+nburnin = 0; % How Many Burn-in Samples?
+nsamples = 5000; % How Many Recorded Samples?
+wj.setup_params(nchains, nburnin, nsamples);
+d = load(fullfile('../data/all', 'bayesdata.mat')).(['bayesdata']);
+d.modelname = '2noisemodel_2cond';
+wj.setup_data_dir(d, outputdir);
+wjinfo = EEbayes_analysis(d, nchains);
+wj.setup(wjinfo.modelfile, wjinfo.params, struct, 'DetRanNoiseR1_convergence');
+wj.run;
+%% main bayes (base classes)
+% set of models
+% 1cond vs 2cond
+% dI vs dIvar
+suffix = {'', '_dIvar', '_2cond', '_2cond_dIvar', '_2cond_dIvar_both', '_2cond_dRonly', '_2cond_nobias', ...
+    'B_2cond','C_2cond','D_2cond','E_2cond','F_2cond','_2cond_nonhierarchical'};% reviewer fit non-hierarchical
+for fi = 1:length(suffix)
+    JAGS_setup_R1;
+    d = load(fullfile('../data/all', 'bayesdata.mat')).(['bayesdata']);
+    d.modelname = sprintf('2noisemodel%s', suffix{fi});
+    wj.setup_data_dir(d, outputdir);
+    wjinfo = EEbayes_analysis(d, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoiseR1' suffix{fi}]);
+    wj.run;
+end
+%% explained variance
+sp = W.load('../bayesoutput/allR1/HBI_DetRanNoiseR1_2cond_samples.mat');
+%% get explained variance
+det = []; ran = [];
+vardet = [];
+for i = 1:2
+    det(i,:) = reshape(sp.NoiseDet(:,:,i),1,[]);
+    ran(i,:) = reshape(sp.NoiseRan(:,:,i),1,[]);
+    vardet(i,:) = det(i,:).^2./(det(i,:).^2 + ran(i,:).^2);
+end
+ddet = reshape(sp.dNoiseDet,1,[]);
+dran = reshape(sp.dNoiseRan,1,[]);
+vard = (ddet.^2./(ddet.^2+dran.^2));
[…]

### 2025-04-19 codes/main_8_6model_2cond.m

@@ -1,0 +1,216 @@
+JAGS_setup_R1;
+%% fit the 6 models
+data = importdata('../data/all/bayesdata.mat');
+for numi = 1:6
+    W.print('model: %d', numi);
+    data.modelname = sprintf('2noisemodel%s_2cond', char(numi + 'A' -1));
+    wj.setup_data_dir(data, '../bayesoutput/6modelR1');
+    wjinfo = EEbayes_analysis(data, nchains);
+    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['R1_fit_' char(numi + 'A' -1)]);
+    wj.run;
+end
+%% simulate & fit with modelA (unused)
+for repi = 1:10
+    for numi = 2:6
+        paramsub = load(fullfile('../bayesoutput/6modelR1', [HBItest 'HBI_R1_fit_' char(numi + 'A' -1) '_stat.mat'])).stats.mean;
+        if ~isfield(paramsub, 'NoiseRan_sub')
+            paramsub.NoiseRan_sub = zeros(size(paramsub.NoiseDet_sub));
+        end
+        if ~isfield(paramsub, 'NoiseDet_sub')
+            paramsub.NoiseDet_sub = zeros(size(paramsub.NoiseRan_sub));
+        end
+        if ~exist(sprintf('../bayesoutput/6modelR1/R1_simugame_%s_%d.mat', char(numi + 'A' -1), repi), 'file')
+            simugame = EEsimulate_bayes_2noise_2cond(data, paramsub.Infobonus_sub, ...
+                paramsub.bias_sub, paramsub.NoiseRan_sub, paramsub.NoiseDet_sub);
+            W.save(sprintf('../bayesoutput/6modelR1/R1_simugame_%s_%d.mat', char(numi + 'A' -1), repi), ...
+                'simugame', simugame);
+        else
+            simugame = W.load(sprintf('../bayesoutput/6modelR1/R1_simugame_%s_%d.mat', char(numi + 'A' -1), repi));
+        end
+        % fit
+        simugame.modelname = '2noisemodelA_2cond';
+        wj.setup_data_dir(simugame, '../bayesoutput/6modelR1');
+        wjinfo = EEbayes_analysis(simugame, nchains);
+        savename = sprintf('R1_fitA_simugame_%s_%d', char(numi + 'A' -1), repi);
[…]

### 2025-04-19 codes_original_submission/OLD/UnusedCodes/EEplot_2noise_parameter_recovery_hyperprior.m

@@ -1,57 +1,0 @@
-function plt = EEplot_2noise_parameter_recovery_hyperprior(plt, sp, sp2)
-names = {'Random noise - \sigma_{ran}','Random noise - \sigma_{ran}',...
-    'Deterministic noise - \sigma_{det}','Deterministic noise - \sigma_{det}'};
-plt.figure(2,2,'title');
-plt.setup_pltparams('fontsize_leg', 20, 'linewidth', 2);
-plt.setfig('xlim', {[-1 22],[-3 12], [-1 22], [-3 12]}, 'ylim', [0 1.2], ...
-    'ytick', [0:0.2:1.2], 'xtick',{0:4:20,-3:3:12,0:4:20,-3:3:12});
-color = {{'AZblue','AZblue50','AZred','AZred50'},{'black','black50'}, ...
-    {'AZblue','AZblue50','AZred','AZred50'},{'black','black50'}};
-fns = {'NoiseRan', 'NoiseDet'};
-stepsize = 0.02;
-xbins = [-10:stepsize:30];
-lg1 = {'original, h = 1','recovered, h = 1','original, h = 6', 'recovered, h = 6'};
-lg2 = {'original, change', 'recovered change'};
-plt.setfig('color', color, 'legend', ...
-    {lg1,lg2,lg1,lg2},...
-    'ylabel', {'posterior density','','posterior density',''}, ...
-    'xlabel', {'','','noise standard deviation [points]', 'noise standard deviation [points]'}, ...
-    'title', names);
-ymax = 1.2;
-plt.param_fig.colordelete = 1;
-for i = 1:2
-    fn = fns{i};
-    plt.new;
-    for hi = 1:2
-        td = sp.(fn);
-        td = squeeze(td(:,:,hi));
-        td = reshape(td, 1, []);
-        tl = hist(td, xbins)/(length(td)*stepsize);
-        plt.lineplot(tl, [], xbins);
-        
-        td = sp2.(fn);
-        td = squeeze(td(:,:,hi));
-        td = reshape(td, 1, []);
-        tl = hist(td, xbins)/(length(td)*stepsize);
-        plt.lineplot(tl, [], xbins);
-    end
-    hold on;
-    plot([0 0],[0 ymax], '--k','LineWidth', plt.param_figsetting.linewidth/2);
-    hold off;
-    
-    plt.new;
-    td = sp.(['d' fn]);
-    td = reshape(td, 1, []);
[…]

### 2025-04-19 codes_original_submission/OLD/UnusedCodes/draft_main_2_bayes2noise.m

@@ -1,92 +1,0 @@
-%% simplistic case (only noise, no dR, dI, bias, choice = dQ)
-% load fitted params
-vv = '';
-vv2 = '';
-paramsub = load(fullfile(outputdir, [HBItest 'HBI_DetRanNoise_' vv 'stat.mat'])).stats.mean;
-[nh, nsub] = size(paramsub.NoiseDet_sub);
-if ~isfield(paramsub, 'bias_sub')
-    paramsub.bias_sub = zeros(nh, nsub);
-end
-% simulate behavior
-data = load(fullfile(fullfile('../data',ver), ['bayes_2noise' vv2 '.mat'])).(['bayes_data' vv2]);
-data.modelname = '2noiseONLY';
-for numi = 1:1
-    % zero det noise
-    simugame = EEsimulate_bayes_ONLYnoise(data, paramsub.NoiseRan_sub', zeros(nsub, nh));
-    % fit
-    wj.setup_data_dir(simugame, outputdir);
-    wjinfo = EEbayes_analysis(simugame, nchains, mdpath);
-    wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_' vv 'ONLYNOISE_ranonly_' num2str(numi)]);
-    wj.run;
-    % zero ran noise
-    simugame = EEsimulate_bayes_ONLYnoise(data, zeros(nsub, nh), paramsub.NoiseDet_sub');
-    % fit
-    wj.setup_data_dir(simugame, outputdir);
-    wjinfo = EEbayes_analysis(simugame, nchains, mdpath);
-    wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_' vv 'ONLYNOISE_detonly_' num2str(numi)]);
-    wj.run;
-end
-%%
-vvs = {'','paironly_'};
-vv2s = {'','_paironly'};
-for i = 1:1 % just the original
-    %% parameter recovery
-    vv = vvs{i}; vv2 = vv2s{i};
-    %% parameter recovery (with only random noise and only deterministic noise)
-    % load fitted params
-    paramsub = load(fullfile(outputdir, [HBItest 'HBI_DetRanNoise_' vv 'stat.mat'])).stats.mean;
-    [nh, nsub] = size(paramsub.NoiseDet_sub);
-    if ~isfield(paramsub, 'bias_sub')
-        paramsub.bias_sub = zeros(nh, nsub);
-    end
-    % simulate behavior
[…]

### 2025-04-19 codes_original_submission/OLD/UnusedCodes/mainBOTH_FIGURE_6model.m

@@ -1,91 +1,0 @@
-%% setup
-ver = 'all';
-verout = 'all_50chains';
-% ver = 'paironly';
-% verout = ver;
-outputdir = fullfile('../bayesoutput/', verout);
-figdir = '../figures';
-plt = W_plt;
-plt.setup_W_plt('fig_dir', fullfile(figdir,verout), 'fig_suffix', '','fig_projectname', 'RanDetNoise');
-%% 6-model comparison (simulation)
-% data = load(fullfile('../data', ver, 'bayes_2noise.mat')).(['bayes_data']);
-% name = {'','B_','C_','D_','E_','F_'};
-% simu = {};
-% isoverwrite = true;
-% for repi = 1:50
-%     disp(repi);
-%     tfile = fullfile('../data', verout, sprintf('bayes_6model_simu_rep%d.mat', repi));
-%     if ~exist(tfile) || isoverwrite
-%         for mi = 1:6
-%             tparam = load(fullfile(outputdir, ['HBI_DetRanNoise_' name{mi} 'stat.mat'])).stats.mean;
-%             if ~isfield(tparam, 'NoiseRan_sub')
-%                 tparam.NoiseRan_sub = zeros(size(tparam.NoiseDet_sub));
-%             end
-%             if ~isfield(tparam, 'NoiseDet_sub')
-%                 tparam.NoiseDet_sub = zeros(size(tparam.NoiseRan_sub));
-%             end
-%             simu{mi} = EEsimulate_bayes_2noise(data, tparam.Infobonus_sub', ...
-%                 tparam.bias_sub', tparam.NoiseRan_sub', tparam.NoiseDet_sub');
-%         end
-%         save(tfile,'simu');
-%     end
-% end
-%%
-% basic analysis of simulated behavior
-% files = dir(fullfile('../data', ver, sprintf('data_%s.csv', ver)));
-% filename = fullfile(files.folder, files.name);
-% game0 = readtable(filename, 'Delimiter', ',');
-% idx = load(fullfile('../data', ver, 'idxsub_exclude'));
-% tid = idx.idxsub(idx.id);
-% 
-% sub = readtable(fullfile('../data', ver, 'SUB_EE_blinkCuriosity_2016S.csv'),'Delimiter', ',');
-% gp = W_sub.analysis_1group(sub(idx.id,:));
-% 
-% temp = {};
-% for repi = 1:50
-%     disp(repi);
[…]

### 2025-04-19 codes_original_submission/OLD/UnusedCodes/old/oldmain_ANALYSIS_bayes2noise.m

@@ -1,161 +1,0 @@
-cd(W.funcout('fileparts', 1, matlab.desktop.editor.getActiveFilename));
-outputdir = '../bayesoutput';
-outputdir = GetFullPath(outputdir);
-wj = W_JAGS;
-%% set up chains
-nchains = 4; % How Many Chains?
-nburnin = 5000; % How Many Burn-in Samples?
-nsamples = 5000; % How Many Recorded Samples?
-HBItest = '';
-wj.setup_params(nchains, nburnin, nsamples);
-%% test
-% HBItest = 'test_';
-% wj.setup_params;
-%%
-nchains = wj.bayes_params.nchains;
-%% main bayes
-files = {'bayes_2noise.mat', 'bayes_2noise_all.mat'};
-suffix = {'','_all'};
-for fi = 1:2
-    d = load(fullfile('../data', files{fi})).(['bayes_data', suffix{fi}]);
-    wj.setup_data_dir(d, outputdir);
-    wjinfo = EEbayes_analysis(d, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise' suffix{fi}]);
-    wj.run;
-end
-%% model no bias
-data = load(fullfile('../data', 'bayes_2noise.mat')).(['bayes_data']);
-data.modelname = ['2noisemodel_nobias'];
-wj.setup_data_dir(data, outputdir);
-wjinfo = EEbayes_analysis(data, nchains);
-wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_nobias']);
-wj.run;
-%% model paironly
-data = load(fullfile('../data', 'bayes_2noise_paironly.mat')).(['bayes_data_paironly']);
-data.modelname = ['2noisemodel'];
-wj.setup_data_dir(data, outputdir);
-wjinfo = EEbayes_analysis(data, nchains);
-wj.setup(wjinfo.modelfile, wjinfo.params, wjinfo.init0, ['DetRanNoise_paironly']);
-wj.run;
-%% 6 model comparison (rest of the 5 models)
-data = load(fullfile('../data', 'bayes_2noise.mat')).(['bayes_data']);
-for mi = 2:6
-    data.modelname = ['2noisemodel' char(64 + mi)];
-    
-    wj.setup_data_dir(data, outputdir);
-    wjinfo = EEbayes_analysis(data, nchains);
[…]

### 2025-04-19 codes_original_submission/main_1_behavior.m

@@ -1,75 +1,0 @@
-%% main analysis 2016
-addpath(genpath('../codes_support'))
-rawdatadir = '../data';
-savedir = '../data/all';
-files = dir(fullfile(rawdatadir,'Imported_EE_blinkCuriosity_2016S.csv'));
-filename = fullfile(files.folder, files.name);
-%% basic analysis
-W.library_wang('Wang_EEHorizon')
-opt_sub = {'subjectID', 'date', 'time'};
-opt_preprocess = 'EEpreprocess_game_basic';
-opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};
-opt_analysis = {'EEanalysis_sub_basic', 'EEanalysis_sub_MLEbasic'};
-[sub0, game, idxsub] = W.analysis_pipeline(filename, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, savedir);
-W.writetable(game, fullfile(savedir,'data_all.csv'));
-%% exclude participants
-sub0 = readtable('../data/all/SUB_EE_blinkCuriosity_2016S.csv', 'Delimiter', ',');
-load('../data/all/idxsub_EE_blinkCuriosity_2016S.mat');
-[suball, idall] = W.excludesubject(sub0, 'age', [18 NaN]);
-[sub, id] = W.excludesubject(sub0, 'age', [18 NaN], 'pvalue_ac', [NaN 0.05]);
-save(fullfile(savedir, 'idxsub_exclude.mat'), 'id', 'idall','idxsub');
-%% analyze group
-clear gp;
-additionalcompare = {{'p_inconsistent13'},{'p_inconsistent22'},...
-    {'p_inconsistent13','p_inconsistent13_randomtheory_byR'},{'p_inconsistent22','p_inconsistent22_randomtheory_byR'}};
-gp{1} = W.analysis_1group(sub, true, additionalcompare);
-gp{2} = W.analysis_1group(suball, true, additionalcompare);
-suffix = {'','_all'};
-%% behavioral plots
-plt = W_plt('savedir', '../figures', 'savepfx', 'RanDetNoise', 'isshow', true, ...
-    'issave', true, 'extension',{'svg', 'jpg'});
-%%
-for gi = 1:2
-    plt.set_pltsetting('savesfx', suffix{gi});
-    %% model-free plots
-    plt.figure(2,2,'is_title',1);
[…]

### 2025-04-19 codes_revision_PlosCB/EEbayes_analysis.m

@@ -1,109 +1,0 @@
-function info = EEbayes_analysis(bayes_data, nchains)
-%     paths = strsplit(path, ';');
-%     pid = find(contains(string(paths), 'HBI_models'));
-%     mdpath = paths{pid};
-    mdpath = dir('./HBI_models');
-    mdpath = mdpath(1).folder;
-    if ~exist('nchains', 'var')
-        nchains = 4;
-    end
-    md = bayes_data.modelname;
-    switch md
-        case {'2noisemodel','2noisemodelB','2noisemodelC','2noisemodelD',...
-                '2noisemodel_2cond_nonhierarchical', ...
-                '2noisemodel_dIonly', ...
-                '2noisemodel_2cond','2noisemodelA_2cond', '2noisemodelB_2cond','2noisemodelC_2cond','2noisemodelD_2cond',...
-                '2noisemodel_2cond_dIvar','2noisemodel_2cond_dIvar_both', ...
-                '2noisemodel_v2', '2noisemodel_v2_2cond','2noisemodel_v2_2cond_dIvar','2noisemodel_v2_2cond_dIvar_both', ...
-                }
-            params = {'NoiseDet_k_p','NoiseDet_lambda_p','NoiseDet',...
-                'NoiseRan_k_p', 'NoiseRan_lambda_p','NoiseRan', ...
-                'InfoBonus_mu_n','InfoBonus_sigma_p','bias_mu_n','bias_sigma_p', ...
-                'NoiseRan_sub','NoiseDet_sub', 'Infobonus_sub','bias_sub',...
-                'dInfoBonus','dbias','dNoiseRan','dNoiseDet','dNoiseRan_sub','dNoiseDet_sub'};
-            init0 = get_bayesinit(params, nchains, 'simple');
-            modelfile = ['model_2noise' md(12:end) '.txt'];
-
-        case {'2noisemodel_nobias', '2noisemodel_2cond_nobias'}
-            params = {'NoiseDet_k_p','NoiseDet_lambda_p','NoiseDet',...
-                'NoiseRan_k_p', 'NoiseRan_lambda_p','NoiseRan', ...
-                'InfoBonus_mu_n','InfoBonus_sigma_p', ...
-                'NoiseRan_sub','NoiseDet_sub', 'Infobonus_sub',...
-                'dInfoBonus','dNoiseRan','dNoiseDet','dNoiseRan_sub','dNoiseDet_sub'};
[…]

### 2025-04-19 codes_revision_PlosCB/EEbayes_reformat.m

@@ -1,362 +1,0 @@
-function bayesdata = EEbayes_reformat(data, idxsub, modelname, maxtrial)
-    if ~exist('maxtrial', 'var')
-        maxtrial = Inf;
-    end
-    data = W.tab_autofieldcombine(data);
-    bayesdata.modelname = modelname;
-    switch modelname
-        case 'EE2noise'
-            bayesdata.nHorizon = 2;
-            bayesdata.nSubject = length(idxsub);
-            nT = cellfun(@(x)length(x), idxsub);
-            LEN = min(max(nT),maxtrial);
-            bayesdata.nForcedTrials = 4;
-            bayesdata.nCond = 2;
-            for si = 1:bayesdata.nSubject
-                gd = data(idxsub{si},:);
-                nT = min(size(gd,1), LEN);
-                bayesdata.nTrial(si,1) = nT;
-                bayesdata.horizon(si,:) = W.extend(ceil(gd.cond_horizon'/5), LEN); 
-                % 1 is short, 2 is long
-                bayesdata.dI(si,:) = W.extend(gd.dI',LEN); % -1, 0, 1
-                bayesdata.choice(si,:) = W.extend((gd.choice(:,5)' == 2) + 0, LEN); 
-                % 1 is right, 0 is left
-                bayesdata.dR(si,:) = W.extend(gd.dR(:,4)',LEN);
-                [~, ~, ranking] = unique(gd.repeat_id);
-                bayesdata.nrepeatID(si,1) = max(ranking);
-                bayesdata.repeatID(si,:) = W.extend(ranking',LEN);
-
-                bayesdata.infocond(si,:) = W.nan_equal(bayesdata.dI(si,:), 0) + 1;
-                % 1 - [1 3], 2 - [2 2]
-                bayesdata.rpIDHorizon(si,:) = W.extend(W.arrayfun(@(x)unique(bayesdata.horizon(si,(ranking == x))), 1:bayesdata.nrepeatID(si,1)), LEN);
-                bayesdata.rpIDCond(si,:) = W.extend(W.arrayfun(@(x)unique(bayesdata.infocond(si,(ranking == x))), 1:bayesdata.nrepeatID(si,1)), LEN);
-                
-                bayesdata.choice4(si,1:LEN,:) = W.extend((gd.choice(:, 1:4)' == 2) + 0, LEN)';
-                bayesdata.reward4(si,1:LEN,:) = W.extend(gd.reward(:, 1:4)', LEN)'; 
-                
[…]

### 2025-04-19 codes_revision_PlosCB/EEplot_2noise_modelcomparison.m

@@ -1,40 +1,0 @@
-function EEplot_2noise_modelcomparison(plt, gp, simugp)
-    plt.figure(6,4,'fontsize_label', 18, 'fontsize_leg',15, 'fontsize_axes',18, 'fontsize_title', 20, ...
-        'gapH_custom', [10 0 0 0 0 0 10], 'gapW_custom', [10 0 0 0 0]);   
-    plt.setfig('xlabel', [repmat({''},1,20), repmat({'horizon'},1,4)], ...
-        'ylabel', [{'\sigma^{ran}_{horizon}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{horizon}, \sigma^{det}_{}'} repmat({''},1,3), ...
-        {'\sigma^{ran}_{}, \sigma^{det}_{horizon}'} repmat({''},1,3), {'\sigma^{ran}_{}, \sigma^{det}_{}'} repmat({''},1,3), ...
-        {'\sigma^{ran}_{horizon}'} repmat({''},1,3), {'\sigma^{det}_{horizon}'} repmat({''},1,3)], ...
-        'title', [{'p(high info)', 'p(low mean)', 'p(inconsistent), [1 3]', 'p(inconsistent), [2 2]', ...
-        } repmat({''}, 1, 20)], 'xlim', [0.5 2.5], 'xtick', [1 2], 'xticklabel', [1 6], ...
-        'ylim', repmat({[0.35 0.65],[0 0.35],[0 0.45],[0 0.45]},1,6), ...
-        'ytick', repmat({[0.4:0.1:0.6],[0:.1:0.3],[0:.1:0.5],[0:.1:0.5]},1,6));
-    plt.setfig_all('legloc', 'SE');
-    for mi = 1:6
-        tgp = simugp.(['model' char(64+mi)]);
-        plt.ax(mi,1);
-        plt.plot(1:2,tgp.GPav_GPav_p_hi13, tgp.GPav_GPste_p_hi13,'line','color','AZblue');
-        plt.plot(1:2,gp.GPav_p_hi13, gp.GPste_p_hi13,'line','color','AZred');
-        plt.setfig_ax('legend',{'model','data'});
-        plt.ax(mi,2);
-        plt.plot(1:2,tgp.GPav_GPav_p_lm22, tgp.GPav_GPste_p_lm22,'line','color','AZblue');
[…]

### 2025-04-19 codes_revision_PlosCB/Figures_6model.m

@@ -1,75 +1,0 @@
-%%
-plt = W_plt('savedir', '../figures_revision', 'savepfx', 'RDBayes', 'isshow', true, ...
-    'issave', true, 'extension',{'svg', 'jpg'});
-%% load simulations
-W.library_wang('Wang_EEHorizon');
-game0 = readtable('../data/all/data_all.csv', 'Delimiter', ',');
-idx = load('../data/all/idxsub_exclude');
-tid = idx.idxsub(idx.id);
-game0 = game0(vertcat(tid{:}),:);
-
-opt_sub = {'subjectID', 'date', 'time'};
-opt_preprocess = 'EEpreprocess_game_basic';
-opt_game_sub = {'EEpreprocess_game_sub_repeatedgame'};
-opt_analysis = {'EEanalysis_sub_basic'};
-[sub, ~, idxsub] = W.analysis_pipeline(game0, opt_sub, opt_preprocess, opt_game_sub, opt_analysis, []);
-gp = W.analysis_1group(sub, [], ...
-            {{'p_inconsistent13','p_inconsistent13_randomtheory_byR'},{'p_inconsistent22','p_inconsistent22_randomtheory_byR'},...
-             {'p_inconsistent13'},{'p_inconsistent22'}});
-%% single out simu
-gamenow = game0;
-for si = 1:length(idxsub)
-    gamenow.choice_5(idxsub{si}) = simu.choice(si,1:simu.nTrial(si))' + 1;
-end
-[tsimu_sub] = W.analysis_pipeline(gamenow, opt_sub, opt_preprocess, ...
-    opt_game_sub, opt_analysis, [], idxsub);
-ttgp = W.analysis_1group(tsimu_sub, [], ...
-    {{'p_inconsistent13','p_inconsistent13_randomtheory_byR'},{'p_inconsistent22','p_inconsistent22_randomtheory_byR'},...
-    {'p_inconsistent13'},{'p_inconsistent22'}});
-sgp.modelA = W.analysis_1group(vertcat(ttgp, ttgp), false);
-%%
-global muteprint
-muteprint = 1;
-simugp = cell(10,6);
-for repi = 1:10
-    W.disp(repi,'');
-    simu = load(fullfile('../bayesoutput_revision/simu6modelMAP/', sprintf('bayes_6model_simu_rep%d.mat', repi))).simu;
-    for mi = 1:6
-        gamenow = game0;
-        for si = 1:length(idxsub)
[…]

### 2025-04-19 codes_revision_PlosCB/MAINNNN_6model.m

@@ -1,58 +1,0 @@
-JAGS_setup;
-%%
-data = importdata('../data/all_revision/bayesdata.mat');
-for numi = 1:6
-    W.print('model: %d', numi);
-    data.modelname = sprintf('2noisemodel%s_2cond', char(numi + 'A' -1));
-    wj.setup_data_dir(data, '../bayesoutput_revision/R1_6model');
-    wjinfo = EEbayes_analysis(data, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['R1_fit_' char(numi + 'A' -1)]);
-    wj.run;
-end
-%% simulate 
-for repi = 1:10
-    for numi = 2:6
-        paramsub = load(fullfile('../bayesoutput_revision/R1_6model', [HBItest 'HBI_R1_fit_' char(numi + 'A' -1) '_stat.mat'])).stats.mean;
-        if ~isfield(paramsub, 'NoiseRan_sub')
-            paramsub.NoiseRan_sub = zeros(size(paramsub.NoiseDet_sub));
-        end
-        if ~isfield(paramsub, 'NoiseDet_sub')
-            paramsub.NoiseDet_sub = zeros(size(paramsub.NoiseRan_sub));
-        end
-        simugame = EEsimulate_bayes_2noise_2cond(data, paramsub.Infobonus_sub, ...
-            paramsub.bias_sub, paramsub.NoiseRan_sub, paramsub.NoiseDet_sub);
-        W.save(sprintf('../bayesoutput_revision/R1_6model/R1_simugame_%s_%d.mat', char(numi + 'A' -1), repi), ...
-            'simugame', simugame);
-        % fit
-        simugame.modelname = '2noisemodelA_2cond';
-        wj.setup_data_dir(simugame, '../bayesoutput_revision/R1_6model');
-        wjinfo = EEbayes_analysis(simugame, nchains);
-        savename = sprintf('R1_fitA_simugame_%s_%d', char(numi + 'A' -1), repi);
-        wj.setup(wjinfo.modelfile, wjinfo.params, struct, savename);
-        wj.run;
-    end
-end
-%% posterior checks
-game0 = readtable('../data/all/data_all.csv', 'Delimiter', ',');
-idx = load('../data/all/idxsub_exclude');
-idxsub = idx.idxsub(idx.id);
-tid = idx.idxsub(idx.id);
[…]

### 2025-04-19 codes_revision_PlosCB/getMAP.m

@@ -1,37 +1,0 @@
-function st = getMAP(sp)
-    varnames = {'Infobonus_sub', 'bias_sub', 'NoiseRan_sub', 'NoiseDet_sub'};
-    for vi = 1:length(varnames)
-        vn = varnames{vi};
-        tval = sp.(vn);
-        if vi <=2
-            [~,~,sz1,sz2] = size(tval);
-            tout = nan(sz1,sz2);
-            for i = 1:sz1
-                for j = 1:sz2
-                    tout(i,j) = findMAP(tval(:,:,i,j));
-                end
-            end
-        else
-            [~,~,sz1,sz2,sz3] = size(tval);
-            tout = nan(sz1,sz2,sz3);
-            for i = 1:sz1
-                for j = 1:sz2
-                    for k = 1:sz3
-                        tout(i,j,k) = findMAP(tval(:,:,i,j,k));
-                    end
-                end
-            end
-        end
-        st.(vn) = tout;
-    end
-end
-function x = findMAP(v)
-    v = reshape(v, 1, []);
-    vx = linspace(min(v), max(v), 1000);
-    dx = mean(diff(vx), 'all');
-    stv = std(v);
-    one = ones(length(v),1);
-    val = sum1d(one, v', stv, dx, min(v), max(v));
-    [~, tid] = max(val);
-    x = vx(tid);
-end

### 2025-04-19 codes_revision_PlosCB/main_2_preparedata.m

@@ -1,26 +1,0 @@
-addpath(genpath('../codes_support/'))
-%% prepare data
-load('../data/all/idxsub_exclude.mat');
-game = readtable('../data/all/data_all.csv', 'Delimiter', ',');
-g = W.tab_autofieldcombine(game);
-dI = arrayfun(@(x)getdI_onegame(g.choice(x,:), g.reward(x,:)), 1:size(g,1));
-game.dIvar = dI';
-
-bayesdata = EEbayes_reformat(game, idxsub(id), 'EE2noise');
-save('../data/all_revision/bayesdata.mat', 'bayesdata');
-
-
-% bayes_data_2condKF = EEbayes_reformat(game, idxsub(id), '2noisemodel_2condKF');
-% save('../data/all_revision/bayes_2noise_2condKF.mat', 'bayes_data_2condKF');
-% %%
-
-% p_mu_obs = generate_likelihood();
-% g = W.tab_autofieldcombine(game);
-% dI = arrayfun(@(x)getdI_onegame(g.choice(x,:), g.reward(x,:), p_mu_obs), 1:size(g,1));
-% game.dIvar = dI';
-% bayes_data_2cond_dIvar = EEbayes_reformat(game, idxsub(id), '2noisemodel_2cond_dIvar');
-% save('../data/all_revision/bayes_2noise_2cond_dIvar.mat', 'bayes_data_2cond_dIvar');
-%%
-% 
-% bayes_data_2cond_dIvar2 = EEbayes_reformat(game, idxsub(id), '2noisemodel_2cond_dIvar');
-% save('../data/all_revision/bayes_2noise_2cond_dIvar2.mat', 'bayes_data_2cond_dIvar2');

### 2025-04-19 codes_revision_PlosCB/main_3_fit_Bayes_models.m

@@ -1,71 +1,0 @@
-%% test convergence
-JAGS_setup;
-nchains = 4; % How Many Chains?
-nburnin = 0; % How Many Burn-in Samples?
-nsamples = 5000; % How Many Recorded Samples?
-wj.setup_params(nchains, nburnin, nsamples);
-d = load(fullfile(fullfile('../data',ver), 'bayesdata.mat')).(['bayesdata']);
-d.modelname = '2noisemodel_2cond';
-wj.setup_data_dir(d, outputdir);
-wjinfo = EEbayes_analysis(d, nchains);
-wj.setup(wjinfo.modelfile, wjinfo.params, struct, 'DetRanNoiseR1_convergence');
-wj.run;
-%% main bayes (base classes)
-% set of models
-% 1cond vs 2cond
-% dI vs dIvar
-suffix = {'', '_2cond', '_2cond_dIvar', '_2cond_dIvar_both', '_2cond_dRonly', '_2cond_nobias', ...
-    'B_2cond','C_2cond','D_2cond','E_2cond','F_2cond','_2cond_nonhierarchical'};% reviewer fit non-hierarchical
-for fi = 1:length(suffix)
-    JAGS_setup;
-    d = load(fullfile(fullfile('../data',ver), 'bayesdata.mat')).(['bayesdata']);
-    d.modelname = sprintf('2noisemodel%s', suffix{fi});
-    wj.setup_data_dir(d, outputdir);
-    wjinfo = EEbayes_analysis(d, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['DetRanNoiseR1' suffix{fi}]);
-    wj.run;
-end
-%% check convergence
-% load('..\bayesoutput_revision\all_revision\HBI_DetRanNoiseR1_convergence_samples.mat');
-%% 
-% plt = W_plt('savedir', '../figures', 'savepfx', 'RanDetNoise', 'isshow', true, ...
-%     'issave', false, 'extension',{'svg', 'jpg'});
-% plt.figure(6,10);
-% % id = 100:10000;
-% id = 1:5000;
-% for i = 1:10
-%     plt.ax(1, i);
-%     plt.plot([], samples.InfoBonus_mu_n(i,:,1), [], 'line');
-%     plt.plot([], samples.InfoBonus_mu_n(i,:,2), [], 'line');
-%     plt.dashY(2000)
-%     plt.ax(2, i);
[…]

### 2025-04-19 codes_revision_PlosCB/main_compare_hierarchical_vs_non_hierarchical.m

@@ -1,224 +1,0 @@
-JAGS_setup;
-outputdir = '../bayesoutput_revision/hierarchical';
-%% load data
-data = W.load('../data/all_revision/bayesdata');
-%%
-modelnames = {'model_2noise_2cond_nonhierarchical','model_2noise_2cond'};
-suffix = {'_2cond_nonhierarchical','_2cond'};
-for mi = 1:length(modelnames)
-    modelname = modelnames{mi};
-    %% fit models to data
-    d = data;
-    d.modelname = sprintf('2noisemodel%s', suffix{mi});
-    wj.setup_data_dir(d, outputdir);
-    wjinfo = EEbayes_analysis(d, nchains);
-    wj.setup(wjinfo.modelfile, wjinfo.params, struct, ['fit' suffix{mi}]);
-    wj.run;
-    %% load fitted data
-    st0 = W.load(fullfile(outputdir, ['HBI_fit' suffix{mi} '_stat']));
-    st0 = st0.stats.mean;
-    sp = W.load(fullfile(outputdir, ['HBI_fit' suffix{mi} '_samples']));
-    stmap = getMAP(sp);
-    save(fullfile(outputdir, ['HBI_fit' suffix{mi} '_MAP']), 'stmap');
-    %% simulate data from fitted model
-    st = st0;
-    simu = EEsimulate_bayes_2noise_2cond(data, st.Infobonus_sub, ...
-        st.bias_sub, st.NoiseRan_sub, st.NoiseDet_sub);
-    %% simulate data from shuffled model
-    nsub = size(st.Infobonus_sub, 2);
-    id1 = randperm(nsub);
-    id2 = randperm(nsub);
-    st_shuffled = st;
-    st_shuffled.NoiseRan_sub = st.NoiseRan_sub(:,:,id1);
-    st_shuffled.NoiseDet_sub = st.NoiseDet_sub(:,:,id2);
-    simu_shuffled = EEsimulate_bayes_2noise_2cond(data, st_shuffled.Infobonus_sub, ...
-        st_shuffled.bias_sub, st_shuffled.NoiseRan_sub, st_shuffled.NoiseDet_sub);
-    save(fullfile(outputdir, ['simugames' suffix{mi}]), 'simu', 'simu_shuffled', 'st_shuffled');
[…]

### 2025-04-19 codes_revision_PlosCB/plot_det_vs_ran.m

@@ -1,17 +1,0 @@
-function plt = plot_det_vs_ran(plt, st1, st2)
-    plt.figure(2,4, 'is_title', 'all');
-    cond = {'[1 3]', '[2 2]'};
-    for ci = 1:2
-        for hi = 1:2
-            plt.ax(ci, hi);
-            ylab = W.iif(hi == 1, {cond{ci}, 'Deterministic'}, {'Deterministic'});
-            str = plt.scatter(squeeze(st1.NoiseRan_sub(ci, hi,:)), squeeze(st1.NoiseDet_sub(ci, hi,:)), 'corr');
-            plt.setfig_ax('xlabel', 'Random', 'ylabel', ylab, 'title', sprintf('data, H = %d', hi), 'legend', str)
-
-            plt.ax(ci, hi + 2);
-            str = plt.scatter(squeeze(st2.NoiseRan_sub(ci, hi,:)), squeeze(st2.NoiseDet_sub(ci, hi,:)), 'corr');
-            plt.setfig_ax('xlabel', 'Random', 'ylabel', 'Deterministic', 'title', sprintf('shuffled, H = %d', hi), 'legend', str)
-        end
-    end
-    plt.update;
-end

### 2025-04-19 codes_revision_PlosCB/plot_parameter_recovery.m

@@ -1,39 +1,0 @@
-function plt = plot_parameter_recovery(plt, st1, st2, option)
-    if option == 1
-        plt.figure(2,2, 'is_title', 'all');
-        for hi = 1:2
-            plt.ax(1, hi);
-            ylab = {'recovered'};
-                tx = squeeze(st1.NoiseRan_sub(hi,:));
-                ty = squeeze(st2.NoiseRan_sub(hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-            plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-
-            plt.ax(2, hi);
-                tx = squeeze(st1.NoiseDet_sub(hi,:));
-                ty = squeeze(st2.NoiseDet_sub(hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-            plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-        end
-    else
-        plt.figure(2,4, 'is_title', 'all');
-        cond = {'[1 3]', '[2 2]'};
-        for ci = 1:2
-            for hi = 1:2
-                plt.ax(ci, hi);
-                ylab = W.iif(hi == 1, {cond{ci}, 'recovered'}, {'recovered'});
-                tx = squeeze(st1.NoiseRan_sub(ci, hi,:));
-                ty = squeeze(st2.NoiseRan_sub(ci, hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-                plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-    
-                plt.ax(ci, hi + 2);
-                tx = squeeze(st1.NoiseDet_sub(ci, hi,:));
-                ty = squeeze(st2.NoiseDet_sub(ci, hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-                plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-            end
-        end
-    end
-    plt.update;
-end

### 2025-04-19 Plos Comp Bio - revision v1/manuscript_DeterministicRandomNoise.tex
u

@@
 \PassOptionsToPackage{textsize=scriptsize}{todonotes}
-\usepackage[]{changes}
+\usepackage[final]{changes}
 \usepackage{changes}
@@
 In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
+\added[id=siyu]{To ensure that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we also calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa), and they do not correlate (Supplementary Figure S9).}
 	% $\frac{\mbox{\small Recovered } \Delta \sigma_{det}}{\mbox{\small Simulated } \Delta \sigma_{det}} = 0.74$
@@
 That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is very similar to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
+\added[id=siyu]{We also fit alternative models that fit random and deterministic noises separately for [1 3] and [2 2] conditions, and a model with an alternative definition of information bonus $dI$.
+These results are presented in the supplemental materials (Supplemental Figure.
+S17, S18).
+}
 
@@
 As shown in Figure \ref{fig:mb3}, the model can account for all qualitative patterns in the data --- the increase in p(high info), p(low mean), and p(inconsistent) with horizon, and that p(inconsistent) is in between pure random and pure deterministic noise.
-The quantitative agreement is almost perfect for p(high info) and for p(inconsistent) in the [1 3] condition, but the model seems to systematically overestimate p(low mean) and p(inconsistent) in [2 2] conditions, although the discrepancy is relatively small (overestimating p(low mean) by 0.054 or 31.37\%, and p(inconsistent) by 0.049 or 27.83\% in [2 2] condition).
+\added[id=siyu]{The quantitative agreement is almost perfect for p(high info) and for p(inconsistent), but the model slightly overestimate p(low mean) in [2 2] conditions.
+This has to do with the skewness of the subject-level posterior distribution (see Supplemental Materials).
+}
+%The quantitative agreement is almost perfect for p(high info) and for p(inconsistent) in the [1 3] condition, but the model seems to systematically overestimate p(low mean) and p(inconsistent) in [2 2] conditions, although the discrepancy is relatively small (overestimating p(low mean) by 0.054 or 31.37\%, and p(inconsistent) by 0.049 or 27.83\% in [2 2] condition).
 
-As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these reduced models fail to capture all qualitative patterns (Supplementary Figure S13).
+As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these reduced models fail to capture all qualitative patterns (Supplementary Figure S14, S15).
 Full details of this analysis can be found in Supplementary Materials section 3.2.

### 2025-04-19 Plos Comp Bio - revision v1/supplemental/DetRanNoise_Supplementary.tex
u

@@
 
+	In general, increasing random noise leads to higher p(low mean) and lower p(consistent) between repeated games, and increasing deterministic noise leads to higher p(low mean) and higher p(consistent).
+
 	\begin{figure}[H]
@@
 			\caption{Same as Figure \ref{fig:paramrecover_session}, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.}
+			\label{fig:paramrecover_sub}
+		\end{center}
+	\end{figure}
+
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.65\textwidth]{figures/R1_shuffle_shuffled_recovery.jpg}
+			\caption{Parameter recovery for shuffled data.
+To show that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa).
+Ground-truth values are shuffled best-fit parameters.
+As expected, ground-truth random values do not correlate with recovered deterministic noises, showing that the increase of deterministic noise with horizon is genuine and not a by-product of increase of random noise, and vice versa.
+}
 			\label{fig:paramrecover_sub}
@@
 
+	In our model, the subject-level noises are assumed to follow a gamma distribution (to ensure positiveness), the posteriors are right skewed (maximal likelihood estimation or mode is smaller than the mean), Because of the skewness, simulating data with the “mean” is nosier than simulating with the true distribution.
+Simulating from the true distribution requires taking expected value over all possible noise values, for ease of implementation, we simulated data by taking random samples from the posterior distribution (instead of using the mean).
+The simulation was repeated 50 times and then averaged (Supplementary Figure \ref{fig:posteriorcheck6models}).
+Simulation using the "mode" or maximal likelihood estimation was also shown here (Supplementary Figure \ref{fig:posteriorcheck6models2}).
+
 	\begin{figure}[hp]
@@
 		\begin{center}
-			\includegraphics[width=0.8\textwidth]{figures/RDBayes_2noise_6modelcomparison.jpg}
+			\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_dist.jpg}
 			\caption{
@@
 			\label{fig:posteriorcheck6models}
+		\end{center}
+	\end{figure}
+	\begin{figure}[hp]
+	\begin{center}
+		\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_map.jpg}
+		\caption{
+			Model comparison (using maximal likelihood estimation).
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+		\label{fig:posteriorcheck6models2}
 	\end{center}
@@
 	\end{figure}
+
+	\newpage
+	\subsection{Alternative models: separate noises for [1 3] and [2 2] condition}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior.jpg}
+			\caption{Model based analysis from a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions.
+The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, E) and deterministic (C,G) noises are nonzero (A, C, E, G) and change with horizon (B, D, F, H).
+However, random noise has both a greater magnitude overall (A, E) and a greater change with horizon (B, F) than deterministic noise.
+Both random and deterministic noises are higher in [1 3] compared to [2 2] conditions.}
+			\label{fig:s8}
+		\end{center}
+	\end{figure}
+	\newpage
+	\subsection{Alternative models: use variance differences as dI}
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior_dIvar.jpg}
+			\caption{Model based analysis from a model that uses variance differences as dI.
+The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
+			\label{fig:s8}
+		\end{center}
+	\end{figure}
+
 \end{document}

### 2025-04-19 codes/codedraft_revision_PlosCB/plot_parameter_recovery.m

@@ -1,54 +1,0 @@
-function plt = plot_parameter_recovery(plt, st1, st2, option, ax, tlt)
-    if option == 1
-        plt.figure(2,2, 'is_title', 'all');
-        for hi = 1:2
-            plt.ax(1, hi);
-            ylab = {'recovered'};
-                tx = squeeze(st1.NoiseRan_sub(hi,:));
-                ty = squeeze(st2.NoiseRan_sub(hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-            plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-
-            plt.ax(2, hi);
-                tx = squeeze(st1.NoiseDet_sub(hi,:));
-                ty = squeeze(st2.NoiseDet_sub(hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-            plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-        end
-    elseif option == 2
-        plt.figure(2,4, 'is_title', 'all');
-        cond = {'[1 3]', '[2 2]'};
-        for ci = 1:2
-            for hi = 1:2
-                plt.ax(ci, hi);
-                ylab = W.iif(hi == 1, {cond{ci}, 'recovered'}, {'recovered'});
-                tx = squeeze(st1.NoiseRan_sub(ci, hi,:));
-                ty = squeeze(st2.NoiseRan_sub(ci, hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-                plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-    
-                plt.ax(ci, hi + 2);
-                tx = squeeze(st1.NoiseDet_sub(ci, hi,:));
-                ty = squeeze(st2.NoiseDet_sub(ci, hi,:));
-                str = plt.scatter(tx, ty, 'diag');
-                plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
-            end
-        end
[…]

### 2025-04-19 codes/plot_nonhierarchical.m

@@ -1,0 +1,54 @@
+function plt = plot_nonhierarchical(plt, st1, st2, option, ax, tlt)
+    % if option == 1
+    %     plt.figure(2,2, 'is_title', 'all');
+    %     for hi = 1:2
+    %         plt.ax(1, hi);
+    %         ylab = {'recovered'};
+    %             tx = squeeze(st1.NoiseRan_sub(hi,:));
+    %             ty = squeeze(st2.NoiseRan_sub(hi,:));
+    %             str = plt.scatter(tx, ty, 'diag');
+    %         plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+    % 
+    %         plt.ax(2, hi);
+    %             tx = squeeze(st1.NoiseDet_sub(hi,:));
+    %             ty = squeeze(st2.NoiseDet_sub(hi,:));
+    %             str = plt.scatter(tx, ty, 'diag');
+    %         plt.setfig_ax('xlabel', 'simulated', 'ylabel', 'recovered', 'title', sprintf('Det, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+    %     end
+    % elseif option == 2
+    %     plt.figure(2,4, 'is_title', 'all');
+    %     cond = {'[1 3]', '[2 2]'};
+    %     for ci = 1:2
+    %         for hi = 1:2
+    %             plt.ax(ci, hi);
+    %             ylab = W.iif(hi == 1, {cond{ci}, 'recovered'}, {'recovered'});
+    %             tx = squeeze(st1.NoiseRan_sub(ci, hi,:));
+    %             ty = squeeze(st2.NoiseRan_sub(ci, hi,:));
+    %             str = plt.scatter(tx, ty, 'diag');
+    %             plt.setfig_ax('xlabel', 'simulated', 'ylabel', ylab, 'title', sprintf('Ran, H = %d, %.2f', hi, mean(tx < ty)), 'legend', str)
+    % 
+    %             plt.ax(ci, hi + 2);
+    %             tx = squeeze(st1.NoiseDet_sub(ci, hi,:));
+    %             ty = squeeze(st2.NoiseDet_sub(ci, hi,:));
+    %             str = plt.scatter(tx, ty, 'diag');
[…]

### 2025-04-21 Plos Comp Bio - revision v1/manuscript_DeterministicRandomNoise.tex
u

@@
 For purely deterministic noise this computation is simple because people should make the exact same decisions each time in repeated games, meaning that $p(\mbox{inconsistent}) = 0$ in this case.
-For purely random noise, the two games should be treated independently, allowing us to compute the choice inconsistency in terms of the probability of choosing the low mean option, $p(\mbox{low mean})$, as
+For purely random noise, the two games should be treated independently.
+\added[id=siyu]{Since repeated decisions mean that participants either choose the low-mean option twice, or choose the high-mean option twice, we could predict} the choice inconsistency, $p(\mbox{inconsistent})$, based on the probability of choosing the low mean option, $p(\mbox{low mean})$, as
 \begin{equation*}
@@
 In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.
+\added[id=siyu]{While an increase in either random or deterministic noises could lead to higher $p(\mbox{low mean})$, an increase in random noise predicts higher $p(\mbox{inconsistent})$ while an increase in deterministic noise predicts lower $p(\mbox{inconsistent})$.
+}
 
@@
 Model fitting was performed using the MATJAGS and JAGS software \citep{jags, matjags} with full details given in the Methods.
+
+\added[id=siyu]{We also fit a series of reduced and alternative models to the data.
+This includes reduced models that assume only deterministic or random noises.
+We also fit models in which the standard deviation of random and deterministic noises $\sigma_{det}$ and $\sigma_{ran}$ are estimated separately for [1 3] and [2 2] information conditions.
+Lastly, we fit a model with an alternative definition of $\Delta I$ that $\Delta I$ is defined to be difference between the variances of rewards shown on the forced-choice trials.
+Results on these model variants are presented in the Supplementary Materials.}
 
@@
 
-Parameter recover in this task was good for this model, with fit values for $\sigma_{ran}$ and $\sigma_{det}$ showing strong correlations with their simulated values in both horizon (H) conditions (For $\sigma_{ran}$, $R = 0.91 (\mbox{H} = 1) \mbox{ and } 0.84 (\mbox{H} = 6), p < 0.001$, For $\sigma_{det}$, $R = 0.60 (\mbox{H} = 1) \mbox{ and } 0.59 (\mbox{H} = 6), p < 0.001$).
+Parameter recover in this task was good for this model (Figure \ref{fig:paramrecover_main}, Supplementary Figure S7, S8), with fit values for $\sigma_{ran}$ and $\sigma_{det}$ showing strong correlations with their simulated values in both horizon (H) conditions (For $\sigma_{ran}$, $R = 0.91 (\mbox{H} = 1) \mbox{ and } 0.84 (\mbox{H} = 6), p < 0.001$, For $\sigma_{det}$, $R = 0.60 (\mbox{H} = 1) \mbox{ and } 0.59 (\mbox{H} = 6), p < 0.001$).
 However, while the relationship was near perfect for random noise ($\frac{\mbox{\small Recovered } \sigma_{ran}}{\mbox{\small Simulated } \sigma_{ran}} = 1.01$), there was a systematic bias to underestimate the level of deterministic noise by about 32\% ($\frac{\mbox{\small Recovered } \sigma_{det}}{\mbox{\small Simulated } \sigma_{det}} = 0.68$).
@@
 In particular, while we generate two samples of random noise for each repeated game pair, we only generate one sample of deterministic noise, which by definition is the same in both of the repeated games.
-\added[id=siyu]{To ensure that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we also calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa), and they do not correlate (Supplementary Figure S9).}
 	% $\frac{\mbox{\small Recovered } \Delta \sigma_{det}}{\mbox{\small Simulated } \Delta \sigma_{det}} = 0.74$
 
-In addition to the conventional subject-level parameter recovery analysis presented here, we also performed parameter recovery analysis that examined how faithful the full posterior distribution of group-level parameters can be recovered in simulated data (Supplementary Figure S5, S6).
+In addition to the conventional subject-level parameter recovery analysis presented here, we also performed parameter recovery analysis that examined how faithful the full posterior distribution of group-level parameters can be recovered in simulated data (Supplementary Figure S5, S6, S9, S10).
 Qualitatively, we also showed that our way of modeling deterministic noise is capable of capturing known deterministic processes intentionally omitted from the full model (Supplementary Figure S4).
@@
 That is, when we compute the relative increase in deterministic noise with horizon, $\sigma^{det}_{horizon6}/\sigma^{det}_{horizon1}$, it is very similar to the relative increase in random noise with horizon $\sigma^{ran}_{horizon6}/\sigma^{ran}_{horizon1}$.
-\added[id=siyu]{We also fit alternative models that fit random and deterministic noises separately for [1 3] and [2 2] conditions, and a model with an alternative definition of information bonus $dI$.
-These results are presented in the supplemental materials (Supplemental Figure.
-S17, S18).
+
+\added[id=siyu]{Similar results are found in other variants of our model.
+Results for a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions, and a model with an alternative definition of information bonus $dI$, can be found in the Supplemental Materials (Supplemental Figure S12, S13).
 }
+
+\added[id=siyu]{To ensure that the joint increase of random and deterministic noises is genuine and not an artifact from the fitting procedure, we computed the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa), and they do not correlate (Supplementary Figure S14).
+Furthermore, we simulated data from a series of reduced models with known random and deterministic noise values in which either random or deterministic noise does not change with Horizon, and fit our model to the simulated data.
+Our model detects and only detects a change in random/deterministic noise with horizon, when the change is present in the model that simulates the data (Supplementary Figure S15, S16).}
 
@@
 
-As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these reduced models fail to capture all qualitative patterns (Supplementary Figure S14, S15).
-Full details of this analysis can be found in Supplementary Materials section 3.2.
+As a control, we also applied posterior predictive checks on alternative models that consider only deterministic or only random noise, and these reduced models fail to capture all qualitative patterns (Supplementary Figure S17, S18).
+Full details of this analysis can be found in Supplementary Materials section 3.5.
 
@@
 For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}.
-\added[id=siyu]{Cross-game deterministic strategies and stimuli from previous games (e.g., memory of previously seen game) were also treated as random noise in our model \citep{WYART2016, Wyart_2018}.} In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
+\added[id=siyu]{Another limitation is that deterministic noise is defined based on the stimulus within a game, between-game deterministic strategies and stimuli from previous games (e.g., memory of previously seen game) were also treated as random noises in our model, although our model could be potentially extended by considering deterministic noise over a sequence of stimuli across games \citep{WYART2016, Wyart_2018}.} In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
 As a result, from both a conceptual and methodological perspective, it is possible that the remaining 86\% of the decision noise that is not stimulus-driven noise, could be deterministic.
@@
 
-\added[id=siyu]{We will show that changes in reward processing can be equivalent to changes in noise in decision models (see also \cite{Cinotti19}).}.
+\added[id=siyu]{We show first that a change in noise is mathematically equivalent to a change in reward signal strength in our decision model (see also \cite{Cinotti19}).}.
 To show how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}.

### 2025-04-21 Plos Comp Bio - revision v1/supplemental/DetRanNoise_Supplementary.tex
u

@@
 
-	\subsection{Model-free analysis with simulated choice from a pure random noise model}
+	\subsection{Control analysis: model-free analysis with simulated choice from a random-noise only model}
 	\begin{figure}[H]
@@
 This suggests that our model is capable of detecting missing deterministic processes.
-
-	In general, increasing random noise leads to higher p(low mean) and lower p(consistent) between repeated games, and increasing deterministic noise leads to higher p(low mean) and higher p(consistent).
 
@@
 
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=0.65\textwidth]{figures/R1_shuffle_shuffled_recovery.jpg}
-			\caption{Parameter recovery for shuffled data.
-To show that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa).
-Ground-truth values are shuffled best-fit parameters.
-As expected, ground-truth random values do not correlate with recovered deterministic noises, showing that the increase of deterministic noise with horizon is genuine and not a by-product of increase of random noise, and vice versa.
-}
-			\label{fig:paramrecover_sub}
-		\end{center}
-	\end{figure}
-
 	\newpage	
-	\subsection{Parameter recovery over a bigger range of arbitrary combinations of deterministic and random noises \label{gridrecovery}}
+	\subsection{Parameter recovery of arbitrary combinations of deterministic and random noises \label{gridrecovery}}
 
@@
 	\newpage
+	\subsection{Alternative model: fit separate random and deterministic noises for [1 3] and [2 2] conditions}
+
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior.jpg}
+			\caption{Model based analysis from a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions.
+The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, E) and deterministic (C,G) noises are nonzero (A, C, E, G) and change with horizon (B, D, F, H).
+However, random noise has both a greater magnitude overall (A, E) and a greater change with horizon (B, F) than deterministic noise.
+Moreover, both random and deterministic noises have a greater magnitude in [1 3] compared to [2 2] conditions.}
+			\label{fig:s8}
+		\end{center}
+	\end{figure}
+	\newpage
+	\subsection{Alternative model: use variance differences as dI}
+	Here, we fit a model with an alternative definition of $\Delta I$ that $\Delta I$ is defined to be difference between the variances of rewards shown on the forced-choice trials.
+	\begin{figure}[H]
+		\begin{center}
+			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior_dIvar.jpg}
+			\caption{Model based analysis from a model that uses variance differences as dI.
+The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
+			\label{fig:s8}
+		\end{center}
+	\end{figure}
+	\newpage
+
+	\subsection{Examine the co-variance of random vs deterministic noise}
+	\begin{figure}[hp]
+		\begin{center}
+			\includegraphics[width=0.65\textwidth]{figures/R1_shuffle_shuffled_recovery.jpg}
+			\caption{Parameter recovery for shuffled data.
+To show that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa).
+Ground-truth values are shuffled best-fit parameters.
+As expected, ground-truth random values do not correlate with recovered deterministic noises, showing that the increase of deterministic noise with horizon is genuine and not a by-product of increase of random noise, and vice versa.
+}
+			\label{fig:paramrecover_sub}
+		\end{center}
+	\end{figure} 
+	\newpage
+
 	\subsection{Alternative models: variations of the two-noise model}	
-	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we also built and fit five additional versions of the model.
-These models varied in whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+	In this section, we examined 6 model variants that varied in whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
 Specifically, we tested the following 6 models (Note that the $\sigma^{ran}_{horizon},\sigma^{det}_{horizon}$ model is our original full model).
@@
 
-	We again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
-As shown in Figure \ref{fig:posteriorcheck6models}, only one of these alternative models, where random noise is horizon dependent but deterministic noise is not, can capture the full qualitative pattern of behavior.
-However, the quantitative fit to the data is not as good (Figure \ref{fig:posteriorcheck6models}).% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
-
-	In our model, the subject-level noises are assumed to follow a gamma distribution (to ensure positiveness), the posteriors are right skewed (maximal likelihood estimation or mode is smaller than the mean), Because of the skewness, simulating data with the “mean” is nosier than simulating with the true distribution.
-Simulating from the true distribution requires taking expected value over all possible noise values, for ease of implementation, we simulated data by taking random samples from the posterior distribution (instead of using the mean).
-The simulation was repeated 50 times and then averaged (Supplementary Figure \ref{fig:posteriorcheck6models}).
-Simulation using the "mode" or maximal likelihood estimation was also shown here (Supplementary Figure \ref{fig:posteriorcheck6models2}).
-
 	\begin{figure}[hp]
@@
 
-	\begin{figure}[hp]
-		\begin{center}
-			\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_dist.jpg}
-			\caption{
-				Model comparison.
-A-D.
-both deterministic and random noise are horizon dependent, E-H.
-only random noise is horizon dependent, I-L.
-only deterministic noise is horizon dependent, M-P.
-neither random nor deterministic noise is horizon dependent, Q-T.
-only deterministic noise is assumed to be present, U-X.
-only random noise is assumed to be present.}
-			\label{fig:posteriorcheck6models}
-		\end{center}
-	\end{figure}
-	\begin{figure}[hp]
-	\begin{center}
-		\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_map.jpg}
-		\caption{
-			Model comparison (using maximal likelihood estimation).
-A-D.
-both deterministic and random noise are horizon dependent, E-H.
-only random noise is horizon dependent, I-L.
-only deterministic noise is horizon dependent, M-P.
-neither random nor deterministic noise is horizon dependent, Q-T.
-only deterministic noise is assumed to be present, U-X.
-only random noise is assumed to be present.}
-		\label{fig:posteriorcheck6models2}
-	\end{center}
-	\end{figure}
-
-	Moreover, we examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+	We then examined if our model can indeed qualitatively capture whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
 To test this, we simulated choices from each of the 6 models, and then fit the simulated choices with our original full model.
@@
 
-	\newpage
-	\subsection{Alternative models: separate noises for [1 3] and [2 2] condition}
-	\begin{figure}[H]
+	%we also built and fit five additional versions of the model.
+
+	To check whether all aspects of the model were necessary to reproduce the qualitative pattern of findings, we again simulated choices using fitted parameters from these models and repeated the model-free analysis on the simulated data.
+As shown in Figure \ref{fig:posteriorcheck6models}, only one of these alternative models, where random noise is horizon dependent but deterministic noise is not, can capture most of the qualitative features, including the increase in p(high info), p(low mean) and p(consistent).
+However, the quantitative fit to the data is not as good (Figure \ref{fig:posteriorcheck6models}).
+Moreover, when comparing p(consistent) with the theoretically predicted p(consistent) for pure random noise (see analysis in Figure 3), only the fully model is statistically significant in both horizon conditions.
+Out of 50 simulations, the best alternative model only shows statistical significance in half of the simulations in Horizon 6, and fails on the other 50\% of the times.
+Because deterministic noise is not allowed to change with horizon in this alternative model, in horizon 6, deterministic noise is not large enough compared to the magnitude of random noise, to make p(inconsistent) lower than the random-noise-only prediction.
+% SIYU says: they actually look quite similar, without horizon dependent deterministic noise, the fixed det noise model is only marginally worse on p(inconsistent) in [2 2] condition, as it overlaps with the pure random noise line more.
+
+	\begin{figure}[hp]
 		\begin{center}
-			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior.jpg}
-			\caption{Model based analysis from a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions.
-The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
-Both random (A, E) and deterministic (C,G) noises are nonzero (A, C, E, G) and change with horizon (B, D, F, H).
-However, random noise has both a greater magnitude overall (A, E) and a greater change with horizon (B, F) than deterministic noise.
-Both random and deterministic noises are higher in [1 3] compared to [2 2] conditions.}
-			\label{fig:s8}
+			\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_dist.jpg}
+			\caption{
+				Model comparison.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+			\label{fig:posteriorcheck6models}
 		\end{center}
@@
 	\newpage
-	\subsection{Alternative models: use variance differences as dI}
-	\begin{figure}[H]
+
+	In our model, the subject-level noises are assumed to follow a gamma distribution (to ensure positiveness), the posteriors are right skewed (maximal likelihood estimation or mode is smaller than the mean), Because of the skewness, simulating data with the “mean” is noisier than simulating with the true distribution.
+Simulating from the true distribution requires taking expected value over all possible noise values, for ease of implementation, we simulated data by taking random samples from the posterior distribution (instead of using the mean).
+The simulation was repeated 50 times and then averaged (Supplementary Figure \ref{fig:posteriorcheck6models}).
+Simulation using the "mode" or maximal likelihood estimation was also shown here (Supplementary Figure \ref{fig:posteriorcheck6models2}) and produces closer fit to the data.
+
+	\begin{figure}[hp]
 	\begin{center}
-			\includegraphics[width=0.7\textwidth]{figures/R1_hyperprior_dIvar.jpg}
-			\caption{Model based analysis from a model that uses variance differences as dI.
-The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
-Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
-However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.}
-			\label{fig:s8}
+		\includegraphics[width=0.8\textwidth]{figures/R1_2noise_6modelcomparison_map.jpg}
+		\caption{
+			Model comparison (using maximal likelihood estimation).
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.}
+		\label{fig:posteriorcheck6models2}
 	\end{center}

### 2025-04-22 Plos Comp Bio - revision v1/manuscript_DeterministicRandomNoise.tex
noise paper revision

@@
 \title{Separating random and deterministic sources of computational noise in explore-exploit decisions}
-\author[1,\textcurrency]{Siyu Wang}
+\author[1]{Siyu Wang}%,\textcurrency
 \author[1,2,3]{Robert C. Wilson}
@@
 \affil[3]{Cognitive Science Program, University of Arizona, Tucson AZ, USA}
-\affil[ \textcurrency]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}
+%\affil[ \textcurrency]{Current Address: Laboratory of Neuropsychology, National Institute of Mental Health, National Institutes of Health, Bethesda MD, USA}
 
@@
 		Human decision making is inherently variable.
-While this variability is often seen as a sign of suboptimal behavior, \added[id=siyu]{both theoretical work in machine learning and empirical human studies} suggest that variability can actually be adaptive.
+While this variability is often seen as a sign of suboptimal behavior, \added{both theoretical work in machine learning and empirical human studies} suggest that variability can actually be adaptive.
 An example arises when we must choose between exploring unknown options or exploiting options we know well.
@@
 	\end{abstract}
+
+	\newpage
+	\section*{Author Summary}
+	\added{
+	Human decisions often seem random.
+Even simple decisions like what food to order at a restaurant can be difficult to predict ahead of time.
+This randomness in our decisions can be beneficial, effectively allowing us to explore new options.
+One outstanding question is where the randomness in our decisions comes from.
+Sometimes, our seemingly random decisions are driven by predictable external factors, like what the guest at the next table ordered could influence what we order.
+Other times, our decisions are not driven by external factors but are instead made by random thoughts within our brain.
+In this work, we developed a computational method that quantifies the extent to which the apparent randomness in our decisions can be explained by deterministic sources of variability in the external stimuli, or random variability unexplained by the stimuli.
+We found evidence that randomness in exploratory decisions can be explained by both random (up to 86\%) and deterministic (more than 14\%) sources of variability.
+Moreover, our results suggest that both sources of variability are adaptive, which enables humans to explore more when it is more beneficial to explore.
+The joint adaptation of random and deterministic noises also suggests a common noise-gating mechanism for exploration.
+}
+
 	\newpage
@@
 
-	It has recently been shown that humans appear to use random exploration and can increase decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014}, \added[id=siyu]{as has also been suggested in computational models of animal behavior \citep{DOYA2002495, KHAMASSI2013441}}.
+	It has recently been shown that humans appear to use random exploration and can increase decision noise when it is more beneficial to explore \citep{Gershman2018, wilson2014}, \added{as has also been suggested in computational models of animal behavior \citep{DOYA2002495, KHAMASSI2013441}}.
 In one of these tasks, known as the Horizon Task \citep{wilson2014}, the key manipulation is the horizon condition, i.e.
@@
 Avoiding an ex is a deterministic process, but if we fail to take the ex's presence into account as scientists modeling the decision, then over a series of such decisions where the ex is present or not, we would mistakenly attribute the ensuing `variability' in choice to randomness.
-\added[id=siyu]{To dissociate a missing deterministic component from a true random process, choice consistency between repeated decisions can be utilized to decompose behavioral variability into predictable deterministic components and unpredictable random components \citep{WYART2016, Wyart_2018, Findling19, Findling21}.}
+\added{To dissociate a missing deterministic component from a true random process, choice consistency between repeated decisions can be utilized to decompose behavioral variability into predictable deterministic components and unpredictable random components \citep{WYART2016, Wyart_2018, Findling19, Findling21}.}
 
@@
 In both analyses, we focus on just the first free-choice trial in each game, where the only thing that differs between the horizon conditions is the number of choices that participants will make in the future.
-\added[id=siyu]{Subsequent choices in Horizon 6 games were not analyzed.}
+\added{Subsequent choices in Horizon 6 games were not analyzed.}
 
@@
 
-%\deleted[id=siyu]{For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6}
+%\deleted{For [1 3] condition, t(64) = 13.72, p $<$ 0.001 for horizon 1, t(64) = 16.71, p $<$ 0.001 for horizon 6; For [2 2] condition, t(64) = 9.55, p $<$ 0.001 for horizon 1, t(64) = 17.93, p $<$ 0.001 for horizon 6}
 \begin{figure}[H]
@@
 For purely random noise, the two games should be treated independently.
-\added[id=siyu]{Since repeated decisions mean that participants either choose the low-mean option twice, or choose the high-mean option twice, we could predict} the choice inconsistency, $p(\mbox{inconsistent})$, based on the probability of choosing the low mean option, $p(\mbox{low mean})$, as
+\added{Since repeated decisions mean that participants either choose the low-mean option twice, or choose the high-mean option twice, we could predict} the choice inconsistency, $p(\mbox{inconsistent})$, based on the probability of choosing the low mean option, $p(\mbox{low mean})$, as
 \begin{equation*}
@@
 In symbolic terms, for repeated games $i$ and $j$,  $n_{det}^i=n_{det}^j$  and $n_{ran}^i \neq n_{ran}^j$.
-\added[id=siyu]{While an increase in either random or deterministic noises could lead to higher $p(\mbox{low mean})$, an increase in random noise predicts higher $p(\mbox{inconsistent})$ while an increase in deterministic noise predicts lower $p(\mbox{inconsistent})$.
+\added{While an increase in either random or deterministic noises could lead to higher $p(\mbox{low mean})$, an increase in random noise predicts higher $p(\mbox{inconsistent})$ while an increase in deterministic noise predicts lower $p(\mbox{inconsistent})$.
 }
@@
 
-\added[id=siyu]{We also fit a series of reduced and alternative models to the data.
+\added{We also fit a series of reduced and alternative models to the data.
 This includes reduced models that assume only deterministic or random noises.
@@
 
-\added[id=siyu]{Similar results are found in other variants of our model.
+\added{Similar results are found in other variants of our model.
 Results for a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions, and a model with an alternative definition of information bonus $dI$, can be found in the Supplemental Materials (Supplemental Figure S12, S13).
@@
 
-\added[id=siyu]{To ensure that the joint increase of random and deterministic noises is genuine and not an artifact from the fitting procedure, we computed the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa), and they do not correlate (Supplementary Figure S14).
+\added{To ensure that the joint increase of random and deterministic noises is genuine and not an artifact from the fitting procedure, we computed the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa), and they do not correlate (Supplementary Figure S14).
 Furthermore, we simulated data from a series of reduced models with known random and deterministic noise values in which either random or deterministic noise does not change with Horizon, and fit our model to the simulated data.
@@
 As shown in Figure \ref{fig:mb3}, the model can account for all qualitative patterns in the data --- the increase in p(high info), p(low mean), and p(inconsistent) with horizon, and that p(inconsistent) is in between pure random and pure deterministic noise.
-\added[id=siyu]{The quantitative agreement is almost perfect for p(high info) and for p(inconsistent), but the model slightly overestimate p(low mean) in [2 2] conditions.
+\added{The quantitative agreement is almost perfect for p(high info) and for p(inconsistent), but the model slightly overestimate p(low mean) in [2 2] conditions.
 This has to do with the skewness of the subject-level posterior distribution (see Supplemental Materials).
@@
 For example, such deterministic noise could be driven by differences in where people look, or for how long they look, or by whether they were fidgeting or scratching their nose \citep{Musall2019}.
-\added[id=siyu]{Another limitation is that deterministic noise is defined based on the stimulus within a game, between-game deterministic strategies and stimuli from previous games (e.g., memory of previously seen game) were also treated as random noises in our model, although our model could be potentially extended by considering deterministic noise over a sequence of stimuli across games \citep{WYART2016, Wyart_2018}.} In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
+\added{Another limitation is that deterministic noise is defined based on the stimulus within a game, between-game deterministic strategies and stimuli from previous games (e.g., memory of previously seen game) were also treated as random noises in our model, although our model could be potentially extended by considering deterministic noise over a sequence of stimuli across games \citep{WYART2016, Wyart_2018}.} In addition to this conceptual limitation in measuring deterministic noise, parameter recovery simulations suggest that our estimation method also slightly underestimates deterministic noise (see Figure \ref{fig:paramrecover_main}, Supplementary Figure S5, S6).
 As a result, from both a conceptual and methodological perspective, it is possible that the remaining 86\% of the decision noise that is not stimulus-driven noise, could be deterministic.
@@
 
-\added[id=siyu]{We show first that a change in noise is mathematically equivalent to a change in reward signal strength in our decision model (see also \cite{Cinotti19}).}.
+\added{We show first that a change in noise is mathematically equivalent to a change in reward signal strength in our decision model (see also \cite{Cinotti19}).}.
 To show how a change in reward processing could affect random and deterministic noise, consider the simple decision model we introduced in Equation \ref{eq:newmodel}.
@@
 Conversely, our estimate of deterministic noise is a lower bound.
-\added[id=siyu]{Future work is needed to identify these additional sources of deterministic noise that are not controlled in our work, for example by tracking people's pose, body movements and eye movements during the experiment.}
+\added{Future work is needed to identify these additional sources of deterministic noise that are not controlled in our work, for example by tracking people's pose, body movements and eye movements during the experiment.}
 
@@
 
-% \added[id=siyu]{The deterministic noise hypothesis is in line with works in which neural variability can be accounted for by fluctuations in sensory inputs.
+% \added{The deterministic noise hypothesis is in line with works in which neural variability can be accounted for by fluctuations in sensory inputs.
 For example, }MT neurons were shown to have a reproducible temporal modulation in response to a fixed random motion stimuli \citep{Bair1996}.
@@
 
-% \added[id=siyu]{The random noise hypothesis, on the other hand, is consistent with findings of \cite{drugowitsch16}.} In particular these authors show that randomness in behavior could arise from imperfections in mental inference, which happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
+% \added{The random noise hypothesis, on the other hand, is consistent with findings of \cite{drugowitsch16}.} In particular these authors show that randomness in behavior could arise from imperfections in mental inference, which happen inside the brain, rather than in peripheral processes such as sensory processing and response selection.
 This suggests that a large proportion of variability in behavior may arise from computational errors in computing the correct strategy.
-\added[id=siyu]{Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition.
+\added{Although suboptimal inference is different from simply adding random noise to neural circuitry\citep{Pouget12}, as long as the suboptimality in neural computation is not a deterministic function of the stimuli, it is a form of random noise in our definition.
 }In the context of the Horizon Task, such computational errors would likely be larger in the long horizon condition as the correct course of action in these cases is much harder to compute.
@@
 %Taken at face value, the horizon-dependent increase in random noise is consistent with the idea that random exploration is driven by intrinsic variability in the brain. 
-% \added[id=siyu]{
+% \added{
 % % Regardless of whether the remaining 85\% is deterministic or random, the fact that the horizon change in the two noises are proportional to each other (Figure \ref{fig:ratio}) suggests a possible mechanism for random exploration.
@@
 % \end{equation*}
-% \added[id=siyu]{In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
+% \added{In this equivalent form, $\beta$ determines the weight given to the reward in the decision.
 From a model fitting perspective, these two forms are equivalent.
@@
 \subsection*{Data and code}
-Behavioral data as well as MATLAB codes to recreate the main figures from this paper will be made available upon publication.
-%website at https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/CZT6EE.
+Behavioral data as well as MATLAB codes to recreate the main figures from this paper will be made available upon publication at \url{https://github.com/wangxsiyu/RW_RandomDeterministicNoise.git}
+
+\section*{Supporting information}
+
+% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+\paragraph*{S1 Fig.}
+\label{S1_Fig}
+{\bf Replication of previous findings with data from all participants (i.e.
+no exclusions).} (A) model-free measure of behavioral variability, $p(\mbox{low mean})$, increases with horizon.
+(B) model-free measure of information seeking, $p(\mbox{high info})$, increases with horizon.
+(C) model-based measure of behavioral variability, decision noise $\sigma$, increases with horizon.
+(D) model-based measure of information seeking, information bonus $A$, increases with horizon.
+
+\paragraph*{S2 Fig.}
+\label{S2_Fig}
+{\bf Model-free analysis with data from all participants (i.e.
+no exclusions) suggests that both deterministic and random noise contribute to the choice variability in random exploration.
+} For both the [1 3] (A) and [2 2] (B) condition, people show greater choice inconsistency in horizon 6 than horizon 1.
+However, the extent to which their choices are inconsistent lies between what is predicted by purely deterministic and random noise, suggesting that both noise sources influence the decision..
+
+\paragraph*{S3 Fig.}
+\label{S3_Fig}
+{\bf Model-free analysis with simulated choices from a model that has only random noise validates our prediction of p(inconsistent) for pure random noise.} The extent to which simulated choices are inconsistent completely overlaps with our pure random noise prediction(p $>$ 0.05).
+This suggests that when choice inconsistency lies below the pure random noise prediction indeed provides evidence that deterministic noise exists in random exploration (Figure 3).
+
+\paragraph*{S4 Fig.}
+\label{S4_Fig}
+{\bf Deterministic noise can recover known deterministic processes that's intentionally omitted by the model.} In the reduced model where the deterministic effect of uncertainty condition is omitted from the model, deterministic noise is higher compared to the full model that accounts for the effect of uncertainty.
+Random noise remains unchanged between the two models.
+
+\paragraph*{S5 Fig.}
+\label{S5_Fig}
+{\bf Hyperprior recovery.}
+Parameter recovery over the posterior distribution of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices.
+Lighter color shades represent the re-fitted posterior to the simulated choices.
+Our model fitting procedure faithfully recovers the non-stimulus-driven random noise (A, B), but systematically underestimates deterministic noise in both horizons (D, E).
+The horizon differences in random noise is also faithfully recovered (C).
+The horizon differences in deterministic noise is also underestimated but not significant (F).
+
+\paragraph*{S6 Fig.}
+\label{S6_Fig}
+{\bf Frequentist coverage analysis.}
+Parameter recovery over the mean estimates of random and deterministic noise standard deviations $\sigma_{det}$ and $\sigma_{ran}$.
+Solid lines are true posterior used to simulate choices, dashed black line is the mean of the true posterior.
+Histograms represent the mean estimates of the respective parameters in the refitting to the simulated data.
+(A) and (B) are random noise at H = 1 and H = 6, respectively.
+(C) is the random noise differences between horizons.
+(D) and (E) are deterministic noise at H = 1 and H = 6, respectively.
+(F) is the deterministic noise differences between horizons.
+
+\paragraph*{S7 Fig.}
+\label{S7_Fig}
+{\bf Parameter recovery.}
+Parameter recovery over the subject-level means of information bonus, $A$, spatial bias, $b$, random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for horizon 1 (left column) and horizon 6 (right column) games.
+
+\paragraph*{S8 Fig.}
+\label{S8_Fig}
+{\bf Parameter recovery (200 repetitions).}
+Same as Figure \ref{S7_Fig}{S7}, except that the recovered parameters were averaged across 200 repetitions and then compared to the original parameters.
+
+\paragraph*{S9 Fig.}
+\label{S9_Fig}
+{\bf Parameter recovery with 0 random noise or 0 deterministic noise.} Parameter recovery over the posterior of random noise standard deviation, $\sigma_{ran}$, and deterministic noise standard deviation, $\sigma_{det}$, for purely random noise (top row) and purely deterministic noise (bottom row) games.
+
+\paragraph*{S10 Fig.}
+\label{S10_Fig}
+{\bf Parameter recovery on arbitrary combinations of random and deterministic noises.} A.
+Recovered posterior distributions of random noise.
+B.
+Recovered posterior distributions of deterministic noise.
+For both A and B, from the top row to the bottom row, the true noise standard deviation that is used in the simulations go from 0 to 10.
+The y limit of each panel is 4 (+/- 2 from the true value).
+Our model did a relatively good job in recovering all combinations of deterministic and random noises.
+
+\paragraph*{S11 Fig.}
+\label{S11_Fig}
+{\bf Model based analysis with data from all participants (i.e.
+no exclusions)} showing the posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.
+
+\paragraph*{S12 Fig.}
+\label{S12_Fig}
+{\bf Model based analysis from a model that estimates random and deterministic noises separately for [1 3] and [2 2] conditions.} The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, E) and deterministic (C,G) noises are nonzero (A, C, E, G) and change with horizon (B, D, F, H).
+However, random noise has both a greater magnitude overall (A, E) and a greater change with horizon (B, F) than deterministic noise.
+Moreover, both random and deterministic noises have a greater magnitude in [1 3] compared to [2 2] conditions.
+
+\paragraph*{S13 Fig.}
+\label{S13_Fig}
+{\bf Model based analysis from a model that uses variance differences as dI.} The posterior distributions over the group-level mean of the standard deviations of  random and deterministic noise.
+Both random (A, B) and deterministic (C,D) noises are nonzero (A, C) and change with horizon (B, D).
+However, random noise has both a greater magnitude overall (A, C) and a greater change with horizon (B, D) than deterministic noise.
+
+\paragraph*{S14 Fig.}
+\label{S14_Fig}
+{\bf Parameter recovery for shuffled data.} To show that the joint increase of random and deterministic sources of noise is not caused by a limitation of the fitting procedure, we calculated the correlation between ground-truth values of random noise, and best-fitting values of deterministic noise (and vice versa).
+Ground-truth values are shuffled best-fit parameters.
+As expected, ground-truth random values do not correlate with recovered deterministic noises, showing that the increase of deterministic noise with horizon is genuine and not a by-product of increase of random noise, and vice versa.
+
+\paragraph*{S15 Fig.}
+\label{S15_Fig}
+{\bf Model based analysis with reduced models.} Each row is one model.
+These models varied in whether deterministic $\sigma^{det}$ and random noise $\sigma^{ran}$ are present or not and whether either types of noise is dependent on horizon (subscript denotes the dependence on horizon).
+
+\paragraph*{S16 Fig.}
+\label{S16_Fig}
+{\bf Hyperprior recovery of reduced models.} Our model qualitatively captures whether deterministic and random noise are present or not and whether either types of noise is dependent on horizon.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.
+
+\paragraph*{S17 Fig.}
+\label{S17_Fig}
+{\bf Posterior checks for reduced models}
+Model comparison.
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.
+
+\paragraph*{S18 Fig.}
+\label{S18_Fig}
+{\bf Posterior checks for reduced models (maximal likelihood estimation)}
+Model comparison (using maximal likelihood estimation).
+A-D.
+both deterministic and random noise are horizon dependent, E-H.
+only random noise is horizon dependent, I-L.
+only deterministic noise is horizon dependent, M-P.
+neither random nor deterministic noise is horizon dependent, Q-T.
+only deterministic noise is assumed to be present, U-X.
+only random noise is assumed to be present.
+
+\paragraph*{S1 Table.}
+\label{S1_Table}
+{\bf Variants of the model.}
+
+\paragraph*{S1 File.}
+\label{S1_File}
+{\bf Supplemental Materials containing additional analyses.}
 
